### 2020-09-30 doc/draft-doc/main.tex
Add draft document for thesis' topic and content.

%-------------------------------------------------------------------------------
\section{FM Radio}

Die Masterarbeit soll sich mit der Entwicklung eines FM Radios beschäftigen.
Die Implementierung soll in High Level Synthesis (HLS), als auch direkt in VHDL umgesetzt werden. Die jeweiligen Ergebnisse werden anhand verschiedener Metriken verglichen.
Die Implementierung umfasst sowohl Sender, als auch Empfänger.

%-------------------------------------------------------------------------------
\subsection{Idee}

Die Idee hinter dem Thema liegt darin, dass in einem FM Radio sämtliche Konzepte der digitalen Signalverarbeitung stecken. Im Studium HSD/ESD werden diese Inhalte in der Theorie vermittelt, jedoch nie wirklich in Software, bzw. Hardware umgesetzt.\\
\\
Die Arbeit soll sich aus diesem Grund mit digitaler Signalverarbeitung (DSP) in der Praxis beschäftigen. Anhand von FM lässt sich dies aufgrund der relativ geringem Komplexität des Signales vergleichsweise einfach umsetzen, wobei trotzdem noch die meisten DSP Prinzipien vorkommen.\\
\\
Ein HLS Tool kann beispielsweise high-level C\texttt{++} Code in low-level VHDL Code transformieren. Hier ergibt sich die Fragestellung, ob HLS in Zukunft das händische Schreiben von VHDL ersetzen könnte. Diese Frage soll anhand verschiedener Metriken, wie Implementierungsaufwand und -dauer, oder Hardwareverbrauch beantwortet werden.

%-------------------------------------------------------------------------------
\subsection{Hardware}

\textit{\underline{Hinweis:} Hier gilt es noch zu evaluieren, wie und ob HLS in Intel Quartus unterstützt wird. Möglicherweise muss hierfür auf ein Xilinx Development Board umgestiegen werden. Sinngemäß gilt folgender Absatz jedoch für beide Umgebungen gleichermaßen.}\\
\\
Als Hardware Plattform wird vorerst das DE1-SoC Board angestrebt, da dort vieles an benötigter Hardware bereits vorhanden ist, welche auf anderen Development Boards erst zusätzlich angebracht werden müsste.\\
\\
Die wichtigsten Hardware Elemente hierfür sind in folgender Liste angeführt.

\begin{itemize}
  \item User Interface (Switches, Buttons, 7-Segments)
  \item Line In/Out Buchsen
  \item Audio Codec für Line In/Out
  \item externer ADC Chip (Details noch zu evaluieren)
  \item externer DAC Chip (Details noch zu evaluieren)
\end{itemize}

Das Ziel für die Sender- und Empfängerarchitektur ist es, möglichst wenig, oder keine externe Hardware zu brauchen. Die Architektur soll möglichst ohne analoge externe Schaltungen auskommen. Hierbei gilt es zu evaluieren, ob Signale im FM Frequenzbereich im FPGA zum Senden erzeugt werden können, beziehungsweise ob der VGA DAC Chip dafür geeignet ist. Dieselbe Überlegung gilt für den Empfängerteil, beziehungsweise die Eigenschaften des externen ADCs.

\includepicture [0.9] [0] {DE1-SoC Board} {img/de1soc}

%-------------------------------------------------------------------------------
\subsection{Geplantes System}

Im Zuge der Arbeit sollen FM Sender und Empfänger entwickelt werden.
Ein grober Entwurf der geplanten Anwendung ist im folgenden Blockdiagramm dargestellt. Die blau markierten Blöcke sind jene, die in HLS und VHDL verglichen werden sollen.

\includepicture [0.9] [0] {Blockdiagramm der geplanten Anwendung} {img/fmradio-block-diagram}

%-------------------------------------------------------------------------------
\subsection{Matlab}

Alle Prinzipien sollen vorerst in Matlab erarbeitet werden.\\
\\
Das dient einerseits dazu, mit der Materie vertraut zu werden. Andererseits werden so die Architektur und sämtliche Parameter bereits in Simulation festgelegt, was die spätere Implementierung im FPGA deutlich beschleunigen wird.

%-------------------------------------------------------------------------------
\subsection{Theorie}

Die folgende Themen könnten in der theoretischen Ausarbeitung behandelt werden.

\begin{itemize}
  \item FM Modulation
  \item FM De-Modulation
  \item ADC / DAC Sampling (Quantisierung, Rauschen, etc.)
  \item Funktionsweise von HLS (Datenabhängigkeit, Scheduling, etc)
  \item ...
\end{itemize}

%-------------------------------------------------------------------------------
\section{Referenzen / Recherche}

\subsection{Rechtliches}

\url{https://www.conrad.at/de/o/fm-transmitter-1310051.html#faq} (Conrad FAQ)\\
\url{https://de.wikipedia.org/wiki/Minisender} (Wikipedia Minisender)

\subsection{Ähnliche Projekte}
Links to similar projects are listed here.\\
\\
\url{https://marsohod.org/projects/marsohod2/246-fmradio} (Project Documentation)
\\
\url{https://www.youtube.com/watch?v=PDdIK5zrPTY} (Receiver with analog antenna circuit)
\\
\url{https://www.youtube.com/watch?v=9Udbidwqr3s} (DE1-SoC Sender)

### 2020-10-20 sim/gnuradio/stream-live-rtl2832u.sh
add recording script

@@ -1,1 +1,8 @@
+#!/usr/bin/env bash
+################################################################################
+# File        : stream-live-rtl2832u.sh
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Streams a FM radio station from RTL2832u.
+################################################################################
+
 rtl_fm -f 88.8e6 -s 200000 -r 48000 | aplay -r 48000 -f S16_LE

### 2020-10-21 doc/paper/paper.tex
incorporate feedback

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{FM Radio -\\Comparing HLS versus manual VHDL implementation}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Michael Wurm}
\IEEEauthorblockA{University of Applied Sciences Upper Austria\\
Campus Hagenberg\\
Softwarepark 11, 4232 Hagenberg\\
Email: michael.wurm@students.fh-hagenberg.at}
%\and
%\IEEEauthorblockN{Homer Simpson}
%\IEEEauthorblockA{Twentieth Century Fox\\
%Springfield, USA\\
%Email: homer@thesimpsons.com}
%\and
%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
%\IEEEauthorblockA{Starfleet Academy\\
%San Francisco, California 96678--2391\\
%Telephone: (800) 555--1212\\
%Fax: (888) 555--1212}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The aim of this master thesis is to compare an FM radio receiver implementation in HLS versus manual VHDL. Multiple metrics, such as time effort in implementation, hardware utilization in the FPGA, or efficiency are taken into account. As a first step, the receiver structure is simulated using Matlab to evaluate different architectures and techniques to demodulate a FM radio signal. This evaluation is especially targeting a setup that requires the least possible amount of analog hardware. Once a suitable strategy is found, a prototype is implemented in hardware on a FPGA.\\
\end{abstract}

\begin{IEEEkeywords}
  Audio signal processing, FM radio frequency, FPGA, HLS, VHDL, Matlab
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
This demo file is intended to serve as a ``starter file''
for IEEE conference papers produced under \LaTeX\ using
IEEEtran.cls version 1.8b and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
I wish you the best of success.

\hfill mds

\hfill August 26, 2015

\subsection{Subsection Heading Here}
Subsection text here.

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
\begin{figure}[!h]
  \centering
    \includegraphics[width=2.5in]{img/fmradio-block-diagram.png}
  \caption{Block diagram of the entire system.}
  \label{fig_sim}
\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}
The conclusion goes here.

% conference papers do not normally have an appendix

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% that's all folks
\end{document}

### 2020-11-20 doc/thesis/chapters/introduction.tex
add empty HgbThesisEN template

\chapter{Introduction}
\label{cha:Introduction}

### 2020-11-20 doc/thesis/front/abstract.tex
Add HgbThesisEN template. (#1)

\chapter{Abstract}

This should be a 1-page (maximum) summary of your work in English.

### 2020-11-20 doc/thesis/front/kurzfassung.tex
Add HgbThesisEN template. (#1)

\chapter{Kurzfassung}

\begin{german}
An dieser Stelle steht eine Zusammenfassung der Arbeit, Umfang
max.\ 1 Seite. 
...
\end{german}

### 2020-11-20 sim/matlab/record-rtl2832u.sh
add recording script

@@ -1,0 +1,8 @@
+#!/usr/bin/env bash
+################################################################################
+# File        : record-rtl2832u.sh
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Records IQ data from RTL2832u to file.
+################################################################################
+
+rtl_sdr -f 93.3e6 -s 1e6 -n 5e6 fm_record.bin

### 2020-12-04 sim/matlab/fm_demodulate.m
beautify

@@ -14,35 +14,38 @@ addpath(genpath('./helpers/auto-arrange-figs/'));
 fc_oe3 = 98.1e6;
 
 %% Load recording
-f_center = 98.0e6;
-fs = 1e6;
-n_seconds = 10;
+fc    = 98.0e6;
+fs    = 1e6;
+n_sec = 10;
 
 y = loadFile('fm_record.bin');
 
-assert(size(y,1) == n_seconds*fs, ...
-    'Recording is corrupted. Expected %d samples, but the file only contains %d.', n_seconds*fs, size(y(:,1)))
+assert(size(y,1) == n_sec*fs, ...
+    'Recording is corrupted. Expected %d samples, but the file only contains %d.', ...
+    n_sec*fs, size(y(:,1)))
 n_samples = size(y,1);
 
 %% Plot the entire recorded spectrum
 range_s = 0.001;
 
 figure();
-freqz(y(1:round(n_samples*range_s)),1,[-4E6:.01E6:4E6],fs);
+freqz(y(1:round(n_samples*range_s)),1,(-4e6:.01e6:4e6),fs);
+title('Periodic spectrum')
 
 %% Plot closer around the recorded center frequency
-plot_FFT_IQ(y,1,range_s*fs,fs,f_center);
+plot_FFT_IQ(y,1,range_s*fs,fs,fc);
 
 len_section = 200000;
-n_overlap = 1500;
-%figure();
-%spectrogram(y,len_section,n_overlap,[-1.25E6:.02E6:1.25E6],fs,'yaxis');
-%title('Power Spectrum')
+n_overlap   = 1500;
+
+figure();
+spectrogram(y,len_section,n_overlap,(-1.25E6:.02E6:1.25E6),fs,'yaxis');
+title('Power Spectrum')
 
 %% Shift the recording from IF down to baseband
-delta_f = f_center - fc_oe3;
-
-y_shifted = y .* exp(-1j*2*pi*delta_f*[1:1:length(y)]/fs)';
+delta_f = fc - fc_oe3;
+t = (1:length(y))/fs;
+y_shifted = y .* exp(-1j*2*pi*delta_f*t)';
 
 plot_FFT_IQ(y_shifted,1,range_s*fs,fs,fc_oe3);
 
@@ -53,11 +56,9 @@ y_dec = decimate(y_shifted,dec_factor,'fir');
 
[…]

### 2020-12-04 sim/matlab/plot_FFT_IQ.m

@@ -1,4 +1,4 @@
-function plot_FFT_IQ(x,n0,nf,fs,f0,title_of_plot)
+function plot_FFT_IQ(x, n0, nf, fs, f0, title_of_plot)
 
 %
 %  plot_FFT_IQ(x,n0,nf)
@@ -10,43 +10,43 @@ function plot_FFT_IQ(x,n0,nf,fs,f0,title_of_plot)
 %                nf -- block size for transform (signal duration = nf/fs)
 %                fs -- sampling frequency [Hz]
 %                f0 -- center frequency [Hz]
-%                title_of_plot--title of plot (string) (optional)
+%                title_of_plot -- title of plot (string) (optional)
 %
-%-This extracts a segment of x starting at n0, of length nf, and plots the FFT.
+% This extracts a segment of x starting at n0, of length nf, and plots the FFT.
 %
 
-% transform from Hz to MHz
+% Transform from Hz to MHz
 fs = fs/1e6;
 f0 = f0/1e6;
 
+% Extract a small segment of data from signal
+x_segment = x(n0:(n0+nf-1));
+
+% Calculate FFT (normalized to maximum)
+fft_res      = fftshift(fft(x_segment));
+fft_res_norm = 20*log10(abs(fft_res)/max(abs(fft_res)));
+
+% Plot
+Low_freq  = (f0-fs/2); % Lowest frequency to plot
+High_freq = (f0+fs/2); % Highest frequency to plot
+
+N    = length(fft_res_norm);
+freq = (0:1:N-1)*fs/N + Low_freq;
+
 figure();
-x_segment=x(n0:(n0+nf-1)); %extracts a small segment of data from signal
-
-p=fftshift(fft(x_segment)); %find FFT
-z = 20*log10(abs(p)/max(abs(p))); %normalize
-
-Low_freq=(f0-fs/2); %lowest frequency to plot
-High_freq=(f0+fs/2); %highest frequency to plot
-
-N=length(z);
-freq=[0:1:N-1]*(fs)/N+Low_freq;
-
-plot(freq,z);
-axis tight
-xlabel('Frequency [MHz]','FontSize', 14)
-ylabel('Relative amplitude [dB down from max]','FontSize', 14)
-grid on
-set(gcf,'color','white');
-
-if nargin==6
-    title(title_of_plot,'FontSize', 14)
-else
[…]

### 2020-12-04 sim/matlab/FM_IQ_Demod.m
more beautify

@@ -1,14 +1,23 @@
-function [y_FM_demodulated] = FM_IQ_Demod(y)
-%This function demodualtes an FM signal. It is assumed that the FM signal
-%is complex (i.e. an IQ signal) centered at DC and occupies less than 90%
-%of total bandwidth.
+function [y_fm_demod] = FM_IQ_Demod(y)
 
+%
+% y_fm_demod = FM_IQ_Demod(y);
+%
+% This function demodualtes an FM signal. 
+% It is assumed that the FM signal is complex (i.e. an IQ signal), 
+% that is centered at DC and occupies less than 90%
+% of total bandwidth.
+%
 
-b = firls(30,[0 .9],[0 1],'differentiator'); %design differentiater
-d=y./abs(y);%normalize the amplitude (i.e. remove amplitude variations)
-rd=real(d); %real part of normalized siganl.
-id=imag(d); %imaginary part of normalized signal.
-y_FM_demodulated=(rd.*conv(id,b,'same')-id.*conv(rd,b,'same'))./(rd.^2+id.^2); %demodulate!
+% Normalize the amplitude (remove amplitude variations)
+d = y ./ abs(y);
+rd = real(d);
+id = imag(d);
+
+% Design differentiator
+diff = firls(30,[0 .9],[0 1],'differentiator'); 
+
+% Demodulate
+y_fm_demod = (rd.*conv(id,diff,'same') - id.*conv(rd,diff,'same')) ./ (rd.^2 + id.^2); 
 
 end
-

### 2020-12-04 sim/matlab/fm_demodulate.m

@@ -57,7 +57,7 @@ y_dec = decimate(y_shifted,dec_factor,'fir');
 plot_FFT_IQ(y_dec,1,10*range_s*fs_dec,fs_dec,fc_oe3,'Spectrum of decimated signal');
 
 %% Demodulate FM
-[y_fm_demod] = FM_IQ_Demod(y_dec);
+y_fm_demod = FM_IQ_Demod(y_dec);
 plot_FFT_IQ(y_fm_demod,1,20*range_s*fs_dec,fs_dec,0,'Spectrum of demodulated signal');
 
 %% Decimate again for replay on PCs' audio sound card

### 2020-12-04 sim/matlab/loadFile.m

@@ -1,4 +1,6 @@
 function y = loadFile(filename)
+
+%
 %  y = loadFile(filename)
 %
 % reads complex samples from the RTL-SDR file

### 2020-12-04 sim/matlab/plot_FFT_IQ.m

@@ -1,8 +1,6 @@
 function plot_FFT_IQ(x, n0, nf, fs, f0, title_of_plot)
 
 %
-%  plot_FFT_IQ(x,n0,nf)
-%
 %  Plots the FFT of sampled IQ data
 %
 %                x  -- input signal

### 2020-12-04 sim/matlab/fm_demodulate.m
FM Radio Matlab implementation. (#2)

@@ -1,0 +1,74 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%FM-Radio Thesis
+%%Topic : FM-Radio Demodulator
+%%Name  : Michael Wurm
+%%Date  : 11/2020
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Prepare environment
+clear; close all; clc;
+
+addpath(genpath('./helpers/auto-arrange-figs/'));
+
+%% FM station info
+fc_oe3 = 98.1e6;
+
+%% Load recording
+fc    = 98.0e6;
+fs    = 1e6;
+n_sec = 10;
+
+y = loadFile('fm_record.bin');
+
+assert(size(y,1) == n_sec*fs, ...
+    'Recording is corrupted. Expected %d samples, but the file only contains %d.', ...
+    n_sec*fs, size(y(:,1)))
+n_samples = size(y,1);
+
+%% Plot the entire recorded spectrum
+range_s = 0.001;
+
+figure();
+freqz(y(1:round(n_samples*range_s)),1,(-4e6:.01e6:4e6),fs);
+title('Periodic spectrum')
+
+%% Plot closer around the recorded center frequency
+plot_FFT_IQ(y,1,range_s*fs,fs,fc);
+
+len_section = 200000;
+n_overlap   = 1500;
+
+figure();
+spectrogram(y,len_section,n_overlap,(-1.25E6:.02E6:1.25E6),fs,'yaxis');
+title('Power Spectrum')
+
+%% Shift the recording from IF down to baseband
+delta_f = fc - fc_oe3;
+t = (1:length(y))/fs;
+y_shifted = y .* exp(-1j*2*pi*delta_f*t)';
+
+plot_FFT_IQ(y_shifted,1,range_s*fs,fs,fc_oe3);
+
+%% Decimate (low-pass filter + decimate)
+dec_factor = 4;
+fs_dec = fs / dec_factor;
+y_dec = decimate(y_shifted,dec_factor,'fir');
+
+plot_FFT_IQ(y_dec,1,10*range_s*fs_dec,fs_dec,fc_oe3,'Spectrum of decimated signal');
+
+%% Demodulate FM
+y_fm_demod = FM_IQ_Demod(y_dec);
[…]

### 2020-12-04 sim/matlab/loadFile.m

@@ -1,0 +1,17 @@
+function y = loadFile(filename)
+
+%
+%  y = loadFile(filename)
+%
+% reads complex samples from the RTL-SDR file
+%
+
+% Read binary
+fid = fopen(filename,'rb');
+
+% Convert from uint8 to double in signed range
+y = fread(fid,'uint8=>double');
+y = y-127.5;
+
+% Convert interleaved I/Q values to complex values
+y = y(1:2:end) + 1j*y(2:2:end);

### 2020-12-04 sim/matlab/plot_FFT_IQ.m

@@ -1,0 +1,50 @@
+function plot_FFT_IQ(x, n0, nf, fs, f0, title_of_plot)
+
+%
+%  Plots the FFT of sampled IQ data
+%
+%                x  -- input signal
+%                n0 -- first sample (start time = n0/fs)
+%                nf -- block size for transform (signal duration = nf/fs)
+%                fs -- sampling frequency [Hz]
+%                f0 -- center frequency [Hz]
+%                title_of_plot -- title of plot (string) (optional)
+%
+% This extracts a segment of x starting at n0, of length nf, and plots the FFT.
+%
+
+% Transform from Hz to MHz
+fs = fs/1e6;
+f0 = f0/1e6;
+
+% Extract a small segment of data from signal
+x_segment = x(n0:(n0+nf-1));
+
+% Calculate FFT (normalized to maximum)
+fft_res      = fftshift(fft(x_segment));
+fft_res_norm = 20*log10(abs(fft_res)/max(abs(fft_res)));
+
+% Plot
+Low_freq  = (f0-fs/2); % Lowest frequency to plot
+High_freq = (f0+fs/2); % Highest frequency to plot
+
+N    = length(fft_res_norm);
+freq = (0:1:N-1)*fs/N + Low_freq;
+
+figure();
+plot(freq, fft_res_norm);
+xlabel('Frequency [MHz]')
+ylabel('Relative amplitude [dB down from max]')
+axis tight;
+grid on;
+hold on;
+
+% Add vertical line
+xline(f0, 'r','linewidth',2);
+
+% Add title if it was specified
+if nargin==6
+    title(title_of_plot)
+else
+    title({'Spectrum',['Center frequency = ' num2str(f0) ' MHz'] })
+end

### 2020-12-04 sim/matlab/record-rtl2832u.sh

@@ -1,0 +1,25 @@
+#!/usr/bin/env bash
+################################################################################
+# File        : record-rtl2832u.sh
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Records IQ data from RTL2832u to file.
+################################################################################
+
+# Record 10s of IQ signal data.
+# NOTE:
+#   Saalfelden: OE3 98.1 MHz
+
+center_freq=98.0e6
+sample_freq=1000000
+n_seconds=10
+out_file="./fm_record.bin"
+
+n_samples=$(( $sample_freq * $n_seconds))
+echo "Recording $n_samples @ $sample_freq SPS ($n_seconds seconds)."
+
+cmd="rtl_sdr -f $center_freq -s $sample_freq -n $n_samples $out_file"
+
+echo "Using this command:"
+echo -e "$cmd \n"
+
+eval $cmd

### 2020-12-04 doc/thesis/chapters/fm_radio.tex
start table of contents

%%%%%%%%%%%%%%%%%%%%
\chapter{Signal Processing Theory}
\label{cha:TheThesis}

  \section{Up-Downconversion to/from RF}

  \section{Frequency Modulation}
  see literature/Present\_lec6\_AM\_FM.pdf\\
  see lectures out of HSD/ESD

  \section{Algorithms for Digital FM Demodulation}
    see literature/FmDemodulator.pdf (Sect. 3.3)
    see literature/00476180 Digital FM Demodulator for FM, TV, and Wireless.pdf (Sect. II and III)

    \subsection{Baseband Delay Demodulator}
    \subsection{Phase-Adapter Demodulator}
    \subsection{Phase-Locked Loop (PLL)}
    \subsection{Mixed Demodulator}

%%%%%%%%%%%%%%%%%%%%
\chapter{High-Level Synthesis}
  \section{Introduction / State of the Art}
  \section{Functionality (high level code to HDL)}
  \section{Variants}
    \subsection{C++}
    \subsection{SystemC}
  \section{Advantages / Disadvantages}

%%%%%%%%%%%%%%%%%%%%
\chapter{Implementation}
  \section{Matlab Simulation}
  in fixed point, close to hardware level algorithm

  \section{VHDL}
    \subsection{Channel Selection (IF to Channel-BB)}
    \subsection{Phase Detector}
    \subsection{other Elements}

  \section{HLS}
    \subsection{Channel Selection (IF to Channel-BB)}
    \subsection{Phase Detector}
    \subsection{other Elements}

  \section{Comparison of Metrics}
    \subsection{Implementation effort/time}
    \subsection{Hardware Utilization}
    \subsection{Others}

%%%%%%%%%%%%%%%%%%%%
\chapter{Deployment on Hardware}
  \section{Hardware Platform}
    \subsection{RTL2832u Dongle}
    \subsection{ZedBoard}

### 2020-12-04 doc/thesis/chapters/introduction.tex
start table of contents

@@
 \label{cha:Introduction}
+
+Thesis Design Decisions: \\
+\\
+Main focus on system design. \\
+\\
+How to achieve the same result in 3 (4?) different levels of abstraction?
+
+\begin{enumerate}
+  \item GnuRadio (?)
+  \item Matlab
+  \item HLS\\
+      Implementation targeting Xilinx' ZedBoard.\\
+      Optionally: \\
+      Intel as a comparison (synth only, no HW) - "How platform-independent is HLS?"
+  \item VHDL
+\end{enumerate}
+\vspace{.5 cm}
+Compare:
+- Implementation effort (difficulty)
+- Implementation time
+- Usefulness on Target Systems (uC, SoC, RasPi, etc)
+- others?

### 2020-12-07 doc/thesis/chapters/fm_radio.tex
some more table of content

@@
 
-  \section{Up-Downconversion to/from RF}
+  \section{Modulation from Baseband to RF}
+  \section{Modulation from RF to Baseband}
 
-  \section{Frequency Modulation}
+  \section{Frequency Modulation (FM)}
   see literature/Present\_lec6\_AM\_FM.pdf\\
   see lectures out of HSD/ESD
+    \subsection{Mathematical description}
+    \subsection{Frequency Band, Channel allocation/distribution}
 
@@
   \section{Introduction / State of the Art}
-  \section{Functionality (high level code to HDL)}
-  \section{Variants}
+  \section{Functionality (transform high-level code to HDL)}
+  \section{Language Support}
     \subsection{C++}
     \subsection{SystemC}
+  \section{Coding}
+    \subsection{Compiler Directives (\#pragma's)}
+    \subsection{Data types}
+    \subsection{Functions}
+    \subsection{Loops}
+    \subsection{Conditional statements}
   \section{Advantages / Disadvantages}
+    easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV.
+However, it is less optimized, and a developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.
 
@@
 \chapter{Implementation}
-  \section{Matlab Simulation}
+  \section{Matlab/Python Model}
   in fixed point, close to hardware level algorithm
@@
 
+  \section{Common Testbench}
+    \subsection{Architecture (same tb for VHDL and HLS-generated HDL)}
+    \subsection{Framework cocotb, with ghdl compiler}
+        Instantiate both HDL models in the testbench.\\
+        Display a direct comparison of outputs in graphs. This is practical, since the cocotb framework runs in python and graphs can be generated using Python's matplotlib.
+
+  \section{Develop HLS compiler independent code (optional)}
+    \subsection{Challenges}
+     Different compilers use different \#pragmas, etc.\\
+     The \#pragmas need to be within the code, at the appropriate positions, which makes it difficult to write compiler independent code....
+
 %%%%%%%%%%%%%%%%%%%%

### 2020-12-07 doc/thesis/chapters/introduction.tex
some more table of content

@@
 \label{cha:Introduction}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+\section{Motivation}
+
+FM is a common RF signal that's available everywhere.\\
+The demodulated signal is an audio signal, that can be listened to. This is subjectively more attractive than a generic data stream.
+
+%%%%%%%%%%%%%%%%%%%%%%%
+\section{Broadcast Radio}
+Evolution from AM to FM because of certain advantages, etc.\\
+Explain FM usage/existence nowadays (geographical; frequencies; devices; new standard DAB, etc)
+
+\section{Goal}
+put all knowledge that was accumulated during studies (HSD, ESD, mention subjects such as DSP) together in a project.\\
+practical usage of DSP in an FPGA\\
+develop an FM receiver and explain it, so that it can be followed in a tutorial with an affordable budget in hardware.\\
+show multiple ways of how to implement the same thing, in different levels of abstraction. Weigh each ways' field of application, efficiency and applicability.
+
+%%%%%%%%%%%%%%%%%%%%%%%
+% NOTES:
+%%%%%%%%%%%%%%%%%%%%%%%
 
@@
   \item GnuRadio (?)
-  \item Matlab
+  \item Matlab/Python
   \item HLS\\
@@
 \vspace{.5 cm}
+
 Compare:

### 2020-12-11 doc/paper/paper.tex
prepare structure with section headlines

@@
 % Do not put math or special symbols in the title.
-\title{FM Radio -\\Comparing HLS versus manual VHDL implementation}
+\title{Signal Processing for Broadcast FM Radio Demodulation}
 
@@
 \begin{abstract}
-The aim of this master thesis is to compare an FM radio receiver implementation in HLS versus manual VHDL.
+
+  %The aim of this master thesis is to compare an FM radio receiver implementation in HLS versus manual VHDL.
 Multiple metrics, such as time effort in implementation, hardware utilization in the FPGA, or efficiency are taken into account.
@@
 \begin{IEEEkeywords}
-  Audio signal processing, FM radio frequency, FPGA, HLS, VHDL, Matlab
+%  Audio signal processing, FM radio frequency, FPGA, HLS, VHDL, Matlab
+  Audio signal processing, FM radio frequency, Matlab, Simulation
 \end{IEEEkeywords}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-% no \IEEEPARstart
-This demo file is intended to serve as a ``starter file''
-for IEEE conference papers produced under \LaTeX\ using
-IEEEtran.cls version 1.8b and later.
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
-I wish you the best of success.
 
-\hfill mds
+  \subsection{Motivation}
 
-\hfill August 26, 2015
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Signal Processing Theory}
 
-\subsection{Subsection Heading Here}
-Subsection text here.
+  \subsection{Modulation from Baseband to RF}
 
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
+  \subsection{Modulation from RF to Baseband}
+
+  \subsection{Frequency Modulation (FM)}
+
+  \subsection{Frequency Band}
+  %may be used freely, with limited power
+
+  \subsection{Mathematical Description of FM}
+
+  \subsubsection{FM Channel Allocation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{FM Signal Demodulation}
+
+  \subsection{Baseband Delay Demodulator}
+  \subsection{Phase-Adapter Demodulator}
+  \subsection{Phase-Locked Loop (PLL)}
+  \subsection{Mixed Demodulator}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Matlab Simulation}
+
+  \subsection{Data Acquisition}
+    % RTL2832u dongle
+    % Linux driver
+    % IQ interleaved data stream
+
+  \subsection{Demodulation Steps}
+    \subsubsection{Recorded IF spectrum}
+    \subsubsection{Modulate to FM channel center frequency}
+    \subsubsection{Decimation}
+    \subsubsection{Demodulation}
+    \subsubsection{Verification}
+      % is done subjectively only, by listening to the audio stream via a PCs' soundcard.
+
+\section{Conclusion}
+  The conclusion goes here.
 
@@
 %
-\begin{figure}[!h]
-  \centering
-    \includegraphics[width=2.5in]{img/fmradio-block-diagram.png}
-  \caption{Block diagram of the entire system.}
-  \label{fig_sim}
-\end{figure}
+%\begin{figure}[!h]
+%  \centering
+%    \includegraphics[width=2.5in]{img/fmradio-block-diagram.png}
+%  \caption{Block diagram of the entire system.}
+%  \label{fig_sim}
+%\end{figure}
 
@@
 
-\section{Conclusion}
-The conclusion goes here.
-
 % conference papers do not normally have an appendix
@@
 % use section* for acknowledgment
-\section*{Acknowledgment}
+% \section*{Acknowledgment}
 
-The authors would like to thank...
+% The authors would like to thank...
 

### 2020-12-11 doc/paper/paper.tex
wip

@@
 % Do not put math or special symbols in the title.
-\title{Signal Processing for Broadcast FM Radio Demodulation}
+\title{Signal Processing to Demodulate FM Radio}
 
@@
 \section{Introduction}
+
+  This paper describes the necessary parts of signal processing, in order to listen to a radio station that is broadcast in the FM frequency band.
+Frequency modulation (FM) is a widely used standard to provide [...]
 

### 2020-12-11 doc/paper/paper.tex
update

--- 2020-12-11 doc/paper/paper.tex
@@
 \begin{abstract}
+  This paper describes the necessary parts of digital signal processing, in order to be able to listen to a radio station that is broadcast in the FM frequency band.
+  Fundamental knowledge about digital signal processing is presented.
+  The processes of signal modulation up to radio frequency and down to baseband are described, as well as frequency modulation as a modulation type to transmit data.
+  A single FM radio channel in the FM radio band consists of multiple parts in the frequency spectrum, which is an important factor to be able to demodulate an audio stream correctly.
+  This demodulation itself can be done in different algorithms, of which a selection is presented.
+  The entire demodulation chain is implemented in a Matlab simulation script, to demonstrate the functionality of an FM radio receiver.
 
@@
 %  Audio signal processing, FM radio frequency, FPGA, HLS, VHDL, Matlab
-  Audio signal processing, FM radio frequency, Matlab, Simulation
+  Audio signal processing, FM radio frequency, Demodulation, Matlab, Simulation
 \end{IEEEkeywords}
@@
 
-  \subsection{Motivation}
+  Frequency modulation (FM) is a widely used standard to transmit data streams.
+  The probably best known usecase therefore is commercial broadcast radio, where an audio stream is transmitted.
+  Devices to receive these streams are available starting from affordable low prices to the public.
+  Since this is such a common technology which so many people are using on a daily basis, and which is included in all vehicles, a more detailed look into the system is very interesting.
+  Multiple courses in the program Embedded Systems Design at the University of Applied Sciences Upper Austria, Campus Hagenberg, cover the topic of digital signal processing.
+  This paper tries to accumulate all this knowledge in a single paper and give an overview about the FM radio technology.
+  A thesis for the Masters' program will be elaborated based on this topic.
 
@@
 
+  \subsection{Mathematical Description of FM}
+
   \subsection{Frequency Band}
   %may be used freely, with limited power
-
-  \subsection{Mathematical Description of FM}
 

### 2020-12-12 doc/paper/paper.tex
write some stuff

@@
 
-  \subsection{Frequency Modulation (FM)}
+  \subsection{Frequency Modulation}
+    Frequency modulation (FM) is a modulation method, in which the information content is encoded in the frequency of a signal.
+    To be exact, the information is encoded in the frequencies' deviation from the carrier frequency.
+    The technology is widely used in FM broadcast radio.
+    One of its main advantages over amplitude modulation (AM) is the better signal-to-noise ration (SNR).
+    This is achieved, because of the constant amplitude in FM, which makes it less sensitive to additive noise in the transmission channel.
+    Additionally, this fact makes it easier for a receiver to synchronize on the received signal.
 
@@
   \subsection{Frequency Band}
-  %may be used freely, with limited power
+    The frequency band that is used for FM sound broadcasting spans from 87.5 to 108 MHz in Europe. %https://www.itu.int/en/ITU-R/terrestrial/broadcast/Pages/FMTV.aspx
+    This range is located within the so-called Very-High-Frequency (VHF) band.
+    Anyone may use the specified frequency range freely, with a limited transmission power, as long as neighboring receivers are not disturbed in receiving an existing channel.
+    Austria allowed the legal usage in 2006. %CHECK
+    As an example, the German Federal Network Agency specifies the power limit as 50 nW, or -43 dBm, of effective radiated power (ERP).
+    In a practical usecase, this means that a transmitter needs to select a channel that is not already occupied by an official sender.
+    If a free channel is found, the transmitter may start sending a FM signal with the mentioned power limitations.
+    %check Wikipedias' sources for my source references/links
+    %https://www.conrad.at/de/o/fm-transmitter-1310051.html
+    %https://de.wikipedia.org/wiki/FM-Transmitter
+    %https://www.ris.bka.gv.at/GeltendeFassung.wxe?Abfrage=Bundesnormen&Gesetzesnummer=20008807
+    %https://www.rtr.at/medien/was_wir_tun/frequenzverwaltung/Frequenzen.en.html
 
   \subsubsection{FM Channel Allocation}
+    %each channel:
+    % spacing
+    % bandwidth
+    % specific part of spectrum (see Fig.1)
+    % TODO: make Fig1 smaller (cut top/bottom)
+    \begin{figure}[!h]
+      \centering
+        \includegraphics[width=7cm]{img/fm-channel-freqs.png}
+      \caption{Block diagram of the entire system.}
+      \label{fig_sim}
+    \end{figure}
 

### 2020-12-13 doc/paper/paper.tex
update

@@
   \subsection{Frequency Band}
-    The frequency band that is used for FM sound broadcasting spans from 87.5 to 108 MHz in Europe. %https://www.itu.int/en/ITU-R/terrestrial/broadcast/Pages/FMTV.aspx
+    The frequency band that is used for FM sound broadcasting spans from 87.5 to 108 MHz in Austria.
+    This is also true worldwide, except that some countries only partially use the band. %http://search.itu.int/history/HistoryDigitalCollectionDocLibrary/1.43.48.en.101.pdf
     This range is located within the so-called Very-High-Frequency (VHF) band.
@@
     Austria allowed the legal usage in 2006. %CHECK
-    As an example, the German Federal Network Agency specifies the power limit as 50 nW, or -43 dBm, of effective radiated power (ERP).
+    As an example, the German Federal Network Agency specifies the power limit as 50 nW, or -43 dBm, of effective radiated power (ERP). %https://tech.ebu.ch/docs/r/r120.pdf
     In a practical usecase, this means that a transmitter needs to select a channel that is not already occupied by an official sender.
@@
   \subsubsection{FM Channel Allocation}
+    %https://www.electronics-notes.com/articles/audio-video/broadcast-audio/vhf-fm-frequency-modulation-basics.php
+    The entire FM band frequency range divides into multiple channels which can be used.
+    In Europe, the European Telecommunications Standards Institute (ETSI) sets the respective standards for the usage of these frequencies.
+    The main specifications for a single FM broadcasting channels are explained in this section.
+    Each channel may allocate a bandwidth of 200 kHz.
+    Within a channel, the maximum deviation from the center frequency shall not exceed 75 kHz.
+    This leaves a guard band of 25 kHz on either side, to minimize interference with adjacent channels.
+    Center frequencies may be allocated on multiples of 100 kHz.
     %each channel:
@@
       \centering
-        \includegraphics[width=7cm]{img/fm-channel-freqs.png}
-      \caption{Block diagram of the entire system.}
+        \includegraphics[width=7cm]{img/fm-channel-baseband.png}
+      \caption{Allocation of frequencies in an FM channel}
       \label{fig_sim}

### 2020-12-13 doc/paper/paper.tex
update

@@
     Center frequencies may be allocated on multiples of 100 kHz.
-    %each channel:
-    % spacing
-    % bandwidth
-    % specific part of spectrum (see Fig.1)
-    % TODO: make Fig1 smaller (cut top/bottom)
+    A sum of the left and right audio channel is located starting from 30 Hz, up to 15 kHz.
+    The lower limit at 30 Hz is to prevent the transmission of a direct current (DC) part, which would require a large amount of power in transmission.
+This effect is one of the downsides of amplitude modulation (AM), where the carrier is located at the channels' center frequeny and requires a high transmission power.
+% TODO: check this fact
+    The upper limit at 15 kHz is chosen, to maintain a sufficient spacing to the first subcarrier.
+    This subcarrier is allocated at an offset of 19 kHz from the carrier.
+    It is also called the 'pilot tone', since it is a continuous signal.
+    The pilot tone is used to synchronize the left and right channel of an audio signal, to generate stereo audio.
+    The center frequency for stereo audio is located at 38 kHz, which is an integer multiple of 19 kHz for practical reasons.
+    A signal that is constructed of the difference between left and right audio channel is modulated on the 38 kHz subcarrier.
+    The bandwidth for the difference-signal spans 15 kHz on either side.
+
     \begin{figure}[!h]

### 2020-12-13 doc/paper/paper.tex
update

@@
 
-  \subsubsection{FM Channel Allocation}
+  \subsection{Channel Frequency Spectrum}
     %https://www.electronics-notes.com/articles/audio-video/broadcast-audio/vhf-fm-frequency-modulation-basics.php
@@
     Center frequencies may be allocated on multiples of 100 kHz.
+    \ref{sec_fm_sig_demod} shows the upper half-bands' frequency spectrum of a single channel.
     A sum of the left and right audio channel is located starting from 30 Hz, up to 15 kHz.
+    The summation of left and right audio channel resembles the mono audio stream.
     The lower limit at 30 Hz is to prevent the transmission of a direct current (DC) part, which would require a large amount of power in transmission.
@@
     The center frequency for stereo audio is located at 38 kHz, which is an integer multiple of 19 kHz for practical reasons.
-    A signal that is constructed of the difference between left and right audio channel is modulated on the 38 kHz subcarrier.
-    The bandwidth for the difference-signal spans 15 kHz on either side.
+    However, the 38 kHz carrier is suppressed and thus not visible in the received spectrum.
+    It can still be recovered, since it is phase coherent with the 19 kHz pilot tone per definition.
+    A signal that is constructed by the difference between left and right audio channel is modulated on a 38 kHz subcarrier.
+    The bandwidth for this difference-signal spans 15 kHz on either side of the subcarrier.
+    This signal is used to generate a stereo audio signal, in combination with the mono signal.
+    This process is explained into more detail in chapter \ref{sec_fm_sig_demod}.
+    Considering all these parts in the spectrum, there is still bandwidth free to use, up to the maximum channel bandwitdh 100 kHz in one sideband.
+    Because of that, additional services were added to the audio transmission, to provide additional data services and information.
+    Services that were implemented are the Data Radio Channel (DARC), which is mostly used in Japan and the USA, the Subsidary Communication Authorization (SCA) and the Radio Data System (RDS).
+    Out of these, RDS is the most significant service in Europe.
+    It is used to transmit additional information about the channel, such as the radio stations' name, currently playing songs' title or traffic information.
 
@@
       \caption{Allocation of frequencies in an FM channel}
-      \label{fig_sim}
+      \label{fig_channel_baseband_freqs}
     \end{figure}
@@
 \section{FM Signal Demodulation}
-
+\label{sec_fm_sig_demod}
+    % mono:   just LP on 15 kHz
+    % stereo: sum (L+R) + (L-R) = 2 L, (L+R) - (L-R) = 2R
   \subsection{Baseband Delay Demodulator}

### 2020-12-15 doc/paper/paper.tex
write math. description of FM

@@
 %
-%\usepackage{cite}
+\usepackage{cite}
 % cite.sty was written by Donald Arseneau
@@
 %
-%\usepackage{url}
+\usepackage{url}
 % url.sty was written by Donald Arseneau. It provides better support for
@@
   \subsection{Mathematical Description of FM}
+    The signal of an FM baseband signal can be described in the following formulas.
+    As already mentioned in earlier in this paper, FM encodes the information in the frequency. 
+    Since the frequency has a direct relationship to the angle, FM belongs to the group of angle modulated signals.
+
+    The instantanious frequency $f_i$ can be described as 
+    \begin{equation}
+      f_i = f_c + \Delta f \cdot m(t)
+    \end{equation}
+    where $f_c$ is the carrier frequency, $\Delta f$ is the maximum frequency deviation and $m(t)$ is the information or message signal that is to be transmitted.
+    Simply looking at this equation, the frequency deviation, which is also called the swing, varies in the range between $f_c \pm \Delta f$.
+
+    Considering the relationship between angular frequency and angle and after some simple substitutions, which will not be described into detail here, the equation for a generic frequency modulated wave is the following.
+    \begin{equation}
+      s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int m(t) dt \Big)
+    \end{equation}
+    Where $A_c$ is the amplitude of the resulting FM signal and $k_f$ is the frequency sensitivity factor.
+
+    An example application of FM is transmitting audio streams. 
+    An audio stream can be described as a cosine wave.
+    If the information signal that is to be transmitted is a cosine wave in the form of 
+    \begin{equation}
+      m(t) = A_m\ cos(2 \pi f_m t)
+    \end{equation}
+    the final equation for an FM signal transforms into the following formula.
+    \begin{equation}
+      y(t) = A_c\ cos \Big(2 \pi f_c t + \beta\ sin(2 \pi f_m t)\Big )
+    \end{equation}
+
+    The variable $\beta$ is called the modulation index,
+    \begin{equation}
+      \beta = \frac{\Delta f}{f_m} = \frac{k_f A_m}{f_m}
+    \end{equation}
+    where $f_m$ is the highest existing frequency, or the bandwidth, of the information signal. 
+    Several formulas, as well as their derivations were described according to \cite{FMMaths1}\cite{FMMaths2}.
 
@@
 
+  \bibitem{FMMaths1}
+    \url{https://fas.org/man/dod-101/navy/docs/es310/FM.htm}, opened on Dec. 12th, 2020
+
+    \bibitem{FMMaths2}
+    \url{https://www.tutorialspoint.com/analog_communication/analog_communication_angle_modulation.htm}
+
   \end{thebibliography}

### 2020-12-15 doc/paper/paper.tex
cite and newlines

@@
   Devices to receive these streams are available starting from affordable low prices to the public.
-  Since this is such a common technology which so many people are using on a daily basis, and which is included in all vehicles, a more detailed look into the system is very interesting.
+  Since this is such a common technology which so many people are using on a daily basis, and which is included in virtually every automobile, a more detailed look into this signal transmission system is very interesting.
+
   Multiple courses in the program Embedded Systems Design at the University of Applied Sciences Upper Austria, Campus Hagenberg, cover the topic of digital signal processing.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Signal Processing Theory}
+%\section{Signal Processing Theory}
+%
+%  A comprehensive set of signal processing techniques is required, in order to %achieve a data transmission with frequency modulation.
+%  In traditional commercial senders and receivers...
+%
+%  \subsection{Modulation from Baseband to RF}
+%
+%
+%  \subsection{Modulation from RF to Baseband}
+%  \subsection{Frequency Modulation}
 
-  \subsection{Modulation from Baseband to RF}
-
-  \subsection{Modulation from RF to Baseband}
-
-  \subsection{Frequency Modulation}
+  \section{Frequency Modulation}
     Frequency modulation (FM) is a modulation method, in which the information content is encoded in the frequency of a signal.
@@
     The technology is widely used in FM broadcast radio.
-    One of its main advantages over amplitude modulation (AM) is the better signal-to-noise ration (SNR).
+    One of its main advantages over amplitude modulation (AM) is the higher signal-to-noise ration (SNR).
     This is achieved, because of the constant amplitude in FM, which makes it less sensitive to additive noise in the transmission channel.
-    Additionally, this fact makes it easier for a receiver to synchronize on the received signal.
+    Additionally, this fact makes it easier for a receiver to synchronize on the received signal \cite{FM_Maths_Info_1}.
 
-  \subsection{Mathematical Description of FM}
+  \subsection{Mathematical Description}
     The signal of an FM baseband signal can be described in the following formulas.
@@
     where $f_m$ is the highest existing frequency, or the bandwidth, of the information signal.
-    Several formulas, as well as their derivations were described according to \cite{FMMaths1}\cite{FMMaths2}.
+    Several formulas, as well as their derivations were described according to \cite{FM_Maths_Info_1}\cite{FMMaths2}.
 
   \subsection{Frequency Band}
-    The frequency band that is used for FM sound broadcasting spans from 87.5 to 108 MHz in Austria.
-    This is also true worldwide, except that some countries only partially use the band. %http://search.itu.int/history/HistoryDigitalCollectionDocLibrary/1.43.48.en.101.pdf
-    This range is located within the so-called Very-High-Frequency (VHF) band.
-    Anyone may use the specified frequency range freely, with a limited transmission power, as long as neighboring receivers are not disturbed in receiving an existing channel.
-    Austria allowed the legal usage in 2006. %CHECK
-    As an example, the German Federal Network Agency specifies the power limit as 50 nW, or -43 dBm, of effective radiated power (ERP). %https://tech.ebu.ch/docs/r/r120.pdf
-    In a practical usecase, this means that a transmitter needs to select a channel that is not already occupied by an official sender.
-    If a free channel is found, the transmitter may start sending a FM signal with the mentioned power limitations.
-    %check Wikipedias' sources for my source references/links
-    %https://www.conrad.at/de/o/fm-transmitter-1310051.html
-    %https://de.wikipedia.org/wiki/FM-Transmitter
+    The frequency band that is used for FM broadcasting is defined worldwide and spans from 87.5 to 108 MHz.
+    However, some countries only partially use the band \cite{itu_regulations}.
+    The range is located within the so-called Very-High-Frequency (VHF) band.
+    In terms of usage, the band is open to anyone.
+    This means, that a transmitter may use the specified frequency range freely.
+    Austria allowed the legal usage in 2006.
+    However, the transmission power needs to be limited, so that neighboring receivers are not disturbed in receiving any existing channels.
+    As an example, the German Federal Network Agency specifies the power limit as 50 nW, or -43 dBm, of effective radiated power (ERP).
+    In a practical usecase, this means that a transmitter needs to select a channel, or frequency range that is not already occupied by an official sender.
+    If a free range is found, the transmitter may start sending an FM signal with the mentioned power limitations \cite{ebu_fm_modulators_in_europe_power_regulations}.
+
     %https://www.ris.bka.gv.at/GeltendeFassung.wxe?Abfrage=Bundesnormen&Gesetzesnummer=20008807
@@
   \subsection{Channel Frequency Spectrum}
-    %https://www.electronics-notes.com/articles/audio-video/broadcast-audio/vhf-fm-frequency-modulation-basics.php
     The entire FM band frequency range divides into multiple channels which can be used.
@@
     The main specifications for a single FM broadcasting channels are explained in this section.
+
     Each channel may allocate a bandwidth of 200 kHz.
@@
     Center frequencies may be allocated on multiples of 100 kHz.
-    \ref{sec_fm_sig_demod} shows the upper half-bands' frequency spectrum of a single channel.
+
+    Fig.\ref{fig_channel_baseband_freqs} shows the upper half-bands' frequency spectrum of a single channel.
     A sum of the left and right audio channel is located starting from 30 Hz, up to 15 kHz.
@@
     Considering all these parts in the spectrum, there is still bandwidth free to use, up to the maximum channel bandwitdh 100 kHz in one sideband.
-    Because of that, additional services were added to the audio transmission, to provide additional data services and information.
+    Because of that, additional services were added to the pure audio transmission, to provide additional data services and information.
     Services that were implemented are the Data Radio Channel (DARC), which is mostly used in Japan and the USA, the Subsidary Communication Authorization (SCA) and the Radio Data System (RDS).
@@
     It is used to transmit additional information about the channel, such as the radio stations' name, currently playing songs' title or traffic information.
+    The information about the frequency spectrum was found in \cite{fm_broadcast_tutorial_and_basics}.
 
@@
         \includegraphics[width=7cm]{img/fm-channel-baseband.png}
-      \caption{Allocation of frequencies in an FM channel}
+      \caption{Allocation of frequencies in an FM channel \cite{fig_channel_freqs}}
       \label{fig_channel_baseband_freqs}
@@
 \begin{thebibliography}{1}
+  \bibitem{itu_regulations}
+    International Telecommunication Union (ITU), \emph{Radio Regulation Articles, Edition of 2020}, Vol. I-EA5
 
-  \bibitem{IEEEhowto:kopka}
-    H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-      0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
+  \bibitem{ebu_fm_modulators_in_europe_power_regulations}
+    European Broadcasting Union (EBU), \emph{Low-power FM modulators in Europe}, EBU-R120-2007, \url{https://tech.ebu.ch/docs/r/r120.pdf}, pg.5
 
-  \bibitem{FMMaths1}
+  \bibitem{fm_broadcast_tutorial_and_basics}
+    Electronics Notes, \emph{Broadcast VHF FM Tutorial \& Basics}, \url{https://www.electronics-notes.com/articles/audio-video/broadcast-audio/vhf-fm-frequency-modulation-basics.php}, opened Dec.
+10th, 2020
+
+  \bibitem{FM_Maths_Info_1}
     \url{https://fas.org/man/dod-101/navy/docs/es310/FM.htm}, opened on Dec. 12th, 2020
@@
   \bibitem{FMMaths2}
-    \url{https://www.tutorialspoint.com/analog_communication/analog_communication_angle_modulation.htm}
+    \url{https://www.tutorialspoint.com/analog_communication/analog_communication_angle_modulation.htm}, opened on Dec. 12th, 2020
 
+  \bibitem{fig_channel_freqs}
+    Fig.\ref{fig_channel_baseband_freqs}, Baseband Signal of an FM channel, \url{https://de.wikipedia.org/wiki/UKW-Rundfunk#Technische_Details}, opened on Dec. 11th, 2020
 \end{thebibliography}

### 2020-12-16 doc/paper/paper.tex
notes
start FM demod

@@
 \label{sec_fm_sig_demod}
+  An FM signal that is received by an antenna needs to be demodulated, in order to decode its actual data content.
+  The radio-frequency antenna signal usually needs to be amplified, bandpass-filtered, and down-converted to baseband, using subsampling and quadrature-mixing, or similar strategies.
+  This part of the signal processing chain is assumed to be working correctly and is out of the scope of this paper.
+  The FM signal that is evaluated here, is assumed to be a quadrature-mixed signal in baseband, which means that inphase and quadrature (I/Q-) signals are available.
+  %TODO: show downconversion path (maybe as a figure?)
+  %https://www.veron.nl/wp-content/uploads/2014/01/FmDemodulator.pdf
+
   % mono:   just LP on 15 kHz
   % stereo: sum (L+R) + (L-R) = 2 L, (L+R) - (L-R) = 2R
-  \subsection{Baseband Delay Demodulator}
-  \subsection{Phase-Adapter Demodulator}
+  % page 30: https://www.hs-schmalkalden.de/fileadmin/portal/Dokumente/Fakult%C3%A4t_ET/Personal/Roppel/Buch/Analoge_Modulationsverfahren.pdf
+
+  \subsection{FM-AM-Demodulator}
+    convert FM to AM
+    %https://www.tutorialspoint.com/analog_communication/analog_communication_fm_demodulators.htm
+    %sect 3.3 : https://www.hs-schmalkalden.de/fileadmin/portal/Dokumente/Fakult%C3%A4t_ET/Personal/Roppel/Buch/Analoge_Modulationsverfahren.pdf
+
   \subsection{Phase-Locked Loop (PLL)}
-  \subsection{Mixed Demodulator}
 

### 2020-12-16 doc/paper/paper.tex
wip

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{FM Signal Demodulation}
+\section{Demodulation of a Generic FM Signal}
 \label{sec_fm_sig_demod}
@@
   The FM signal that is evaluated here, is assumed to be a quadrature-mixed signal in baseband, which means that inphase and quadrature (I/Q-) signals are available.
+
   %TODO: show downconversion path (maybe as a figure?)
@@
 
-  % mono:   just LP on 15 kHz
-  % stereo: sum (L+R) + (L-R) = 2 L, (L+R) - (L-R) = 2R
-  % page 30: https://www.hs-schmalkalden.de/fileadmin/portal/Dokumente/Fakult%C3%A4t_ET/Personal/Roppel/Buch/Analoge_Modulationsverfahren.pdf
+  In the following sections, two digital FM demodulator variants are described.
 
-  \subsection{FM-AM-Demodulator}
-    convert FM to AM
+  \subsection{Frequency Discriminator, Slope detector}
+    A frequency discriminator can be defined as a black box, with an output that is direclty proportional to the frequency of the input FM signal.
+
+    different strategies in the black box.
+    Here, the slope detector (differnetiator??) is used.
+
+    This type of demodulator converts the FM signal into an AM signal, which still has an FM content.
+
   %https://www.tutorialspoint.com/analog_communication/analog_communication_fm_demodulators.htm
@@
   \subsection{Phase-Locked Loop (PLL)}
+
+\section{Demodulation of Broadcast FM}
+  % mono:   just LP on 15 kHz
+  % stereo: sum (L+R) + (L-R) = 2 L, (L+R) - (L-R) = 2R
+
+  % page 30: https://www.hs-schmalkalden.de/fileadmin/portal/Dokumente/Fakult%C3%A4t_ET/Personal/Roppel/Buch/Analoge_Modulationsverfahren.pdf
 

### 2020-12-16 doc/paper/paper.tex
pics; wip

--- 2020-12-16 doc/paper/paper.tex
@@
       s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int m(t) dt \Big)
+      \label{equ_fm_func}
     \end{equation}
@@
         \includegraphics[width=7cm]{img/fm-channel-baseband.png}
-      \caption{Allocation of frequencies in an FM channel \cite{fig_channel_freqs}}
+      \caption{Allocation of frequencies in an FM channel \cite{ref_fig_channel_freqs}}
       \label{fig_channel_baseband_freqs}
@@
 
-  % mono:   just LP on 15 kHz
-  % stereo: sum (L+R) + (L-R) = 2 L, (L+R) - (L-R) = 2R
-  % page 30: https://www.hs-schmalkalden.de/fileadmin/portal/Dokumente/Fakult%C3%A4t_ET/Personal/Roppel/Buch/Analoge_Modulationsverfahren.pdf
+  In the following sections, two digital FM demodulator variants are described.
 
-  \subsection{FM-AM-Demodulator}
-    convert FM to AM
+  \subsection{Frequency Discriminator}
+    % TODO: check if this is also called a Slope Detector
+    %       --> check the frequency response of a differentiator for that
+    A frequency discriminator can be defined as a black box component, with an output that is directly proportional to the frequency of the input FM signal.
+    Different strategies can be implemented in this black box component, the discriminator.
+    Here, a differentiator with a subsequent envelope detector is used, which is shown in Fig.\ref{fig_freq_discriminator_bd}.
+
+    \begin{figure}[!h]
+      \centering
+        \includegraphics[width=8.7cm]{img/discrimination_method_bd2.png}
+      \caption{Frequency Discriminator Block Diagram \cite{ref_fig_freq_discriminator}}
+      \label{fig_freq_discriminator_bd}
+    \end{figure}
+
+    This type of demodulator converts the FM signal into an AM signal.
+    However, this signal still contains an FM content.
+    The transformation that happens here can be expressed in a formula, by simply differentiating equation \ref{equ_fm_func}.
+    This operation, with simple algebraic transformations, results in
+    \begin{equation}
+      \frac{d s(t)}{dt} = A_c 2 \pi \Big[f_c + k_f m(t) \Big] sin \Big(2 \pi f_c t + 2 \pi k_f \int m(t) dt -pi \Big)
+    \end{equation}
+    It can clearly be seen, that the signal now contains an AM and an FM content.
+    The FM part is described by the sine function.
+    More importantly the AM part, which resembles the envelope of the signal, is described by the term within the square brackets.
+    The resulting signal in time-domain is shown in the top left diagram in Fig.\ref{fig_time_domain_envelope_detect}.
+    This signal needs to be fed into an envelope detector, as shown in Fig.\ref{fig_freq_discriminator_bd}.
+    It extracts the envelope by removing the high frequency portion.
+    The method implemented here is one of the most simple ones and is usually referred to as 'Asynchronous Half-Wave Envelope Detector'\cite{ref_envelope_detector}.
+    It consists of a thresholding unit to remove negative values and a conventional lowpass filter.
+    In analog implementations, the thresholding unit is a diode.
+    The lowpass' cutoff frequency needs to be adapted to the envelope signals' maximum frequency, since it needs to be able to follow the signal, but also sufficiently smooth out the rectified signal.
+    Different methods may be implemented for the envelope detector.
+    For example, an 'Asynchronous Full-Wave Envelope Detector' can be used.
+    Therefore, the mentioned method only needs to swap the thresholding unit with an unit that calculates an absolut value.
+    In that architecture, a full-wave rectification is performed on the signal, which significantly improves the envelope detection accuracy \cite{ref_envelope_detector}.
+    The envelope is followed more exactly, because of the higher power that is available in the signal, since the entire signal is taken into account, and not only the positive half, as in the previous method.
+
+    %TODO: talk about normalization
+
+    \begin{figure}[!h]
+      \centering
+        \includegraphics[width=8.8cm]{img/envelope-detect-time-domain.jpg}
+      \caption{Time-Domain Signals in Envelope Detection \cite{ref_time_domain_envelope_detect}}
+      \label{fig_time_domain_envelope_detect}
+    \end{figure}
+
   %https://www.tutorialspoint.com/analog_communication/analog_communication_fm_demodulators.htm
@@
 
-  \bibitem{fig_channel_freqs}
-    Fig.\ref{fig_channel_baseband_freqs}, Baseband Signal of an FM channel, \url{https://de.wikipedia.org/wiki/UKW-Rundfunk#Technische_Details}, opened on Dec. 11th, 2020
+  \bibitem{ref_fig_channel_freqs}
+    Figure: Baseband Signal of an FM channel, \url{https://de.wikipedia.org/wiki/UKW-Rundfunk#Technische_Details}, opened on Dec. 11th, 2020
+
+  \bibitem{ref_fig_freq_discriminator}
+    Figure: Frequency Discriminator Block Diagram, \url{https://www.tutorialspoint.com/analog_communication/analog_communication_fm_demodulators.htm}, opened on Dec. 16th, 2020
+
+  \bibitem{ref_envelope_detector}
+    Rick Lyons, \emph{Digital Envelope Detection: The Good, the Bad, and the Ugly}, April 3, 2016, \url{https://www.dsprelated.com/showarticle/938.php}, opened on Dec. 16th, 2020
+
+  \bibitem{ref_time_domain_envelope_detect}
+    C.Roppel, \emph{Analoge Modulationsverfahren und Rundfunktechnik} Begleitmaterial zum Buch 'Grundlagen der digitalen Kommunikationstechnik, Übertragungstechnik - Signalverarbeitung - Netze', \url{https://www.hs-schmalkalden.de/fileadmin/portal/Dokumente/Fakult%C3%A4t_ET/Personal/Roppel/Buch/Analoge_Modulationsverfahren.pdf}, opened on Dec.
+16th, 2020
 \end{thebibliography}

### 2020-12-16 doc/paper/paper.tex
done with FM demod

@@
 %
-%\usepackage{amsmath}
+\usepackage{amsmath}
 % A popular package from the American Mathematical Society that provides
@@
   In the following sections, two digital FM demodulator variants are described.
+  In general, a FM demodulator serves the purpose to transform FM to AM, so that after the FM demodulator regular AM signal processing techniques can be applied.
+  % TODO: add img/discrimination_method_bd2.png here
 
   \subsection{Frequency Discriminator}
-    % TODO: check if this is also called a Slope Detector
-    %       --> check the frequency response of a differentiator for that
-    A frequency discriminator can be defined as a black box component, with an output that is directly proportional to the frequency of the input FM signal.
+    A frequency discriminator can be defined as a black box component, that generates an output that is directly proportional to the frequency of the input FM signal.
     Different strategies can be implemented in this black box component, the discriminator.
@@
       \centering
-        \includegraphics[width=8.7cm]{img/discrimination_method_bd2.png}
+        \includegraphics[width=7cm]{img/discrimination_method_bd1.jpg}
       \caption{Frequency Discriminator Block Diagram \cite{ref_fig_freq_discriminator}}
@@
     \begin{equation}
-      \frac{d s(t)}{dt} = A_c 2 \pi \Big[f_c + k_f m(t) \Big] sin \Big(2 \pi f_c t + 2 \pi k_f \int m(t) dt -pi \Big)
+      \frac{d s(t)}{dt} = A_c 2 \pi \Big[f_c + k_f m(t) \Big] sin \Big(\omega_c t + 2 \pi k_f \int m(t) dt -pi \Big)
     \end{equation}
+
+    A differentiation in hardware is simply performed by substracting two consecutive samples, like
+    \begin{equation}
+      \frac{d s(t)}{dt} = s(t) - s(t-\Delta t)
+      \label{equ_differentiator_two_samples}
+    \end{equation}
+    where $\Delta t$ is the inverse of the sample frequency.
+
     It can clearly be seen, that the signal now contains an AM and an FM content.
@@
 
-    %TODO: talk about normalization
-
     \begin{figure}[!h]
@@
 
-  %https://www.tutorialspoint.com/analog_communication/analog_communication_fm_demodulators.htm
-  %sect 3.3 : https://www.hs-schmalkalden.de/fileadmin/portal/Dokumente/Fakult%C3%A4t_ET/Personal/Roppel/Buch/Analoge_Modulationsverfahren.pdf
+    An important requirement for the differentiator is, that the input signal is of a constant amplitude \cite{ref_schnyder_haller}.
+    The transmitted signal is subject to random additive noise on the transmission channel.
+    The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Equation \ref{equ_differentiator_two_samples}.
+    To achieve this, the complex signal $s(t)$ can be normalized to a value of one.
+    \begin{equation}
+      s_{norm} = \frac{s}{|s|} = \frac{A(n)\ e^{j\phi_{FM}(n)}}{|{A(n)\ e^{j\phi_{FM}(n)}|}} = e^{j\phi_{FM}(n)} = |1|
+    \end{equation}
 
-  \subsection{Phase-Locked Loop (PLL)}
+  \subsection{Phase-Locked Loop}
+    A Phase-Locked Loop (PLL) can also be used to transform an FM signal to AM.
+    A PLL is a feedback loop that is usually used to generate an output signal that has a fixed phase reference to its input.
+    In the case of FM demodulation, the loop filter is configured to be able to follow the frequency variations on the input, which is directly related to the encoded information.
+    The output of the PLL directly delivers an AM signal, which corresponds to the transmitted information. \cite{ref_schnyder_haller}
+
+    Due to the limited space in this paper, FM demodulation with a PLL is not described into more detail.
 
@@
 16th, 2020
+
+  \bibitem{ref_schnyder_haller}
+    F.Schnyder, C.Haller, \emph{Implementation of FM Demodulator
+    Algorithms on a High Performance Digital Signal Processor}, 2002, \url{https://www.veron.nl/wp-content/uploads/2014/01/FmDemodulator.pdf}
 \end{thebibliography}

### 2020-12-16 doc/paper/paper.tex
demod mono, stereo

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Demodulation of Broadcast FM}
-  % mono:   just LP on 15 kHz
-  % stereo: sum (L+R) + (L-R) = 2 L, (L+R) - (L-R) = 2R
+  This chapter is describing the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
+  The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec_fm_sig_demod} is already done correctly.
+  A frequency spectrum of an FM channel is shown in Fig.\ref{fig_channel_baseband_freqs} above.
+
+  \subsection{Mono}
+    The first part of the channel frequency spectrum is the summation signal of the left and right audio channel.
+    A mono receiver can thus simply replay the so-called multiplex signal, which is generated by the FM demodulation described in chapter \ref{sec_fm_sig_demod}.
+The 19 kHz pilot tone will not be audible by most people, because it is outside of the range of a human ear.
+Besides that, any used speaker may also be unable to produce a frequency in this range.
+This explanation applies to several higher frequency components in the spectrum as well.
+    In order to be able to decimate the signal to a lower samplerate, for example to store it in a file with 44.1 kHz samplerate, a lowpass needs to be inserted before the decimation and output.
+    The lowpass' cutoff frequency needs to be configured after 15 kHz.
+    The filter needs to reach a sufficient attenuation before the pilot tone at 19 kHz.
+
+  \subsection{Stereo}
+    Demodulating an FM channel to recover a stereo audio signal, requires a more sophisticated approach.
+    In the channel frequency spectrum, there are two main parts of the audio signal, the left-right sum and -difference signals.
+    To combine these to a stereo signal, the following equations can be applied.
+    \begin{equation*}
+      L = (L+R) + (L-R) = (2)L
+    \end{equation*}
+    \begin{equation*}
+      R = (L+R) - (L-R) = (2)R
+    \end{equation*}
 
@@
 \section{Matlab Simulation}
+  % TODO: insert entire block diagram of demodulation
+  %       HF => abs() => discriminator => AM demod => audio
 

### 2020-12-16 doc/paper/paper.tex
almost entirely done with it

@@
     Different strategies can be implemented in this black box component, the discriminator.
-    Here, a differentiator with a subsequent envelope detector is used, which is shown in Fig.\ref{fig_freq_discriminator_bd}.
+    Here, a differentiator with a subsequent envelope detector is used, which is shown in Fig.\ref{fig_bd_freq_discriminator}.
 
@@
         \includegraphics[width=7cm]{img/discrimination_method_bd1.jpg}
-      \caption{Frequency Discriminator Block Diagram \cite{ref_fig_freq_discriminator}}
-      \label{fig_freq_discriminator_bd}
+      \caption{Frequency discriminator block diagram \cite{ref_fig_freq_discriminator}}
+      \label{fig_bd_freq_discriminator}
     \end{figure}
@@
     However, this signal still contains an FM content.
-    The transformation that happens here can be expressed in a formula, by simply differentiating equation \ref{equ_fm_func}.
+    The transformation that happens here can be expressed in a formula, by simply differentiating Eq.\ref{equ_fm_func}.
     This operation, with simple algebraic transformations, results in
@@
     The resulting signal in time-domain is shown in the top left diagram in Fig.\ref{fig_time_domain_envelope_detect}.
-    This signal needs to be fed into an envelope detector, as shown in Fig.\ref{fig_freq_discriminator_bd}.
+    This signal needs to be fed into an envelope detector, as shown in Fig.\ref{fig_bd_freq_discriminator}.
     It extracts the envelope by removing the high frequency portion.
@@
         \includegraphics[width=8.8cm]{img/envelope-detect-time-domain.jpg}
-      \caption{Time-Domain Signals in Envelope Detection \cite{ref_time_domain_envelope_detect}}
+      \caption{Time-domain signals in envelope detection \cite{ref_roppel}}
       \label{fig_time_domain_envelope_detect}
@@
     The transmitted signal is subject to random additive noise on the transmission channel.
-    The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Equation \ref{equ_differentiator_two_samples}.
+    The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Eq.\ref{equ_differentiator_two_samples}.
     To achieve this, the complex signal $s(t)$ can be normalized to a value of one.
@@
   \subsection{Mono}
+  \label{demod_mono}
     The first part of the channel frequency spectrum is the summation signal of the left and right audio channel.
@@
   \subsection{Stereo}
-    Demodulating an FM channel to recover a stereo audio signal, requires a more sophisticated approach.
-    In the channel frequency spectrum, there are two main parts of the audio signal, the left-right sum and -difference signals.
+    Demodulating an FM channel to recover a stereo audio signal requires a more sophisticated approach.
+    In the channel frequency spectrum, there are two main parts of the audio signal - the sum and difference of the left and right channel audio signals.
     To combine these to a stereo signal, the following equations can be applied.
@@
     \end{equation*}
-    \begin{equation*}
+    \begin{equation}
       R = (L+R) - (L-R) = (2)R
-    \end{equation*}
+      \label{equ_stereo_from_sum_diff}
+    \end{equation}
 
-  % page 30: https://www.hs-schmalkalden.de/fileadmin/portal/Dokumente/Fakult%C3%A4t_ET/Personal/Roppel/Buch/Analoge_Modulationsverfahren.pdf
+    The block diagram in Fig.\ref{fig_bd_stereo_demod} illustrates the signal processing chain for stereo audio.
+In the block diagram, $x(t)$ represents the multiplex signal, which comes from the FM demodulator.
+Signals $x_l(t)$ and $x_r(t)$ describe the left and right audio signal, respectively.
+
+    \begin{figure}[!h]
+      \centering
+        \includegraphics[width=8.8cm]{img/fm-demod-stereo-audio.png}
+      \caption{Block diagram of FM stereo audio demodulation \cite{ref_roppel}}
+      \label{fig_bd_stereo_demod}
+    \end{figure}
+
+    The central branch selects the pilot tone at 19 kHz with a bandpass filter.
+    This bandpass needs to have a frequency response that is sharp enough to have a sufficient attenuation at $\pm$4 kHz, since this is where the summation and difference signals frequencies are allocated.
+    Afterwards, the pilot tone frequency is doubled via a PLL to achieve a 38 kHz to later demodulate the difference signal phase-coherently.
+    The PLL lock indicator can be re-used as an indicator, whether a pilot tone is existent.
+    In the upper branch, a bandpass filter selects the difference signal with ranges from 23 to 53 kHz.
+    It is then modulated to baseband by the previously generated 38 kHz subcarrier.
+    Subsequently, a lowpass limits the signal to the audio signal bandwidth of 15 kHz.
+    The lower branch lowpass-filters the summation signal.
+    The rightmost part in the diagram performs the combination of summation and difference signals, according to the Eq.\ref{equ_stereo_from_sum_diff}.
 
@@
 \section{Matlab Simulation}
+  Matlab is used to simulate an FM demodulator.
+  The objective is to demodulate a previously recorded data stream, so that one can listen to a mono audio signal.
+  Several steps that are required therefore are described briefly in the following sections.
+
   % TODO: insert entire block diagram of demodulation
@@
   \subsection{Data Acquisition}
-    % RTL2832u dongle
-    % Linux driver
-    % IQ interleaved data stream
+    A data stream is recorded, using a cheap RTL-SDR dongle.
+    The device can be plugged into USB on a host computer.
+    It features an RTL2832U chipset.
+    In combination with a host software, this basically assembles a software defined radio (SDR).
+    The recorded data stream is stored as a file with interleaved IQ baseband data.
 
   \subsection{Demodulation Steps}
-    \subsubsection{Recorded IF spectrum}
+    The SDR is used to record a data stream with a sampling frequency of 1 MHz, at a center frequency of 98.0 MHz. A dataset of 10 seconds is recorded.
+    The aim is to listen to the Austrian radio station OE3, which is broadcasted at a center frequency of 98.1 MHz in the region of Saalfelden.
+
+    \subsubsection{Recorded spectrum}
+      The file was recorded at a center frequency $f_c$ of 98 MHz, while the desired station is located at 98.1 MHz.
+      This means, that the recorded spectrum is allocated at an intermediate frequency (IF) of 0.1 MHz.
+
     \subsubsection{Modulate to FM channel center frequency}
+      Because the signal needs to be in baseband, to be able to demodulate it, the spectrum needs to be shifted by 0.1 MHz.
+      This is achieved by modulating with a complex signal with exactly that frequency.
+
     \subsubsection{Decimation}
+      The signal still has a sample rate of 1 MHz, which is an overkill for the application, since a channels' bandwith only ranges up to 100 kHz.
+      Keeping in mind the Nyquist theorem, the sample rate therefore needs to be at least 200 kHz.
+      Thus, the signal is decimated by an arbitrary factor of 4 % TODO: more? see Nyquist.. work this out in script
+      , which still leaves a rate of x M/kHz.
+
     \subsubsection{Demodulation}
+      Demodulation is done by normalizing the signal amplitude to one.
+      A FIR filter is used as a differentiator.
+      Envelope detection is omitted for now, for simplicity and because of the mentioned effects in section \ref{demod_mono}.
     \subsubsection{Verification}
-      % is done subjectively only, by listening to the audio stream via a PCs' soundcard.
+      Verification of correct functionality is only done subjectively, by listening to the demodulated audio signal via the PCs' soundcard.
+      The OE3 stations' signal can be heard clearly.
 
 \section{Conclusion}
-  The conclusion goes here.
+  FM is a well-known technology, that is established in anyones daily life.
+  Looking at it from a signal processing perspective is an interesting task, since it requires many principles of digital signal processing, but is still comparably simple to demodulate.
 
@@
 
-  \bibitem{ref_time_domain_envelope_detect}
+  \bibitem{ref_roppel}
     C.Roppel, \emph{Analoge Modulationsverfahren und Rundfunktechnik} Begleitmaterial zum Buch 'Grundlagen der digitalen Kommunikationstechnik, Übertragungstechnik - Signalverarbeitung - Netze', \url{https://www.hs-schmalkalden.de/fileadmin/portal/Dokumente/Fakult%C3%A4t_ET/Personal/Roppel/Buch/Analoge_Modulationsverfahren.pdf}, opened on Dec.

### 2020-12-17 sim/matlab/fm_demodulate.m
adapt matlab code

@@ -35,11 +35,11 @@ title('Periodic spectrum')
 %% Plot closer around the recorded center frequency
 plot_FFT_IQ(y,1,range_s*fs,fs,fc);
 
-len_section = 200000;
-n_overlap   = 1500;
+len_section = round( length(y)/100   ); % calculate 100 sections
+n_overlap   = round( len_section/100 ); % overlap 1%
 
 figure();
-spectrogram(y,len_section,n_overlap,(-1.25E6:.02E6:1.25E6),fs,'yaxis');
+spectrogram(y,len_section,n_overlap,(-1.25e6:.02e6:1.25e6),fs,'yaxis');
 title('Power Spectrum')
 
 %% Shift the recording from IF down to baseband
@@ -49,24 +49,17 @@ y_shifted = y .* exp(-1j*2*pi*delta_f*t)';
 
 plot_FFT_IQ(y_shifted,1,range_s*fs,fs,fc_oe3);
 
-%% Decimate (low-pass filter + decimate)
-dec_factor = 4;
-fs_dec = fs / dec_factor;
-y_dec = decimate(y_shifted,dec_factor,'fir');
-
-plot_FFT_IQ(y_dec,1,10*range_s*fs_dec,fs_dec,fc_oe3,'Spectrum of decimated signal');
-
 %% Demodulate FM
-y_fm_demod = FM_IQ_Demod(y_dec);
-plot_FFT_IQ(y_fm_demod,1,20*range_s*fs_dec,fs_dec,0,'Spectrum of demodulated signal');
+y_fm_demod = FM_IQ_Demod(y_shifted);
+plot_FFT_IQ(y_fm_demod,1,20*range_s*fs,fs,0,'Spectrum of demodulated signal');
 
 %% Decimate again for replay on PCs' audio sound card
-dec_factor_audio = 10;
-fs_dec_audio = fs_dec / dec_factor_audio;
+dec_factor_audio = 20;
+fs_dec_audio = fs / dec_factor_audio;
 
 y_fm_demod_dec = decimate(y_fm_demod,dec_factor_audio,'fir');
 
-plot_FFT_IQ(y_fm_demod_dec,1,10*range_s*fs_dec_audio,fs_dec_audio,0,'Spectrum of demod+dec signal');
+plot_FFT_IQ(y_fm_demod_dec,1,20*range_s*fs_dec_audio,fs_dec_audio,0,'Spectrum of demod+dec signal');
 
 sound(y_fm_demod_dec, fs_dec_audio);
 

### 2020-12-17 doc/paper/paper.tex
adapt text for matlab

@@
     \subsubsection{Decimation}
-      The signal still has a sample rate of 1 MHz, which is an overkill for the application, since a channels' bandwith only ranges up to 100 kHz.
-      Keeping in mind the Nyquist theorem, the sample rate therefore needs to be at least 200 kHz.
-      Thus, the signal is decimated by an arbitrary factor of 4 % TODO: more? see Nyquist.. work this out in script
-      , which still leaves a rate of x M/kHz.
-
+      The signal still has a sample rate of 1 MHz at this point.
+      Keeping in mind the Nyquist theorem, and a single channels' bandwith of only up to 100 kHz, the sample rate therefore needs to be at least 200 kHz.
+      However, the final hardware implementation is targeting an FPGA, where processing of samples at 1 MHz is easily achievable. Therefore, a decimation is not done at this point.
     \subsubsection{Demodulation}

### 2020-12-17 doc/paper/paper.tex
final adaptions

@@
   This paper describes the necessary parts of digital signal processing, in order to be able to listen to a radio station that is broadcast in the FM frequency band.
-  Fundamental knowledge about digital signal processing is presented.
-  The processes of signal modulation up to radio frequency and down to baseband are described, as well as frequency modulation as a modulation type to transmit data.
+  Fundamental knowledge about frequency modulation and its application in audio broadcast is presented.
   A single FM radio channel in the FM radio band consists of multiple parts in the frequency spectrum, which is an important factor to be able to demodulate an audio stream correctly.
-  This demodulation itself can be done in different algorithms, of which a selection is presented.
+  This demodulation itself can be done in different algorithms, of which a selection is introduced.
   The entire demodulation chain is implemented in a Matlab simulation script, to demonstrate the functionality of an FM radio receiver.
@@
 %  Audio signal processing, FM radio frequency, FPGA, HLS, VHDL, Matlab
-  Audio signal processing, FM radio frequency, Demodulation, Matlab, Simulation
+  Audio signal processing, FM radio frequency, FM demodulation, Digital FM receiver, Matlab, Simulation
 \end{IEEEkeywords}
@@
   Devices to receive these streams are available starting from affordable low prices to the public.
-  Since this is such a common technology which so many people are using on a daily basis, and which is included in virtually every automobile, a more detailed look into this signal transmission system is very interesting.
+  Since this is such a common technology that so many people are using on a daily basis - and which is included in virtually every automobile - a more detailed look into this signal transmission system is very interesting.
 
-  Multiple courses in the program Embedded Systems Design at the University of Applied Sciences Upper Austria, Campus Hagenberg, cover the topic of digital signal processing.
-  This paper tries to accumulate all this knowledge in a single paper and give an overview about the FM radio technology.
-  A thesis for the Masters' program will be elaborated based on this topic.
+%  Multiple courses in the program Embedded Systems Design at the University of Applied Sciences Upper Austria, Campus Hagenberg, cover the topic of digital signal processing.
+%  This paper tries to accumulate all this knowledge in a single paper and give an overview about the FM radio technology.
+  This paper tries to give an overview about the receivers' side of the FM radio technology.
+  A Masters' thesis will be elaborated based on this topic.
 
@@
     This is achieved, because of the constant amplitude in FM, which makes it less sensitive to additive noise in the transmission channel.
-    Additionally, this fact makes it easier for a receiver to synchronize on the received signal \cite{FM_Maths_Info_1}.
+    Additionally, this fact makes it easier for a receiver to synchronize on the received signal \cite{ref_FM_Maths_Info_1}.
 
   \subsection{Mathematical Description}
-    The signal of an FM baseband signal can be described in the following formulas.
-    As already mentioned in earlier in this paper, FM encodes the information in the frequency.
-    Since the frequency has a direct relationship to the angle, FM belongs to the group of angle modulated signals.
+    The signal of an FM baseband signal can be described with the formulas presented in this section.
+    FM encodes the information content in its instantanious frequency, and since frequency has a direct relationship to an angle, FM belongs to the group of angle modulated signals.
 
@@
 
-    Considering the relationship between angular frequency and angle and after some simple substitutions, which will not be described into detail here, the equation for a generic frequency modulated wave is the following.
+    Considering the relationship between angular frequency and angle and after some simple substitutions, which will not be described into detail here, the equation for a generic frequency modulated wave is the following
     \begin{equation}
@@
     \end{equation}
-    Where $A_c$ is the amplitude of the resulting FM signal and $k_f$ is the frequency sensitivity factor.
+    where $A_c$ is the amplitude of the resulting FM signal and $k_f$ is the frequency sensitivity factor.
 
@@
     \end{equation}
-
     The variable $\beta$ is called the modulation index,
@@
     \end{equation}
-    where $f_m$ is the highest existing frequency, or the bandwidth, of the information signal.
-    Several formulas, as well as their derivations were described according to \cite{FM_Maths_Info_1}\cite{FMMaths2}.
+    where $f_m$ is the highest existing frequency, or the bandwidth of the information signal.
+    Several formulas, as well as their derivations are described according to \cite{ref_FM_Maths_Info_1}\cite{ref_FMMaths2}.
 
@@
     The frequency band that is used for FM broadcasting is defined worldwide and spans from 87.5 to 108 MHz.
-    However, some countries only partially use the band \cite{itu_regulations}.
+    However, some countries only partially use the band \cite{ref_itu_regulations}.
     The range is located within the so-called Very-High-Frequency (VHF) band.
-    In terms of usage, the band is open to anyone.
+    In terms of usage the band is open to the public.
     This means, that a transmitter may use the specified frequency range freely.
-    Austria allowed the legal usage in 2006.
+    Austria allowed the legal usage in 2006 \cite{ref_austria_rundfunkgesetz_2006}.
     However, the transmission power needs to be limited, so that neighboring receivers are not disturbed in receiving any existing channels.
-    As an example, the German Federal Network Agency specifies the power limit as 50 nW, or -43 dBm, of effective radiated power (ERP).
+    The European Commision specifies the power limit as 50 nW of effective radiated power (ERP) \cite{ref_eu_commision_radio_spectrum}.
     In a practical usecase, this means that a transmitter needs to select a channel, or frequency range that is not already occupied by an official sender.
-    If a free range is found, the transmitter may start sending an FM signal with the mentioned power limitations \cite{ebu_fm_modulators_in_europe_power_regulations}.
+    If a free range is found, the transmitter may start sending an FM signal with the mentioned power limitations \cite{ref_ebu_fm_modulators_in_europe_power_regulations}.
 
     %https://www.ris.bka.gv.at/GeltendeFassung.wxe?Abfrage=Bundesnormen&Gesetzesnummer=20008807
-    %https://www.rtr.at/medien/was_wir_tun/frequenzverwaltung/Frequenzen.en.html
 
   \subsection{Channel Frequency Spectrum}
+   % More info on modulation index: https://fas.org/man/dod-101/navy/docs/es310/FM.htm
     The entire FM band frequency range divides into multiple channels which can be used.
     In Europe, the European Telecommunications Standards Institute (ETSI) sets the respective standards for the usage of these frequencies.
-    The main specifications for a single FM broadcasting channels are explained in this section.
+    The main specifications for a single FM broadcasting channel are explained in this section.
 
@@
     Fig.\ref{fig_channel_baseband_freqs} shows the upper half-bands' frequency spectrum of a single channel.
-    A sum of the left and right audio channel is located starting from 30 Hz, up to 15 kHz.
-    The summation of left and right audio channel resembles the mono audio stream.
+    A sum of the left and right audio channel is located between 30 Hz and 15 kHz.
+    This summation of left and right audio channel resembles the mono audio stream.
     The lower limit at 30 Hz is to prevent the transmission of a direct current (DC) part, which would require a large amount of power in transmission.
@@
     It is also called the 'pilot tone', since it is a continuous signal.
-    The pilot tone is used to synchronize the left and right channel of an audio signal, to generate stereo audio.
-    The center frequency for stereo audio is located at 38 kHz, which is an integer multiple of 19 kHz for practical reasons.
+    The pilot tone is used to be able to demodulate the left and right channel of an audio signal correctly, to generate stereo audio.
+    A signal that is constructed by the difference between left and right audio channel is modulated on a 38 kHz subcarrier.
+    This subcarrier is an integer multiple of 19 kHz for practical reasons.
     However, the 38 kHz carrier is suppressed and thus not visible in the received spectrum.
     It can still be recovered, since it is phase coherent with the 19 kHz pilot tone per definition.
-    A signal that is constructed by the difference between left and right audio channel is modulated on a 38 kHz subcarrier.
     The bandwidth for this difference-signal spans 15 kHz on either side of the subcarrier.
-    This signal is used to generate a stereo audio signal, in combination with the mono signal.
+    It is used to generate a stereo audio signal, in combination with the mono signal.
     This process is explained into more detail in chapter \ref{sec_fm_sig_demod}.
-    Considering all these parts in the spectrum, there is still bandwidth free to use, up to the maximum channel bandwitdh 100 kHz in one sideband.
+
+    Considering all these parts in the spectrum, there is still bandwidth free to use, up to the maximum channel bandwitdh of 100 kHz in one sideband.
     Because of that, additional services were added to the pure audio transmission, to provide additional data services and information.
-    Services that were implemented are the Data Radio Channel (DARC), which is mostly used in Japan and the USA, the Subsidary Communication Authorization (SCA) and the Radio Data System (RDS).
+    Services that were implemented are the Data Radio Channel (DARC), which is mostly used in Japan and the USA, the Subsidary Communication Authorization (SCA) and the Radio Data System (RDS) \cite{ref_rohde_u_schwarz}.
     Out of these, RDS is the most significant service in Europe.
     It is used to transmit additional information about the channel, such as the radio stations' name, currently playing songs' title or traffic information.
-    The information about the frequency spectrum was found in \cite{fm_broadcast_tutorial_and_basics}.
+    The information about the frequency spectrum was found in \cite{ref_fm_broadcast_tutorial_and_basics}.
 
@@
   This part of the signal processing chain is assumed to be working correctly and is out of the scope of this paper.
-  The FM signal that is evaluated here, is assumed to be a quadrature-mixed signal in baseband, which means that inphase and quadrature (I/Q-) signals are available.
+  The FM signal that is evaluated here is assumed to be a quadrature-mixed signal in baseband, which means that inphase and quadrature (I/Q) signals are available.
 
@@
   In the following sections, two digital FM demodulator variants are described.
-  In general, a FM demodulator serves the purpose to transform FM to AM, so that after the FM demodulator regular AM signal processing techniques can be applied.
+  In general, an FM demodulator serves the purpose to transform FM to AM, so that AM signal processing techniques can be applied afterwards.
   % TODO: add img/discrimination_method_bd2.png here
@@
 
-    This type of demodulator converts the FM signal into an AM signal.
-    However, this signal still contains an FM content.
+    This type of demodulator converts the FM signal into an AM signal, but still leaves an FM content.
     The transformation that happens here can be expressed in a formula, by simply differentiating Eq.\ref{equ_fm_func}.
@@
 
-    It can clearly be seen, that the signal now contains an AM and an FM content.
+    It can clearly be seen, that the differentiated signal consist of an AM and an FM content.
     The FM part is described by the sine function.
@@
     The resulting signal in time-domain is shown in the top left diagram in Fig.\ref{fig_time_domain_envelope_detect}.
-    This signal needs to be fed into an envelope detector, as shown in Fig.\ref{fig_bd_freq_discriminator}.
+    This signal needs to be fed into an envelope detector, as seen in Fig.\ref{fig_bd_freq_discriminator}.
     It extracts the envelope by removing the high frequency portion.
@@
     For example, an 'Asynchronous Full-Wave Envelope Detector' can be used.
-    Therefore, the mentioned method only needs to swap the thresholding unit with an unit that calculates an absolut value.
+    Therefore, the previous method only needs to swap the thresholding unit with an unit that calculates an absolut value.
     In that architecture, a full-wave rectification is performed on the signal, which significantly improves the envelope detection accuracy \cite{ref_envelope_detector}.
@@
     The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Eq.\ref{equ_differentiator_two_samples}.
-    To achieve this, the complex signal $s(t)$ can be normalized to a value of one.
+    To achieve a constant amplitude, the complex signal $s(t)$ can be normalized to a value of one.
     \begin{equation}
@@
 This explanation applies to several higher frequency components in the spectrum as well.
-    In order to be able to decimate the signal to a lower samplerate, for example to store it in a file with 44.1 kHz samplerate, a lowpass needs to be inserted before the decimation and output.
-    The lowpass' cutoff frequency needs to be configured after 15 kHz.
-    The filter needs to reach a sufficient attenuation before the pilot tone at 19 kHz.
+    In case the multiplex signal should be stored it in a file with 44.1 kHz samplerate, a lowpass with a cutoff frequency of 15 kHz and sufficient attenutation before 19 kHz needs to be inserted before the decimation and file output.
 
@@
     Demodulating an FM channel to recover a stereo audio signal requires a more sophisticated approach.
-    In the channel frequency spectrum, there are two main parts of the audio signal - the sum and difference of the left and right channel audio signals.
+    In the channel frequency spectrum, there are two main parts of the audio signal - the sum and difference of the left and right audio channel signals.
     To combine these to a stereo signal, the following equations can be applied.
@@
     The block diagram in Fig.\ref{fig_bd_stereo_demod} illustrates the signal processing chain for stereo audio.
-In the block diagram, $x(t)$ represents the multiplex signal, which comes from the FM demodulator.
+In the block diagram, $x(t)$ represents the multiplex signal from the FM demodulator.
 Signals $x_l(t)$ and $x_r(t)$ describe the left and right audio signal, respectively.
@@
     The lower branch lowpass-filters the summation signal.
-    The rightmost part in the diagram performs the combination of summation and difference signals, according to the Eq.\ref{equ_stereo_from_sum_diff}.
+    The rightmost part in the diagram performs the combination of summation and difference signals, according to Eq.\ref{equ_stereo_from_sum_diff}.
 
@@
 \section{Matlab Simulation}
-  Matlab is used to simulate an FM demodulator.
+  Matlab is used to simulate an FM receiver.
   The objective is to demodulate a previously recorded data stream, so that one can listen to a mono audio signal.
@@
   \subsection{Data Acquisition}
-    A data stream is recorded, using a cheap RTL-SDR dongle.
+    A data stream is recorded, using a low priced RTL-SDR dongle.
     The device can be plugged into USB on a host computer.
@@
     \subsubsection{Recorded spectrum}
-      The file was recorded at a center frequency $f_c$ of 98 MHz, while the desired station is located at 98.1 MHz.
+      The file was recorded at a center frequency of 98 MHz, while the desired station is located at 98.1 MHz.
       This means, that the recorded spectrum is allocated at an intermediate frequency (IF) of 0.1 MHz.
@@
     \subsubsection{Modulate to FM channel center frequency}
-      Because the signal needs to be in baseband, to be able to demodulate it, the spectrum needs to be shifted by 0.1 MHz.
-      This is achieved by modulating with a complex signal with exactly that frequency.
+      Because the signal needs to be in baseband to be able to demodulate it, the spectrum needs to be shifted by 0.1 MHz.
+      This is achieved by modulating with a complex signal with that exact frequency.
 
@@
       The signal still has a sample rate of 1 MHz at this point.
-      Keeping in mind the Nyquist theorem, and a single channels' bandwith of only up to 100 kHz, the sample rate therefore needs to be at least 200 kHz.
+      Keeping in mind the Nyquist theorem, and a single channels' bandwith of only up to 100 kHz, the sample rate therefore only needs to be at least 200 kHz.
       However, the final hardware implementation is targeting an FPGA, where processing of samples at 1 MHz is easily achievable. Therefore, a decimation is not done at this point.
@@
       Demodulation is done by normalizing the signal amplitude to one.
-      A FIR filter is used as a differentiator.
-      Envelope detection is omitted for now, for simplicity and because of the mentioned effects in section \ref{demod_mono}.
+      Then, an FIR filter is used as a differentiator.
+      Envelope detection is omitted for now, for simplicity and because of the mentioned 'natural' filter effects in section \ref{demod_mono}.
     \subsubsection{Verification}
-      Verification of correct functionality is only done subjectively, by listening to the demodulated audio signal via the PCs' soundcard.
+      The correct functionality is only verified subjectively, by listening to the demodulated audio signal via the PCs' soundcard.
       The OE3 stations' signal can be heard clearly.
@@
 \section{Conclusion}
-  FM is a well-known technology, that is established in anyones daily life.
-  Looking at it from a signal processing perspective is an interesting task, since it requires many principles of digital signal processing, but is still comparably simple to demodulate.
+  FM is a well-known technology, that is established in peoples' daily lifes.
+  Looking at it from a signal processing perspective is an interesting task, since it requires many principles of digital signal processing, but is still comparably simple to demodulate and implement.
 
@@
 \begin{thebibliography}{1}
-  \bibitem{itu_regulations}
+  \bibitem{ref_itu_regulations}
     International Telecommunication Union (ITU), \emph{Radio Regulation Articles, Edition of 2020}, Vol. I-EA5
 
-  \bibitem{ebu_fm_modulators_in_europe_power_regulations}
+  \bibitem{ref_ebu_fm_modulators_in_europe_power_regulations}
     European Broadcasting Union (EBU), \emph{Low-power FM modulators in Europe}, EBU-R120-2007, \url{https://tech.ebu.ch/docs/r/r120.pdf}, pg.5
 
-  \bibitem{fm_broadcast_tutorial_and_basics}
+  \bibitem{ref_eu_commision_radio_spectrum}
+    European Commision, \emph{Commission Decision on harmonisation of the radio spectrum for use by short-range devices}, 13.08.2019, 2006/771/EC, \url{https://eur-lex.europa.eu/eli/dec/2006/771(2)/2019-08-13}, pg.11
+
+  \bibitem{ref_austria_rundfunkgesetz_2006}
+    Bundesministerium für Verkehr, Innovation und Technologie, \emph{Frequenznutzungsverordnung 2013 - FNV 2013}, BGBl.
+II Nr.
+63/2014, Anlage 2, \url{https://www.ris.bka.gv.at/Dokumente/Bundesnormen/NOR40219692/II_390_2016_Anlage_2.pdf}, pg.99
+
+  \bibitem{ref_rohde_u_schwarz}
+  Rohde\&Schwarz, \emph{Messungen an FM-Sendern für Abnahme, Inbetriebnahme oder Wartung}, \url{https://cdn.rohde-schwarz.com/pws/dl_downloads/dl_application/application_notes/7bm105/7BM105_0D.pdf}, opened Dec.
+12th, 2020
+
+  \bibitem{ref_fm_broadcast_tutorial_and_basics}
     Electronics Notes, \emph{Broadcast VHF FM Tutorial \& Basics}, \url{https://www.electronics-notes.com/articles/audio-video/broadcast-audio/vhf-fm-frequency-modulation-basics.php}, opened Dec.
@@
 
-  \bibitem{FM_Maths_Info_1}
+  \bibitem{ref_FM_Maths_Info_1}
     \url{https://fas.org/man/dod-101/navy/docs/es310/FM.htm}, opened on Dec. 12th, 2020
 
-  \bibitem{FMMaths2}
+  \bibitem{ref_FMMaths2}
     \url{https://www.tutorialspoint.com/analog_communication/analog_communication_angle_modulation.htm}, opened on Dec. 12th, 2020

### 2020-12-17 doc/paper/paper.tex
Write a paper about FM demodulation. (#4)

@@
   Rohde\&Schwarz, \emph{Messungen an FM-Sendern für Abnahme, Inbetriebnahme oder Wartung}, \url{https://cdn.rohde-schwarz.com/pws/dl_downloads/dl_application/application_notes/7bm105/7BM105_0D.pdf}, opened Dec.
-12th, 2020
+17th, 2020
 
@@
   \bibitem{ref_FMMaths2}
-    \url{https://www.tutorialspoint.com/analog_communication/analog_communication_angle_modulation.htm}, opened on Dec. 12th, 2020
+    TutorialsPoint, \emph{Analog Communication - Angle Modulation}, \url{https://www.tutorialspoint.com/analog_communication/analog_communication_angle_modulation.htm}, opened on Dec. 12th, 2020
 

### 2021-02-15 sim/gnuradio/fm_radio_receiver.py
GnuRadio FM demodulator. (#3)

@@ -1,0 +1,296 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+
+#
+# SPDX-License-Identifier: GPL-3.0
+#
+# GNU Radio Python Flow Graph
+# Title: fm_radio_receiver
+# GNU Radio version: 3.8.2.0
+
+from distutils.version import StrictVersion
+
+if __name__ == '__main__':
+    import ctypes
+    import sys
+    if sys.platform.startswith('linux'):
+        try:
+            x11 = ctypes.cdll.LoadLibrary('libX11.so')
+            x11.XInitThreads()
+        except:
+            print("Warning: failed to XInitThreads()")
+
+from PyQt5 import Qt
+from gnuradio import qtgui
+from gnuradio.filter import firdes
+import sip
+from gnuradio import analog
+from gnuradio import audio
+from gnuradio import blocks
+from gnuradio import filter
+from gnuradio import gr
+import sys
+import signal
+from argparse import ArgumentParser
+from gnuradio.eng_arg import eng_float, intx
+from gnuradio import eng_notation
+from gnuradio.qtgui import Range, RangeWidget
+import osmosdr
+import time
+
+from gnuradio import qtgui
+
+class fm_radio_receiver(gr.top_block, Qt.QWidget):
+
+    def __init__(self):
+        gr.top_block.__init__(self, "fm_radio_receiver")
+        Qt.QWidget.__init__(self)
+        self.setWindowTitle("fm_radio_receiver")
+        qtgui.util.check_set_qss()
+        try:
+            self.setWindowIcon(Qt.QIcon.fromTheme('gnuradio-grc'))
+        except:
+            pass
+        self.top_scroll_layout = Qt.QVBoxLayout()
+        self.setLayout(self.top_scroll_layout)
+        self.top_scroll = Qt.QScrollArea()
+        self.top_scroll.setFrameStyle(Qt.QFrame.NoFrame)
+        self.top_scroll_layout.addWidget(self.top_scroll)
+        self.top_scroll.setWidgetResizable(True)
+        self.top_widget = Qt.QWidget()
+        self.top_scroll.setWidget(self.top_widget)
+        self.top_layout = Qt.QVBoxLayout(self.top_widget)
+        self.top_grid_layout = Qt.QGridLayout()
+        self.top_layout.addLayout(self.top_grid_layout)
+
[…]

### 2021-02-15 sim/gnuradio/fm_radio_receiver_2.py

@@ -1,0 +1,293 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+
+#
+# SPDX-License-Identifier: GPL-3.0
+#
+# GNU Radio Python Flow Graph
+# Title: fm_radio_receiver_2
+# Author: mike
+# GNU Radio version: 3.8.2.0
+
+from distutils.version import StrictVersion
+
+if __name__ == '__main__':
+    import ctypes
+    import sys
+    if sys.platform.startswith('linux'):
+        try:
+            x11 = ctypes.cdll.LoadLibrary('libX11.so')
+            x11.XInitThreads()
+        except:
+            print("Warning: failed to XInitThreads()")
+
+from PyQt5 import Qt
+from gnuradio import qtgui
+from gnuradio.filter import firdes
+import sip
+from gnuradio import analog
+from gnuradio import audio
+from gnuradio import blocks
+from gnuradio import filter
+from gnuradio import gr
+import sys
+import signal
+from argparse import ArgumentParser
+from gnuradio.eng_arg import eng_float, intx
+from gnuradio import eng_notation
+from gnuradio.qtgui import Range, RangeWidget
+import osmosdr
+import time
+
+from gnuradio import qtgui
+
+class fm_radio_receiver_2(gr.top_block, Qt.QWidget):
+
+    def __init__(self):
+        gr.top_block.__init__(self, "fm_radio_receiver_2")
+        Qt.QWidget.__init__(self)
+        self.setWindowTitle("fm_radio_receiver_2")
+        qtgui.util.check_set_qss()
+        try:
+            self.setWindowIcon(Qt.QIcon.fromTheme('gnuradio-grc'))
+        except:
+            pass
+        self.top_scroll_layout = Qt.QVBoxLayout()
+        self.setLayout(self.top_scroll_layout)
+        self.top_scroll = Qt.QScrollArea()
+        self.top_scroll.setFrameStyle(Qt.QFrame.NoFrame)
+        self.top_scroll_layout.addWidget(self.top_scroll)
+        self.top_scroll.setWidgetResizable(True)
+        self.top_widget = Qt.QWidget()
+        self.top_scroll.setWidget(self.top_widget)
+        self.top_layout = Qt.QVBoxLayout(self.top_widget)
+        self.top_grid_layout = Qt.QGridLayout()
+        self.top_layout.addLayout(self.top_grid_layout)
+
[…]

### 2021-02-15 sim/gnuradio/stream-live-rtl2832u.sh

@@ -1,0 +1,8 @@
+#!/usr/bin/env bash
+################################################################################
+# File        : stream-live-rtl2832u.sh
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Streams a FM radio station from RTL2832u.
+################################################################################
+
+rtl_fm -f 88.8e6 -s 200000 -r 48000 | aplay -r 48000 -f S16_LE

### 2021-02-15 sim/matlab/fm_transceiver.m
move files; start implementation transmitter

@@ -1,0 +1,68 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% File   : fm_transceiver.m
+% Author : Michael Wurm <wurm.michael95@gmail.com>
+% Topic  : FM-Radio Sender and Receiver
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Prepare environment
+clear; close all; clc;
+%restoredefaultpath;
+
+addpath(genpath('./helpers/auto-arrange-figs/'));
+
+%% Settings
+% Channel
+fc_oe3 = 98.1e6;
+
+fs    = 1e6;
+n_sec = 1;
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Sender
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Audio stream data
+t = 0:1/fs:n_sec-1;
+
+audioFreqLHz = 500;
+audioDataL   = 1 * sin(2*pi*audioFreqLHz/fs*t);
+
+audioFreqRHz = 500;
+audioDataR   = 1 * sin(2*pi*audioFreqRHz/fs*t);
+
+audioData = audioDataL + audioDataR;
+
+%% 19kHz pilot tone
+
+pilotFreqHz = 19000;
+pilotTone = 1 * sin(2*pi*pilotFreqHz/fs);
+
+%% Difference signal (for stereo)
+
+audioDiff = audioDataL - audioDataR;
+
+% Modulate it to 38 kHz
+audioDiffMod = audioDiff * 0.5; %TODO
+
+%% Radio Data Signal (RDS)
+% TODO
+
+%% FM channel
+% Sum up all signal parts
+
+fmChannel = audioData + pilotTone + audioDiffMod;
+
+%% FM channel spectrum
+
+
+
+
+figure();
+
+
+
+
+
+
+%% Arrange all plots on the display
+autoArrangeFigures(4,4,1);

### 2021-02-15 sim/matlab/from_tutorial/fm_demodulate.m

@@ -8,7 +8,7 @@
 %% Prepare environment
 clear; close all; clc;
 
-addpath(genpath('./helpers/auto-arrange-figs/'));
+addpath(genpath('../helpers/auto-arrange-figs/'));
 
 %% FM station info
 fc_oe3 = 98.1e6;
@@ -18,7 +18,7 @@ fc    = 98.0e6;
 fs    = 1e6;
 n_sec = 10;
 
-y = loadFile('fm_record.bin');
+y = loadFile('../recordings/fm_record.bin');
 
 assert(size(y,1) == n_sec*fs, ...
     'Recording is corrupted. Expected %d samples, but the file only contains %d.', ...

### 2021-02-15 sim/matlab/record-rtl2832u.sh

@@ -12,7 +12,7 @@
 center_freq=98.0e6
 sample_freq=1000000
 n_seconds=10
-out_file="./fm_record.bin"
+out_file="./recordings/fm_record.bin"
 
 n_samples=$(( $sample_freq * $n_seconds))
 echo "Recording $n_samples @ $sample_freq SPS ($n_seconds seconds)."

### 2021-02-15 sim/matlab/fm_transceiver.m
comments

@@ -17,6 +17,9 @@ fc_oe3 = 98.1e6;
 fs    = 1e6;
 n_sec = 1;
 
+% Options
+EnableTrafficInfoTrigger = false;
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% Sender
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@ -47,16 +50,24 @@ audioDiffMod = audioDiff * 0.5; %TODO
 %% Radio Data Signal (RDS)
 % TODO
 
+%% Hinz-Triller (traffic info trigger)
+% TODO
+% https://de.wikipedia.org/wiki/Autofahrer-Rundfunk-Information#Hinz-Triller
+hinzTriller = 0;
+if EnableTrafficInfoTrigger
+    %TODO
+end
+
 %% FM channel
 % Sum up all signal parts
 
-fmChannel = audioData + pilotTone + audioDiffMod;
+fmChannel = audioData + pilotTone + audioDiffMod + hinzTriller;
 
 %% FM channel spectrum
 
+% FFT
 
-
-
+% Plot
 figure();
 
 

### 2021-05-05 doc/thesis/chapters/dsp.tex
split into separate files

--- 2020-12-07 doc/thesis/chapters/fm_radio.tex
@@
 %%%%%%%%%%%%%%%%%%%%
-\chapter{High-Level Synthesis}
-  \section{Introduction / State of the Art}
-  \section{Functionality (transform high-level code to HDL)}
-  \section{Language Support}
-    \subsection{C++}
-    \subsection{SystemC}
-  \section{Coding}
-    \subsection{Compiler Directives (\#pragma's)}
-    \subsection{Data types}
-    \subsection{Functions}
-    \subsection{Loops}
-    \subsection{Conditional statements}
-  \section{Advantages / Disadvantages}
-    easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV.
-However, it is less optimized, and a developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.
-
-%%%%%%%%%%%%%%%%%%%%
 \chapter{Implementation}
@@
 
-  \section{VHDL}
+  \section{VHDL (no shortcut)}
     \subsection{Channel Selection (IF to Channel-BB)}
@@
 
-  \section{HLS}
+  \section{High-Level Synthesis}
     \subsection{Channel Selection (IF to Channel-BB)}
@@
     \subsection{other Elements}
-
-  \section{Comparison of Metrics}
-    \subsection{Implementation effort/time}
-    \subsection{Hardware Utilization}
-    \subsection{Others}
 
@@
 
+\chapter{Test Results}
+  \section{Functionality}
+  % Functionality same?
+  % Sim speed of x min of recorded FM signal
+  % lines of code
+  \subsection{Implementation effort/time}
+  \subsection{Hardware Utilization}
+  \subsection{Others}
+
+  % TODO Überlegungen in diesem Kapitel
+  % kann Demodulator geändert werden? Bsp. von FM zu DAB+, QAM, etc..
+  % - von matlab heraus VHDL generieren, etc
+
 %%%%%%%%%%%%%%%%%%%%

### 2021-05-05 doc/thesis/chapters/hls.tex
split into separate files

--- 2020-12-04 doc/thesis/chapters/fm_radio.tex
@@
 %%%%%%%%%%%%%%%%%%%%
-\chapter{Signal Processing Theory}
-\label{cha:TheThesis}
+\chapter[HLS]{High-Level Synthesis}
+\label{cha:HLS}
 
-  \section{Up-Downconversion to/from RF}
+\section{Introduction / State of the Art}
 
-  \section{Frequency Modulation}
-  see literature/Present\_lec6\_AM\_FM.pdf\\
-  see lectures out of HSD/ESD
+\section{Functionality (transform high-level code to HDL)}
 
-  \section{Algorithms for Digital FM Demodulation}
-    see literature/FmDemodulator.pdf (Sect. 3.3)
-    see literature/00476180 Digital FM Demodulator for FM, TV, and Wireless.pdf (Sect. II and III)
+\section{Language Support}
 
-    \subsection{Baseband Delay Demodulator}
-    \subsection{Phase-Adapter Demodulator}
-    \subsection{Phase-Locked Loop (PLL)}
-    \subsection{Mixed Demodulator}
+  \subsection{C++}
 
-%%%%%%%%%%%%%%%%%%%%
-\chapter{High-Level Synthesis}
-  \section{Introduction / State of the Art}
-  \section{Functionality (high level code to HDL)}
-  \section{Variants}
-    \subsection{C++}
   \subsection{SystemC}
+
+\section{Coding}
+
+  \subsection{Compiler Directives (\#pragma's)}
+
+  \subsection{Data types}
+
+  \subsection{Functions}
+
+  \subsection{Loops}
+
+  \subsection{Conditional statements}
+
 \section{Advantages / Disadvantages}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Implementation}
-  \section{Matlab Simulation}
-  in fixed point, close to hardware level algorithm
-
-  \section{VHDL}
-    \subsection{Channel Selection (IF to Channel-BB)}
-    \subsection{Phase Detector}
-    \subsection{other Elements}
-
-  \section{HLS}
-    \subsection{Channel Selection (IF to Channel-BB)}
-    \subsection{Phase Detector}
-    \subsection{other Elements}
-
-  \section{Comparison of Metrics}
-    \subsection{Implementation effort/time}
-    \subsection{Hardware Utilization}
-    \subsection{Others}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Deployment on Hardware}
-  \section{Hardware Platform}
-    \subsection{RTL2832u Dongle}
-    \subsection{ZedBoard}
+    easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV.
+However, it is less optimized, and a developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.

### 2021-05-05 doc/thesis/chapters/system-architecture-and-concept.tex
split into separate files

--- 2020-12-04 doc/thesis/chapters/fm_radio.tex
@@
 %%%%%%%%%%%%%%%%%%%%
-\chapter{Signal Processing Theory}
-\label{cha:TheThesis}
+\chapter{System Architecture/Concept}
+\label{cha:SystemArchitecture/Concept}
 
-  \section{Up-Downconversion to/from RF}
+  \section{Block Diagram (with details)}
+    \subsection{describe blocks}
 
-  \section{Frequency Modulation}
-  see literature/Present\_lec6\_AM\_FM.pdf\\
-  see lectures out of HSD/ESD
+  \section{Which parts in HLS and VHDL}
 
-  \section{Algorithms for Digital FM Demodulation}
-    see literature/FmDemodulator.pdf (Sect. 3.3)
-    see literature/00476180 Digital FM Demodulator for FM, TV, and Wireless.pdf (Sect. II and III)
-
-    \subsection{Baseband Delay Demodulator}
-    \subsection{Phase-Adapter Demodulator}
-    \subsection{Phase-Locked Loop (PLL)}
-    \subsection{Mixed Demodulator}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{High-Level Synthesis}
-  \section{Introduction / State of the Art}
-  \section{Functionality (high level code to HDL)}
-  \section{Variants}
-    \subsection{C++}
-    \subsection{SystemC}
-  \section{Advantages / Disadvantages}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Implementation}
-  \section{Matlab Simulation}
-  in fixed point, close to hardware level algorithm
-
-  \section{VHDL}
-    \subsection{Channel Selection (IF to Channel-BB)}
-    \subsection{Phase Detector}
-    \subsection{other Elements}
-
-  \section{HLS}
-    \subsection{Channel Selection (IF to Channel-BB)}
-    \subsection{Phase Detector}
-    \subsection{other Elements}
-
-  \section{Comparison of Metrics}
-    \subsection{Implementation effort/time}
-    \subsection{Hardware Utilization}
-    \subsection{Others}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Deployment on Hardware}
-  \section{Hardware Platform}
-    \subsection{RTL2832u Dongle}
-    \subsection{ZedBoard}
+  \section{Test Environment}

### 2021-05-05 doc/thesis/chapters/dsp.tex
more splitting into files

--- 2020-12-04 doc/thesis/chapters/fm_radio.tex
@@
 
-  \section{Up-Downconversion to/from RF}
+  \section{Modulation from Baseband to RF}
+  \section{Modulation from RF to Baseband}
 
-  \section{Frequency Modulation}
+  \section{Frequency Modulation (FM)}
   see literature/Present\_lec6\_AM\_FM.pdf\\
   see lectures out of HSD/ESD
+
+  Frequency modulation (FM) is a widely used standard to transmit data streams.
+  The probably best known usecase therefore is commercial broadcast radio, where an audio stream is transmitted.
+  Devices to receive these streams are available starting from affordable low prices to the public.
+
+    \subsection{Mathematical description}
+    \subsection{Frequency Band, Channel allocation/distribution}
 
@@
     \subsection{Mixed Demodulator}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{High-Level Synthesis}
-  \section{Introduction / State of the Art}
-  \section{Functionality (high level code to HDL)}
-  \section{Variants}
-    \subsection{C++}
-    \subsection{SystemC}
-  \section{Advantages / Disadvantages}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Implementation}
-  \section{Matlab Simulation}
-  in fixed point, close to hardware level algorithm
-
-  \section{VHDL}
-    \subsection{Channel Selection (IF to Channel-BB)}
-    \subsection{Phase Detector}
-    \subsection{other Elements}
-
-  \section{HLS}
-    \subsection{Channel Selection (IF to Channel-BB)}
-    \subsection{Phase Detector}
-    \subsection{other Elements}
-
-  \section{Comparison of Metrics}
-    \subsection{Implementation effort/time}
-    \subsection{Hardware Utilization}
-    \subsection{Others}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Deployment on Hardware}
-  \section{Hardware Platform}
-    \subsection{RTL2832u Dongle}
-    \subsection{ZedBoard}

### 2021-05-05 doc/thesis/chapters/implementation.tex
more splitting into files

--- 2020-12-07 doc/thesis/chapters/fm_radio.tex
@@
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Signal Processing Theory}
-\label{cha:TheThesis}
-
-  \section{Modulation from Baseband to RF}
-  \section{Modulation from RF to Baseband}
-
-  \section{Frequency Modulation (FM)}
-  see literature/Present\_lec6\_AM\_FM.pdf\\
-  see lectures out of HSD/ESD
-    \subsection{Mathematical description}
-    \subsection{Frequency Band, Channel allocation/distribution}
-
-  \section{Algorithms for Digital FM Demodulation}
-    see literature/FmDemodulator.pdf (Sect. 3.3)
-    see literature/00476180 Digital FM Demodulator for FM, TV, and Wireless.pdf (Sect. II and III)
-
-    \subsection{Baseband Delay Demodulator}
-    \subsection{Phase-Adapter Demodulator}
-    \subsection{Phase-Locked Loop (PLL)}
-    \subsection{Mixed Demodulator}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{High-Level Synthesis}
-  \section{Introduction / State of the Art}
-  \section{Functionality (transform high-level code to HDL)}
-  \section{Language Support}
-    \subsection{C++}
-    \subsection{SystemC}
-  \section{Coding}
-    \subsection{Compiler Directives (\#pragma's)}
-    \subsection{Data types}
-    \subsection{Functions}
-    \subsection{Loops}
-    \subsection{Conditional statements}
-  \section{Advantages / Disadvantages}
-    easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV.
-However, it is less optimized, and a developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.
-
 %%%%%%%%%%%%%%%%%%%%
 \chapter{Implementation}
+\label{cha:Implementation}
 \section{Matlab/Python Model}
@@
 
-  \section{VHDL}
+  \section{VHDL (no shortcut)}
   \subsection{Channel Selection (IF to Channel-BB)}
@@
 
-  \section{HLS}
+  \section{High-Level Synthesis}
   \subsection{Channel Selection (IF to Channel-BB)}
@@
   \subsection{other Elements}
-
-  \section{Comparison of Metrics}
-    \subsection{Implementation effort/time}
-    \subsection{Hardware Utilization}
-    \subsection{Others}
 
@@
   The \#pragmas need to be within the code, at the appropriate positions, which makes it difficult to write compiler independent code....
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Deployment on Hardware}
-  \section{Hardware Platform}
-    \subsection{RTL2832u Dongle}
-    \subsection{ZedBoard}

### 2021-05-05 doc/thesis/chapters/dsp.tex
dsp

@@
 
+%%%%%%%%%%%%%%%%%%%%
 \section{Modulation from Baseband to RF}
+
+%%%%%%%%%%%%%%%%%%%%
 \section{Modulation from RF to Baseband}
 
-  \section{Frequency Modulation (FM)}
-  see literature/Present\_lec6\_AM\_FM.pdf\\
-  see lectures out of HSD/ESD
+%%%%%%%%%%%%%%%%%%%%
+\section{Frequency Modulation in Broadcasting}
+  %see literature/Present\_lec6\_AM\_FM.pdf\\
+  %see lectures out of HSD/ESD
 
@@
 The probably best known usecase therefore is commercial broadcast radio, where an audio stream is transmitted.
-  Devices to receive these streams are available starting from affordable low prices to the public.
+Devices to receive these streams are available for low prices to the public.
 
+This chaper describes the main properties of broadcast FM, such as the mathematical description, frequency bands that are used, or the specific frequency parts within a channel spectrum.
+
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Mathematical description}
+
+The mathematical description of an FM baseband signal can be described with the formulas presented in this section.
+FM encodes the information content in its instantanious frequency.
+This means that the measured frequency at any moment in time represents a specific value of a transmitted message.
+For a general classification, FM belongs to the group of angle, or phase modulated signals (PM).
+The simple reason therefor is, that a frequency has a direct relationship to an angle, if the signal is seen on a unit circle.\\
+
+The instantanious frequency $f_i$ of an FM signal can be described as
+\begin{equation}
+  f_i = f_c + \Delta f \cdot m(t)
+\end{equation}
+where $f_c$ is the carrier- or center frequency, $\Delta f$ is the maximum frequency deviation and $m(t)$ is the information or message signal that is to be transmitted.
+Simply looking at this equation, the frequency deviation, which is sometimes also called the swing, varies in the range between $f_c \pm \Delta f$.
+
+Considering the relationship between frequency and angle, as described above, and after some simple substitutions which will not be described into detail here, the equation for a generic frequency modulated wave is the following
+\begin{equation}
+  s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int m(t) dt \Big)
+  \label{equ:fm_func}
+\end{equation}
+where $A_c$ is the amplitude of the resulting FM signal and $k_f$ is the frequency sensitivity factor.
+This sensitivity factor has a direct relationship with the modulation index $\beta$.
+\begin{equation}
+  \beta = \frac{\Delta f}{f_m} = \frac{k_f A_m}{f_m}
+\end{equation}
+where $f_m$ is the highest existing frequency, or the bandwidth of the information signal.
+
+Formula \ref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
+A widely used example application of FM is broadcast radio, where audio streams are transmitted.
+An audio stream can be described as a cosine wave.
+Therefor, the information signal that is to be transmitted in broadcast radio can be represented as a cosine wave in the form of
+\begin{equation}
+  m(t) = A_m\ cos(2 \pi f_m t).
+\end{equation}
+By inserting this message signal into the generic FM formula \ref{equ:fm_func}, the final equation for an FM signal transforms into the following form.
+\begin{equation}
+  y(t) = A_c\ cos \Big(2 \pi f_c t + \beta\ sin(2 \pi f_m t)\Big )
+\end{equation}
+Several formulas, as well as their derivations are described according to \cite{ref_FM_Maths_Info_1}\cite{ref_FMMaths2}.
+
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Frequency Band, Channel allocation/distribution}
 
+%%%%%%%%%%%%%%%%%%%%
 \section{Algorithms for Digital FM Demodulation}
@@
 
+  %%%%%%%%%%%%%%%%%%%%
   \subsection{Baseband Delay Demodulator}
+  %%%%%%%%%%%%%%%%%%%%
   \subsection{Phase-Adapter Demodulator}
+  %%%%%%%%%%%%%%%%%%%%
   \subsection{Phase-Locked Loop (PLL)}
+  %%%%%%%%%%%%%%%%%%%%
   \subsection{Mixed Demodulator}

### 2021-05-05 doc/thesis/chapters/dsp.tex
another chapter; from paper

@@
 where $f_m$ is the highest existing frequency, or the bandwidth of the information signal.
+% More info on modulation index: https://fas.org/man/dod-101/navy/docs/es310/FM.htm
 
@@
 %%%%%%%%%%%%%%%%%%%%
-\subsection{Frequency Band, Channel allocation/distribution}
+\subsection{Frequency Band}
+
+The frequency band that is used for FM broadcasting is defined worldwide and spans from 87.5 to 108 MHz.
+However, some countries only partially use the band \cite{ref_itu_regulations}.
+The range is located within the so-called Very-High-Frequency (VHF) band, which is open to the public in terms of usage.
+This means, that any transmitter may use the specified frequency range freely.
+Austria allowed the legal usage in 2006 \cite{ref_austria_rundfunkgesetz_2006}.
+However, the transmission power needs to be limited, so that neighboring receivers are not disturbed in receiving any existing channels.
+The European Commision specifies the power limit as 50 nW of effective radiated power (ERP) \cite{ref_eu_commision_radio_spectrum}.
+Another limitation is, that a transmitter must be capable to transmit on multiple center frequencies within the broadcasting band.
+It must not be fixed to a single center frequency.
+Through this specification, a transmitter is capable of switching its transmission center frequency, in order to avoid interference with another transmitters signal.
+
+In a practical usecase this means that a transmitter needs to select a channel or frequency range, that is not already occupied by an official transmitter.
+If a free range is found, the transmitter may start sending an FM signal with the mentioned power limitations \cite{ref_ebu_fm_modulators_in_europe_power_regulations}.
+
+%https://www.ris.bka.gv.at/GeltendeFassung.wxe?Abfrage=Bundesnormen&Gesetzesnummer=20008807
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Frequency Spectrum of a Broadcast FM Channel}
+
+The entire FM broadcasting band from 87.5 to 108 MHz divides into multiple channels which can be used.
+In Europe, the European Telecommunications Standards Institute (ETSI) sets the respective standards for the usage of these frequencies.
+The main specifications for a single FM broadcasting channel are explained in this section.\\
+
+Each channel may allocate a bandwidth of 200 kHz.
+Within one of these channels, the maximum deviation from the center frequency shall not exceed 75 kHz.
+This leaves a guard band of 25 kHz on either side, to minimize interference with adjacent channels.
+Center frequencies may be allocated on multiples of 100 kHz.
+However, neighboring channels need to be considered therefor, since this may result in an overlap in the spectrum.\\
+
+Fig.\ref{fig_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
+For a better overview, only the positive spectum part is shown.
+The negative part is an exact mirror, because the transmitted and received signals from the antenna are real signals - they do not have an imaginary part.
+The spectrum consists of multiple parts, which are described in the following paragraphs.
+
+\begin{figure}[!h]
+  \centering
+    \includegraphics[width=7cm]{img/fm-channel-baseband.png}
+  \caption{Allocation of frequencies in an FM channel \cite{ref_fig_channel_freqs}}
+  \label{fig_channel_baseband_freqs}
+\end{figure}
+
+\subsubsection{Mono Audio Part}
+
+The mono audio stream is located between 30 Hz and 15 kHz.
+This signal is built by the sum of the left and right audio channels.
+%The lower limit at 30 Hz is to prevent the transmission of a direct current (DC) part, which would require a large amount of power in transmission.
+%This effect is one of the downsides of amplitude modulation (AM), where the carrier is located at the channels' center frequeny and requires a high transmission power. % TODO: check this fact
+The upper limit at 15 kHz is chosen to maintain a sufficient spacing to the first subcarrier.
+For audio streaming, as in FM broadcasting, this is not a limitation, since it is unlikely to have frequencies higher than 15 kHz in an audio stream.
+Also, this already reaches the upper limit of the human ears' bandwith, so higher spectral parts will not be heard anyway.
+
+\subsubsection{Pilot Tone}
+
+The first subcarrier is allocated at an offset of 19 kHz from the center frequency.
+It is also called the 'pilot tone', since it is a continuous signal.
+The pilot tone is used to be able to demodulate the left and right channel of an audio signal correctly, to generate stereo audio.
+
+\subsubsection{Stereo Audio Difference}
+
+A signal that is constructed by the difference between left and right audio channel is modulated on a 38 kHz subcarrier.
+This subcarrier is an integer multiple of 19 kHz for practical reasons.
+However, the 38 kHz carrier is suppressed and thus not visible in the received spectrum.
+It can still be recovered, since it is phase coherent with the 19 kHz pilot tone per definition.
+The bandwidth for this difference-signal spans 15 kHz on either side of the subcarrier.
+It is used to generate a stereo audio signal, in combination with the mono signal.
+This process is explained into more detail in chapter \ref{sec_fm_sig_demod}.
+
+\subsubsection{Additional Services}
+
+Considering all these parts in the spectrum, there is still bandwidth free to use, up to the maximum channel bandwitdh of 100 kHz in one sideband.
+Because of that, additional services were added to the pure audio transmission, to provide additional data services and information.
+Services that were implemented are the Data Radio Channel (DARC), which is mostly used in Japan and the USA, the Subsidary Communication Authorization (SCA) and the Radio Data System (RDS) \cite{ref_rohde_u_schwarz}.
+Out of these, RDS is the most significant service in Europe.
+It is used to transmit additional information about the channel, such as the radio stations' name, currently playing songs' title or traffic information.\\
+
+\noindent
+The information about the frequency spectrum was taken from \cite{ref_fm_broadcast_tutorial_and_basics}.
 

### 2021-05-05 doc/thesis/chapters/dsp.tex
update

@@
   \centering
-    \includegraphics[width=7cm]{img/fm-channel-baseband.png}
+    \includegraphics[width=9cm]{img/fm-channel-baseband.png}
   \caption{Allocation of frequencies in an FM channel \cite{ref_fig_channel_freqs}}
@@
 The upper limit at 15 kHz is chosen to maintain a sufficient spacing to the first subcarrier.
+
 For audio streaming, as in FM broadcasting, this is not a limitation, since it is unlikely to have frequencies higher than 15 kHz in an audio stream.
@@
 The first subcarrier is allocated at an offset of 19 kHz from the center frequency.
-It is also called the 'pilot tone', since it is a continuous signal.
-The pilot tone is used to be able to demodulate the left and right channel of an audio signal correctly, to generate stereo audio.
+This subcarrier is also called the 'pilot tone', since it is a continuous signal.
+It is independent of any varying message signal content.
+The pilot tone is used for stereo audio demodulation.
+To regenerate a stereo audio signal, the left and right audio channels need to be recovered correctly.
 
@@
 
-A signal that is constructed by the difference between left and right audio channel is modulated on a 38 kHz subcarrier.
-This subcarrier is an integer multiple of 19 kHz for practical reasons.
+A signal that is constructed by the difference between the left and right audio channel is modulated on a 38 kHz subcarrier.
+This subcarrier is an integer multiple of the pilot tone at 19 kHz, for practical reasons.
 However, the 38 kHz carrier is suppressed and thus not visible in the received spectrum.
-It can still be recovered, since it is phase coherent with the 19 kHz pilot tone per definition.
+The modulation technique that is used for this spectral part is called dual-sideband suppressed-carrier (DSB-SC).
+Even though the carrier is suppressed, it can still be recovered, because it is phase coherent with the 19 kHz pilot tone per definition.
+
 The bandwidth for this difference-signal spans 15 kHz on either side of the subcarrier.
@@
 
-Considering all these parts in the spectrum, there is still bandwidth free to use, up to the maximum channel bandwitdh of 100 kHz in one sideband.
+Considering all these parts in the spectrum, there is still free bandwidth available to use up to the maximum channel bandwitdh of 100 kHz in one sideband.
 Because of that, additional services were added to the pure audio transmission, to provide additional data services and information.
+
 Services that were implemented are the Data Radio Channel (DARC), which is mostly used in Japan and the USA, the Subsidary Communication Authorization (SCA) and the Radio Data System (RDS) \cite{ref_rohde_u_schwarz}.
 Out of these, RDS is the most significant service in Europe.
-It is used to transmit additional information about the channel, such as the radio stations' name, currently playing songs' title or traffic information.\\
+It is used to transmit additional information about the channel, such as the radio stations' name, the currently playing songs' title or traffic information.\\
 

### 2021-05-05 doc/thesis/chapters/dsp.tex
update

@@
 \section{Algorithms for Digital FM Demodulation}
-  see literature/FmDemodulator.pdf (Sect. 3.3)
-  see literature/00476180 Digital FM Demodulator for FM, TV, and Wireless.pdf (Sect. II and III)
+  %see literature/FmDemodulator.pdf (Sect. 3.3)
+  %see literature/00476180 Digital FM Demodulator for FM, TV, and Wireless.pdf (Sect. II and III)
+
+An FM signal that is received, for example from an antenna, needs to be demodulated in order to decode its actual data content.
+In general, an FM demodulator serves the purpose to transform a frequency modulated signal (FM) into an amplitude modulated signal (AM), so that AM signal processing techniques can be applied afterwards.\\
+
+The radio frequency antenna signal usually needs to be amplified, bandpass-filtered, and down-converted to baseband, using subsampling and quadrature-mixing, or similar strategies.
+This part of the signal processing chain is assumed to be working correctly and is out of the scope of this chapter. %TODO: write up/down conversion in other chapter?
+The FM signal that is evaluated here is assumed to be a quadrature-mixed signal in baseband, which means that inphase and quadrature (I/Q) signals are available.
+
+%TODO: show downconversion path (maybe as a figure?)
+%https://www.veron.nl/wp-content/uploads/2014/01/FmDemodulator.pdf
+
+In the following sections, two digital FM demodulator variants are described.
+
+% TODO: add img/discrimination_method_bd2.png here (?)
 
 %%%%%%%%%%%%%%%%%%%%
-  \subsection{Baseband Delay Demodulator}
+\subsection{Frequency Discriminator}
+
+A frequency discriminator can be defined as a black box component, that generates an output that is directly proportional to the frequency of the input FM signal.
+Different strategies can be implemented in this black box component - the discriminator.
+Here, a differentiator with a subsequent envelope detector is used, which is shown in Fig.\ref{fig_bd_freq_discriminator}.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=7cm]{img/discrimination_method_bd1.jpg}
+  \caption{Frequency discriminator block diagram \cite{ref_fig_freq_discriminator}}
+  \label{fig_bd_freq_discriminator}
+\end{figure}
+
+% TODO: check which one i'm describing
 %%%%%%%%%%%%%%%%%%%%
-  \subsection{Phase-Adapter Demodulator}
+\subsubsection{this one}
 %%%%%%%%%%%%%%%%%%%%
-  \subsection{Phase-Locked Loop (PLL)}
+%\subsubsection{Baseband Delay Demodulator}
 %%%%%%%%%%%%%%%%%%%%
-  \subsection{Mixed Demodulator}
+%\subsubsection{Phase-Adapter Demodulator}
+%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Phase-Locked Loop (PLL)}
+%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Mixed Demodulator}
+
+This type of demodulator converts the FM signal into an AM signal, but still leaves an FM content.
+The transformation that happens here can be expressed in a formula, by simply differentiating Eq.\ref{equ_fm_func}.
+This operation, with simple algebraic transformations, results in
+\begin{equation}
+  \frac{d s(t)}{dt} = A_c 2 \pi \Big[f_c + k_f m(t) \Big] sin \Big(\omega_c t + 2 \pi k_f \int m(t) dt -pi \Big)
+  \label{equ:fm_demod_discriminator}
+\end{equation}
+
+An important requirement for the differentiator is, that the input signal is of a constant amplitude \cite{ref_schnyder_haller}.
+The transmitted signal is subject to random additive noise on the transmission channel.
+The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Eq.\ref{equ:differentiator_two_samples}.
+To achieve a constant amplitude, the received complex baseband signal $s(t)$ can be normalized to a value of one.
+\begin{equation}
+  s_{norm} = \frac{s}{|s|} = \frac{A(n)\ e^{j\phi_{FM}(n)}}{|{A(n)\ e^{j\phi_{FM}(n)}|}} = e^{j\phi_{FM}(n)}
+\end{equation}
+\begin{equation}
+  e^{j\phi_{FM}(n)} = |1|
+\end{equation}
+
+A differentiation in hardware is simply performed by substracting two consecutive samples, like
+\begin{equation}
+  \frac{d s(t)}{dt} = s(t) - s(t-\Delta t)
+  \label{equ:differentiator_two_samples}
+\end{equation}
+where $\Delta t$ is the inverse of the sample frequency.\\ %TODO: check this (i'm using a 3-tap FIR in FPGA)
+
+The differentiated signal Eq.\ref{equ:fm_demod_discriminator} consist of an AM and an FM content.
+The FM part is described by the sine function.
+The more important part is the AM part, which resembles the message signal.
+It is described by the term within the square brackets.
+
+The differentiated signal Eq.\ref{equ:fm_demod_discriminator} in time-domain is shown in the top left diagram in Fig.\ref{fig_time_domain_envelope_detect}.
+This signal needs to be fed into an envelope detector, as seen in Fig.\ref{fig_bd_freq_discriminator}.
+The envelope detector extracts the envelope by removing the high frequency portion.
+The method implemented in Fig.\ref{fig_time_domain_envelope_detect} is one of the most simple ones and is usually referred to as 'Asynchronous Half-Wave Envelope Detector'\cite{ref_envelope_detector}.
+It consists of a thresholding unit to remove negative values and a conventional lowpass filter.
+In analog implementations, the thresholding unit is a diode.
+The lowpass' cutoff frequency needs to be adapted to the envelope signals' maximum frequency, since it needs to be able to follow the signal, but also sufficiently smooth out the rectified signal.
+
+Different methods may be implemented for the envelope detector.
+For example, an 'Asynchronous Full-Wave Envelope Detector' can be used.
+Therefore, the previous method only needs to swap the thresholding unit with an unit that calculates an absolut value.
+In that architecture, a full-wave rectification is performed on the signal, which significantly improves the envelope detection accuracy \cite{ref_envelope_detector}.
+The envelope is followed more exactly, because of the higher power that is available in the signal, since the entire signal is taken into account, and not only the positive half, as in the previous half-wave method.
+
+\begin{figure}[!h]
+  \centering
+    \includegraphics[width=8.8cm]{img/envelope-detect-time-domain.jpg}
+  \caption{Time-domain signals in envelope detection \cite{ref_roppel}}
+  \label{fig_time_domain_envelope_detect}
+\end{figure}
+
+\subsection{Phase-Locked Loop}
+A Phase-Locked Loop (PLL) can also be used to transform an FM signal to AM.
+A PLL is a feedback loop that is usually used to generate an output signal that has a fixed phase reference to its input.
+In the case of FM demodulation, the loop filter is configured to be able to follow the frequency variations on the input, which is directly related to the encoded information.
+The output of the PLL directly delivers an AM signal, which corresponds to the transmitted information. \cite{ref_schnyder_haller}
+
+Due to the limited space in this paper, FM demodulation with a PLL is not described into more detail.

### 2021-05-05 doc/thesis/chapters/dsp.tex
wip

@@
 
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Phase-Locked Loop}
-A Phase-Locked Loop (PLL) can also be used to transform an FM signal to AM.
+
+A Phase-Locked Loop (PLL) can also be used to transform an FM signal to an AM signal.
 A PLL is a feedback loop that is usually used to generate an output signal that has a fixed phase reference to its input.
-In the case of FM demodulation, the loop filter is configured to be able to follow the frequency variations on the input, which is directly related to the encoded information.
+In the case of FM demodulation, the loop filter in the feedback branch is configured to be able to follow the frequency variations on the input, which is directly related to the encoded information.
 The output of the PLL directly delivers an AM signal, which corresponds to the transmitted information. \cite{ref_schnyder_haller}
 
-Due to the limited space in this paper, FM demodulation with a PLL is not described into more detail.
+FM demodulation with a PLL is not described into more detail, since this thesis is focussing on the concept of an entire system architecture.
+
+%%%%%%%%%%%%%%%%%%%%
+\section{Demodulation of Broadcast FM}
+
+This chapter describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
+The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec_fm_sig_demod} is already done correctly.
+A frequency spectrum of an FM channel is shown in Fig.\ref{fig_channel_baseband_freqs} above.
+
+explaine multiplex signal somewhere
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Mono}
+\label{subsec:demod_mono}
+  The first part of the channel frequency spectrum is the mono audio signal. It is summation signal of the left and right audio channel.
+  A mono receiver can thus simply replay the so-called multiplex signal, which is generated by the FM demodulation described in chapter \ref{sec_fm_sig_demod}.
+The 19 kHz pilot tone will not be audible by most people, because it is outside of the range of a human ear.
+Besides that, any used speaker may also be unable to produce a frequency in this range.
+This explanation applies to several higher frequency components in the spectrum as well.
+  In case the multiplex signal should be stored it in a file with 44.1 kHz samplerate, a lowpass with a cutoff frequency of 15 kHz and sufficient attenutation before 19 kHz needs to be inserted before the decimation and file output.
+
+\subsection{Stereo}
+  Demodulating an FM channel to recover a stereo audio signal requires a more sophisticated approach.
+  In the channel frequency spectrum, there are two main parts of the audio signal - the sum and difference of the left and right audio channel signals.
+  To combine these to a stereo signal, the following equations can be applied.
+  \begin{equation*}
+    L = (L+R) + (L-R) = (2)L
+  \end{equation*}
+  \begin{equation}
+    R = (L+R) - (L-R) = (2)R
+    \label{equ_stereo_from_sum_diff}
+  \end{equation}
+
+  The block diagram in Fig.\ref{fig_bd_stereo_demod} illustrates the signal processing chain for stereo audio.
+In the block diagram, $x(t)$ represents the multiplex signal from the FM demodulator.
+Signals $x_l(t)$ and $x_r(t)$ describe the left and right audio signal, respectively.
+
+  \begin{figure}[!h]
+    \centering
+      \includegraphics[width=8.8cm]{img/fm-demod-stereo-audio.png}
+    \caption{Block diagram of FM stereo audio demodulation \cite{ref_roppel}}
+    \label{fig_bd_stereo_demod}
+  \end{figure}
+
+  The central branch selects the pilot tone at 19 kHz with a bandpass filter.
+  This bandpass needs to have a frequency response that is sharp enough to have a sufficient attenuation at $\pm$4 kHz, since this is where the summation and difference signals frequencies are allocated.
+  Afterwards, the pilot tone frequency is doubled via a PLL to achieve a 38 kHz to later demodulate the difference signal phase-coherently.
+  The PLL lock indicator can be re-used as an indicator, whether a pilot tone is existent.
+  In the upper branch, a bandpass filter selects the difference signal with ranges from 23 to 53 kHz.
+  It is then modulated to baseband by the previously generated 38 kHz subcarrier.
+  Subsequently, a lowpass limits the signal to the audio signal bandwidth of 15 kHz.
+  The lower branch lowpass-filters the summation signal.
+  The rightmost part in the diagram performs the combination of summation and difference signals, according to Eq.\ref{equ_stereo_from_sum_diff}.

### 2021-05-06 doc/thesis/chapters/dsp.tex
update.

@@
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter{Signal Processing Theory}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Modulation from Baseband to RF}
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Modulation from RF to Baseband}
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Frequency Modulation in Broadcasting}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Mathematical description}
@@
 
-Formula \ref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
+Formula \eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
 A widely used example application of FM is broadcast radio, where audio streams are transmitted.
@@
 \end{equation}
-By inserting this message signal into the generic FM formula \ref{equ:fm_func}, the final equation for an FM signal transforms into the following form.
+By inserting this message signal into the generic FM formula \eqref{equ:fm_func}, the final equation for an FM signal transforms into the following form.
 \begin{equation}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Frequency Band}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Frequency Spectrum of a Broadcast FM Channel}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Algorithms for Digital FM Demodulation}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Frequency Discriminator}
 
-A frequency discriminator can be defined as a black box component, that generates an output that is directly proportional to the frequency of the input FM signal.
-Different strategies can be implemented in this black box component - the discriminator.
-Here, a differentiator with a subsequent envelope detector is used, which is shown in Fig.\ref{fig_bd_freq_discriminator}.
+A frequency discriminator generates an output that is directly proportional to the frequency of the input FM signal.
+In other words, it converts FM to AM.
+The discriminator consists of two main parts - a differentiator and an envelope detector.
+Different strategies can be implemented for both parts.
+A block diagram for a frequency discriminator with its two main parts is shown in Fig.\ref{fig:bd_freq_discriminator}.
 
-\begin{figure}[!h]
-  \centering
-  \includegraphics[width=7cm]{img/discrimination_method_bd1.jpg}
-  \caption{Frequency discriminator block diagram \cite{ref_fig_freq_discriminator}}
-  \label{fig_bd_freq_discriminator}
-\end{figure}
+\includepicture [0.6] [0] {Frequency discriminator block diagram \cite{ref_fig_freq_discriminator}} {bd_freq_discriminator} {img/discrimination_method_bd1.jpg}
 
-% TODO: check which one i'm describing
-%%%%%%%%%%%%%%%%%%%%
-\subsubsection{this one}
-%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Baseband Delay Demodulator}
-%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Phase-Adapter Demodulator}
-%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Phase-Locked Loop (PLL)}
-%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Mixed Demodulator}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Differentiator}
 
-This type of demodulator converts the FM signal into an AM signal, but still leaves an FM content.
-The transformation that happens here can be expressed in a formula, by simply differentiating Eq.\ref{equ_fm_func}.
-This operation, with simple algebraic transformations, results in
+The differentiator already converts the FM signal into an AM signal, but still leaves an FM content.
+The transformation that happens here can be expressed in a formula, by simply differentiating Eq.\eqref{equ:fm_func}.
+This operation, with some additional simple algebraic transformations, results in
 \begin{equation}
@@
 
+This formula shows the AM and an FM contents clearly.
+The FM part is described by the sine function, with its argument.
+The more important part is the AM part, which resembles the message signal.
+It is described by the term within the square brackets.
+
+A time-domain example for this signal is shown in the top left diagram in Fig.\ref{fig:time_domain_envelope_detect}.\\
+
 An important requirement for the differentiator is, that the input signal is of a constant amplitude \cite{ref_schnyder_haller}.
 The transmitted signal is subject to random additive noise on the transmission channel.
-The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Eq.\ref{equ:differentiator_two_samples}.
+The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Eq.\eqref{equ:differentiator_two_samples}.
 To achieve a constant amplitude, the received complex baseband signal $s(t)$ can be normalized to a value of one.
@@
 
-A differentiation in hardware is simply performed by substracting two consecutive samples, like
+In a hardware implementation, a differentiation is simply performed by substracting two consecutive samples, like
 \begin{equation}
@@
 
-The differentiated signal Eq.\ref{equ:fm_demod_discriminator} consist of an AM and an FM content.
-The FM part is described by the sine function.
-The more important part is the AM part, which resembles the message signal.
-It is described by the term within the square brackets.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Envelope Detector}
 
-The differentiated signal Eq.\ref{equ:fm_demod_discriminator} in time-domain is shown in the top left diagram in Fig.\ref{fig_time_domain_envelope_detect}.
-This signal needs to be fed into an envelope detector, as seen in Fig.\ref{fig_bd_freq_discriminator}.
-The envelope detector extracts the envelope by removing the high frequency portion.
-The method implemented in Fig.\ref{fig_time_domain_envelope_detect} is one of the most simple ones and is usually referred to as 'Asynchronous Half-Wave Envelope Detector'\cite{ref_envelope_detector}.
+The differentiated signal still consists of a FM part, as explained above, using Eq.\eqref{equ:fm_demod_discriminator}.
+In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.
+A block diagram is shown in Fig.\ref{fig:bd_freq_discriminator}.
+
+The method implemented in Fig.\ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as `Asynchronous Half-Wave Envelope Detector'\cite{ref_envelope_detector}.
 It consists of a thresholding unit to remove negative values and a conventional lowpass filter.
 In analog implementations, the thresholding unit is a diode.
-The lowpass' cutoff frequency needs to be adapted to the envelope signals' maximum frequency, since it needs to be able to follow the signal, but also sufficiently smooth out the rectified signal.
+The cutoff frequency of the lowpass needs to be adapted to the envelope signals' maximum frequency.
+It needs to be able to follow the signal, but also sufficiently smooth out the rectified signal.\\
 
 Different methods may be implemented for the envelope detector.
-For example, an 'Asynchronous Full-Wave Envelope Detector' can be used.
-Therefore, the previous method only needs to swap the thresholding unit with an unit that calculates an absolut value.
+For example, an `Asynchronous Full-Wave Envelope Detector' can be used.
+Therefore, the previous method only needs to swap the thresholding unit with an unit that calculates the absolut value.
 In that architecture, a full-wave rectification is performed on the signal, which significantly improves the envelope detection accuracy \cite{ref_envelope_detector}.
@@
 
-\begin{figure}[!h]
-  \centering
-    \includegraphics[width=8.8cm]{img/envelope-detect-time-domain.jpg}
-  \caption{Time-domain signals in envelope detection \cite{ref_roppel}}
-  \label{fig_time_domain_envelope_detect}
-\end{figure}
+\includepicture [0.75] [0] {Time-domain signals in envelope detection \cite{ref_roppel}} {time_domain_envelope_detect} {img/envelope-detect-time-domain.jpg}
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Phase-Locked Loop}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Baseband Delay Demodulator}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Phase-Adapter Demodulator}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Mixed Demodulator}
 
+%TODO: explain a PLL in some more detail (block diagram)
 A Phase-Locked Loop (PLL) can also be used to transform an FM signal to an AM signal.
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Demodulation of Broadcast FM}
@@
 
-explaine multiplex signal somewhere
+explain multiplex signal somewhere
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Mono}
@@
   To combine these to a stereo signal, the following equations can be applied.
-  \begin{equation*}
-    L = (L+R) + (L-R) = (2)L
-  \end{equation*}
+
   \begin{equation}
+    \begin{split}
+      L = (L+R) + (L-R) = (2)L \\
       R = (L+R) - (L-R) = (2)R
       \label{equ_stereo_from_sum_diff}
+    \end{split}
   \end{equation}
@@
   The lower branch lowpass-filters the summation signal.
-  The rightmost part in the diagram performs the combination of summation and difference signals, according to Eq.\ref{equ_stereo_from_sum_diff}.
+  The rightmost part in the diagram performs the combination of summation and difference signals, according to Eq.\eqref{equ_stereo_from_sum_diff}.

### 2021-05-06 doc/thesis/chapters/dsp.tex
update

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Oversampling / Sample Rate Reduction}
+%TODO: explain DSP functions that are actually used in Matlab FM demod
+Example: \\
+In case the multiplex signal needs to be stored in a file with 44.1 kHz samplerate, in order to avoid aliasing, a lowpass with a cutoff frequency of 15 kHz and sufficient attenutation before 19 kHz needs to be inserted before the decimation and file output.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Frequency Modulation in Broadcasting}
@@
 Frequency modulation (FM) is a widely used standard to transmit data streams.
-The probably best known usecase therefore is commercial broadcast radio, where an audio stream is transmitted.
+The probably best known usecase therefor is commercial broadcast radio, where an audio stream is transmitted.
 Devices to receive these streams are available for low prices to the public.
@@
 An audio stream can be described as a cosine wave.
-Therefor, the information signal that is to be transmitted in broadcast radio can be represented as a cosine wave in the form of
+Therefore, the information signal that is to be transmitted in broadcast radio can be represented as a cosine wave in the form of
 \begin{equation}
@@
 Center frequencies may be allocated on multiples of 100 kHz.
-However, neighboring channels need to be considered therefor, since this may result in an overlap in the spectrum.\\
+However, neighboring channels need to be considered therefore, since this may result in an overlap in the spectrum.\\
 
@@
 The negative part is an exact mirror, because the transmitted and received signals from the antenna are real signals - they do not have an imaginary part.
-The spectrum consists of multiple parts, which are described in the following paragraphs.
+The spectrum consists of multiple parts. Therefore, it is also referred to as `multiplex signal' in the literature. Several parts are described in the following paragraphs.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Mono Audio Part}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Pilot Tone}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Stereo Audio Difference}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Additional Services}
@@
 \section{Algorithms for Digital FM Demodulation}
+\label{sec:algorithms-for-digital-fm-demodulation}
   %see literature/FmDemodulator.pdf (Sect. 3.3)
@@
 For example, an `Asynchronous Full-Wave Envelope Detector' can be used.
-Therefore, the previous method only needs to swap the thresholding unit with an unit that calculates the absolut value.
+Therefor, the previous method only needs to swap the thresholding unit with an unit that calculates the absolut value.
 In that architecture, a full-wave rectification is performed on the signal, which significantly improves the envelope detection accuracy \cite{ref_envelope_detector}.
@@
 This chapter describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
-The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec_fm_sig_demod} is already done correctly.
-A frequency spectrum of an FM channel is shown in Fig.\ref{fig_channel_baseband_freqs} above.
-
-explain multiplex signal somewhere
+The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly.
+A frequency spectrum of the multiplex signal of an FM channel is shown in Fig.\ref{fig_channel_baseband_freqs}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Mono}
+\subsection{Mono Audio}
 \label{subsec:demod_mono}
-  The first part of the channel frequency spectrum is the mono audio signal. It is summation signal of the left and right audio channel.
-  A mono receiver can thus simply replay the so-called multiplex signal, which is generated by the FM demodulation described in chapter \ref{sec_fm_sig_demod}.
-The 19 kHz pilot tone will not be audible by most people, because it is outside of the range of a human ear.
-Besides that, any used speaker may also be unable to produce a frequency in this range.
-This explanation applies to several higher frequency components in the spectrum as well.
-  In case the multiplex signal should be stored it in a file with 44.1 kHz samplerate, a lowpass with a cutoff frequency of 15 kHz and sufficient attenutation before 19 kHz needs to be inserted before the decimation and file output.
 
-\subsection{Stereo}
-  Demodulating an FM channel to recover a stereo audio signal requires a more sophisticated approach.
-  In the channel frequency spectrum, there are two main parts of the audio signal - the sum and difference of the left and right audio channel signals.
-  To combine these to a stereo signal, the following equations can be applied.
+The first part of the channel frequency spectrum is the mono audio signal.
+It is the summation signal of the left and right audio channel.
+Thus, it contains enough information to replay an audio stream in mono quality.
+A mono receiver can thus simply replay the so-called multiplex signal that is produced by the FM demodulator, without any further filtering or processing.
+The 19 kHz pilot tone, and any higher frequency components, will not be audible by most people, because it is outside of the range of a human ear.
+Besides that, any used speaker may also be unable to produce a frequency in this range.\\
+
+Consequently, in order to demodulate this spectral part, a lowpass filter is required.
+The filter needs to have a cutoff frequency of 15 kHz, to allow the entire mono audio spectrum to pass.
+Sufficient attenutation must be available before 19 kHz, in order to suppress the pilot tone.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Stereo Audio}
+
+The demodulation of an FM channel to recover a stereo audio signal, requires a more sophisticated approach and multiple steps in signal processing.
+Two parts need to be combined in order to achieve stereo.
+The sum and the difference of the left and right audio channel signals - the mono and the stereo difference part, respectively.
+The following equations need to be applied, to compute a left and a right channel.
 
@@
   \begin{split}
-      L = (L+R) + (L-R) = (2)L \\
-      R = (L+R) - (L-R) = (2)R
+    (L+R) + (L-R) = (2)L \\
+    (L+R) - (L-R) = (2)R
     \label{equ_stereo_from_sum_diff}
@@
 
+where $(L+R)$ represents the mono part and $(L-R)$ the stereo difference.\\
+
 The block diagram in Fig.\ref{fig_bd_stereo_demod} illustrates the signal processing chain for stereo audio.
 In the block diagram, $x(t)$ represents the multiplex signal from the FM demodulator.
-Signals $x_l(t)$ and $x_r(t)$ describe the left and right audio signal, respectively.
+The signals $x_l(t)$ and $x_r(t)$ describe the left and right audio signal, respectively.\\
 
@@
 The central branch selects the pilot tone at 19 kHz with a bandpass filter.
-  This bandpass needs to have a frequency response that is sharp enough to have a sufficient attenuation at $\pm$4 kHz, since this is where the summation and difference signals frequencies are allocated.
-  Afterwards, the pilot tone frequency is doubled via a PLL to achieve a 38 kHz to later demodulate the difference signal phase-coherently.
-  The PLL lock indicator can be re-used as an indicator, whether a pilot tone is existent.
+This bandpass needs to have a frequency response that is sharp enough to have a sufficient attenuation at $\pm$4 kHz, since this is where the mono and difference signals' frequency spectra are allocated.
+Afterwards, the pilot tone frequency is doubled to achieve a 38 kHz subcarrier.
+This guarantees a phase-coherency between pilot and subcarrier, which is required to correctly shift the difference signal down to baseband.
+
+The frequency doubling can be implemented with multiple methods.
+The pilot tone can be modulated by simply multiplying it with itself.
+Another option is to use a PLL and configure it to generate the double frequency at its output.
+The PLL lock indicator can then be re-used as an indicator, whether a pilot tone is existent.\\
+
 In the upper branch, a bandpass filter selects the difference signal with ranges from 23 to 53 kHz.
 It is then modulated to baseband by the previously generated 38 kHz subcarrier.
-  Subsequently, a lowpass limits the signal to the audio signal bandwidth of 15 kHz.
-  The lower branch lowpass-filters the summation signal.
-  The rightmost part in the diagram performs the combination of summation and difference signals, according to Eq.\eqref{equ_stereo_from_sum_diff}.
+Subsequently, a lowpass limits signal bandwidth to 15 kHz, which removes the generated modulation artifacts.\\
+
+The lower branch lowpass-filters the summation, or mono signal, as described in paragraph \ref{subsec:demod_mono}.\\
+
+The rightmost part in the diagram performs the combination of the mono and the difference signal, according to Eq.\eqref{equ_stereo_from_sum_diff}.

### 2021-05-06 doc/thesis/chapters/dsp.tex
update

@@
     \includegraphics[width=9cm]{img/fm-channel-baseband.png}
-  \caption{Allocation of frequencies in an FM channel \cite{ref_fig_channel_freqs}}
+  \caption{Allocation of frequencies in an FM channel \cite{ref_fig_channel_freqs}.}
   \label{fig_channel_baseband_freqs}
@@
 
-\includepicture [0.6] [0] {Frequency discriminator block diagram \cite{ref_fig_freq_discriminator}} {bd_freq_discriminator} {img/discrimination_method_bd1.jpg}
+\includepicture [0.6] [0] {Frequency discriminator block diagram \cite{ref_fig_freq_discriminator}.} {bd_freq_discriminator} {img/discrimination_method_bd1.jpg}
 
@@
 
-\includepicture [0.75] [0] {Time-domain signals in envelope detection \cite{ref_roppel}} {time_domain_envelope_detect} {img/envelope-detect-time-domain.jpg}
+\includepicture [0.75] [0] {Time-domain signals in envelope detection \cite{ref_roppel}.} {time_domain_envelope_detect} {img/envelope-detect-time-domain.jpg}
 
@@
     \includegraphics[width=8.8cm]{img/fm-demod-stereo-audio.png}
-  \caption{Block diagram of FM stereo audio demodulation \cite{ref_roppel}}
+  \caption{Block diagram of FM stereo audio demodulation \cite{ref_roppel}.}
   \label{fig_bd_stereo_demod}

### 2021-05-06 doc/thesis/chapters/dsp.tex
first graphic with draw.io

@@
 
-\includepicture [0.6] [0] {Frequency discriminator block diagram \cite{ref_fig_freq_discriminator}.} {bd_freq_discriminator} {img/discrimination_method_bd1.jpg}
+\includepicture [0.7] [0] {Frequency discriminator block diagram \cite{ref_fig_freq_discriminator}.} {bd_freq_discriminator} {img/bd_freq_discriminator}
 

### 2021-05-11 doc/thesis/chapters/dsp.tex
diagram

@@
 
-\includepicture [0.7] [0] {Frequency discriminator block diagram \cite{ref_fig_freq_discriminator}.} {bd_freq_discriminator} {img/bd_freq_discriminator}
+\includepicture [0.75] [0] {Frequency discriminator block diagram \cite{ref_fig_freq_discriminator}.} {bd_freq_discriminator} {img/draw.io/bd_freq_discriminator}
 

### 2021-05-19 doc/thesis/img/matlab/do_plot.m
plots wip

@@ -1,0 +1,24 @@
+%-------------------------------------------------------------------------
+% File        : do_plot.m
+% Author      : Michael Wurm <wurm.michael95@gmail.com>
+% Description : Plot helper.
+%-------------------------------------------------------------------------
+
+function [fig] = do_plot(fig)
+
+ax = gca;
+ax.XAxisLocation = 'origin';
+ax.YAxisLocation = 'origin';
+%box on;
+title(fig_title, 'FontSize',fontsize);
+Ylm=ylim();                          % get x, y axis limits 
+Xlm=xlim();                          % so can position relative instead of absolute
+Xlb=mean(Xlm);                       % set horizontally at midpoint
+Ylb=0.99*Ylm(1);                     % and just 1% below minimum y value
+xlabel('time [s]','FontSize',fontsize, 'Position',[Xlb Ylb],'HorizontalAlignment','center');
+ylabel('amplitude','FontSize',fontsize);
+set(gca,'XTickLabel',[],'YTickLabel',[])
+legend();
+saveas(fig_time_rf, sprintf("%s/%s",dir_output, "fig_time_rf.png"));
+
+end

### 2021-05-19 doc/thesis/img/matlab/envelope_detect.m

@@ -1,0 +1,146 @@
+%-------------------------------------------------------------------------
+% File        : envelope_detect.m
+% Author      : Michael Wurm <wurm.michael95@gmail.com>
+% Description : Envelope detector time diagrams
+%-------------------------------------------------------------------------
+
+%% Prepare environment
+clear;
+close all;
+clc;
+
+addpath(genpath('../../../../sim/matlab/helpers/'));
+
+%=========================================================================
+%% Settings
+
+% Simulation Options
+EnableFilterAnalyzeGUI = false;
+
+% Common
+fs = 1e6;
+
+n_sec = 0.01;
+tn    = (0:1:fs*n_sec-1).';
+
+% Modulation carrier
+fc  = 15e3;
+A_c = 1;
+
+%=========================================================================
+%% Message signal
+
+fmsg = 1e3;
+Amsg = 1;
+msg  = 1.4 + Amsg * cos(2*pi*fmsg/fs*tn);
+
+%=========================================================================
+%% Modulate
+
+carrier = A_c * cos(2*pi*fc/fs*tn);
+
+rf_signal = msg .* carrier;
+
+%=========================================================================
+%% Thresholding
+
+rf_signal_thresh = rf_signal;             % copy entire signal
+rf_signal_thresh(rf_signal_thresh<0) = 0; % threshold
+
+%=========================================================================
+%% Filter (lowpass)
+
+% Create filter
+Nfilt = 150;
+wcut = 2*fmsg/fs;
+filter_lp = fir1(Nfilt, wcut);
+if EnableFilterAnalyzeGUI
+    fvtool(filter_lp,1);
+end
+
+% Filter
+envelope = filter(filter_lp,1, rf_signal_thresh);
+
+%=========================================================================
+%% Compensate filter delay
+
+grp_delay = (length(filter_lp)-1)/2;
+msg_del = [zeros(grp_delay,1); msg(1:end-grp_delay)];
+
+%=========================================================================
+%% Analysis
+set(0,'defaulttextinterpreter','latex')
+
+% Plots
+%-------------------------------------------------------------------------
+dir_output = './out';
[…]

### 2021-05-19 doc/thesis/img/matlab/envelope_detect.m
done plot

@@ -51,7 +51,7 @@ rf_signal_thresh(rf_signal_thresh<0) = 0; % threshold
 %% Filter (lowpass)
 
 % Create filter
-Nfilt = 150;
+Nfilt = 116;
 wcut = 2*fmsg/fs;
 filter_lp = fir1(Nfilt, wcut);
 if EnableFilterAnalyzeGUI
@@ -59,7 +59,7 @@ if EnableFilterAnalyzeGUI
 end
 
 % Filter
-envelope = filter(filter_lp,1, rf_signal_thresh);
+envelope = filter(filter_lp,1, rf_signal_thresh) * 2;
 
 %=========================================================================
 %% Compensate filter delay
@@ -80,59 +80,30 @@ fig_title = 'RF signal';
 fig_time_rf = figure('Name',fig_title);
 hold on;
 plot(tn/fs, msg,       'r--', 'LineWidth',2, 'DisplayName', 'msg');
-plot(tn/fs, rf_signal, 'b', 'LineWidth',1, 'DisplayName', 'rf\_signal');
-%grid on;
+plot(tn/fs, rf_signal, 'b',   'LineWidth',1, 'DisplayName', 'rf\_signal');
 xlim([-0.1/fmsg,3/fmsg]);
 ylim([-3,3]);
-ax = gca;
-ax.XAxisLocation = 'origin';
-ax.YAxisLocation = 'origin';
-%box on;
-title(fig_title, 'FontSize',fontsize);
-xlabel('time [s]','FontSize',fontsize);
-ylabel('amplitude','FontSize',fontsize);
-set(gca,'XTickLabel',[],'YTickLabel',[])
-legend();
-saveas(fig_time_rf, sprintf("%s/%s",dir_output, "fig_time_rf.png"));
+fig_time_rf = format_plot(fig_time_rf,fig_title,fontsize);
+exportgraphics(fig_time_rf, sprintf("%s/%s",dir_output, "fig_time_rf.png"));
 
 fig_title = 'Half-wave rectified';
 fig_time_rf_thresh = figure('Name',fig_title);
 hold on;
-plot(tn/fs, rf_signal_thresh, 'b', 'DisplayName', 'rf\_signal\_thresh');
-%grid on;
-ax = gca;
-ax.XAxisLocation = 'origin';
-ax.YAxisLocation = 'origin';
+plot(tn/fs, msg,              'r--', 'LineWidth',2, 'DisplayName', 'msg');
[…]

### 2021-05-19 doc/thesis/img/matlab/format_plot.m

@@ -1,24 +1,26 @@
 %-------------------------------------------------------------------------
-% File        : do_plot.m
+% File        : format_plot.m
 % Author      : Michael Wurm <wurm.michael95@gmail.com>
-% Description : Plot helper.
+% Description : Plot format helper.
 %-------------------------------------------------------------------------
 
-function [fig] = do_plot(fig)
+function [fig] = format_plot(fig,fig_title,fontsize)
 
 ax = gca;
 ax.XAxisLocation = 'origin';
 ax.YAxisLocation = 'origin';
-%box on;
+%box off;
+
+% get axis limits for label positions
+Ylm=ylim();
+Xlm=xlim();
+Xlb=mean(Xlm);
+Ylb=0.99*Ylm(1);
+
 title(fig_title, 'FontSize',fontsize);
-Ylm=ylim();                          % get x, y axis limits 
-Xlm=xlim();                          % so can position relative instead of absolute
-Xlb=mean(Xlm);                       % set horizontally at midpoint
-Ylb=0.99*Ylm(1);                     % and just 1% below minimum y value
 xlabel('time [s]','FontSize',fontsize, 'Position',[Xlb Ylb],'HorizontalAlignment','center');
 ylabel('amplitude','FontSize',fontsize);
 set(gca,'XTickLabel',[],'YTickLabel',[])
-legend();
-saveas(fig_time_rf, sprintf("%s/%s",dir_output, "fig_time_rf.png"));
+%legend();
 
 end

### 2021-05-19 doc/thesis/img/matlab/envelope_detect.m
more plot

@@ -16,6 +16,7 @@ addpath(genpath('../../../../sim/matlab/helpers/'));
 
 % Simulation Options
 EnableFilterAnalyzeGUI = false;
+EnableShowFigures      = false;
 
 % Common
 fs = 1e6;
@@ -74,7 +75,7 @@ set(0,'defaulttextinterpreter','latex')
 % Plots
 %-------------------------------------------------------------------------
 dir_output = './out';
-fontsize = 16;
+fontsize = 22;
 
 fig_title = 'RF signal';
 fig_time_rf = figure('Name',fig_title);
@@ -84,7 +85,8 @@ plot(tn/fs, rf_signal, 'b',   'LineWidth',1, 'DisplayName', 'rf\_signal');
 xlim([-0.1/fmsg,3/fmsg]);
 ylim([-3,3]);
 fig_time_rf = format_plot(fig_time_rf,fig_title,fontsize);
-exportgraphics(fig_time_rf, sprintf("%s/%s",dir_output, "fig_time_rf.png"));
+saveas(fig_time_rf, sprintf("%s/%s",dir_output, "fig_1_time_rf.svg"),'svg');
+exportgraphics(fig_time_rf, sprintf("%s/%s",dir_output, "fig_1_time_rf.png"));
 
 fig_title = 'Half-wave rectified';
 fig_time_rf_thresh = figure('Name',fig_title);
@@ -94,7 +96,8 @@ plot(tn/fs, rf_signal_thresh, 'b', 'LineWidth',1, 'DisplayName', 'rf\_signal\_th
 xlim([-0.1/fmsg,3/fmsg]);
 ylim([-1,3]);
 fig_time_rf_thresh = format_plot(fig_time_rf_thresh,fig_title,fontsize);
-exportgraphics(fig_time_rf_thresh, sprintf("%s/%s",dir_output, "fig_time_rf_thresh.png"));
+saveas(fig_time_rf_thresh, sprintf("%s/%s",dir_output, "fig_2_time_rf_thresh.svg"),'svg');
+exportgraphics(fig_time_rf_thresh, sprintf("%s/%s",dir_output, "fig_2_time_rf_thresh.png"));
 
 fig_title = 'Detected envelope';
 fig_time_envelope = figure('Name',fig_title);
@@ -103,14 +106,16 @@ plot(tn/fs, envelope, 'r', 'LineWidth',2, 'DisplayName', 'envelope');
 xlim([-0.1/fmsg,3/fmsg]);
 ylim([-1,3]);
[…]

### 2021-05-19 doc/thesis/img/matlab/envelope_detect.m
better alignment of axis titles

@@ -16,7 +16,9 @@ addpath(genpath('../../../../sim/matlab/helpers/'));
 
 % Simulation Options
 EnableFilterAnalyzeGUI = false;
-EnableShowFigures      = false;
+EnableShowFigures      = true;
+EnablePlotSaveSVG      = true;
+EnablePlotSavePNG      = false;
 
 % Common
 fs = 1e6;
@@ -85,8 +87,8 @@ plot(tn/fs, rf_signal, 'b',   'LineWidth',1, 'DisplayName', 'rf\_signal');
 xlim([-0.1/fmsg,3/fmsg]);
 ylim([-3,3]);
 fig_time_rf = format_plot(fig_time_rf,fig_title,fontsize);
-saveas(fig_time_rf, sprintf("%s/%s",dir_output, "fig_1_time_rf.svg"),'svg');
-exportgraphics(fig_time_rf, sprintf("%s/%s",dir_output, "fig_1_time_rf.png"));
+if EnablePlotSaveSVG saveas(fig_time_rf, sprintf("%s/%s",dir_output, "fig_1_time_rf.svg"),'svg'); end
+if EnablePlotSavePNG exportgraphics(fig_time_rf, sprintf("%s/%s",dir_output, "fig_1_time_rf.png")); end
 
 fig_title = 'Half-wave rectified';
 fig_time_rf_thresh = figure('Name',fig_title);
@@ -96,8 +98,8 @@ plot(tn/fs, rf_signal_thresh, 'b', 'LineWidth',1, 'DisplayName', 'rf\_signal\_th
 xlim([-0.1/fmsg,3/fmsg]);
 ylim([-1,3]);
 fig_time_rf_thresh = format_plot(fig_time_rf_thresh,fig_title,fontsize);
-saveas(fig_time_rf_thresh, sprintf("%s/%s",dir_output, "fig_2_time_rf_thresh.svg"),'svg');
-exportgraphics(fig_time_rf_thresh, sprintf("%s/%s",dir_output, "fig_2_time_rf_thresh.png"));
+if EnablePlotSaveSVG saveas(fig_time_rf_thresh, sprintf("%s/%s",dir_output, "fig_2_time_rf_thresh.svg"),'svg'); end
+if EnablePlotSavePNG exportgraphics(fig_time_rf_thresh, sprintf("%s/%s",dir_output, "fig_2_time_rf_thresh.png")); end
 
 fig_title = 'Detected envelope';
 fig_time_envelope = figure('Name',fig_title);
[…]

### 2021-05-19 doc/thesis/img/matlab/format_plot.m

@@ -6,21 +6,50 @@
 
 function [fig] = format_plot(fig,fig_title,fontsize)
 
-ax = gca;
+%ax = gca;
+ax=fig.CurrentAxes;
 ax.XAxisLocation = 'origin';
 ax.YAxisLocation = 'origin';
 %box off;
 
-% get axis limits for label positions
-Ylm=ylim();
-Xlm=xlim();
-Xlb=mean(Xlm);
-Ylb=0.99*Ylm(1);
+fig.Color='w';
+fig.OuterPosition=fig.InnerPosition;
 
 title(fig_title, 'FontSize',fontsize);
-xlabel('time [s]','FontSize',fontsize, 'Position',[Xlb Ylb],'HorizontalAlignment','center');
-ylabel('amplitude','FontSize',fontsize);
-set(gca,'XTickLabel',[],'YTickLabel',[])
+xlabel('time [s]', 'FontSize',fontsize, 'Position',getLabelPosition('x'),'HorizontalAlignment','center');
+ylabel('amplitude','FontSize',fontsize, 'Position',getLabelPosition('y'),'HorizontalAlignment','center','VerticalAlignment','bottom','Rotation',90);
+
+% integer ticks
+%xticks(unique(round(ax.XTick)));
+yticks(unique(round(ax.YTick)));
+% no tick labels
+set(gca,'XTickLabel',[],'YTickLabel',[]);
+
 %legend();
+
+end
+
+function [pos] = getLabelPosition(label_axis)
+%getLabelPosition - Description
+%
+% Syntax: [x,y] = getLabelPosition(label_axis)
+%
+%         label_axis ... x | y
+%
+
+Ylm=ylim();
+Xlm=xlim();
+
+if label_axis == 'x'
+    x = mean(Xlm);
+    y = 0.99*Ylm(1);
+elseif label_axis == 'y'
+    y = mean(Ylm);
+    x = 0.99*Xlm(1);    
+else
+    error('unknown parameter for label_axis');
+end
+
+pos = [x,y]
 
 end

### 2021-05-20 doc/thesis/chapters/dsp.tex
new bd diag
more plot
overview

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Modulation from Baseband to RF}
+This chapter describes techniques of signal processing, which are required to receive and decode an FM broadcast radio signal.
+The entire signal processing chain from the antenna, through the analog front-end, to the digital processing parts is explained.
+However, the main focus is put on the digital signal processing part, since this is also the main focus of the practical implementation of this thesis.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Modulation from RF to Baseband}
+\section{Overview}
+
+Wireless transmission systems usually transmit their signals over the air.
+Various frequency bands and different types of modulation are being used for this purpose.
+The choice of frequency, as well as the choice of the modulation type, have impacts on the transmission characteristics.
+Depending on these choices, positive effects can be achieved in the amount of data that can be sent over the transmission channel, for example.
+Another advantage is the ability to adapt the ongoing transmission to the changing characteristics of the channel.
+This is especially useful for mobile receivers, since the signals path is inherently changing over time there, which leads to challenging conditions for the receiver.
+Additionally, varying signal strength, or multipath propagation may occur.
+Multipath propagation is caused by reflections of the signal, which - for example - can occur on a house facade, or a large vehicle that is passing by.
+The receiver then gets a signal, that is a summation of the direct path and the reflected path - a signal from multiple paths.
+Besides that, environmental impacts, such as rain or fog have an effect on the transmission channel and thus the received signal.\\
+
+Many more factors could be listed, and an elaboration of the various modulation types can fill entire books.
+However, with all these different facts, there are things that they all have in common.
+On the transmitter side, all these signals are transmitted in a band-limited, high frequency signal.
+The band limitation needs to be given at the transmitting antenna, to minimize interference to other bands in the frequency spectrum.
+The frequency needs to be high, to radiate the signal by an antenna.
+
+On the receiver side, once the signal reaches the antenna of a receiver, there are further similarities.
+In order to decode the actual data content of the signals, they need to be converted down to become a baseband signal.
+Again, there are multiple ways to achieve that.
+In baseband, the signal can then be further processed, until the data content of interest is visible.
+
+The block diagram in Fig.\ref{fig:bd_dsp_overview} gives a high-level overview of a receiver structure.
+
+\includepicture [0.5] [0] {An overview of the signal processing chain.} {bd_dsp_overview} {img/draw.io/bd_dsp_overview}
+
+% TODO: block diagram of receiver
+% Antenna -->
+% --> Analog Frontend (gain, filter, mixer to baseband)
+% --> Digital Signal Processing
+% --> Output (e.g. speakers)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Antenna signal to }
+
+block diagram of superheterodyne
 
@@
 
+\includepicture [1.0] [0] {Time-domain signals in envelope detection \cite{ref_roppel}.} {time_domain_envelope_detect} {img/draw.io/envelope-detect-time-domain}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Demodulation of Broadcast FM}
+%TODO: once FM is converted to AM, traditional AM DSP techniques can be employed
 
@@
   \centering
-    \includegraphics[width=8.8cm]{img/fm-demod-stereo-audio.png}
+    \includegraphics[width=1.0\textwidth]{img/draw.io/bd_fm_demod_stereo_audio}
   \caption{Block diagram of FM stereo audio demodulation \cite{ref_roppel}.}

### 2021-05-20 doc/thesis/chapters/dsp.tex
move overview

@@
 On the receiver side, once the signal reaches the antenna of a receiver, there are further similarities.
-In order to decode the actual data content of the signals, they need to be converted down to become a baseband signal.
-Again, there are multiple ways to achieve that.
-In baseband, the signal can then be further processed, until the data content of interest is visible.
+The antenna signal needs to be amplified and filtered.
+Both steps depend on the antenna characteristics, but usually the received signal is of relatively low power, and the antenna receives a wider range of frequencies than required.
+The filter therefor selects the frequency area that is of interest.
+Another reason for the filter is to avoid spectral replicas in the subsequent modulation, or down-conversion.
+Here, the front-end often down-converts to a low intermediate frequency, instead of directly going to baseband.
+The intermediate frequency is then modulated, or mixed, down to baseband later in the digital domain, which brings advantages in the downstream processing chain.
+%TODO check this or explain more which advantages (HF to BB directly --> i/q complex signal?
+; HF to IF to BB --> real signal ) ??
 
+These three steps - amplification, filtering and down-conversion - are usually done in the analog domain, since digital circuits cannot work with such high frequencies.
+After the analog signal is down-converted to baseband, it needs to be digitized.
+This is done by an analog-digital-converter (ADC).
+
+After the signal is processed to digital baseband, the received signal can finally be decoded to get the actual data content.
+This decoding process depends on the respective modulation type that is used.\\
+
+\noindent
 The block diagram in Fig.\ref{fig:bd_dsp_overview} gives a high-level overview of a receiver structure.
 
-\includepicture [0.5] [0] {An overview of the signal processing chain.} {bd_dsp_overview} {img/draw.io/bd_dsp_overview}
-
-% TODO: block diagram of receiver
-% Antenna -->
-% --> Analog Frontend (gain, filter, mixer to baseband)
-% --> Digital Signal Processing
-% --> Output (e.g. speakers)
+\includepicture [1.0] [0] {An overview of the signal processing chain.} {bd_dsp_overview} {img/draw.io/bd_dsp_overview}
 

### 2021-05-20 doc/thesis/chapters/dsp.tex
new headlines
update

@@
 On the receiver side, once the signal reaches the antenna of a receiver, there are further similarities.
-The antenna signal needs to be amplified and filtered.
+The antenna signal needs to be amplified and filtered at first.
 Both steps depend on the antenna characteristics, but usually the received signal is of relatively low power, and the antenna receives a wider range of frequencies than required.
@@
 Here, the front-end often down-converts to a low intermediate frequency, instead of directly going to baseband.
-The intermediate frequency is then modulated, or mixed, down to baseband later in the digital domain, which brings advantages in the downstream processing chain.
+The intermediate frequency is then shifted down to baseband by a modulator later in the digital domain, which brings advantages in the downstream processing chain.
 %TODO check this or explain more which advantages (HF to BB directly --> i/q complex signal?
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Antenna signal to }
+\section{Receiver types}
 
-block diagram of superheterodyne
+\subsection{heterodyne}
+block diagram
+
+\subsection{super-heterodyne}
+block diagram
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Down-Conversion}
+
+from HF to BB\\
+math. formulas
 

### 2021-05-20 sim/matlab/fm_transceiver.m
fix paths in matlab

@@ -108,7 +108,7 @@ if EnableWriteDataFiles
 
     % Filter coefficients
     disp('--- Filter coefficients to VHDL');
-    folder = '../../hardware/vhdl/src/packages/';
+    folder = '../../hardware/vhdl/rtl/packages';
     writeFilterCoeffsToVHDLFile(filter_bp_pilot, 'filter_bp_pilot',  folder, fp_config);
     writeFilterCoeffsToVHDLFile(filter_lp_mono,  'filter_lp_mono',   folder, fp_config);
     writeFilterCoeffsToVHDLFile(filter_bp_lrdiff,'filter_bp_lrdiff', folder, fp_config);
@@ -123,7 +123,7 @@ if EnableWriteDataFiles
     disp('--- Constants to VHDL/C++/Python');
     writeConstantsToPythonFile('../../hardware/common/tb/packages/fm_global/fm_global.py', 'fm_global_spec', ...
         fp_config, fs, fs_rx, fs_audio, osr_rx, osr_audio, pilot_scale_factor, rx_carrier_38kHz_offset);
-    writeConstantsToVHDLFile(  '../../hardware/vhdl/src/packages/fm_global_spec_pkg.vhd','fm_global_spec', ...
+    writeConstantsToVHDLFile(  '../../hardware/vhdl/rtl/packages/fm_global_spec_pkg.vhd','fm_global_spec', ...
         fp_config, fs, fs_rx, fs_audio, osr_rx, osr_audio, pilot_scale_factor, rx_carrier_38kHz_offset);
     writeConstantsToCPPFile(  '../../hardware/hls/src/fm_global_spec.hpp',               'fm_global_spec', ...
         fp_config, fs, fs_rx, fs_audio, osr_rx, osr_audio, pilot_scale_factor, rx_carrier_38kHz_offset);

### 2021-05-20 sim/matlab/helpers/writeConstantsToCPPFile.m

@@ -40,12 +40,12 @@ fprintf(fileID, "#define FP_WIDTH      ((uint32_t)%d + 1)\n", fp_config.width);
 fprintf(fileID, "#define FP_WIDTH_FRAC ((uint32_t)%d + 1)\n", fp_config.width_frac);
 fprintf(fileID, "#define FP_WIDTH_INT  (FP_WIDTH - FP_WIDTH_FRAC)\n");
 fprintf(fileID, "\n");
-fprintf(fileID, "#define FS            ((uint32_t)%d)\n", fs);
-fprintf(fileID, "#define FS_RX         ((uint32_t)%d)\n", fs_rx);
-fprintf(fileID, "#define FS_AUDIO      ((uint32_t)%d)\n", fs_audio);
+fprintf(fileID, "#define FS       ((uint32_t)%d)\n", fs);
+fprintf(fileID, "#define FS_RX    ((uint32_t)%d)\n", fs_rx);
+fprintf(fileID, "#define FS_AUDIO ((uint32_t)%d)\n", fs_audio);
 fprintf(fileID, "\n");
-fprintf(fileID, "#define OSR_RX        ((uint32_t)%d)\n", osr_rx);
-fprintf(fileID, "#define OSR_AUDIO     ((uint32_t)%d)\n", osr_audio);
+fprintf(fileID, "#define OSR_RX    ((uint32_t)%d)\n", osr_rx);
+fprintf(fileID, "#define OSR_AUDIO ((uint32_t)%d)\n", osr_audio);
 fprintf(fileID, "\n");
 
 fprintf(fileID, "// IP specific\n");

### 2021-05-20 sim/matlab/helpers/writeConstantsToCPPFile.m
work with bitwidths

@@ -36,9 +36,9 @@ fprintf(fileID, [ ...
 fprintf(fileID, "// Global constants for FM Receiver IP and testbench\n");
 
 fprintf(fileID, "\n// General\n");
-fprintf(fileID, "#define FP_WIDTH      ((uint32_t)%d + 1)\n", fp_config.width);
-fprintf(fileID, "#define FP_WIDTH_FRAC ((uint32_t)%d + 1)\n", fp_config.width_frac);
-fprintf(fileID, "#define FP_WIDTH_INT  (FP_WIDTH - FP_WIDTH_FRAC)\n");
+fprintf(fileID, "#define FP_WIDTH      ((uint32_t)%d )\n", fp_config.width);
+fprintf(fileID, "#define FP_WIDTH_FRAC ((uint32_t)%d )\n", fp_config.width_frac);
+fprintf(fileID, "#define FP_WIDTH_INT  ((uint32_t)%d )\n", fp_config.width - fp_config.width_frac);
 fprintf(fileID, "\n");
 fprintf(fileID, "#define FS       ((uint32_t)%d)\n", fs);
 fprintf(fileID, "#define FS_RX    ((uint32_t)%d)\n", fs_rx);

### 2021-05-20 sim/matlab/helpers/writeConstantsToVHDLFile.m

@@ -35,7 +35,7 @@ fprintf(fileID, [ ...
 fprintf(fileID, "  -- General\n");
 fprintf(fileID, "  constant fp_width_spec_c      : natural := %d;\n", fp_config.width);
 fprintf(fileID, "  constant fp_width_frac_spec_c : natural := %d;\n", fp_config.width_frac);
-fprintf(fileID, "  constant fp_width_int_spec_c  : natural := %d;\n", fp_config.width - fp_config.width_frac - 1);
+fprintf(fileID, "  constant fp_width_int_spec_c  : natural := %d;\n", fp_config.width - fp_config.width_frac);
 fprintf(fileID, "\n");
 fprintf(fileID, "  constant fs_spec_c            : natural := %d;\n", fs);
 fprintf(fileID, "  constant fs_rx_spec_c         : natural := %d;\n", fs_rx);

### 2021-07-02 doc/thesis/chapters/introduction.tex
write stuff

--- 2020-12-04 doc/thesis/chapters/introduction.tex
@@
 \label{cha:Introduction}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+\section{Motivation}
+
+In the master's degree program Embedded Systems Design a wide range of topics relate to digital signal processing, chip design and software development.
+All of these disciplines find their application in an embedded system.
+Multiple different methods of implementation are introduced for each topic and their usage is tested in practical sessions.
+However, in the setting of a university course it is not always possible to look at the currently handled topic as a part of a larger, integrated system.
+Instead, it is often treated as a standalone part that fulfills its own, specific functionality.
+
+Therefore, one of the main motivation points of this thesis is to combine several disciplines and multiple implementation methods in a common context of a single project.
+
+In a more general perspective, the current electronics market situation is a huge motivational factor as well.
+Smart devices, smart sensors, Internet of Things (IoT) and many similar terms are well-known and are becoming more and more present in the publics daily lives.
+In some degree, all of these devices represent embedded systems, since they all consist of processing units, combined with sensors and interfaces to communicate with their outside world.
+In the future, the number of devices and the range of applications in this area is likely going to grow.
+This will require a lot of engineering work to advance the state of technology we have today and explore new areas of application.
+Since the development of an embedded system combines so many displines, it is of advantage to have knowledge of as many as possible of them.
+However, not only understanding single parts, but to have an overview and understanding of the entire, integrated system and the interaction between those single parts is important in an embedded system.
+
+Those major reasons lead to the choice of topic for this thesis - understanding all the single parts by themselves, while maintaining the understanding over an interated embedded system that fulfills a task as a whole.
+
+%%%%%%%%%%%%%%%%%%%%%%%
+\section{Choice of Application}
+
+In order to elaborate on the theoretical topics of this thesis, a comprehensive practical project is developed along with it.
+The project of choice is a digital FM broadcast radio receiver.
+It combines several disciplines that appear in the master's program Embedded Systems Design, such as digital signal processing (DSP), software development and system design and thus is a perfect candidate for this project.
+
+FM broadcast radio is chosen as the input signal, because it is a common radio frequency signal that is available almost everywhere.
+Furthermore, the demodulated signal is an audio signal, usually music or speech, that can be listened to.
+This is subjectively more attractive than a generic data stream.
+Additionally, it is a comparably simple signal to demodulate, but still requires several techniques of signal processing to decode it.
+
+\section{Goal}
+
+The main goal of this thesis is to elaborate a system design for an embedded system, that combines all the knowledge of different disciplines and topics that are accumulated over the time of studies in a single project.
+
+This includes the development of a system model in a tool like Matlab, the practical usage and implementation of DSP in an FPGA, the development of a digital receiver and the deployment on actual hardware.
+
+Furthermore, the receiver implementation on the FPGA is to be done in two variants.
+The first variant is traditional, manually written hardware description language (HDL) in the programming language VHDL.
+For the second variant, high-level synthesis (HLS) is employed, to generate the low-level HDL from high-level C++ code.
+These two methods are implemented and compared on the basis of usability and some metrics.
+
+develop an FM receiver and explain it, so that it can be followed in a tutorial with an affordable budget in hardware.\\
+show multiple ways of how to implement the same thing, in different levels of abstraction. Weigh each ways' field of application, efficiency and applicability.
+
+%%%%%%%%%%%%%%%%%%%%%%%
+% NOTES:
+%%%%%%%%%%%%%%%%%%%%%%%
 

### 2021-07-02 doc/thesis/chapters/introduction.tex
introduction done for now

@@
 
-\section{Goal}
+\section{Goals}
 
@@
 
-develop an FM receiver and explain it, so that it can be followed in a tutorial with an affordable budget in hardware.\\
-show multiple ways of how to implement the same thing, in different levels of abstraction. Weigh each ways' field of application, efficiency and applicability.
+In the view of a system design approach, another focus is put on a sustainable testing and development strategy for the entire system.
+Above mentioned implementation variants should optimally share as much as possible of verification code, in order to save time and effort in implementation.
+Also, as many tasks as possible should be automated by scripts, to save time in manual labor for the software engineers.
+All that enables an efficient development cycle of the product.
+
+By the end of this thesis, the implementation of an FM radio receiver in multiple methods and abstraction levels should be achieved, as well as a good understanding of the overall system level design.
+Knowledge of each methods' preferred field of application, efficiency and application should be gained.
+
+%develop an FM receiver and explain it, so that it can be followed in a tutorial with an affordable budget in hardware.\\
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%
-
-Thesis Design Decisions: \\
-\\
-Main focus on system design. \\
-\\
-How to achieve the same result in 3 (4?) different levels of abstraction?
-
-\begin{enumerate}
-  \item GnuRadio (?)
-  \item Matlab/Python
-  \item HLS\\
-      Implementation targeting Xilinx' ZedBoard.\\
-      Optionally: \\
-      Intel as a comparison (synth only, no HW) - "How platform-independent is HLS?"
-  \item VHDL
-\end{enumerate}
-\vspace{.5 cm}
-
-Compare:
-- Implementation effort (difficulty)
-- Implementation time
-- Usefulness on Target Systems (uC, SoC, RasPi, etc)
-- others?
+%
+%Thesis Design Decisions: \\
+%\\
+%Main focus on system design. \\
+%\\
+%How to achieve the same result in 3 (4?) different levels of abstraction?
+%
+%\begin{enumerate}
+%  \item GnuRadio (?)
+%  \item Matlab/Python
+%  \item HLS\\
+%      Implementation targeting Xilinx' ZedBoard.\\
+%      Optionally: \\
+%      Intel as a comparison (synth only, no HW) - "How platform-independent is HLS?"
+%  \item VHDL
+%\end{enumerate}
+%\vspace{.5 cm}
+%
+%Compare:
+%- Implementation effort (difficulty)
+%- Implementation time
+%- Usefulness on Target Systems (uC, SoC, RasPi, etc)
+%- others?

### 2021-07-02 doc/thesis/chapters/dsp.tex
write stuff
minor adaptions; write heterodyne

@@
 The entire signal processing chain from the antenna, through the analog front-end, to the digital processing parts is explained.
-However, the main focus is put on the digital signal processing part, since this is also the main focus of the practical implementation of this thesis.
+However, the main focus is put on the digital signal processing part, since this is a major part of the practical implementation of this thesis.
 
@@
 \section{Overview}
+%TODO: this is copied from introduction:
+%Evolution from AM to FM because of certain advantages, etc.\\
+%Explain FM usage/existence nowadays (geographical; frequencies; devices; new %standard DAB, etc)
 
@@
 Another advantage is the ability to adapt the ongoing transmission to the changing characteristics of the channel.
-This is especially useful for mobile receivers, since the signals path is inherently changing over time there, which leads to challenging conditions for the receiver.
+This is especially useful for mobile receivers, since the signals' path is inherently changing over time there, which leads to challenging conditions for the receiver.
 Additionally, varying signal strength, or multipath propagation may occur.
@@
 
-Many more factors could be listed, and an elaboration of the various modulation types can fill entire books.
+Many more factors could be listed, and elaborations over the various modulation types fill entire books.
 However, with all these different facts, there are things that they all have in common.
@@
 Another reason for the filter is to avoid spectral replicas in the subsequent modulation, or down-conversion.
-Here, the front-end often down-converts to a low intermediate frequency, instead of directly going to baseband.
+Here, the front-end often down-converts to a low intermediate frequency (IF), instead of directly going to baseband.
 The intermediate frequency is then shifted down to baseband by a modulator later in the digital domain, which brings advantages in the downstream processing chain.
@@
 
-These three steps - amplification, filtering and down-conversion - are usually done in the analog domain, since digital circuits cannot work with such high frequencies.
+These three steps - amplification, filtering and down-conversion to an IF - are often done in the analog domain, since digital circuits cannot work with such high frequencies.
 After the analog signal is down-converted to baseband, it needs to be digitized.
@@
 
-\includepicture [1.0] [0] {An overview of the signal processing chain.} {bd_dsp_overview} {img/draw.io/bd_dsp_overview}
+\includepicture [1.0] [0] {High-level block diagram of a signal processing chain.} {bd_dsp_overview} {img/draw.io/bd_dsp_overview}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Receiver types}
+\section{Superheterodyne Receiver}
 
-\subsection{heterodyne}
-block diagram
+Multiple different receiver structures exist and are implemented in receivers.
+A very common structure is the so-called superheterodyne receiver.
+In the literature, short forms like \textit{superhet}, or simply \textit{super} are found.
 
-\subsection{super-heterodyne}
-block diagram
+The main characteristic of a superhet is, that it converts the radio-frequency (RF) antenna signal to an IF in a first stage, instead of going to baseband directly.\\
+
+The IF is usually set to a fixed frequency.
+This results in the advantage, that the subsequent IF filter can be optimized, since it does not have to be flexible in frequency.
+
+Fig.\ref{fig:bd_superhet} shows a block diagram of a superhet receiver.
+The antenna signal first runs through an RF amplifier and filter.
+It selects the desired frequency band and filters out any unwanted signals outside of this band.
+The next step is the mixer, which is fed by a local oscillator.
+This shifts, or down-converts, the antenna signal down to the IF.
+Afterwards, the IF signal is amplified and filtered again.
+This is to remove the side-products of the mixer and may also be called as image rejection filter.
+The demodulator stage can now perform any further signal processing on the signal, until the required signal is retrieved.
+In the shown case, which may be an FM broadcast receiver, an audio output is generated.
+
+\includepicture [1.0] [0] {Block design of a superhetorodyne receiver.} {bd_superhet} {img/superheterodyne-receiver-block-diagram.png}
+%https://www.electronics-notes.com/articles/radio/superheterodyne-receiver/block-diagram.php
 

### 2021-07-02 doc/thesis/chapters/dsp.tex
receiver types (see literature)

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Superheterodyne Receiver}
+\section{Receiver types}
 
-Multiple different receiver structures exist and are implemented in receivers.
-A very common structure is the so-called superheterodyne receiver.
-In the literature, short forms like \textit{superhet}, or simply \textit{super} are found.
+Multiple different receiver architectures exist and are implemented in receivers.
+Two common examples are explained in the following sections.
 
+\subsection{Direct-Conversion Receiver}
+
+A direct-conversion receiver, as its name suggests, converts an RF signal to baseband directly.
+It can also be called homodyne, synchrodyne, or zero-IF receiver.
+
+In order to achieve direct down-conversion, the local oscillator of the mixer needs to be set exactly to the center frequency of the desired signal.
+
+\subsection{Superheterodyne Receiver}
+
+The superheterodyne receiver is often found in the literature with short forms like \textit{superhet}, or simply \textit{super}.
 The main characteristic of a superhet is, that it converts the radio-frequency (RF) antenna signal to an IF in a first stage, instead of going to baseband directly.\\
@@
 
-from HF to BB\\
-math. formulas
+The process of down-conversion is used to convert an RF signal to a signal of lower frequency, while preserving its information content.
+This is the explanation in the time domain.
+However, it may be easier look at the process in the frequency domain.
+In the frequency domain, the band-limited RF signal is shifted down to a lower frequency.
+This lower frequency can either be zero, i.e. baseband, or an intermediate frequency, such as in the application of a superheterodyne receiver.
+
+math. formulas\\
+digital: meet nyquist criteria, just like any other DSP (show in spectrum)\\
 

### 2021-07-05 doc/thesis/chapters/dsp.tex
superhet and zero-if

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Superheterodyne Receiver}
+
+The superheterodyne receiver is often referred to as \textit{superhet}, or \textit{super} in the literature.
+The main characteristic of a superhet is, that it converts the radio-frequency (RF) antenna signal to an IF in a first stage, instead of going to baseband directly.\\
+
+The IF is usually set to a fixed frequency, which results in the advantage, that the subsequent IF filter can be optimized, since it does not have to be adapted in frequency.
+Therefore, a superhet receiver is able to select narrow band signals, even if the bands' surrounding frequencies are being used by other signals.
+This is one of the superhets' main advantages.
+A disadvantage is that the signal noise around the image frequency band is translated into the IF output band.
+Thus, the filter requirements are relatively high, since they need to deliver a strong attenuation in order to suppress noise outside the desired band.
+
+Fig.\ref{fig:bd_superhet} shows a block diagram of a superhet receiver.
+The antenna signal first runs through a band selection filter.
+It selects the desired frequency band and filters out any unwanted signals outside of this band.
+A low-noise amplifier (LNA) may be used to boost the signal strenght.
+The next step is the mixer, which is fed by a local oscillator (LO).
+This shifts, or down-converts, the antenna signal down to the IF.
+Afterwards, the IF signal is amplified and filtered again to remove the side-products of the mixer, namely the image replicas.
+Thus, it is also called an image rejection filter.
+The demodulator stage can now perform any further signal processing on the signal, until the required signal is retrieved \cite{PassosFábio2020AHSo}.\\
+
+The depicted block diagram may represent an FM broadcast receiver, where an audio output is generated from the antenna input.
+
+\includepicture [1.0] [0] {Block design of a superhetorodyne receiver \cite[Fig.2.1]{PassosFábio2020AHSo}} {bd_superhet} {img/superheterodyne-receiver-block-diagram-2.png}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Direct-Conversion Receiver}
@@
 
-\subsection{Superheterodyne Receiver}
+The main advantage of direct-conversion receivers is, that their output is an image-free signal.
+This is due to the properties of the mixer.
+The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer output, and the flicker noise, also called $\frac{1}{f}$-noise, which induces strong noise at low frequencies \cite{PassosFábio2020AHSo}.
+These effects are not explained into more detail here, since this is not the focus of this thesis.
 
-The superheterodyne receiver is often found in the literature with short forms like \textit{superhet}, or simply \textit{super}.
-The main characteristic of a superhet is, that it converts the radio-frequency (RF) antenna signal to an IF in a first stage, instead of going to baseband directly.\\
-
-The IF is usually set to a fixed frequency.
-This results in the advantage, that the subsequent IF filter can be optimized, since it does not have to be flexible in frequency.
-
-Fig.\ref{fig:bd_superhet} shows a block diagram of a superhet receiver.
-The antenna signal first runs through an RF amplifier and filter.
-It selects the desired frequency band and filters out any unwanted signals outside of this band.
-The next step is the mixer, which is fed by a local oscillator.
-This shifts, or down-converts, the antenna signal down to the IF.
-Afterwards, the IF signal is amplified and filtered again.
-This is to remove the side-products of the mixer and may also be called as image rejection filter.
-The demodulator stage can now perform any further signal processing on the signal, until the required signal is retrieved.
-In the shown case, which may be an FM broadcast receiver, an audio output is generated.
-
-\includepicture [1.0] [0] {Block design of a superhetorodyne receiver.} {bd_superhet} {img/superheterodyne-receiver-block-diagram.png}
-%https://www.electronics-notes.com/articles/radio/superheterodyne-receiver/block-diagram.php
+\includepicture [0.9] [0] {Block design of a direct-conversion receiver \cite[Fig.2.3]{PassosFábio2020AHSo}.} {bd_direct_conversion} {img/direct-conversion-receiver-block-diagram.png}
 

### 2021-07-05 doc/thesis/chapters/dsp.tex
cite references

@@
 
+(TODO: this is amplitude modulation!) -->see \cite[sec 4.2 AM]{MatinMohammadA2018CSfE} \\
+
 The process of down-conversion is used to convert an RF signal to a signal of lower frequency, while preserving its information content.
@@
 \end{equation}
-Several formulas, as well as their derivations are described according to \cite{ref_FM_Maths_Info_1}\cite{ref_FMMaths2}.
+Several formulas, as well as their derivations are described according to \cite[pg.54-55]{MatinMohammadA2018CSfE}.
 
@@
 The frequency band that is used for FM broadcasting is defined worldwide and spans from 87.5 to 108 MHz.
-However, some countries only partially use the band \cite{ref_itu_regulations}.
+However, some countries only partially use the band \cite[RR5-39]{ItuRadioRegulations}.
 The range is located within the so-called Very-High-Frequency (VHF) band, which is open to the public in terms of usage.
 This means, that any transmitter may use the specified frequency range freely.
-Austria allowed the legal usage in 2006 \cite{ref_austria_rundfunkgesetz_2006}.
+Austria allowed the legal usage in 2006 \cite{AustriaRundfunkgesetz2014}.
 However, the transmission power needs to be limited, so that neighboring receivers are not disturbed in receiving any existing channels.
-The European Commision specifies the power limit as 50 nW of effective radiated power (ERP) \cite{ref_eu_commision_radio_spectrum}.
+The European Commision specifies the power limit as 50 nW of effective radiated power (ERP) \cite{EuCommissionRadioSpectrumHarmonization}.
 Another limitation is, that a transmitter must be capable to transmit on multiple center frequencies within the broadcasting band.
@@
 In a practical usecase this means that a transmitter needs to select a channel or frequency range, that is not already occupied by an official transmitter.
-If a free range is found, the transmitter may start sending an FM signal with the mentioned power limitations \cite{ref_ebu_fm_modulators_in_europe_power_regulations}.
+If a free range is found, the transmitter may start sending an FM signal with the mentioned power limitations \cite{EbuFmRegulators}.
 
@@
 The negative part is an exact mirror, because the transmitted and received signals from the antenna are real signals - they do not have an imaginary part.
-The spectrum consists of multiple parts. Therefore, it is also referred to as `multiplex signal' in the literature. Several parts are described in the following paragraphs.
+The spectrum consists of multiple parts.
+Therefore, it is also referred to as `multiplex signal' in the literature.
+Several parts are described in the following paragraphs \cite{EbuFmRegulators}\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard}.
 

### 2021-07-05 doc/thesis/chapters/dsp.tex
lots more bib refs

@@
 Therefore, it is also referred to as `multiplex signal' in the literature.
-Several parts are described in the following paragraphs \cite{EbuFmRegulators}\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard}.
+Several parts are described in the following paragraphs \cite{EbuFmRegulators}\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard}\cite{ElectronicsNotesFmTutorial}.
 
@@
     \includegraphics[width=9cm]{img/fm-channel-baseband.png}
-  \caption{Allocation of frequencies in an FM channel \cite{ref_fig_channel_freqs}.}
+  \caption{Allocation of frequencies in an FM channel \cite{FigFmChannelFreqSpectrum}.}
   \label{fig_channel_baseband_freqs}
@@
 It is used to generate a stereo audio signal, in combination with the mono signal.
-This process is explained into more detail in chapter \ref{sec_fm_sig_demod}.
+This process is explained into more detail in chapter \ref{sec:demodulation-of-broadcast-fm}.
 
@@
 
-Services that were implemented are the Data Radio Channel (DARC), which is mostly used in Japan and the USA, the Subsidary Communication Authorization (SCA) and the Radio Data System (RDS) \cite{ref_rohde_u_schwarz}.
+Services that were implemented are the Data Radio Channel (DARC), which is mostly used in Japan and the USA, the Subsidary Communication Authorization (SCA) and the Radio Data System (RDS) \cite{RohdeSchwarzFM}.
 Out of these, RDS is the most significant service in Europe.
-It is used to transmit additional information about the channel, such as the radio stations' name, the currently playing songs' title or traffic information.\\
-
-\noindent
-The information about the frequency spectrum was taken from \cite{ref_fm_broadcast_tutorial_and_basics}.
+It is used to transmit additional information about the channel, such as the radio stations' name, the currently playing songs' title or traffic information.
 
@@
 
-\includepicture [0.75] [0] {Frequency discriminator block diagram \cite{ref_fig_freq_discriminator}.} {bd_freq_discriminator} {img/draw.io/bd_freq_discriminator}
+\includepicture [0.75] [0] {Frequency discriminator block diagram \cite{FigBdFreqDiscriminator}.} {bd_freq_discriminator} {img/draw.io/bd_freq_discriminator}
 
@@
 
-An important requirement for the differentiator is, that the input signal is of a constant amplitude \cite{ref_schnyder_haller}.
+An important requirement for the differentiator is, that the input signal is of a constant amplitude \cite{SchnyderHaller2002}.
 The transmitted signal is subject to random additive noise on the transmission channel.
@@
 
-The method implemented in Fig.\ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as `Asynchronous Half-Wave Envelope Detector'\cite{ref_envelope_detector}.
+The method implemented in Fig.\ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as `Asynchronous Half-Wave Envelope Detector' \cite{DSPRelatedEnvelopeDetection}.
 It consists of a thresholding unit to remove negative values and a conventional lowpass filter.
@@
 Therefor, the previous method only needs to swap the thresholding unit with an unit that calculates the absolut value.
-In that architecture, a full-wave rectification is performed on the signal, which significantly improves the envelope detection accuracy \cite{ref_envelope_detector}.
+In that architecture, a full-wave rectification is performed on the signal, which significantly improves the envelope detection accuracy \cite{DSPRelatedEnvelopeDetection}.
 The envelope is followed more exactly, because of the higher power that is available in the signal, since the entire signal is taken into account, and not only the positive half, as in the previous half-wave method.
 
-\includepicture [0.75] [0] {Time-domain signals in envelope detection \cite{ref_roppel}.} {time_domain_envelope_detect} {img/envelope-detect-time-domain.jpg}
+\includepicture [0.75] [0] {Time-domain signals in envelope detection \cite{RoppelBegleitmaterial}.} {time_domain_envelope_detect_orig} {img/envelope-detect-time-domain.jpg}
 
-\includepicture [1.0] [0] {Time-domain signals in envelope detection \cite{ref_roppel}.} {time_domain_envelope_detect} {img/draw.io/envelope-detect-time-domain}
+\includepicture [1.0] [0] {Time-domain signals in envelope detection \cite{RoppelBegleitmaterial}.} {time_domain_envelope_detect} {img/draw.io/envelope-detect-time-domain}
 
@@
 In the case of FM demodulation, the loop filter in the feedback branch is configured to be able to follow the frequency variations on the input, which is directly related to the encoded information.
-The output of the PLL directly delivers an AM signal, which corresponds to the transmitted information. \cite{ref_schnyder_haller}
+The output of the PLL directly delivers an AM signal, which corresponds to the transmitted information. \cite{SchnyderHaller2002}
 
@@
 \section{Demodulation of Broadcast FM}
+\label{sec:demodulation-of-broadcast-fm}
 %TODO: once FM is converted to AM, traditional AM DSP techniques can be employed
@@
     \includegraphics[width=1.0\textwidth]{img/draw.io/bd_fm_demod_stereo_audio}
-  \caption{Block diagram of FM stereo audio demodulation \cite{ref_roppel}.}
+  \caption{Block diagram of FM stereo audio demodulation \cite{RoppelBegleitmaterial}.}
   \label{fig_bd_stereo_demod}

### 2021-07-07 doc/thesis/chapters/dsp.tex
wip

@@
 
-(TODO: this is amplitude modulation!) -->see \cite[sec 4.2 AM]{MatinMohammadA2018CSfE} \\
-
-The process of down-conversion is used to convert an RF signal to a signal of lower frequency, while preserving its information content.
+The process of down-conversion is a technique of amplitude modulation (AM).
+It is used to convert an RF signal to a signal of lower frequency, while preserving its information content.
 This is the explanation in the time domain.
@@
 
-math. formulas\\
-digital: meet nyquist criteria, just like any other DSP (show in spectrum)\\
+In order to achieve this, the incoming signal needs to be multiplicated with a carrier signal, that is generated locally in the receiver, by a local oscillator (LO).
+Depending on the frequency $f_{LO}$ of this local carrier, the resulting signal will be shifted to baseband, or an IF.
+
+A block diagram, as well as the signal spectrum of the process is shown in Fig.\ref{fig:bd_down_conversion_spectrum}.
+
+%\includepicture [0.6] [0] {Block diagram of a down-conversion \cite[Fig.4.3]{MatinMohammadA2018CSfE}.} {bd_down_conversion_spectrum} {img/bd-down-conversion-AM-dsb-sc}
+\includepicture [1.0] [0] {Signal spectrum before the down-conversion.} {bd_down_conversion_spectrum} {img/down-conversion-AM}
+
+\subsection{Mathematical description}
+
+The procedure that is used for down-conversion is an amplitude modulation (AM).
+The equation for a generic AM signal is $ S_{AM}(t) = A_m m(t) \cdot A_c c(t) $, where $m(t)$ represents the message signal, and $c(t)$ is the high-frequency carrier.
+Their respective amplitudes are $A_m$ and $A_c$.
+This equation is depicted as a block diagram and signal spectrum in Fig.\ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.
+
+The signal $m(t)$ is a baseband signal.
+It may be antenna signal that was filtered by a bandpass filter.
+The received signal is a real signal, and thus its negative frequency spectrum is an exact mirror of the positive frequencies.
+This fact can be explained in two simple equations.
+
+It is well-known that the real signal $cos(wt) = \frac{1}{2} \Big( e^{jw} + e^{-jw} \Big)$, which follows from $e^{jw} = cos(wt)+jsin(wt)$.
+Looking at the frequencies of the first equation, the negative and positive parts can be seen.\\
+
+The process of down-conversion is described in the following formulas.
+The denotation of the subscripts is assuming a direct conversion to baseband.
+
+\begin{equation}
+  m_{BB}(t) = m(t) \cdot c(t)
+\end{equation}
+
+\includepicture [1.0] [0] {Signal spectrum after the down-conversion.} {down_conversion_spectrum} {img/down-conversion-AM-dsb-sc}
+
+\section{Nyquist criteria}
+
+in digital: any DSP must meet nyquist criteria \\
+show aliasing in spectrum\\
 

### 2021-07-07 doc/thesis/chapters/dsp.tex
nyquist done

@@
 \begin{equation}
-  m_{BB}(t) = m(t) \cdot c(t)
+  m_{BB}(t) = A_m m(t) \cdot A_c c(t)
 \end{equation}
 
-\includepicture [1.0] [0] {Signal spectrum after the down-conversion.} {down_conversion_spectrum} {img/down-conversion-AM-dsb-sc}
+Assuming that $m(t) = cos(\omega_mt)$, $c(t) = cos(\omega_ct)$ and $A_m = A_c = 1$,
 
-\section{Nyquist criteria}
+\begin{equation}
+  m_{BB}(t) = cos(\omega_mt) \cdot cos(\omega_ct)
+\end{equation}
 
-in digital: any DSP must meet nyquist criteria \\
-show aliasing in spectrum\\
+follows.
+By applying the mathematical theorem
+
+\begin{equation}
+  cos(\alpha) \cdot cos(\beta) = \frac{1}{2} \cdot \bigl[ cos(\alpha-\beta) + cos(\alpha+\beta) \bigr]
+\end{equation}
+
+the final formula of the conversion can be represented as
+
+\begin{equation}
+ m_{BB}(t) = \frac{1}{2} \cdot \Bigl[ cos\bigl((w_m-w_c)t\bigr) + cos\bigl((w_m+w_c)t\bigr) \Bigr]
+\end{equation}
+
+The graphical representation in the frequency spectrum is shown in Fig.\ref{fig:down_conversion_spectrum_steps}.
+As mentioned above, the assumption is a direct conversion to baseband, so $w_m = w_c$.
+Looking at the positive original part, it is shown that only half of its energy is translated into the desired baseband.
+However, half of the energy of the original negative part is also shifted to baseband.
+The superposition of the two parts results in an amount of energy that is again equal to the original signal.
+
+\includepicture [1.0] [0] {Signal spectrum, explaining the process of down-conversion.} {down_conversion_spectrum_steps} {img/down-conversion-AM-dsb-sc}
+
+\section{Sampling theorem}
+
+Signal processing is often done in the digital domain, using digital signals.
+A digital signal is obtained by sampling an analog signal in a periodic interval, the so-called sampling frequency $f_s$.
+
+The sampling frequency needs to meet the sampling theorem, in order to correctly represent the analog signal.
+The sampling theorem, by Nyquist and Shannon mainly consists of the following formula.
+
+\begin{equation}
+  f_{s} > 2*f_{signal}
+\end{equation}
+
+In words: The sampling frequency must be higher than twice the frequency of the signals' highest frequency content.
+Any digital signal processing technique must hold the specification of the sampling theorem \cite{AlessioSilviaMaria2016DSPa}.
+
+In case the theorem is violated, the effect of aliasing occurs on the generated signal.
+The sampling process introduces spectral replicas at integer multiples of the sampling frequency $f_s$.
+These replicas overlap, if $f_s$ does not meet the theorem, which results in signal distortion.
+Fig.\ref{fig:aliasing_freq_spectrum} illustrates the effect of aliasing in the frequency spectrum \cite{ThyagarajanK.S2019ItDS}.
+
+\includepicture [0.9] [0] {
+  Signal spectrum, explaining the process of down-conversion.\\
+  Diagram A) $f_s$ is high enough and thus meets the theorem.\\
+  Diagram B) $f_s$ exactly meets the theorem.\\
+  Diagram C) $f_s$ is too low and thus creates overlaps of the spectral replicas - aliasing \cite{ThyagarajanK.S2019ItDS}.
+} {aliasing_freq_spectrum} {img/sampling-aliasing}
 

### 2021-07-07 doc/thesis/chapters/dsp.tex
downsampling

@@
 This is due to the properties of the mixer.
-The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer output, and the flicker noise, also called $\frac{1}{f}$-noise, which induces strong noise at low frequencies \cite{PassosFábio2020AHSo}.
+The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer output, and the flicker noise, also called \textit{1/f}-noise, which induces strong noise at low frequencies \cite{PassosFábio2020AHSo}.
 These effects are not explained into more detail here, since this is not the focus of this thesis.
@@
 In words: The sampling frequency must be higher than twice the frequency of the signals' highest frequency content.
-Any digital signal processing technique must hold the specification of the sampling theorem \cite{AlessioSilviaMaria2016DSPa}.
+Any digital signal processing technique must hold the specification of the sampling theorem \cite[chpt. 4.2]{AlessioSilviaMaria2016DSPa}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Oversampling / Sample Rate Reduction}
+\section{Sample Rate Reduction}
 %TODO: explain DSP functions that are actually used in Matlab FM demod
-Example: \\
-In case the multiplex signal needs to be stored in a file with 44.1 kHz samplerate, in order to avoid aliasing, a lowpass with a cutoff frequency of 15 kHz and sufficient attenutation before 19 kHz needs to be inserted before the decimation and file output.
+
+Sample rate reduction is also called downsampling.
+Downsampling is a way to reduce the amount of data that needs to be processed, without losing any of the signals' information.
+In some applications, the digital signal is oversampled by a large factor.
+However, if the frequency spectrum of the desired signal is a narrow band, a high oversampling rate is not necessary and the signal can be downsampled without losing any information content.
+To guarantee this, the Nyquist sampling theorem must still be met after the downsampling process.
+
+A lower sample rate is especially useful for efficient filter design.
+Factors such as the transition bandwidth or the width of the passband are important there.
+Short transition bandwidths, or a narrow passband require a high order of the filter to fulfill the specifications, which translates into a high computational effort.
+Thus, it is of advantage to downsample the input, in order to generate better conditions for the filter design.
+Downsampling spreads the spectrum of interest, in terms of the relative bandwidth between passband and sampling frequency.
+Therefore, the filter can be designed with wider transition bandwidth and passband, which significantly reduces the required filter order.
+
+As an example, assuming the signal of interest has a bandwidth of 20 kHz.
+The Nyquist theorem thus requires a remaining sampling rate of at least twice the signals' frequency, so 40 kHz.
+The signal is sampled with an oversampling factor of 10 - meaning a sample rate of 200 kHz.
+A downsampling factor of 4 is now applied, which results in a sample rate of 50 kHz.
+The resulting signal is now downsampled and can be processed at this lower data rate, while none of the original data content is lost.\\
+
+The process consists of two parts: filtering and skipping samples.
+Let us assume downsampling by a factor of $M=4$.
+In the first step, a low-pass filter with a normalized cut-off frequency of \textit{1/M} is applied.
+This is required in order to avoid aliasing later in the process.
+The second step simply skips a number of M-1 samples in the sequence.
+
+More detailed information can be found in \cite[chpt. 6.9]{AlessioSilviaMaria2016DSPa}, \cite[chpt. 4.1]{GopiE.S2018MDSP}
+\cite[chpt. 10.2.2]{ThyagarajanK.S2019ItDS}.
 

### 2021-07-07 sim/matlab/research/decimation_test.m

@@ -49,7 +49,7 @@ y1_dec = decimate(signal, osr, 'fir');
 %% 2: Manual, using filter and pick every nth sample.
 
 % Low-pass filter
-Nfilt = 30;
+Nfilt = 50;
 filter_dec = fir1(Nfilt, 1/osr);
 y2_man_filt = filter(filter_dec,1, signal);
 
@@ -62,7 +62,7 @@ y2_man = y2_man_filt(1:osr:end);
 %=========================================================================
 %% Analysis
 
-skip_end = 3;
+skip_end = 10;
 
 fprintf("norm resample-decimate : %.3f\n", norm(y0_res(1:end-skip_end) - y1_dec(1:end-skip_end)));
 fprintf("norm own-resample      : %.3f\n", norm(y0_res(1:end-skip_end) - y2_man(1:end-skip_end)));

### 2021-07-07 doc/thesis/chapters/dsp.tex
downsampling done

@@
 Factors such as the transition bandwidth or the width of the passband are important there.
-Short transition bandwidths, or a narrow passband require a high order of the filter to fulfill the specifications, which translates into a high computational effort.
-Thus, it is of advantage to downsample the input, in order to generate better conditions for the filter design.
+Short transition bandwidths, or a narrow passband require a high filter order to fulfill the specifications.
+This translates into a high computational effort.
+In such cases, it is of advantage to downsample the input, in order to create better conditions for the filter design.
 Downsampling spreads the spectrum of interest, in terms of the relative bandwidth between passband and sampling frequency.
-Therefore, the filter can be designed with wider transition bandwidth and passband, which significantly reduces the required filter order.
+Therefore, the filter can be designed with wider transition bandwidth and passband, which significantly reduces the required filter order.\\
 
@@
 This is required in order to avoid aliasing later in the process.
-The second step simply skips a number of M-1 samples in the sequence.
+The second step simply skips a number of M-1 samples in the sequence.\\
 
-More detailed information can be found in \cite[chpt. 6.9]{AlessioSilviaMaria2016DSPa}, \cite[chpt. 4.1]{GopiE.S2018MDSP}
-\cite[chpt. 10.2.2]{ThyagarajanK.S2019ItDS}.
+More detailed information can be found in \cite[chpt. 6.9]{AlessioSilviaMaria2016DSPa}, \cite[chpt. 4.1]{GopiE.S2018MDSP} and \cite[chpt. 10.2.2]{ThyagarajanK.S2019ItDS}.
 

### 2021-07-08 doc/thesis/chapters/dsp.tex
update
draw own diagrams

@@
 
-\includepicture [1.0] [0] {Block design of a superhetorodyne receiver \cite[Fig.2.1]{PassosFábio2020AHSo}} {bd_superhet} {img/superheterodyne-receiver-block-diagram-2.png}
+\includepicture [1.0] [0] {Block design of a superhetorodyne receiver \cite[Fig.2.1]{PassosFábio2020AHSo}} {bd_superhet} {img/draw.io/superheterodyne-receiver-block-diagram}
 
@@
 
-\includepicture [0.9] [0] {Block design of a direct-conversion receiver \cite[Fig.2.3]{PassosFábio2020AHSo}.} {bd_direct_conversion} {img/direct-conversion-receiver-block-diagram.png}
+\includepicture [1.0] [0] {Block design of a direct-conversion receiver \cite[Fig.2.3]{PassosFábio2020AHSo}.} {bd_direct_conversion} {img/draw.io/direct-conversion-receiver-block-diagram}
 
@@
 A block diagram, as well as the signal spectrum of the process is shown in Fig.\ref{fig:bd_down_conversion_spectrum}.
+In this example, the frequency of the LO is set to shift the signal from $f_m$ to baseband.
 
-%\includepicture [0.6] [0] {Block diagram of a down-conversion \cite[Fig.4.3]{MatinMohammadA2018CSfE}.} {bd_down_conversion_spectrum} {img/bd-down-conversion-AM-dsb-sc}
-\includepicture [1.0] [0] {Signal spectrum before the down-conversion.} {bd_down_conversion_spectrum} {img/down-conversion-AM}
+\includepicture [1.0] [0] {Block diagram and signal spectrum of down-conversion.} {bd_down_conversion_spectrum} {img/draw.io/down-conversion-AM-spectrum-and-bd}
 
@@
 \begin{equation}
-  f_{s} > 2*f_{signal}
+  f_{s} > 2 \cdot f_{signal}
 \end{equation}

### 2021-07-08 doc/thesis/chapters/dsp.tex
new diagrams

@@
 Depending on the frequency $f_{LO}$ of this local carrier, the resulting signal will be shifted to baseband, or an IF.
+Note, that in the case of down-conversion, the local oscillator frequency $f_{LO}$ is interchangeably denoted as carrier frequency $f_c$.
 
@@
 
-\includepicture [1.0] [0] {Block diagram and signal spectrum of down-conversion.} {bd_down_conversion_spectrum} {img/draw.io/down-conversion-AM-spectrum-and-bd}
+\includepicture [1.3] [0] {Block diagram and signal spectrum of down-conversion.} {bd_down_conversion_spectrum} {img/draw.io/down-conversion-AM-spectrum-and-bd}
 
@@
 
-\includepicture [1.0] [0] {Signal spectrum, explaining the process of down-conversion.} {down_conversion_spectrum_steps} {img/down-conversion-AM-dsb-sc}
+\includepicture [1.0] [0] {Signal spectrum, explaining the process of down-conversion.} {down_conversion_spectrum_steps} {img/draw.io/down-conversion-steps-AM-spectrum}
 

### 2021-07-08 doc/thesis/chapters/dsp.tex
diagrams

@@
 
-\includepicture [1.3] [0] {Block diagram and signal spectrum of down-conversion.} {bd_down_conversion_spectrum} {img/draw.io/down-conversion-AM-spectrum-and-bd}
+\includepicture [1.0] [0] {Block diagram and signal spectrum of down-conversion.} {bd_down_conversion_spectrum} {img/draw.io/down-conversion-AM-spectrum-and-bd}
 
@@
 The sampling theorem, by Nyquist and Shannon mainly consists of the following formula.
-
 \begin{equation}
@@
   Diagram C) $f_s$ is too low and thus creates overlaps of the spectral replicas - aliasing \cite{ThyagarajanK.S2019ItDS}.
-} {aliasing_freq_spectrum} {img/sampling-aliasing}
+} {aliasing_freq_spectrum} {img/draw.io/sampling-aliasing}
 

### 2021-07-08 doc/thesis/chapters/dsp.tex
more diagrams

@@
 \includepicture [0.9] [0] {
-  Signal spectrum, explaining the process of down-conversion.\\
-  Diagram A) $f_s$ is high enough and thus meets the theorem.\\
-  Diagram B) $f_s$ exactly meets the theorem.\\
-  Diagram C) $f_s$ is too low and thus creates overlaps of the spectral replicas - aliasing \cite{ThyagarajanK.S2019ItDS}.
+  Frequency spectrum of the sampled baseband signal with its replicas at integer multiples of $f_s$ \cite{ThyagarajanK.S2019ItDS}.\\
+  A) $f_s > 2f_{signal}$: $f_s$ is high enough and thus meets the theorem.\\
+  B) $f_s = 2f_{signal}$: $f_s$ exactly meets the theorem.\\
+  C) $f_s < 2f_{signal}$: $f_s$ is too low and thus creates overlaps of the replicas - aliasing.
 } {aliasing_freq_spectrum} {img/draw.io/sampling-aliasing}

### 2021-07-08 doc/thesis/chapters/dsp.tex
one more

@@
 
-Fig.\ref{fig_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
+Fig.\ref{fig:fm_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
 For a better overview, only the positive spectum part is shown.
@@
 
-\begin{figure}[!h]
-  \centering
-    \includegraphics[width=9cm]{img/fm-channel-baseband.png}
-  \caption{Allocation of frequencies in an FM channel \cite{FigFmChannelFreqSpectrum}.}
-  \label{fig_channel_baseband_freqs}
-\end{figure}
+\includepicture [0.9] [0] {Allocation of frequencies in an FM channel \cite{FigFmChannelFreqSpectrum}.
+} {fm_channel_baseband_freqs} {img/draw.io/fm-channel-baseband-freqs}
 
@@
 The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly.
-A frequency spectrum of the multiplex signal of an FM channel is shown in Fig.\ref{fig_channel_baseband_freqs}.
+A frequency spectrum of the multiplex signal of an FM channel is shown in Fig.\ref{fig:fm_channel_baseband_freqs}.
 

### 2021-07-09 sim/matlab/helpers/format_plot.m
dsp chapter done, so far

@@ -50,6 +50,6 @@ else
     error('unknown parameter for label_axis');
 end
 
-pos = [x,y]
+pos = [x,y];
 
 end

### 2021-07-09 sim/matlab/research/envelope_detect.m

@@ -9,15 +9,18 @@ clear;
 close all;
 clc;
 
-addpath(genpath('../../../../sim/matlab/helpers/'));
+addpath(genpath('../helpers/'));
 
 %=========================================================================
 %% Settings
+
+dir_output = '../../../doc/thesis/img/matlab';
 
 % Simulation Options
 EnableFilterAnalyzeGUI = false;
 EnableShowFigures      = true;
 EnablePlotSaveSVG      = true;
+EnablePlotSavePDF      = false;
 EnablePlotSavePNG      = false;
 
 % Common
@@ -76,39 +79,41 @@ set(0,'defaulttextinterpreter','latex')
 
 % Plots
 %-------------------------------------------------------------------------
-dir_output = './out';
 fontsize = 22;
 
 fig_title = 'RF signal';
 fig_time_rf = figure('Name',fig_title);
 hold on;
-plot(tn/fs, msg,       'r--', 'LineWidth',2, 'DisplayName', 'msg');
+plot(tn/fs, msg,       'r--', 'LineWidth',3, 'DisplayName', 'msg');
 plot(tn/fs, rf_signal, 'b',   'LineWidth',1, 'DisplayName', 'rf\_signal');
 xlim([-0.1/fmsg,3/fmsg]);
 ylim([-3,3]);
 fig_time_rf = format_plot(fig_time_rf,fig_title,fontsize);
 if EnablePlotSaveSVG saveas(fig_time_rf, sprintf("%s/%s",dir_output, "fig_1_time_rf.svg"),'svg'); end
+if EnablePlotSavePDF saveas(fig_time_rf, sprintf("%s/%s",dir_output, "fig_1_time_rf.pdf"),'pdf'); end
 if EnablePlotSavePNG exportgraphics(fig_time_rf, sprintf("%s/%s",dir_output, "fig_1_time_rf.png")); end
 
 fig_title = 'Half-wave rectified';
 fig_time_rf_thresh = figure('Name',fig_title);
 hold on;
-plot(tn/fs, msg,              'r--', 'LineWidth',2, 'DisplayName', 'msg');
-plot(tn/fs, rf_signal_thresh, 'b', 'LineWidth',1, 'DisplayName', 'rf\_signal\_thresh');
+plot(tn/fs, msg,              'r--', 'LineWidth',3, 'DisplayName', 'msg');
[…]

### 2021-07-09 doc/thesis/chapters/deployment-on-hardware.tex
Thesis chapter "signal processing theory" text. (#20)

--- 2020-12-04 doc/thesis/chapters/fm_radio.tex
@@
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Signal Processing Theory}
-\label{cha:TheThesis}
-
-  \section{Up-Downconversion to/from RF}
-
-  \section{Frequency Modulation}
-  see literature/Present\_lec6\_AM\_FM.pdf\\
-  see lectures out of HSD/ESD
-
-  \section{Algorithms for Digital FM Demodulation}
-    see literature/FmDemodulator.pdf (Sect. 3.3)
-    see literature/00476180 Digital FM Demodulator for FM, TV, and Wireless.pdf (Sect. II and III)
-
-    \subsection{Baseband Delay Demodulator}
-    \subsection{Phase-Adapter Demodulator}
-    \subsection{Phase-Locked Loop (PLL)}
-    \subsection{Mixed Demodulator}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{High-Level Synthesis}
-  \section{Introduction / State of the Art}
-  \section{Functionality (high level code to HDL)}
-  \section{Variants}
-    \subsection{C++}
-    \subsection{SystemC}
-  \section{Advantages / Disadvantages}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Implementation}
-  \section{Matlab Simulation}
-  in fixed point, close to hardware level algorithm
-
-  \section{VHDL}
-    \subsection{Channel Selection (IF to Channel-BB)}
-    \subsection{Phase Detector}
-    \subsection{other Elements}
-
-  \section{HLS}
-    \subsection{Channel Selection (IF to Channel-BB)}
-    \subsection{Phase Detector}
-    \subsection{other Elements}
-
-  \section{Comparison of Metrics}
-    \subsection{Implementation effort/time}
-    \subsection{Hardware Utilization}
-    \subsection{Others}
-
 %%%%%%%%%%%%%%%%%%%%
 \chapter{Deployment on Hardware}
+\label{cha:DeploymentOnHardware}
   \section{Hardware Platform}

### 2021-07-09 doc/thesis/chapters/dsp.tex
Thesis chapter "signal processing theory" text. (#20)

@@
 
-\includepicture [0.9] [0] {
+\includepicture [0.95] [0] {
   Frequency spectrum of the sampled baseband signal with its replicas at integer multiples of $f_s$ \cite{ThyagarajanK.S2019ItDS}.\\
@@
 
-\includepicture [0.9] [0] {Allocation of frequencies in an FM channel \cite{FigFmChannelFreqSpectrum}.
-} {fm_channel_baseband_freqs} {img/draw.io/fm-channel-baseband-freqs}
+\includepicture [1.0] [0] {Allocation of frequencies in an FM channel \cite{FigFmChannelFreqSpectrum}.} {fm_channel_baseband_freqs} {img/draw.io/fm-channel-baseband-freqs}
 
@@
 
-\includepicture [0.75] [0] {Time-domain signals in envelope detection \cite{RoppelBegleitmaterial}.} {time_domain_envelope_detect_orig} {img/envelope-detect-time-domain.jpg}
-
-\includepicture [1.0] [0] {Time-domain signals in envelope detection \cite{RoppelBegleitmaterial}.} {time_domain_envelope_detect} {img/draw.io/envelope-detect-time-domain}
+\includepicture [1.0] [0] {Time-domain signals in envelope detection \cite{RoppelBegleitmaterial}.} {time_domain_envelope_detect} {img/draw.io/envelope-detect-time-domain.png}
 

### 2021-07-09 doc/thesis/chapters/introduction.tex
Thesis chapter "signal processing theory" text. (#20)

@@
 
-Therefore, one of the main motivation points of this thesis is to combine several disciplines and multiple implementation methods in a common context of a single project.
+Therefore, one of the main motivation points of this thesis is to combine several disciplines and multiple implementation methods in a common context of a single project.\\
 
 In a more general perspective, the current electronics market situation is a huge motivational factor as well.
-Smart devices, smart sensors, Internet of Things (IoT) and many similar terms are well-known and are becoming more and more present in the publics daily lives.
+Smart devices, smart sensors, Internet of Things (IoT) and many similar terms are well-known and are becoming more and more present in the publics' daily lives.
 In some degree, all of these devices represent embedded systems, since they all consist of processing units, combined with sensors and interfaces to communicate with their outside world.
-In the future, the number of devices and the range of applications in this area is likely going to grow.
-This will require a lot of engineering work to advance the state of technology we have today and explore new areas of application.
+In the future, the number of devices and the range of applications in this area is going to grow from todays' point of view.
+This will require a lot of engineering work to advance the state of technology we have today and explore new areas of application.\\
+
 Since the development of an embedded system combines so many displines, it is of advantage to have knowledge of as many as possible of them.
-However, not only understanding single parts, but to have an overview and understanding of the entire, integrated system and the interaction between those single parts is important in an embedded system.
+However, not only understanding single parts, but to have an overview and understanding of the entire, integrated system and the interaction between those single parts is important in an embedded system.\\
 
-Those major reasons lead to the choice of topic for this thesis - understanding all the single parts by themselves, while maintaining the understanding over an interated embedded system that fulfills a task as a whole.
+Those major reasons lead to the choice of topic for this thesis - understanding all the single parts by themselves, while maintaining the understanding over an integrated embedded system that fulfills a task as a whole.
 
@@
 The project of choice is a digital FM broadcast radio receiver.
-It combines several disciplines that appear in the master's program Embedded Systems Design, such as digital signal processing (DSP), software development and system design and thus is a perfect candidate for this project.
+It combines several disciplines that appear in the master's program Embedded Systems Design, such as digital signal processing (DSP), software development and system design and thus is a perfect candidate for this project.\\
 
@@
 
-The main goal of this thesis is to elaborate a system design for an embedded system, that combines all the knowledge of different disciplines and topics that are accumulated over the time of studies in a single project.
+The main goal of this thesis is to elaborate a system design for an embedded system, that combines all the knowledge of different disciplines and topics which are accumulated over the time of the university program, in a single project.
 
-This includes the development of a system model in a tool like Matlab, the practical usage and implementation of DSP in an FPGA, the development of a digital receiver and the deployment on actual hardware.
+This includes the development of a system model in a tool like Matlab, the practical usage and implementation of DSP in an FPGA, the development of a digital receiver and the deployment on actual hardware.\\
 
-Furthermore, the receiver implementation on the FPGA is to be done in two variants.
-The first variant is traditional, manually written hardware description language (HDL) in the programming language VHDL.
-For the second variant, high-level synthesis (HLS) is employed, to generate the low-level HDL from high-level C++ code.
+The receiver implementation on the FPGA is to be done in two variants.
+The first variant is traditional, manually written hardware description language (HDL) in the programming language VHDL, as it has been done for years.
+For the second variant, high-level synthesis (HLS) is employed, to generate the low-level HDL from high-level \cplusplus code.
 These two methods are implemented and compared on the basis of usability and some metrics.
@@
 In the view of a system design approach, another focus is put on a sustainable testing and development strategy for the entire system.
-Above mentioned implementation variants should optimally share as much as possible of verification code, in order to save time and effort in implementation.
-Also, as many tasks as possible should be automated by scripts, to save time in manual labor for the software engineers.
-All that enables an efficient development cycle of the product.
+Above mentioned implementation variants should optimally share as much as possible of testing and verification code, in order to save time and effort in implementation.
+Also, as many tasks as possible should be automated by scripts, to save time in manual labor for the engineers working on the development.
+All that enables an efficient development cycle of the product.\\
 

### 2021-07-09 doc/thesis/chapters/results.tex
Thesis chapter "signal processing theory" text. (#20)

--- 2020-12-04 doc/thesis/chapters/fm_radio.tex
@@
 %%%%%%%%%%%%%%%%%%%%
-\chapter{Signal Processing Theory}
-\label{cha:TheThesis}
-
-  \section{Up-Downconversion to/from RF}
-
-  \section{Frequency Modulation}
-  see literature/Present\_lec6\_AM\_FM.pdf\\
-  see lectures out of HSD/ESD
-
-  \section{Algorithms for Digital FM Demodulation}
-    see literature/FmDemodulator.pdf (Sect. 3.3)
-    see literature/00476180 Digital FM Demodulator for FM, TV, and Wireless.pdf (Sect. II and III)
-
-    \subsection{Baseband Delay Demodulator}
-    \subsection{Phase-Adapter Demodulator}
-    \subsection{Phase-Locked Loop (PLL)}
-    \subsection{Mixed Demodulator}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{High-Level Synthesis}
-  \section{Introduction / State of the Art}
-  \section{Functionality (high level code to HDL)}
-  \section{Variants}
-    \subsection{C++}
-    \subsection{SystemC}
-  \section{Advantages / Disadvantages}
-
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Implementation}
-  \section{Matlab Simulation}
-  in fixed point, close to hardware level algorithm
-
-  \section{VHDL}
-    \subsection{Channel Selection (IF to Channel-BB)}
-    \subsection{Phase Detector}
-    \subsection{other Elements}
-
-  \section{HLS}
-    \subsection{Channel Selection (IF to Channel-BB)}
-    \subsection{Phase Detector}
-    \subsection{other Elements}
-
-  \section{Comparison of Metrics}
+\chapter{Test Results}
+\label{cha:Results}
+  \section{Functionality}
+  % Functionality same?
+  % Sim speed of x min of recorded FM signal
+  % lines of code
   \subsection{Implementation effort/time}
@@
 
-%%%%%%%%%%%%%%%%%%%%
-\chapter{Deployment on Hardware}
-  \section{Hardware Platform}
-    \subsection{RTL2832u Dongle}
-    \subsection{ZedBoard}
+  % TODO Überlegungen in diesem Kapitel
+  % kann Demodulator geändert werden? Bsp. von FM zu DAB+, QAM, etc..
+  % - von matlab heraus VHDL generieren, etc

### 2021-07-09 sim/matlab/helpers/format_plot.m

@@ -1,0 +1,55 @@
+%-------------------------------------------------------------------------
+% File        : format_plot.m
+% Author      : Michael Wurm <wurm.michael95@gmail.com>
+% Description : Plot format helper.
+%-------------------------------------------------------------------------
+
+function [fig] = format_plot(fig,fig_title,fontsize)
+
+%ax = gca;
+ax=fig.CurrentAxes;
+ax.XAxisLocation = 'origin';
+ax.YAxisLocation = 'origin';
+%box off;
+
+fig.Color='w';
+fig.OuterPosition=fig.InnerPosition;
+
+title(fig_title, 'FontSize',fontsize);
+xlabel('time [s]', 'FontSize',fontsize, 'Position',getLabelPosition('x'),'HorizontalAlignment','center');
+ylabel('amplitude','FontSize',fontsize, 'Position',getLabelPosition('y'),'HorizontalAlignment','center','VerticalAlignment','bottom','Rotation',90);
+
+% integer ticks
+%xticks(unique(round(ax.XTick)));
+yticks(unique(round(ax.YTick)));
+% no tick labels
+set(gca,'XTickLabel',[],'YTickLabel',[]);
+
+%legend();
+
+end
+
+function [pos] = getLabelPosition(label_axis)
+%getLabelPosition - Description
+%
+% Syntax: [x,y] = getLabelPosition(label_axis)
+%
+%         label_axis ... x | y
+%
+
+Ylm=ylim();
+Xlm=xlim();
+
+if label_axis == 'x'
+    x = mean(Xlm);
+    y = 0.99*Ylm(1);
+elseif label_axis == 'y'
+    y = mean(Ylm);
+    x = 0.99*Xlm(1);    
+else
+    error('unknown parameter for label_axis');
+end
+
+pos = [x,y];
+
+end

### 2021-07-09 sim/matlab/research/dsbsc.m

@@ -39,7 +39,7 @@ A_tx_c = 1;
 
 % Rx carrier
 fc_rx  = 57e3;       % <-- ADAPT FREQUENCY ERROR
-phi_rx = 180*pi/180;  % <-- ADAPT PHASE ERROR
+phi_rx = 160*pi/180;  % <-- ADAPT PHASE ERROR
 A_rx_c = 1;
 
 %=========================================================================
@@ -80,7 +80,7 @@ if EnableEqirippleFIR
     filter_name = sprintf("./filters/lowpass_rx.mat");
     ripple_pass_dB = 0.1;                % Passband ripple in dB
     ripple_stop_db = 50;                 % Stopband ripple in dB
-    cutoff_freqs   = [fmsg*5 fmsg*20];   % Cutoff frequencies
+    cutoff_freqs   = [fmsg 3*fmsg];   % Cutoff frequencies
     
     filter_lp_rx = getLPfilter( ...
         filter_name, ...
@@ -126,6 +126,7 @@ fprintf('fft_bin: %.2f Hz\n', fft_bin);
 fft_freqs = (-Nfft/2:Nfft/2-1)/(Nfft*1/fs);
 
 fft_msg    = abs(fftshift( fft( hanning(length(msg)) .* msg,    Nfft) ))/Nfft;
+fft_msg_rx_demod = abs(fftshift( fft( hanning(length(msg)) .* rx_msg_demod, Nfft) ))/Nfft;
 fft_msg_rx = abs(fftshift( fft( hanning(length(msg)) .* rx_msg, Nfft) ))/Nfft;
 
 
@@ -197,13 +198,14 @@ xline(-fmsg,'k--','-fmsg');
 xline( fmsg,'k--','fmsg');
 yline(abs(cos(phi_rx))*max(fft_msg), 'k--', 'power loss due to Rx phase error')
 h0 = plot(fft_freqs, fft_msg,    'b', 'DisplayName', 'msg tx');
-h1 = plot(fft_freqs, fft_msg_rx, 'r', 'DisplayName', 'msg rx');
+h1 = plot(fft_freqs, fft_msg_rx_demod, 'm', 'DisplayName', 'msg rx demod');
+h2 = plot(fft_freqs, fft_msg_rx, 'r', 'DisplayName', 'msg rx');
 grid on;
-xlim([0,fmsg*2]);
+xlim([-fmsg*4,fmsg*4]);
 title(fig_title);
 xlabel('Frequency [Hz]');
[…]

### 2021-07-09 sim/matlab/research/envelope_detect.m

@@ -1,0 +1,129 @@
+%-------------------------------------------------------------------------
+% File        : envelope_detect.m
+% Author      : Michael Wurm <wurm.michael95@gmail.com>
+% Description : Envelope detector time diagrams
+%-------------------------------------------------------------------------
+
+%% Prepare environment
+clear;
+close all;
+clc;
+
+addpath(genpath('../helpers/'));
+
+%=========================================================================
+%% Settings
+
+dir_output = '../../../doc/thesis/img/matlab';
+
+% Simulation Options
+EnableFilterAnalyzeGUI = false;
+EnableShowFigures      = true;
+EnablePlotSaveSVG      = true;
+EnablePlotSavePDF      = false;
+EnablePlotSavePNG      = false;
+
+% Common
+fs = 1e6;
+
+n_sec = 0.01;
+tn    = (0:1:fs*n_sec-1).';
+
+% Modulation carrier
+fc  = 15e3;
+A_c = 1;
+
+%=========================================================================
+%% Message signal
+
+fmsg = 1e3;
+Amsg = 1;
+msg  = 1.4 + Amsg * cos(2*pi*fmsg/fs*tn);
+
+%=========================================================================
+%% Modulate
+
+carrier = A_c * cos(2*pi*fc/fs*tn);
+
+rf_signal = msg .* carrier;
+
+%=========================================================================
+%% Thresholding
+
+rf_signal_thresh = rf_signal;             % copy entire signal
+rf_signal_thresh(rf_signal_thresh<0) = 0; % threshold
+
+%=========================================================================
+%% Filter (lowpass)
+
+% Create filter
+Nfilt = 116;
+wcut = 2*fmsg/fs;
+filter_lp = fir1(Nfilt, wcut);
+if EnableFilterAnalyzeGUI
+    fvtool(filter_lp,1);
+end
+
+% Filter
+envelope = filter(filter_lp,1, rf_signal_thresh) * 2;
+
+%=========================================================================
+%% Compensate filter delay
+
+grp_delay = (length(filter_lp)-1)/2;
+msg_del = [zeros(grp_delay,1); msg(1:end-grp_delay)];
+
[…]

### 2021-07-09 doc/thesis/chapters/system-architecture-and-concept.tex
write stuff
text update

@@
 %%%%%%%%%%%%%%%%%%%%
-\chapter{System Architecture/Concept}
-\label{cha:SystemArchitecture/Concept}
+\chapter{System Architecture and Concept}
+\label{cha:SystemArchitectureAndConcept}
 
+This chapter introduces the FM radio receiver project that is implemented as a major part of this thesis.
+It specifically talks about the system level approach that is taken in the implementation of the project.
+The main concept, as well as the architecture are explained.
+
+%%%%%%%%%%%%%%%%%%%%
+\section{Overview}
+
+In the development of a project that includes multiple functional parts, it is always of advantage to start with a system level overview.
+Major decisions can be made at this stage, before any implementation has begun.
+This can have a direct impact on the efficiency of the entire development, which concequently has an impact on the final quality of the product.
+A thorough system level design can also prevent from issues that would otherwise arise while the development is already ongoing.
+As an example it could turn out that one of the functional parts is not able to fit into the integrated system, in the way it is implemented.
+This leads to the necessity to re-engineer this entire functional block.
+As a result, it takes more effort to implement the functionality, which directly correlates with the development cost and the time-to-market, in order to deliver the final product.
+Summing up, a comprehensive system level architecture and concept can pave the way for a successful implementation of a project from start to finish.\\
+
+In order to create a system level design, considerations around the high-level, final goal of the product are a good starting point.
+This includes the definition of its functional range.
+Once this is done, the target system or target hardware should be considered, because there are various ways to implement a certain functionality.
+All of these variants will eventually result in the same output, but will, however, heavily differ in their implementation effort, efficiency, or their applicability for the products' aim in general.
+Thus, things like the available hardware platform, implementation time, degree of optimization, output quality, and similar things are important factors at this stage.
+Of course, there are many more factors to consider when a product is to be developed.
+However, only a subset of these is covered here, since this chapter is focused on how this thesis' accompanying project was developed and is thus written in that context.\\
+
+%%%%%%%%%%%%%%%%%%%%
+\section{The project}
+
+The aim of this thesis is to implement an FM radio receiver in multiple different methods and ways of implementation.
+The final goal thereby is always be the same, that is: Listen to the music of a radio station that is transmitted over FM broadcast radio.
+
+In the context of a system level design approach, several steps are necessary to find a solution that is ready to be implemented.
+As a very first step, the inner workings of an FM receiver need to be understood.
+This mainly regards the signal processing theory, to be able to decode the signal that is received by the antenna.
+Chapter \ref{cha:SignalProcessingTheory} describes this part into detail.
+
+As already mentioned above, multiple ways of technology and implementation can lead to a similar result.
+Thus, the next step is to define a set of possible ways, which will actually be implemented.
+For this thesis, the decision was made to implement the FM radio receiver, using the following methods.
+
+\begin{itemize}
+  \item \textbf{GNU Radio} \cite{GnuRadio}\\
+      GNU Radio is a free and open-source software development toolkit to implement a software-defined radio.
+The implementation is done by connecting functional blocks in a block design.
+This abstracts the inner workings of each functional block and thus provides an implementation approach on a very high level of abstraction.
+  \item Matlab
+  \item FPGA: \cplusplus High-Level Synthesis
+  \item FPGA: VHDL
+\end{itemize}
+
+Once the basic theory is understood, implementation can begin.
+
+% list of variants
+%-  implementation of a model and DSP concept in Matlab \\
+%-  implementation in VHDL and HLS \\
+%-  both use the same testbench source \\
+%-  both are integrated into FPGA
+%-  the FPGA output is read back into the firmware, for analysis \\
+
+%%%%%%%%%%%%%%%%%%%%
 \section{Block Diagram (with details)}
+%%%%%%%%%%%%%%%%%%%%
 \subsection{describe blocks}
 
-  \section{Which parts in HLS and VHDL}
-
+%%%%%%%%%%%%%%%%%%%%
 \section{Test Environment}

### 2021-07-09 doc/thesis/chapters/system-architecture-and-concept.tex
more text

@@
 
-The aim of this thesis is to implement an FM radio receiver in multiple different methods and ways of implementation.
+The aim of this thesis is develop a system architecture and concept, that allows the implementation of an FM radio receiver in multiple different ways, while providing an elegant way to compare the different solutions.
 The final goal thereby is always be the same, that is: Listen to the music of a radio station that is transmitted over FM broadcast radio.
@@
 \begin{itemize}
-  \item \textbf{GNU Radio} \cite{GnuRadio}\\
+  \item \textbf{GNU Radio}\\
       GNU Radio is a free and open-source software development toolkit to implement a software-defined radio.
@@
       This abstracts the inner workings of each functional block and thus provides an implementation approach on a very high level of abstraction.
-  \item Matlab
-  \item FPGA: \cplusplus High-Level Synthesis
-  \item FPGA: VHDL
+      GNU Radio supports interfaces to external RF hardware \cite{SoftwareGnuRadio}.
+  \item \textbf{Matlab}\\
+      Matlab is a software platform that can work with matrices and compute numerical solutions.
+      Matlab also includes various toolboxes, that support developers with complex calculations and tasks on datasets.
+      This includes a range of toolboxes for signal processing, such as filter designers or FFT functions, as well as convenient ways to visualize data \cite{SoftwareMatlab}.
+  \item \textbf{FPGA: \cplusplus High-Level Synthesis}\\
+      In order to use an FPGA, a hardware design needs to be developed for it.
+      This can be done using high-level synthesis.
+      High-level synthesis (HLS) is explained in detail in Chapter \ref{cha:HLS}.
+  \item \textbf{FPGA: VHDL}\\
+      Similar to the HLS approach, this is a variant to write a hardware design for an FPGA.
+      The language VHDL is used in this approach.
 \end{itemize}
 
-Once the basic theory is understood, implementation can begin.
+In order to implement an FM radio receiver, all of the listed methods require deep knowledge about the neccessary details of signal processing.
+However, the implementation of the various different methods require more or less knowledge and effort on the exact underlying implementation of each signal processing part.
+This depends on their level of abstraction.
 
-% list of variants
 %-  implementation of a model and DSP concept in Matlab \\
@@
 %%%%%%%%%%%%%%%%%%%%
-\section{Block Diagram (with details)}
+\section{Block Diagram}
+
 %%%%%%%%%%%%%%%%%%%%

### 2021-07-09 doc/thesis/chapters/implementation.tex
write stuff
text update
add some thoughts

@@
 \label{cha:Implementation}
-\section{Matlab/Python Model}
+
+\section{Detailed block diagram of digital signal processing chain}
+
+\includepicture [1.0] [0] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/bd_detailed_dsp}
+
+\section{Matlab Model}
+
 in fixed point, close to hardware level algorithm
+
+\subsection{Transmitter block diagram}
+
+\section{GNU Radio}
+  \subsection{Introduction}
+  \subsection{Transmitter}
+  \subsection{Receiver}
 
@@
   The \#pragmas need to be within the code, at the appropriate positions, which makes it difficult to write compiler independent code....
+
+  TODO: \\
+  - comparison of VHDL vs HLS\\
+  - in the point of view:\\
+  - system integration (IP generation, AXI interfaces),\\
+  - logic utilization\\
+--> see introduction.tex

### 2021-07-09 doc/thesis/chapters/system-architecture-and-concept.tex
add some thoughts

@@
 The aim of this thesis is develop a system architecture and concept, that allows the implementation of an FM radio receiver in multiple different ways, while providing an elegant way to compare the different solutions.
-The final goal thereby is always be the same, that is: Listen to the music of a radio station that is transmitted over FM broadcast radio.
+The final goal thereby is always be the same, that is: Listen to the music of a radio station that is transmitted over FM broadcast radio.\\
 
@@
 As already mentioned above, multiple ways of technology and implementation can lead to a similar result.
-Thus, the next step is to define a set of possible ways, which will actually be implemented.
-For this thesis, the decision was made to implement the FM radio receiver, using the following methods.
+Thus, the next step is to define a set of possible ways which will actually be implemented.
+For this thesis, the decision was made to implement the FM radio receiver by using the following methods.
 
@@
 However, the implementation of the various different methods require more or less knowledge and effort on the exact underlying implementation of each signal processing part.
-This depends on their level of abstraction.
+This depends on their respective level of abstraction.
 

### 2021-07-09 doc/thesis/chapters/dsp.tex
text update
\includepicture function with floating option; new block diagram

@@
 \chapter{Signal Processing Theory}
-\label{cha:TheThesis}
+\label{cha:SignalProcessingTheory}
 
@@
 
-\includepicture [1.0] [0] {High-level block diagram of a signal processing chain.} {bd_dsp_overview} {img/draw.io/bd_dsp_overview}
+\includepicture [1.0] {High-level block diagram of a signal processing chain.} {bd_dsp_overview} {img/draw.io/bd_dsp_overview}
 
@@
 
-\includepicture [1.0] [0] {Block design of a superhetorodyne receiver \cite[Fig.2.1]{PassosFábio2020AHSo}} {bd_superhet} {img/draw.io/superheterodyne-receiver-block-diagram}
+\includepicture [1.0] {Block design of a superhetorodyne receiver \cite[Fig.2.1]{PassosFábio2020AHSo}} {bd_superhet} {img/draw.io/superheterodyne-receiver-block-diagram}
 
@@
 
-\includepicture [1.0] [0] {Block design of a direct-conversion receiver \cite[Fig.2.3]{PassosFábio2020AHSo}.} {bd_direct_conversion} {img/draw.io/direct-conversion-receiver-block-diagram}
+\includepicture [1.0] {Block design of a direct-conversion receiver \cite[Fig.2.3]{PassosFábio2020AHSo}.} {bd_direct_conversion} {img/draw.io/direct-conversion-receiver-block-diagram}
 
@@
 
-\includepicture [1.0] [0] {Block diagram and signal spectrum of down-conversion.} {bd_down_conversion_spectrum} {img/draw.io/down-conversion-AM-spectrum-and-bd}
+\includepicture [1.0] {Block diagram and signal spectrum of down-conversion.} {bd_down_conversion_spectrum} {img/draw.io/down-conversion-AM-spectrum-and-bd}
 
@@
 
-\includepicture [1.0] [0] {Signal spectrum, explaining the process of down-conversion.} {down_conversion_spectrum_steps} {img/draw.io/down-conversion-steps-AM-spectrum}
+\includepicture [1.0] {Signal spectrum, explaining the process of down-conversion.} {down_conversion_spectrum_steps} {img/draw.io/down-conversion-steps-AM-spectrum}
 
@@
 
-\includepicture [0.95] [0] {
+\includepicture [0.95] {
   Frequency spectrum of the sampled baseband signal with its replicas at integer multiples of $f_s$ \cite{ThyagarajanK.S2019ItDS}.\\
@@
 
-\includepicture [1.0] [0] {Allocation of frequencies in an FM channel \cite{FigFmChannelFreqSpectrum}.} {fm_channel_baseband_freqs} {img/draw.io/fm-channel-baseband-freqs}
+\includepicture [1.0] {Allocation of frequencies in an FM channel \cite{FigFmChannelFreqSpectrum}.} {fm_channel_baseband_freqs} {img/draw.io/fm-channel-baseband-freqs}
 
@@
 
-\includepicture [0.75] [0] {Frequency discriminator block diagram \cite{FigBdFreqDiscriminator}.} {bd_freq_discriminator} {img/draw.io/bd_freq_discriminator}
+\includepicture [0.75] {Frequency discriminator block diagram \cite{FigBdFreqDiscriminator}.} {bd_freq_discriminator} {img/draw.io/bd_freq_discriminator}
 
@@
 
-\includepicture [1.0] [0] {Time-domain signals in envelope detection \cite{RoppelBegleitmaterial}.} {time_domain_envelope_detect} {img/draw.io/envelope-detect-time-domain.png}
+\includepicture [1.0] {Time-domain signals in envelope detection \cite{RoppelBegleitmaterial}.} {time_domain_envelope_detect} {img/draw.io/envelope-detect-time-domain.png}
 

### 2021-07-09 doc/thesis/chapters/system-architecture-and-concept.tex
optional floating images
generate new
rename diagrams
system level overview detailed bd

@@
 The final goal thereby is always be the same, that is: Listen to the music of a radio station that is transmitted over FM broadcast radio.\\
+
+\noindent
+A very high-level overview of the entire system is given the block diagram in Fig.\ref{fig:bd_system_overview_highlevel}.
+
+\includepicture [1.0] [0] [H] {High-level overview of the system.} {bd_system_overview_highlevel} {img/draw.io/bd_system_overview_highlevel}
 
@@
 %%%%%%%%%%%%%%%%%%%%
-\section{Block Diagram}
+\section{System concept}
+
+\includepicture [1.0] [0] [H] {Block diagram of the overall system concept.} {bd_system_overview_detailed} {img/draw.io/bd_system_overview_detailed}
 

### 2021-07-11 doc/thesis/chapters/system-architecture-and-concept.tex
system concept text

--- 2021-07-09 doc/thesis/chapters/system-architecture-and-concept.tex
@@
 The final goal thereby is always be the same, that is: Listen to the music of a radio station that is transmitted over FM broadcast radio.\\
+
+\noindent
+A very high-level overview of the entire system is given by the block diagram in Fig.\ref{fig:bd_system_overview_highlevel}.
+It shows the front end, which processes the received antenna signal, so that it can be used in the FM radio receiver block.
+The radio receiver itself is implemented in multiple different variants.
+The produced audio output is sent to a speaker, since the ultimate goal is to listen to a radio stations' audio signal.
+
+\includepicture [1.0] [0] [H] {High-level overview of the system.} {bd_system_overview_highlevel} {img/draw.io/bd_system_overview_highlevel}
 
@@
 %%%%%%%%%%%%%%%%%%%%
-\section{Block Diagram}
+\section{System concept}
 
-%%%%%%%%%%%%%%%%%%%%
-\subsection{describe blocks}
+The overall system concept has to be developed in a way, to fit the respective projects' needs and requirements.
+There should be a clear structure, in order to help to create a working system successfully.
+This structure should include a description of the data source, the data processing blocks and stages, commonly used parameters, as well as a strategy to verify the correct functionality of the system.
+A good system structure tries to share as many parts as possible in the system.
+This is to minimize the points of failure, as well as to reduce the effort of implementation.
+
+In the apparent projects' case, a single block of the system is implemented in multiple ways, which should however, all compute an equal output.
+Thus, it is of advantage to define the data source to be common for all of these variants, to create a unified situation at the input.
+This significantly eases the verification of the functionality in the final step, since all variants are expected to create the same output in the optimum, fully functioning case.
+Not only the data source is important to be common, but also the way of verification.
+In order to achieve that, the verification block needs to be able to compare multiple data outputs with the verification data, which is chosen to serve as the reference.\\
+
+An example of a system concept and its architecture is shown in Fig.\ref{fig:bd_system_overview_detailed}.
+Here, a common data source is provided by the block called \textit{Front End}.
+It delivers data, that the \textit{FM Radio Receiver} block requires at its input.
+Within this block, all the different implementation variants are shown in parallel.
+They all work on this common input source and generate their respective output.
+This output is written to a file.
+Hereby, it is important to note, that the output of the \textit{Matlab} implementation is chosen to be the reference model of 'correct output'.
+Its output is used as the verification data.
+This is because Matlab was used to research decoding algorithms and to develop the DSP chain, which is used in the GNU Radio, HLS and VHDL implementations.
+In addition to the verification data, the Matlab model also writes common parameters to a file.
+These parameters are directly used in the other implementations, to ensure a common setup.
+The final block, is the \textit{Verification} entity, which compares the output of all the respective variants, with respective to the chosen model parameters.
+
+A major point in this system concept is, that it works independently of the respective implementation methodology.
+This is especially important, since there is no possible way to run all these different variants in a single verification environment.
+Instead, each methodology writes an output file, which is a functionality that is available anywhere.
+
+\includepicture [1.0] [0] {Block diagram of the overall system concept.} {bd_system_overview_detailed} {img/draw.io/bd_system_overview_detailed}
 

### 2021-07-11 doc/thesis/chapters/system-architecture-and-concept.tex
minor

@@
 
-In the apparent projects' case, a single block of the system is implemented in multiple ways, which should however, all compute an equal output.
+In the apparent projects' case, a single block of the system is implemented in multiple ways, which should, however, all compute an equal output.
 Thus, it is of advantage to define the data source to be common for all of these variants, to create a unified situation at the input.
+Additionally, the data source may be chosen to be recorded, to have a known, reproducable set of input data.
 This significantly eases the verification of the functionality in the final step, since all variants are expected to create the same output in the optimum, fully functioning case.
@@
 
-An example of a system concept and its architecture is shown in Fig.\ref{fig:bd_system_overview_detailed}.
+An example of a system concept and its architecture is developed and shown in Fig.\ref{fig:bd_system_overview_detailed}.
 Here, a common data source is provided by the block called \textit{Front End}.
-It delivers data, that the \textit{FM Radio Receiver} block requires at its input.
-Within this block, all the different implementation variants are shown in parallel.
+It records data from an antenna and saves it to a file, in a format that the consequent blocks require at their input.
+Therefore, a reproducable input is ensured.
+Within the \textit{FM Radio Receiver} block, all the different implementation variants are shown in parallel.
 They all work on this common input source and generate their respective output.
-This output is written to a file.
-Hereby, it is important to note, that the output of the \textit{Matlab} implementation is chosen to be the reference model of 'correct output'.
-Its output is used as the verification data.
+This output data is again written to a file.
+Hereby it is important to note, that the output of the \textit{Matlab} implementation is chosen to be the 'golden reference model' of correct output.
+Its output is used as the verification data later in the process.
 This is because Matlab was used to research decoding algorithms and to develop the DSP chain, which is used in the GNU Radio, HLS and VHDL implementations.
@@
 These parameters are directly used in the other implementations, to ensure a common setup.
-The final block, is the \textit{Verification} entity, which compares the output of all the respective variants, with respective to the chosen model parameters.
+The final block is the \textit{Verification} entity, which compares the output files of all the variants, with respect to the chosen model parameters.\\
 
@@
 This is especially important, since there is no possible way to run all these different variants in a single verification environment.
-Instead, each methodology writes an output file, which is a functionality that is available anywhere.
+Instead, each methodology writes an output file, which is a supported functionality that is available anywhere.
 

### 2021-07-11 doc/thesis/chapters/dsp.tex
add list of abbrev
structuring

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Receiver types}
+\section{Receiver Types}
 
@@
 
-\subsection{Mathematical description}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Mathematical Description}
 
@@
 
-\section{Sampling theorem}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Sampling Theorem}
 
@@
 
-This chaper describes the main properties of broadcast FM, such as the mathematical description, frequency bands that are used, or the specific frequency parts within a channel spectrum.
+This section describes the main properties of broadcast FM, such as the mathematical description, frequency bands that are used, or the specific frequency parts within a channel spectrum.
 
@@
 
-This chapter describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
+This section describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
 The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly.

### 2021-07-11 doc/thesis/chapters/implementation.tex
\includepicture function with floating option; new block diagram
structuring

--- 2021-05-05 doc/thesis/chapters/implementation.tex
@@
 \label{cha:Implementation}
-\section{Matlab/Python Model}
+
+This chapter describes the most important details of each implementation variant.
+
+\section{Digital Signal Processing Chain}
+
+\includepicture [1.0] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/bd_detailed_dsp}
+
+\section{Matlab Model}
+
 in fixed point, close to hardware level algorithm
 
+\subsection{Transmitter block diagram}
+
+\section{GNU Radio}
+  \subsection{Introduction}
+  \subsection{Transmitter}
+  \subsection{Receiver}
+
 \section{VHDL (no shortcut)}
+  \subsection{Testbench}
   \subsection{Channel Selection (IF to Channel-BB)}
@@
   \section{High-Level Synthesis}
+  \subsection{Testbench}
   \subsection{Channel Selection (IF to Channel-BB)}
@@
   Display a direct comparison of outputs in graphs. This is practical, since the cocotb framework runs in python and graphs can be generated using Python's matplotlib.
-
-  \section{Develop HLS compiler independent code (optional)}
-  \subsection{Challenges}
-  Different compilers use different \#pragmas, etc.\\
-  The \#pragmas need to be within the code, at the appropriate positions, which makes it difficult to write compiler independent code....

### 2021-07-11 doc/thesis/chapters/dsp.tex
emphasis

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Emphasis and De-Emphasis}
+
+TODO
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Algorithms for Digital FM Demodulation}
@@
 \label{sec:demodulation-of-broadcast-fm}
-%TODO: once FM is converted to AM, traditional AM DSP techniques can be employed
 
 This section describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
-The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly.
+The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly, including the de-emphasis of the signal.
+
+So at this point the FM signal is converted to an AM signal.
+Thus, traditional AM DSP techniques can be employed to decode the FM broadcast channel.
+
 A frequency spectrum of the multiplex signal of an FM channel is shown in Fig.\ref{fig:fm_channel_baseband_freqs}.

### 2021-07-11 doc/thesis/chapters/implementation.tex
emphasis
impl details

@@
 
+The theory behind the demodulation of an FM signal is already explained in Chapter \ref{cha:SignalProcessingTheory}.
+The demodulation of an FM broadcast channel specifically, is described in Section \ref{sec:demodulation-of-broadcast-fm}.
+In this section however, the DSP chain is described into more detail and with more information, as it is implemented in the various methods.\\
+
+Fig.\ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram. The following list explains implementation details about the respective DSP blocks.
+
 \includepicture [1.0] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/bd_detailed_dsp}
+
+\begin{itemize}
+  \item \textbf{Data Source}\\
+    Matlab is used to generate the data source file, which is used as an input to the system.
+    An audio file is read and modulated as an FM channel signal.
+  \item \textbf{FM Demodulator}\\
+    The FM demodulator is implemented as a Baseband Delay Modulator, as described in \ref{sec:BasebandDelayDemodulator}.
+  \item \textbf{De-Emphasis Filter}\\
+    The above described Matlab source generation does not include an Emphasis filter.
+    Thus, a de-emphasis filter is not implemented as it is not required.
+  \item \textbf{Downsample}\\
+    Downsampling is implemented in a simplified version, without an anti-aliasing filter.
+    It is not required, because of the following explanation.
+    At this point the sampling frequency is set to 960 kHz.
+    The input is band-limited at 75 kHz, which is the maximum frequency in an FM channel, as explained in Section \ref{sec:FrequencySpectrumOfABroadcastFmChannel}.
+    Thus, the input has an oversampling rate of 12.8, while downsampling is only performed with a factor of 8.
+  \item \textbf{Filter Group Delay Compensation}\\
+    FIR filters introduce a delay on the output signal, with respect to the input.
+    This is due to the property of the FIR filters' linear phase response \cite{GaziOrhan2018UDSP}.
+    Since phase is directly proportional to time, the term \textit{group delay} is often used.
+    This delay needs to be accounted for and compensated in digital signal processing.
+
+\end{itemize}
 
@@
 
-in fixed point, close to hardware level algorithm
+\subsection{Emphasis and De-Emphasis}
+show effects in spectrum, show how to design the filters
+
+\subsection{Fixed Point Arithmetic}
+not developed in actual fixed point, but close to hardware level algorithm
 

### 2021-07-11 sim/matlab/fm_receiver.m

@@ -35,8 +35,10 @@ rx_fm_q = imag(rx_fm_bb_norm);
 
 if EnableProcessingLikeHW
     n_shift = 3;
-    rx_fm_i_diff = rx_fm_i - [zeros(n_shift,1); rx_fm_i(1:end-n_shift)];
-    rx_fm_q_diff = rx_fm_q - [zeros(n_shift,1); rx_fm_q(1:end-n_shift)];
+    rx_fm_i_delayed = [zeros(n_shift,1); rx_fm_i(1:end-n_shift)];
+    rx_fm_q_delayed = [zeros(n_shift,1); rx_fm_q(1:end-n_shift)];
+    rx_fm_i_diff = rx_fm_i - rx_fm_i_delayed;
+    rx_fm_q_diff = rx_fm_q - rx_fm_q_delayed;
 else
     % Design differentiator
     filter_diff = [1,0,-1];
@@ -45,8 +47,9 @@ else
     rx_fm_q_diff = filter(filter_diff,1, rx_fm_q);
     
     % Compensate group delay of filter
-    rx_fm_i_diff = circshift(rx_fm_i_diff,-1);
-    rx_fm_q_diff = circshift(rx_fm_q_diff,-1);
+    filter_diff_grp_delay = (length(filter_diff)-1)/2;
+    rx_fm_i_diff = circshift(rx_fm_i_diff,-filter_diff_grp_delay);
+    rx_fm_q_diff = circshift(rx_fm_q_diff,-filter_diff_grp_delay);
     
     %rx_fm_demod =  ...
     %    (rx_fm_i .* conv(rx_fm_q,filter_diff,'same') -   ...

### 2021-07-11 doc/thesis/chapters/implementation.tex
dsp chain list

@@
 
-\includepicture [1.0] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/bd_detailed_dsp}
+\includepicture [1.0] [0] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/bd_detailed_dsp}
 
@@
       An audio file is read and modulated as an FM channel signal.
+
   \item \textbf{FM Demodulator}\\
       The FM demodulator is implemented as a Baseband Delay Modulator, as described in \ref{sec:BasebandDelayDemodulator}.
+
   \item \textbf{De-Emphasis Filter}\\
@@
       Thus, a de-emphasis filter is not implemented as it is not required.
+
   \item \textbf{Downsample}\\
@@
       Thus, the input has an oversampling rate of 12.8, while downsampling is only performed with a factor of 8.
+
+  \item \textbf{Filters}\\
+      Several filters are implemented as FIR filters.
+      The 15 kHz lowpass (LP) filter is instantiated twice, with the exact same coefficients.
+
   \item \textbf{Filter Group Delay Compensation}\\
       FIR filters introduce a delay on the output signal, with respect to the input.
+      The group delay in number of samples can be computed with the following formula.
+      \begin{equation}
+        group\ delay = \frac{filter\ order}{2} = \frac{(N\ coefficients - 1)}{2}
+      \end{equation}
       This is due to the property of the FIR filters' linear phase response \cite{GaziOrhan2018UDSP}.
       Since phase is directly proportional to time, the term \textit{group delay} is often used.
-    This delay needs to be accounted for and compensated in digital signal processing.
+      This delay needs to be accounted for, depending on where the signal is used subsequently.
 
+      Looking at the block diagram, the group delay needs to be compensated at two places.
+      \begin{itemize}
+        \item \textbf{Between the branches \textit{Recover LR diff} and \textit{Recover Mono}:}\\
+          The \textit{Mono} branch only contains a single LP filter, while the \textit{LR diff} branch contains two filters, namely a BP and an LP filter.
+          The latter branches' signal is therefore delayed, with respect to the \textit{Mono} branch.\\
+          The issue here is, that both signals are added and substracted at the end of the processing chain, to compute the left and right channel signals, respectively.
+          Thus, they need to be aligned in time at that point.
+          Consequently, a delay block needs to be inserted in the \textit{Mono} branch, to compensate the group delay of the BP.
+
+        \item \textbf{Between the branches \textit{Recover LR diff} and \textit{Recover Carriers}:}\\
+          Both branches only contain a single BP filter.
+          However, they have properties, like the transition bandwidth and the passband.
+          Thus, their filter order may be different as well, which leads to a non-matching group delay.
+          Since the two signals are multiplicated in the subsequent mixer, and an important factor is the phase-coherency of the recovered carrier (as explained in Section \ref{sec:demodulation-of-broadcast-fm}), the group delay compensation becomes necessary here.\\
+
+          To overcome the issue at this point in the actual hardware implementation, the filters were manually 'tuned' to have a matching order.
       \end{itemize}
+
+  \item \textbf{Carrier Recovery}\\
+      The 38 kHz carrier can be recovered from the 19 kHz pilot tone.
+      It is important that the recovered carrier is phase-coherent with the received input, as described in Section \ref{sec:demodulation-of-broadcast-fm}.
+      This can be achieved, by simply multiplying the pilot tone with itself, which is the way it is implemented here.
+\end{itemize}
+
+Several of the listed details are implemented, in order to achieve a successful demodulation of the FM signal.
 

### 2021-07-11 doc/thesis/chapters/implementation.tex
new diagram

@@
 
-\includepicture [1.0] [0] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/bd_detailed_dsp}
+\includepicture [1.0] [0] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/draw.io/bd_dsp_detailed}
 
@@
   \item \textbf{FM Demodulator}\\
-      The FM demodulator is implemented as a Baseband Delay Modulator, as described in \ref{sec:BasebandDelayDemodulator}.
+      The FM demodulator is implemented as a Baseband Delay Modulator, as described in Section \ref{sec:BasebandDelayDemodulator}.
 
@@
 
+\noindent
 Several of the listed details are implemented, in order to achieve a successful demodulation of the FM signal.

### 2021-07-11 sim/matlab/research/dsbsc.m

@@ -38,7 +38,7 @@ fc_tx  = 57e3;
 A_tx_c = 1;
 
 % Rx carrier
-fc_rx  = 57e3;       % <-- ADAPT FREQUENCY ERROR
+fc_rx  = 57e3;        % <-- ADAPT FREQUENCY ERROR (e.g.  57.1e3)
 phi_rx = 160*pi/180;  % <-- ADAPT PHASE ERROR
 A_rx_c = 1;
 
@@ -125,9 +125,9 @@ fprintf('fft_bin: %.2f Hz\n', fft_bin);
 %Nfft      = 2^nextpow2(Nfft);
 fft_freqs = (-Nfft/2:Nfft/2-1)/(Nfft*1/fs);
 
-fft_msg    = abs(fftshift( fft( hanning(length(msg)) .* msg,    Nfft) ))/Nfft;
+fft_msg          = abs(fftshift( fft( hanning(length(msg)) .* msg,          Nfft) ))/Nfft;
 fft_msg_rx_demod = abs(fftshift( fft( hanning(length(msg)) .* rx_msg_demod, Nfft) ))/Nfft;
-fft_msg_rx = abs(fftshift( fft( hanning(length(msg)) .* rx_msg, Nfft) ))/Nfft;
+fft_msg_rx       = abs(fftshift( fft( hanning(length(msg)) .* rx_msg,       Nfft) ))/Nfft;
 
 
 % Plots

### 2021-07-11 doc/thesis/chapters/implementation.tex
adapt
matlab transmitter

@@
 
-Fig.\ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram. The following list explains implementation details about the respective DSP blocks.
+Fig.\ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram.
 
-\includepicture [1.0] [0] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/draw.io/bd_dsp_detailed}
+\includepicture [1.02] [0] [H] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/draw.io/bd_dsp_detailed}
 
+The following list explains implementation details about the respective DSP blocks.
 \begin{itemize}
@@
 
+Matlab is used to create a model of the receiver, which is used as a reference for the other implementation variants.
+Here, DSP algorithms are tested and evaluated, to find a method that is suitable to implement in hardware, specifically in an FPGA.
+
+The Matlab script consists of a transmitter, the receiver model, and additional code for analysis.
+
+\subsection{Transmitter}
+
+The transmitter is implemented to provide a reproducable source of data, which can be used to feed the receiver.
+Another option would be to use external hardware as a software-defined radio and record some antenna data.
+However, the main advantage of a locally implemented transmitter over a recording is, that the raw data input is known.
+Thus, the expected output of the receiver model is this raw data, that was previously modulated by the local transmitter.
+In the case of a recording, the original data is not known, which does not allow to actually verify the receiver.\\
+
+In the implemented version, an audio file is modulated into the FM channel.
+The audio file contains data that represents speech with the words \textit{'Left channel... Right channel...'}.
+In terms of audio channels, the words exactly match the respective channels' output - the first part is modulated on the left audio channel only, while the second part only targets the right audio channel.
+This is particulary useful for verification, especially to verify the functionality of channel separation.
+From the experience during development, the channel separation is a good indicator of whether the system is functioning correctly.
+This left-right indication also simplifies the verification process, because a developer can recognize it with their visual and acoustic senses.
+It is possible to audibly verify it by listening to the output.
+It is also possible to just visually inspect the the output data, where a clear separation is visible.
+
+A block diagram of the transmitter structure is depicted in Fig.\ref{fig:bd_matlab_transmitter}.
+
+\includepicture [1.0] [0] [H] {Block diagram of the implemented transmitter in Matlab.} {bd_matlab_transmitter} {img/draw.io/bd_matlab_transmitter}
+
 \subsection{Emphasis and De-Emphasis}
@@
 not developed in actual fixed point, but close to hardware level algorithm
-
-\subsection{Transmitter block diagram}
 

### 2021-07-14 doc/thesis/chapters/implementation.tex
new diagram (matlab transmitter)
matlab transmitter

@@
 It is possible to audibly verify it by listening to the output.
-It is also possible to just visually inspect the the output data, where a clear separation is visible.
+It is also possible to just visually inspect the output data, where a clear separation is visible between the channels.\\
 
-A block diagram of the transmitter structure is depicted in Fig.\ref{fig:bd_matlab_transmitter}.
+The transmitter is implemented according to the structure that is depicted as a block diagram in Fig.\ref{fig:bd_matlab_transmitter}.
+To facilitate an easy understanding of the block diagram, it is of advantage to keep in mind that it exactly resembles the frequency spectrum that is previously shown in Fig.\ref{fig:fm_channel_baseband_freqs}.\\
 
-\includepicture [1.0] [0] [H] {Block diagram of the implemented transmitter in Matlab.} {bd_matlab_transmitter} {img/draw.io/bd_matlab_transmitter}
+\includepicture [1.0] [0] {Block diagram of the implemented transmitter in Matlab.} {bd_matlab_transmitter} {img/draw.io/bd_matlab_transmitter}
 
-\subsection{Emphasis and De-Emphasis}
+The transmitter reads an audio file, including a left and a right channel in the first step.
+The audio file contains data in a sample rate of 48 kHz.
+Next, the sample rate is increased by an oversampling factor of 20, to achieve a sufficient sample rate for the subsequent DSP procedures.
+
+The achieved sample rate of 960 kHz could actually be chosen much lower, since an FM channels' maximum frequency is 75 kHz (see \ref{sec:FrequencySpectrumOfABroadcastFmChannel}), and according to the Nyquist theorem the sample rate thus only needs to have 150 kHz (see \ref{sec:SamplingTheorem}).
+However, the higher rate was chosen, mainly because a higher number of signal samples beautify any plots that are created during analysis.
+Since a thorough optimization of the DSP chain is not the focus of this thesis, this was an acceptable trade-off.
+
+The next step is the pre-emphasis filter, which is implemented, but is disabled in the receiver.
+This is done to reduce some complexity in the hardware implementation.
+The following stage generates the Mono and LR Diff signals, by simply calculating the summation and diffence of the left and right channels, respectively.
+
+The next section in the block diagram represents multiple processes in parallel.
+The LR Diff part is shifted to 38 kHz in the spectrum.
+This is done by modulating the signal with its sub-carrier of 38 kHz.
+A subsequent BP filter serves to remove the artifacts of the modulation, such as the image replica.
+It band-limits the spectrum of the LR Diff part between 23 and 53 kHz.
+The BP filter introduces a constant group delay of the signal.
+In order to maintain the alignment of all signals, this side-effect needs to be accounted for and be compensated in the other signal branches.
+Thus, the Mono signal, as well as the Pilot Tone need to be delayed by the amount of samples defined by the BP filters' group delay.
+The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone [ref TODO].
+
+One part in this block diagram is the so-called Hinz Triller, which was part of the Autofahrer-Rundfunk-Information service (ARI, German for Automotive-Driver's-Broadcasting-Information).
+It was used as an indicator for traffic announcements.
+It is now obsolete and is replaced by the more modern Radio Data System (RDS).
+However, the Hinz Triller is still transmitted in many radio stations and can often be heard before and after the traffic announcements \cite{HinzTriller}.
+In the implemented transmitter, the Hinz Triller is implemented, but is disabled, since it only represents a temporarily transmitted signal.
+
+The Radio Data System (RDS) signal is not implemented in the transmitter, since it consists of a comprehensive protocol and multiple steps of modulation, such as differentially-coded BPSK \cite{IntroFmStereoRdsModulation} and is therefor too time-consuming to implement.
+Furthermore, it would not contribute any additional value to this project, since this is not the main focus.
+
+At this point in the DSP chain, the signal resembles the frequency spectrum of Fig.\ref{fig:fm_channel_baseband_freqs}.
+
+In the final chain of DSP blocks, the sample rate is increased again, to be able to frequency-modulate the signal.
+The resulting frequency modulated signal represents the signal that is received by an antenna - the RF signal of an FM channel.
+This signal is then processed by a complex baseband mixer, which results in inphase-, and quadrature components in the baseband.
+An LP filter band-limits the signal and removes artifacts of the mixer.
+As a final step, the sample rate is decreased again, back to the chosen sampling rate of 960 kHz.\\
+
+This signal can now be processed in a digital FM receiver.
+
+\subsection{Pre-Emphasis and De-Emphasis}
 show effects in spectrum, show how to design the filters

### 2021-07-14 doc/thesis/chapters/implementation.tex
another cite ref

@@
 Thus, the Mono signal, as well as the Pilot Tone need to be delayed by the amount of samples defined by the BP filters' group delay.
-The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone [ref TODO].
+The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone \cite{FmMultiplexingForStereo}.
 
@@
 An LP filter band-limits the signal and removes artifacts of the mixer.
-As a final step, the sample rate is decreased again, back to the chosen sampling rate of 960 kHz.\\
+As a final step, the sample rate is decreased again, back to the chosen sampling rate of 960 kHz.
 

### 2021-07-14 doc/thesis/chapters/implementation.tex
fixed point

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Digital Signal Processing Chain}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Matlab Model}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Transmitter}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Pre-Emphasis and De-Emphasis}
-show effects in spectrum, show how to design the filters
+TODO: show effects in spectrum, show how to design the filters
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Fixed Point Arithmetic}
-not developed in actual fixed point, but close to hardware level algorithm
+
+The target hardware platform of the FM radio receiver project is an FPGA.
+These are devices that have a limited number of logic cells, which can be used to implement logical and computational functions.
+Any kind of computation can be implemented within these logic cells.
+However, depending on its detailed implementation, the amount of required logic cells may differ by a large factor.
+In order to overcome this issue, mathematical operations are often implemented in fixed point arithmetic, rather than floating point.
+This is due to the fact, that fixed point operations can be implemented with much less logic cells than floating point operations.
+The reason therefor is, that the bitwidth of input and output numbers is previously known in fixed point.
+Floating point calculations do not have this assumption and thus need to be more flexible, which requires a larger amount of logic cells.\\
+
+The Matlab model is therefore developed by keeping in mind this limitation.
+Matlab provides special datatypes for fixed point number representation.
+However, these were not used in this implementation.
+Instead, everything is calculated in floating point representation, using the standard \textit{double} datatype.
+Then, in certain positions in the DSP chain, the numbers are rounded to the resolution of a fixed point datatype.
+A fixed point format of \textit{2.14}, meaning 2 bit for integer- and 14 bit for fractional number representation is chosen.
+This is sufficient for the FM receiver and could potentially be lowered to a lower bitwidth.
+Again, an optimization of the DSP chain is not the focus of this thesis.\\
+
+Summing up, the Matlab model is not implemented in fixed point arithmetic, but is implemented in a way, that is close enough to the hardware implementation.
+Therefore, it can serve as a reference for the hardware implementation.
 

### 2021-07-14 doc/thesis/chapters/implementation.tex
analysis

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Receiver}
+
+The receiver is implemented, following the block diagram in Fig.\ref{fig:bd_dsp_detailed}.
+Further assumptions and simplifications, such as the fixed point arithmetic are described into detail above.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Analysis}
+
+The Matlab implementation also includes a comprehensive range of analysis, especially in terms of visual plots.
+These plots are useful during the development of new DSP algorithms, in order to verify the correct thereof.\\
+
+An example of a plot that is created by the analysis is depicted in Fig.\ref{fig:matlab_analysis_time_domain}.
+The diagram shows the respective signals in the time domain.
+Above two timelines display the input data, as it is read from the input audio file.
+The two diagrams in the center plot the Mono and LR Diff signals.
+The bottom diagrams then represent the result after the entire signal processing chain.
+It can be seen that the DSP chain is not able to demodulate the original signal without any error.
+However, the result shows a clear separation of the left and right channels, as resembles the original input very closely, which is good enough for this case study.\\
+
+\includepicture [1.0] [0] {Matlab analysis in the time domain.} {matlab_analysis_time_domain} {img/matlab/matlab_analysis_time_domain}
+
+Further analysis is done in the frequency domain, especially in order to better understand the modulation of the different spectral parts in the FM channel.
+An example thereof is shown in Fig.\ref{fig:matlab_analysis_freq_domain}.
+
+The spectrum of an FM channel, as it is generated by a transmitter before the modulation, is shown in the upper diagram.
+The spectrum after the demodulation through a receiver is shown in the lower diagram.
+Spectral parts, such as the pilot tone at 19 kHz, or the LR Diff part around 38 kHz can be seen clearly.
+The demodulated spectrum shows an artifact at 57 kHz, which does not exist in the original transmitter spectrum.
+This is an artifact of the demodulation process, specifically the frequency-shift of the LR Diff part to baseband.
+The artifact is very low in energy, because it is attenuated by a BP filter in the down-conversion process, and can thus be ignored for further processing, even though it can still be seen in the spectral analysis.
+
+\includepicture [1.0] [0] {Matlab analysis in the frequency domain.} {matlab_analysis_freq_domain} {img/matlab/matlab_analysis_freq_domain}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{GNU Radio}

### 2021-07-14 sim/matlab/fm_analysis.m

@@ -126,7 +126,6 @@ disp('-- Plots');
 
 fig_title = 'Time domain signal';
 fig_audio_time = figure('Name',fig_title);
-title(fig_title);
 ax1='';
 ax2='';
 if EnableSenderSourceCreateSim
@@ -157,8 +156,13 @@ ylim([-ymax,ymax]);
 xlabel('time [s]');
 grid on; legend();
 linkaxes([ax1,ax2,ax3,ax4,ax5,ax6],'x');
+title(ax1,fig_title);
 if EnableSavePlotsToPng
     saveas(fig_audio_time, sprintf("%s%s",dir_output, "time_audio.png"));
+end
+if EnableSavePlotsToPDF
+    fig_audio_time.Position(3:4) = [900 600];
+    exportgraphics(fig_audio_time,sprintf("%s%s",dir_output_doc, "matlab_analysis_time_domain.pdf"),'ContentType','vector')
 end
 
 if EnableSenderSourceCreateSim
@@ -250,26 +254,38 @@ end
 
 fig_title = 'FM channel spectrum';
 fig_tx_spec = figure('Name',fig_title);
-hold on;
+ax1 = subplot(2,1,1);
+title(fig_title);
+hold on; grid on;
 xline(19e3,'k--','19 kHz');
 xline(38e3,'k--','38 kHz');
 xline(57e3,'k--','57 kHz');
 h0 ='';
 if EnableSenderSourceCreateSim
-    h0 = plot(psxx_tx_fmChannelData_f, psxx_tx_fmChannelData, 'b','DisplayName', 'Tx (pre-mod)');
+    h0 = plot(psxx_tx_fmChannelData_f, psxx_tx_fmChannelData, 'b','DisplayName', 'Tx (pre-mod.)');
+    xlim([0 62e3]);
+    legend([h0],'Location','northeast');
 end
-h1 = plot(psxx_rxChannelData_f, psxx_rx_fmChannelData,  'r','DisplayName', 'Rx Demod');
+ax2 = subplot(2,1,2);
+hold on; grid on;
+xline(19e3,'k--','19 kHz');
+xline(38e3,'k--','38 kHz');
+xline(57e3,'k--','57 kHz');
+h1 = plot(psxx_rxChannelData_f, psxx_rx_fmChannelData,  'r','DisplayName', 'Rx (demod.)');
+legend([h1],'Location','northeast');
[…]

### 2021-07-14 sim/matlab/fm_transceiver.m

@@ -26,13 +26,15 @@ end
 %% Settings
 
 % Paths
-dir_filters = "./filters/stored/";
-dir_output  = "./matlab_output/";
+dir_filters    = "./filters/stored/";
+dir_output     = "./matlab_output/";
+dir_output_doc = "../../doc/thesis/img/matlab/";
 
 % Simulation options
 EnableWriteDataFiles = true;
 EnablePlots          = true;
 EnableSavePlotsToPng = true;
+EnableSavePlotsToPDF = true;
 
 EnableSenderSourceRecordedFile = false;
 EnableSenderSourceCreateSim    = true;

### 2021-07-14 doc/thesis/chapters/implementation.tex
gnuradio

@@
 \section{GNU Radio}
+
+This section describes the implementation in GNU Radio.
+A transmitter, as well as a receiver are implemented using this software.
+Both can be deployed on actual hardware, so a continuous transmitter or receiver is built in a real environment.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Introduction}
+
+GNU Radio is a free and open-source software development toolkit to implement a software-defined radio.
+The implementation is done by connecting functional blocks in a block design.
+This abstracts the inner workings of each functional block and thus provides an implementation approach on a very high level of abstraction.
+GNU Radio supports interfaces to external RF hardware \cite{SoftwareGnuRadio}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{External Hardware}
+
+In order to interface with the real world, two different hardware devices are used.
+Both devices are supported by GNU Radio and are briefly described here.
+
+\subsubsection{USRP b200mini}
+
+The USRP b200mini is a device that can be used to create a software-defined radio.
+It resembles the size of a business card and is developed by Ettus Research.
+A signal range from 70 MHz to 6 GHz can be covered by the front end.
+The board further features a Xilinx Spartan-6 FPGA that is user-programmable, which makes the device very flexible for various applications with the received signal \cite{USRPb200Mini}.
+
+GNU Radio supports this device with the blocks \textit{UHD: USRP Sink} and \textit{UHD: USRP Source}.
+Therefore, it can be used as both a receiver and a transmitter.
+
+\subsubsection{RTL-SDR}
+
+The RTL-SDR is a cheap USB dongle, that was originally developed as a DVB-T TV tuner.
+It is based on the RTL2832U chipset, which includes an 8-bit ADC and a digital signal processor.
+The device delivers IQ data via the USB interface \cite{RTLSDR}.
+
+Based on its hardware, the RTL-SDR can only be used as a receiver.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Transmitter}
+
+GNU Radio provides a comprehensive set of DSP blocks, which are available to build various functionalities.
+Multiple blocks are connected to a block diagram, that is very similar to the one in Fig.\ref{fig:bd_matlab_transmitter}.
+
+Again, an audio file is used as an input, which is modulated as an FM channel signal.
+The entire multiplex signal is then modulated onto an actual carrier frequency of 99 MHz, that is within the FM radio broadcast band.
+
+take care in using the FM band - follow regulations in \ref{}!.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Receiver}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{System Level Integration for Verification}
+
+system level view: can be used as a local transmitter for testing and verification.
+sent signal is known, thus the receiver can be verified
 

### 2021-07-14 doc/thesis/chapters/implementation.tex
transmitrer

@@
 Multiple blocks are connected to a block diagram, that is very similar to the one in Fig.\ref{fig:bd_matlab_transmitter}.
+The USRP device is used to broadcast the RF signal over the air.
 
@@
 The entire multiplex signal is then modulated onto an actual carrier frequency of 99 MHz, that is within the FM radio broadcast band.
+This entire DSP chain is being run in the GNU Radio software, on a host PC.
+Only the final signal - FM modulated, and on a carrier of 99 MHz - is then sent to the USRP device in the form of digital signal samples.
+The USRP hardware then transmits the signal via its antenna.
 
-take care in using the FM band - follow regulations in \ref{}!.
+\textit{NOTE:}\\
+Care needs to be taken when transmitting signals over the air.
+The frequency spectrum is strongly limited by the radio regulations and may be monitored by the authorities.
+Please make sure to adapt your setup to the respective local regulations, to avoid any legal issues.
+Some of the regulations are mentioned in \ref{sec:FrequencyBand}, but may differ in various countries, however.
+
+Fig.\ref{fig:bd_gnuradio_transmitter} shows a block diagram of the application.
+
+%\includepicture [1.0] [0] {Block diagram of the GNU Radio transmitter setup.} {matlab_analysis_freq_domain} {img/draw.io/gnuradio_setup_transmitter}
 

### 2021-07-14 doc/thesis/chapters/implementation.tex
new diagrams; gnuradio

@@
 Only the final signal - FM modulated, and on a carrier of 99 MHz - is then sent to the USRP device in the form of digital signal samples.
-The USRP hardware then transmits the signal via its antenna.
+The USRP hardware then transmits the signal via its antenna.\\
 
-\textit{NOTE:}\\
-Care needs to be taken when transmitting signals over the air.
+\noindent
+Special care needs to be taken when transmitting signals over the air.
 The frequency spectrum is strongly limited by the radio regulations and may be monitored by the authorities.
@@
 
-Fig.\ref{fig:bd_gnuradio_transmitter} shows a block diagram of the application.
+Fig.\ref{fig:bd_gnuradio_setup_transmitter} shows a block diagram of the transmitter application.
 
-%\includepicture [1.0] [0] {Block diagram of the GNU Radio transmitter setup.} {matlab_analysis_freq_domain} {img/draw.io/gnuradio_setup_transmitter}
+\includepicture [0.9] [0] [H] {Block diagram of the GNU Radio transmitter setup.} {bd_gnuradio_setup_transmitter} {img/draw.io/bd_gnuradio_setup_transmitter}
 
@@
 
+The receiver is also implemented, by using standard DSP blocks that are available in GNU Radio.
+The implemented structure however does not represent a stereo receiver.
+Instead, a mono receiver is implemented.
+The antenna signal is FM-demodulated, decimated and then directly sent to the speakers as the output, without any further processing.
+This simple receiver was intended as a proof-of-concept and a first example of the cooperation of the external hardware and the GNU Radio software.
+Thus, the mono receiver was sufficient, as the focus of the GNU Radio application is put on the transmitter side.
+
+This receiver works with both devices, the USRP b200mini and the RTL-SDR.
+
+Fig.\ref{fig:bd_gnuradio_setup_receiver} shows a block diagram of the receiver application.
+
+\includepicture [0.9] [0] [H] {Block diagram of the GNU Radio receiver setup.} {bd_gnuradio_setup_receiver} {img/draw.io/bd_gnuradio_setup_receiver}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-system level view: can be used as a local transmitter for testing and verification.
-sent signal is known, thus the receiver can be verified
+The usage of a software like GNU Radio in combination with external hardware is an interesting tool for a use-case like this project.
+An FM receiver is to be implemented in multiple different versions.
+All of them need to be verified somehow - this is where the SDR transmitter created with GNU Radio and external hardware opens up an entire new possibility from the viewpoint of system level design integration.\\
+
+Previously it was explained that it is not possible to use recorded antenna data for verification.
+The reason was, that the underlying original data is not known and therefor a comparison of the original versus the received, decoded data is not feasible.
+Consequently, the Matlab transmitter model was justified by the fact, that its originally encoded audio data is well-known.\\
+
+However, with the current approach of using a GNU Radio SDR transmitter, the encoded data \textit{is} well-known, because it is generated locally.
+Thus, the SDR can be used to transmit a verification signal.
+This signal could be recorded to file by another SDR, or sourced into the DUT (device under test) directly.
 

### 2021-07-15 doc/thesis/chapters/implementation.tex
section comments
testbench

@@
 This chapter describes the most important details of each implementation variant.
+The implementation variants in High-Level Synthesis and VHDL are highlighted especially, since they are the main focus of this thesis.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Transmitter}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Pre-Emphasis and De-Emphasis}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Fixed Point Arithmetic}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Receiver}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Analysis}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Introduction}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{External Hardware}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{System Level Integration for Verification}
@@
 
-\section{VHDL (no shortcut)}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{VHDL (TODO: no shortcut?)}
+
+In this section, the most important details of the VHDL implementation are described.
+This implementation variant is one of the two major implementation focus points of this thesis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Testbench}
+
+In the development of software that runs on a CPU, the tool of step-by-step debugging is commonly used during the development process.
+Various verification and testing libraries exist, in order to create unit tests, for example.
+The advantage is, that the code that is developed there, can be compiled and run on the target CPU directly, while maintaining the direct step-by-step representation of the code.
+
+However, in the development of hardware with any hardware description language (HDL), this is not possible.
+HDL languages provide a syntax and semantic representation that allow to create parallel structures and a dependency on timing, such as a system clock.
+Before this code can be run on the target hardware, i.e. an FPGA, it needs to run through a process called synthesis.
+In synthesis, the HDL code is translated into a netlist and is consequently mapped into hardware logic cells that are available in the target device.
+Various optimizations are done during this process, so that the direct connection to the code representation is lost.
+Additionally, it is impossible to access every single signal within the design during debugging on the hardware, which probably is one of the main issues.
+Synthesis tools like Xilinx Vivado provide support to make debugging possible in hardware, but require a large effort to implement them.
+Another big factor that speaks against debugging in hardware, is that the process of synthesis takes a comparably large amount of time to transform code into the target technology.
+Each time the code is adapted, the entire synthesis process needs to be started over.
+
+Because of all these reasons, debugging HDL code step-by-step as in software is not possible.\\
+
+Thus, in order to allow debugging and verification of HDL code, the technology of simulation was invented.
+Here, the HDL is also translated into a netlist that represents all the parallel processes just as in hardware, but the subsequent step of synthesis is not taken.
+Therefore, simulation comes to a result much faster.
+Since HDL designs actually only consist of electrical wires and some blocks (todo richtiges wort?), it is difficult to verify correct values.
+The testbench environment is a very important factor in any project that includes hardware description language (HDL) code.
+The reason therefor is that debugging
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Channel Selection (IF to Channel-BB)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Phase Detector}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{other Elements}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{High-Level Synthesis}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Testbench}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Channel Selection (IF to Channel-BB)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Phase Detector}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{other Elements}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Common Testbench}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Architecture (same tb for VHDL and HLS-generated HDL)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Framework cocotb, with ghdl compiler}

### 2021-07-15 doc/thesis/chapters/implementation.tex
tb

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Testbench}
+\subsection{Simulation Testbench}
+
+The general significance of a simulation testbench in hardware description language (HDL) design is explained here, as well as an introduction to the used testbench framework that is used in this project.
+
+%%%%%%%%%%%%%
+\subsubsection{General}
 
@@
 
-However, in the development of hardware with any hardware description language (HDL), this is not possible.
+However, in the development of hardware with any HDL, this is not possible.
 HDL languages provide a syntax and semantic representation that allow to create parallel structures and a dependency on timing, such as a system clock.
+The design is based on signals that perform logic operations and transfer values into registers.
+Thus, HDL code is often called register-transfer level, or short, RTL code.
 Before this code can be run on the target hardware, i.e. an FPGA, it needs to run through a process called synthesis.
@@
 Various optimizations are done during this process, so that the direct connection to the code representation is lost.
-Additionally, it is impossible to access every single signal within the design during debugging on the hardware, which probably is one of the main issues.
-Synthesis tools like Xilinx Vivado provide support to make debugging possible in hardware, but require a large effort to implement them.
-Another big factor that speaks against debugging in hardware, is that the process of synthesis takes a comparably large amount of time to transform code into the target technology.
-Each time the code is adapted, the entire synthesis process needs to be started over.
+Additionally, it is impossible to access every single signal within the design during debugging on the hardware, which is probably also one of the main issues.
+Synthesis tools like Xilinx Vivado provide support to make debugging possible in hardware, but require a large effort to implement them, especially in terms of synthesis time.
+Another big factor that speaks against debugging in hardware generally, is that the process of synthesis takes a comparably large amount of time to transform code into the target technology.
+Each time the code is adapted, the entire synthesis process needs to be started over again.
 
@@
 Thus, in order to allow debugging and verification of HDL code, the technology of simulation was invented.
-Here, the HDL is also translated into a netlist that represents all the parallel processes just as in hardware, but the subsequent step of synthesis is not taken.
+Here, the HDL is also translated into a netlist that represents all the parallel processes just like in hardware, but the subsequent step of synthesis is not taken.
 Therefore, simulation comes to a result much faster.
-Since HDL designs actually only consist of electrical wires and some blocks (todo richtiges wort?), it is difficult to verify correct values.
-The testbench environment is a very important factor in any project that includes hardware description language (HDL) code.
-The reason therefor is that debugging
+Since HDL designs actually only consist of electrical wires and some basic block primitives, like Flipflops or lookup tables, it is difficult to verify correct values.
+They would need to be compared in their binary bit-pattern format, at the correct point in time.
+Of course the HDL language provides support with different datatypes, that abstract the binary number representation and make the numbers human-readable.
+However, it remains an enormous effort to manually write testing code to verify a bus interface, for example.
+In order to simplify such complex operations, various testbench frameworks are available to developers.
+
+%%%%%%%%%%%%%
+\subsubsection{Simulator GHDL}
+
+%%%%%%%%%%%%%
+\subsubsection{Framework cocotb}
+
+\cite{cocotbDoc}.
 

### 2021-07-15 doc/thesis/chapters/implementation.tex
cocotb

@@
 
+Multiple different simulators are available on the market, such as Mentor Graphics ModelSim or QuestaSim, Synopsys VCS or the Cadence Simulator.
+The issue with most simulators is, that they are subject to a license fee, depending on the required set of features, or simulation speed.\\
+
+However, there are simulators available, that are free to use.
+An example therefor is GHDL, which is an analyzer, compiler and simulator for VHDL and is developed as an open-source project.
+GHDL translates the VHDL code to a binary file that contains machine code, which can be executed.
+This is supported for a all common operating systems such as Linux, Windows and macOS.
+The simulator further supports a VPI (Verilog Procedural Interface), which can be used to control the simulator \cite{GHDLDoc}.
+
 %%%%%%%%%%%%%
@@
 
-\cite{cocotbDoc}.
+The testbench framework cocotb is developed as a free, open-source project.
+It can be used to verify VHDL and SystemVerilog RTL code, in combination with an external simulator.
+Various different simulators can be used with cocotb - amongst others, GHDL is supported.
+The interface that is used to the simulator is the industry-standard VPI.\\
+
+One of the main advantages of cocotb is, that test cases for the respective RTL code are written in the language Python.
+It is therefore also called a co-simulation framework.
+Using Python provides a large factor of flexibility and possibilities, since it is a comprehensive language, considering all its extension libraries.
+Additionally, Python may be more familiar to engineers in comparison to VHDL and thus, writing testbenches is made simpler with cocotb.
+
+The test cases in Python are implemented as coroutines.
+By the way, the name cocotb is constructed by the words \textit{CO}routine, \textit{CO}simulation and \textit{T}est\textit{B}ench \cite{cocotbDoc}.\\
+
+Fig.\ref{fig:bd_cocotb_interface_simulator} gives an overview of how cocotb interfaces with a simulator.
+The Python test cases are shown as a list of coroutines, which are managed by a scheduler.
+This Python code interfaces with the simulator through standard interfaces like VPI.
+The block called \textit{GPI} is developed by cocotb and abstracts the different simulator interfaces, such as VPI, VHPI or FLI.
+In the simulator, the device under test (DUT) is instantiated.
+The RTL simulation itself is managed by the simulators' scheduler.
+
+\includepicture [1.0] [0]  {Block diagram of the interface between cocotb and a simulator \cite{cocotbDoc}.} {bd_cocotb_interface_simulator} {img/draw.io/bd_cocotb_interface_simulator}
 

### 2021-07-15 doc/thesis/chapters/implementation.tex
tb architecture

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Simulation Testbench}
+\subsection{Testbench Simulator and Framework}
 
@@
 \subsubsection{Framework cocotb}
+\label{sub:FrameworkCocotb}
 
@@
 \includepicture [1.0] [0]  {Block diagram of the interface between cocotb and a simulator \cite{cocotbDoc}.} {bd_cocotb_interface_simulator} {img/draw.io/bd_cocotb_interface_simulator}
+
+%%%%%%%%%%%%%
+\subsection{Testbench Architecture}
+
+The architecture for this testbench is created under the assumptions and according to the guidelines that are developed in Chapter \ref{cha:SystemArchitectureAndConcept}.
+
+Fig.\ref{fig:bd_vhdl_implemented_tb_architecture} shows a block diagram of the implemented architecture.
+The split between the Python environment and the VHDL implementation is clearly visible but is represented from the developers perspective.
+The actual interfaces, as described in \ref{sub:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.
+
+\includepicture [1.0] [0]  {Block diagram of the implemented testbench architecture.} {bd_vhdl_implemented_tb_architecture} {img/draw.io/bd_vhdl_implemented_tb_architecture}
+
+The test cases are using the external verification data, that was previously generated by the Matlab model.
+This data is supplied to the input of the DUT.
+The DUTs' input and output is implemented as an AXI Stream interface, which is an interface that features a valid-, as well as a ready-signal along with the data.
+The cocotb framework supports both an AXI Stream Master, and an AXI Stream Slave class, to write and read the input and output, respectively.
+In that way, the interface signals are abstracted to simple read and write function calls in the Python code, while the actual signal handling is done in the mentioned cocotb library classes.
+A clock generator class is also available from cocotb, which is used to provide the system clock to the DUT.
+The Python classes \textit{tb\_analyzer\_helper}, \textit{tb\_data\_handler} and \textit{fm\_receiver\_model} are implemented to create a better structure of the testbench and to provide code that is re-useable for the testbench of the HLS implementation.
+The \textit{tb\_analyzer\_helper} class provides functions to analyze the data, such as comparison against defined error thresholds or the creation of graphical plots.
+The \textit{tb\_data\_handler} only works as a data storage class, that encapsulates the data that is captured from the DUT.
+The class called \textit{fm\_receiver\_model} serves as the 'golden model reference' that is used to compare the DUT against.
+Internally, it loads the Matlab verification data.
+However, Python can be used to actually implement an FM receiver, just like in Matlab.
+Because the Matlab model already exists, this is not done here - loading the Matlab data is sufficient as it is been seen as 'the single source of truth' in this project.
+At the end of the test cases, the DUTs' output data is written to a file.
+This data is later used to directly compare the different implementation variants against each other.
 

### 2021-07-15 doc/thesis/chapters/implementation.tex
adaptions

@@
 The split between the Python environment and the VHDL implementation is clearly visible but is represented from the developers perspective.
-The actual interfaces, as described in \ref{sub:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.
+The actual interfaces, as described in \ref{sub:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.\\
 
@@
 The cocotb framework supports both an AXI Stream Master, and an AXI Stream Slave class, to write and read the input and output, respectively.
-In that way, the interface signals are abstracted to simple read and write function calls in the Python code, while the actual signal handling is done in the mentioned cocotb library classes.
+In that way, the interface signals are abstracted to simple read- and write-function calls in the Python code, while the actual signal handling is done in the mentioned cocotb library classes.
+
 A clock generator class is also available from cocotb, which is used to provide the system clock to the DUT.
+
 The Python classes \textit{tb\_analyzer\_helper}, \textit{tb\_data\_handler} and \textit{fm\_receiver\_model} are implemented to create a better structure of the testbench and to provide code that is re-useable for the testbench of the HLS implementation.
-The \textit{tb\_analyzer\_helper} class provides functions to analyze the data, such as comparison against defined error thresholds or the creation of graphical plots.
+The \textit{tb\_analyzer\_helper} class provides functions to analyze the data, such as comparison against defined error thresholds or the creation of graphical plots with Pythons' matplotlib library.
 The \textit{tb\_data\_handler} only works as a data storage class, that encapsulates the data that is captured from the DUT.
@@
 Internally, it loads the Matlab verification data.
-However, Python can be used to actually implement an FM receiver, just like in Matlab.
+This could be implemented differently, since Python can be used to actually implement an FM receiver including the entire DSP functionality, just like in Matlab.
 Because the Matlab model already exists, this is not done here - loading the Matlab data is sufficient as it is been seen as 'the single source of truth' in this project.
+
 At the end of the test cases, the DUTs' output data is written to a file.
@@
 \subsection{Architecture (same tb for VHDL and HLS-generated HDL)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Framework cocotb, with ghdl compiler}
-  Instantiate both HDL models in the testbench.\\
-  Display a direct comparison of outputs in graphs. This is practical, since the cocotb framework runs in python and graphs can be generated using Python's matplotlib.

### 2021-07-15 doc/thesis/chapters/implementation.tex
new block diagram - vhdl testbench architecture

@@
 
-Fig.\ref{fig:bd_vhdl_implemented_tb_architecture} shows a block diagram of the implemented architecture.
+Fig.\ref{fig:bd_testbench_architecture_vhdl} shows a block diagram of the implemented architecture.
 The split between the Python environment and the VHDL implementation is clearly visible but is represented from the developers perspective.
@@
 
-\includepicture [1.0] [0]  {Block diagram of the implemented testbench architecture.} {bd_vhdl_implemented_tb_architecture} {img/draw.io/bd_vhdl_implemented_tb_architecture}
+\includepicture [1.0] [0]  {Block diagram of the implemented testbench architecture.} {bd_testbench_architecture_vhdl} {img/draw.io/bd_testbench_architecture_vhdl}
 

### 2021-07-16 doc/thesis/chapters/implementation.tex
testing strategies

@@
 \subsubsection{General}
+\label{sub:VHDL:Testbench:General}
 
@@
 \subsubsection{Framework cocotb}
-\label{sub:FrameworkCocotb}
+\label{sub:VHDL:FrameworkCocotb}
 
@@
 \subsection{Testbench Architecture}
+\label{sub:VHDL:TestbenchArchitecture}
 
@@
 The split between the Python environment and the VHDL implementation is clearly visible but is represented from the developers perspective.
-The actual interfaces, as described in \ref{sub:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.\\
+The actual simulator interfaces, as described in \ref{sub:VHDL:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.\\
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Channel Selection (IF to Channel-BB)}
+\subsection{Testing Strategies}
+
+Common strategies for testing are unit tests and integration tests, which represent different levels of testing.
+Unit tests specifically target a single unit, while integration tests verify the combination of multiple units as a whole.
+Both of these stragegies are important.
+Given all unit tests being successful, it is ensured that all these single units are implemented correctly.
+However, their combination and interaction may still be subject to error - this is where integration tests unfold their potential.
+Implementation errors in the glue logic between the units, as well as the interaction of the respective interfaces can be found in these integration tests.
+
+The testbench architecture presented above represents an integration test, since the DUT is tested as an entire block.
+The advantage thereof is, that it tests the DUT in the same way as it will be integrated within the target hardware, the FPGA.
+However, in case the DUT fails to produce a correct output in the integration test, it is not always visible where exactly the error is located.
+Thus, in addition to the DUTs' interfaces, additional signals are logged from the inside of the DUT.
+that can be accessed through the testbench.
+white box: disadvantage: test needs to change as well, when implementation changes
+--> white box, black box
 
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Phase Detector}
+\subsection{Unit Tests}
+FIR the only unit that is worth the effort (others are only delays, and additions/substractions respectively)
+
+A unit test only targets a single unit of the system, as the name already describes.
+This is in the contrary to an integration test, which tests the combination of several units into an integrated system.
+The testbench architecture that is developed in \ref{sub:VHDL:TestbenchArchitecture} represents such an integration test.
+However, if an integration test fails, it is not always
 
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{other Elements}
+\subsection{TODO: Other Stuff...}
 
@@
 
+The implemented testbench structure, as well as some specific implementation details are explained here.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Channel Selection (IF to Channel-BB)}
+A testbench is equally important in the HLS implementation, as it is for the VHDL implementation, explained in \ref{sub:VHDL:Testbench:General}.
+The testbench simulation is
 
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Phase Detector}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{other Elements}
+\subsection{TODO: Other Stuff...}
 

### 2021-07-16 doc/thesis/chapters/introduction.tex
add a new chapter

@@
 
-\section{Goals}
+\section{Objective}
 
-The main goal of this thesis is to elaborate a system design for an embedded system, that combines all the knowledge of different disciplines and topics which are accumulated over the time of the university program, in a single project.
+The main objective of this thesis is to elaborate a system design for an embedded system, that combines all the knowledge of different disciplines and topics which are accumulated over the time of the university program, in a single project.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % NOTES:
-%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %- others?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FEEDBACK, Meeting 07-16-2021
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%formatting:
+%- absätze überall gleich (mit oder ohne leerzeile)
+%references:
+%- hauptsache überall gleich (Fig.2.3 oder Sec.7.1 oder Eq.3 oder Chpt.4)
+%- literature refs at the beginning of a chapter/section
+%  like "this chapter is based on [1][4][5]."
+%  or at the end of the last sentence like "...the end of this sentence. [7]"
+%
+%Allgemeiner Teil / Theorie:
+% zusätzliches kapitel:
+%   system design ansätze
+%   - from algo to VHDL (book: vom algorithmus zur hardware)
+%   - matlab to VHDL
+%   - systementwuf richtung test driven design
+%   --> design space exploration
+%
+%
+% ganze Arbeit aufhängen an:
+%   main objective: fit all the aspects (algorithm, impl, testing) into
+%                   a single project
+%
+%
+%
+%

### 2021-07-16 doc/thesis/chapters/system-design-process.tex
add a new chapter

%%%%%%%%%%%%%%%%%%%%
\chapter{System Design Process}
\label{cha:SystemDesignProcess}

### 2021-07-19 doc/thesis/chapters/deployment-on-hardware.tex
add some thoughts
\includepicture function with floating option; new block diagram
bd_impl_vivado diagram
structure sections

@@
 \label{cha:DeploymentOnHardware}
+
+%%%%%%%%%%%%%%%%%%%%
+\section{GNU Radio Transmitter}
+
+%%%%%%%%%%%%%%%%%%%%
 \section{Hardware Platform}
-    \subsection{RTL2832u Dongle}
-    \subsection{ZedBoard}
+ZedBoard
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{RTL-SDR}
+RTL2832U USB Dongle
+
+%%%%%%%%%%%%%%%%%%%%
+\section{Additional Implementation}
+
+Chapter \ref{cha:Implementation} already described the development of the main IP.
+In order to embed this IP on actual hardware, a significant additional amount of implementation is neccessary.
+In this section, these additional parts of the implementation, which specifically enable and support the deployment on hardware, are described.
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Block Diagram}
+
+\includepicture [1.0] {Block diagram of the Vivado implementation.} {bd_impl_vivado} {img/draw.io/bd_impl_vivado}
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Software}

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
notes
structure sections
chapter begin

@@
 \label{cha:SystemDesignProcess}
+
+The process of system design is a central topic in the development of any product.
+This chapter describes general techniques for system design.
+However, they can also be directly applied to FPGA design.
+Additionally, tools that are specifically used in FPGA design are introduced.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{General}
+
+The research for this chapter is mainly based on \cite{BahillA.Terry2017TDiS}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The SIMILAR Process}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Design for Reuse}
+%see pg. 51
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The Problem Statement}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Concept Exploration}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Validation and Verification}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{System Design Targeting FPGAs}
+The research for this chapter is mainly based on \cite{GesslerRalf2014EES}.
+%https://link-springer-com.fhooe.idm.oclc.org/content/pdf/10.1007%2F978-3-319-26408-0.pdf
+%chapter 3
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{V-Modell}
+
+%applies to general software development, but can directly be mapped to FPGA development.
+
+%Abbildung 3.5 in
+%https://link-springer-com.fhooe.idm.oclc.org/content/pdf/10.1007%2F978-3-8348-2080-8.pdf
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{High-Level Synthesis}
+%\cplusplus\\
+%SystemC (deprecated for synthesis by Xilinx)\\
+
+A comprehensive list of HLS tools can be found in Section 3.3 of
+%https://link-springer-com.fhooe.idm.oclc.org/content/pdf/10.1007%2F978-3-319-26408-0.pdf
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Implementation with Tool Support}
+%https://link-springer-com.fhooe.idm.oclc.org/content/pdf/10.1007%2F978-3-8348-2080-8.pdf
+
+%Matlab Simulink System Generator\\
+%Matlab Simulink HDL Coder\\
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Direct Implementation}
+
+%direct, or 'manual' implementation\\
+%VHDL/Verilog\\

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
similar process; problem statement

@@
 
-The research for this chapter is mainly based on \cite{BahillA.Terry2017TDiS}.
+The following sections describe a generalized approach to system design.
+A process called SIMILAR is explained, as well as selected topics that are relevant.
+The researched information for this section is mainly based on \cite{BahillA.Terry2017TDiS}.
 
@@
 \subsection{The SIMILAR Process}
+
+The SIMILAR process provides a general approach to problem solving.
+It can be applied to any kind of problem that can be of technical, as well as non-technical nature, in order to successfully achieve the specified requirements.
+The concept was developed by A.T. Bahill and B. Gissing, who evaluated and compared multiple existing processes to find their similarities.\\
+
+The acronym SIMILAR consists of the following components.
+
+\begin{itemize}
+  \item \textbf{S}tate the Problem
+  \item \textbf{I}nvestigate Alternatives
+  \item \textbf{M}odel the System
+  \item \textbf{I}ntegrate Components
+  \item \textbf{L}aunch the System
+  \item \textbf{A}ssess Performance
+  \item \textbf{R}e-evaluate
+\end{itemize}
+
+Each of these terms represents a separate step in the process diagram shown in Fig.\ref{fig:similar_process}.
+The steps can be seen as iterative and parallel, as the process runs during the product development cycle.
+
+\includepicture [0.8] [0] {The SIMILAR process in a graphical representation.} {similar_process} {img/similar_process}
+
+%%%%%
+\subsubsection{State the Problem}
+
+The very first step of the process is to state the problem that is to be solved.
+This can be the initial definition of a new product, but can also be the definition of an improvement that is to be made for an existing product.
+Because of that, input for the problem statement can be generated by various parties, such as customers, product developers, regulatory agencies or manufacturers, just to name a few.
+In the process diagram, all these possible input parties are labeled as stakeholders.\\
+
+An important guideline for the specification of this problem statement is, that each requirement is descibed in the way of \textit{what} needs to be done, not \textit{how} it needs to be done.\\
+
+The problem statement is updated and improved by the re-evaluate step, which is described later.
+
+%%%%%
+\subsubsection{Investigate Alternatives}
+
+%%%%%
+\subsubsection{Model the System}
+
+%%%%%
+\subsubsection{Integrate Components}
+
+%%%%%
+\subsubsection{Launch the System}
+
+%%%%%
+\subsubsection{Assess Performance}
+
+%%%%%
+\subsubsection{Re-evaluate}
 

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
investigate alternatives

@@
 
-The problem statement is updated and improved by the re-evaluate step, which is described later.
+The problem statement is updated and improved by the Re-evaluate step, which is described later.
 
@@
 \subsubsection{Investigate Alternatives}
+
+In almost any kind of problem, there are multiple ways in order to archieve a successful solution.
+However, some solutions may be preferred over others because of certain criteria that are evaluated.
+Such criteria may be cost, risk, schedule or performance.
+Based on the combined evaluation of these factors, a decision in favor of one selected alternative is made.\\
+
+It is important to note that this step is a repeated step.
+This is necessary, since there may only be a few known facts available at the initial intestigation.
+Further design steps, such as a first estimation of the intended design implementation, simulation results, or a prototype can provide more data and facts.
+These can be used to re-evaluate the alternatives and consequently optimize the design process.
 

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
system model

@@
 
+Based on the previous steps, a model for the system is developed here.
+This includes a process model, that is used for stragetic planning towards cost reduction, optimization of the implementation effort, or the schedule of subprojects, and a product model, which supports trade-offs and the identification of risks, or helps to explain the system as a whole.\\
+
+Typically, a coarse model is created for almost all alternative designs that were found in the previous step.
+A more detailed model is developed for the single, selected design.
+The model itself may be represented in many different ways, such as state machines, block diagrams, flow diagrams, simulations or object-oriented diagrams.
+
 %%%%%

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
integrate

@@
 
+This step combines multiple single components into a larger, integrated system, which can then be treated as a whole.
+In order to be able to do that, interfaces between the components and subsystems need to be defined.
+Optimally, these interfaces are designed in a way that the amount of data that needs to be exchanged is kept to a minimum.
+It is of advantage, if a subsystem only transfers finished products that it produced to another subsystem.
+The aim to develop simple interfaces is of special advantage in systems that include a feedback loop.
+There, managing the feedback loop connections is easier if there is a clear interface, so the loop can be connected around an entire subsystem, instead of interconnecting within the subsystems.
+
 %%%%%

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
launch

@@
 
+The system launch actually runs the system in a defined environment, so that it can perform what it is intended and designed to do.
+The defined environment can be the final, operational environment, or just a simulation which provides an output that is detailed enough for an evaluation.
+In a product that involves hardware development, this step may include the installation of commercial-off-the-shelf products with a modified prototype software.
+However, since the SIMILAR process is applicable for any kind of problem, this might also be a business process, where the launch step may be the roll-out of the subprocesses and tasks in the respective departments of a company.
+
 %%%%%

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
assess performance

@@
 
+This step is all about measurements.
+The performance of a system can only be evaluated, if it can be measured by using different metrics.
+Evaluation criteria are used, based on the measurements of the technical performance, which can help to find trade-offs or to mitigate risk in design and manufacturing.
+
 %%%%%

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
re-evaluate

@@
 
+The re-evaluate step is a central building block in the SIMILAR concept.
+All of the other steps have a connection to it, in the form of a feedback loop.
+However, depending on the nature of the project and the point in the timeline of its life cycle, not all of these loops may be exercised all the time.
+The idea of re-evaluation and the subsequent improvement of a process is a widely applied and well-known concept.
+Its main function is to observe an output and to use this information to improve the systems' input.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
design for reuse

@@
 
-\includepicture [0.8] [0] {The SIMILAR process in a graphical representation.} {similar_process} {img/similar_process}
+\includepicture [0.7] [0] {The SIMILAR process in a graphical representation.} {similar_process} {img/similar_process}
 
@@
 The defined environment can be the final, operational environment, or just a simulation which provides an output that is detailed enough for an evaluation.
-In a product that involves hardware development, this step may include the installation of commercial-off-the-shelf products with a modified prototype software.
+In a product that involves hardware development, this step may include the installation of commercial off-the-shelf products with a modified prototype software.
 However, since the SIMILAR process is applicable for any kind of problem, this might also be a business process, where the launch step may be the roll-out of the subprocesses and tasks in the respective departments of a company.
@@
 \subsection{Design for Reuse}
-%see pg. 51
+
+A design principle that is often used is design for reuse.
+It enables a faster and more efficient development of products, which leads to an overall decrease in development cost.
+Additionally, the product is likely less error-prone, since the reused parts of the system are already tested at this point.
+After all these positive arguments, there is a counter argument to it, though: designing a system with reuseable subsystems comes with cost.
+Additional effort is required to design the subsystems in a way so that they can be reused in another project.
+However, this additional effort may potentially already be paid off, when a single subsystem can be reused in a future product and thus, it is worth the effort.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The Problem Statement}
+\subsection{Creation of a Problem Statement}
 

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
validation / verification

@@
 \subsection{The SIMILAR Process}
+%TODO: maybe include additional facts from https://slideplayer.com/slide/14327503/
 
@@
 
+Generally speaking, the problem statement serves the purpose to build a common understanding of the problem between a stakeholder and the developer.
+In case there is a mismatch, an engineer may spend valuable time to develop a sophisticated solution to a problem that was not even a requirement.
+Thus, any time spent to develop this feature is worthless, which highlights the importance of a correct problem statement.\\
+
 An important guideline for the specification of this problem statement is, that each requirement is descibed in the way of \textit{what} needs to be done, not \textit{how} it needs to be done.\\
@@
 
-In almost any kind of problem, there are multiple ways in order to archieve a successful solution.
+In almost any kind of problem, there are multiple ways in order to achieve a successful solution.
 However, some solutions may be preferred over others because of certain criteria that are evaluated.
 Such criteria may be cost, risk, schedule or performance.
-Based on the combined evaluation of these factors, a decision in favor of one selected alternative is made.\\
+Based on the combined evaluation of these factors, a decision in favor of one selected alternative is made.
+This part of the process is also called concept exploration.\\
 
-It is important to note that this step is a repeated step.
+It is important to note that this step is a repeated, iterative step.
 This is necessary, since there may only be a few known facts available at the initial intestigation.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Creation of a Problem Statement}
+\subsection{Validation and Verification}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Concept Exploration}
+The terms validation and verification are closely related, but have a different meaning in system design.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Validation and Verification}
+Validation fulfils the purpose to prove that the developed product does what it is intended to do.
+It strives to build the \textit{right system}.
+Verification on the other hand tries to build the \textit{system right}, by testing its properties against the previously determined specification and requirements.\\
 

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
adapt

@@
 The terms validation and verification are closely related, but have a different meaning in system design.
+Verification tests the properties of the developed product against the previously determined specification and requirements.
+It stives to build the \textit{system right}.
+Validation on the other hand tries to build the \textit{right system}.
+It proves that the product does what it is intended to do.
+This is important in order to develop the product according to what the customer actually needs.
+To avoid to divert from these customer needs, validation is done throughout the products' development cycle, from the beginning to the end.
+Various artifacts can be collected during the process, in order to have a traceable documentation of the decision that were made.\\
 
-Validation fulfils the purpose to prove that the developed product does what it is intended to do.
-It strives to build the \textit{right system}.
-Verification on the other hand tries to build the \textit{system right}, by testing its properties against the previously determined specification and requirements.\\
+Validation may also fail, which means that the chosen system may be the wrong system for the respective customers needs.
+As an example, the system may be very sensitive to variations of a parameter that a customer cannot influence.
+Another example is when a system is more sensitive to certain parameters than to its inputs.
+In both cases, the system may be the wrong system for the specific problem and should urgently be re-evaluated.\\
+
+In order to expose such failures as early as possible, any test engineer should be instructed to always keep an eye on the process of validation, which means to check if the team is building the \textit{right system}.
+Therefore, a test engineer should be familiar with the problem statement.
 

### 2021-07-19 doc/thesis/chapters/system-design-process.tex
v-modell diagram

@@
 \section{System Design Targeting FPGAs}
-The research for this chapter is mainly based on \cite{GesslerRalf2014EES}.
-%https://link-springer-com.fhooe.idm.oclc.org/content/pdf/10.1007%2F978-3-319-26408-0.pdf
-%chapter 3
+
+A wide range of design approaches are available that are targeting FPGAs.
+In this section, some selected approaches are described.
+The research for this section is mainly based on \cite{GesslerRalf2014EES}.
 
@@
 
-%applies to general software development, but can directly be mapped to FPGA development.
+applies to general software development, but can directly be mapped to FPGA development.
 
-%Abbildung 3.5 in
-%https://link-springer-com.fhooe.idm.oclc.org/content/pdf/10.1007%2F978-3-8348-2080-8.pdf
+\includepicture [1.0] [0] [H] {The V-Modell \cite[Abbildung 3.5]{GesslerRalf2014EES}.} {v_modell} {img/draw.io/v_modell}
 
@@
 \subsection{High-Level Synthesis}
-%\cplusplus\\
-%SystemC (deprecated for synthesis by Xilinx)\\
+\cplusplus\\
+SystemC (deprecated for synthesis by Xilinx)\\
 
-A comprehensive list of HLS tools can be found in Section 3.3 of
-%https://link-springer-com.fhooe.idm.oclc.org/content/pdf/10.1007%2F978-3-319-26408-0.pdf
+A comprehensive list of HLS tools can be found in Section 3.3 of \cite{2016FfSP}
+
+HLS is described into more detail in section \ref{cha:HLS}.
 
@@
 \subsection{Implementation with Tool Support}
-%https://link-springer-com.fhooe.idm.oclc.org/content/pdf/10.1007%2F978-3-8348-2080-8.pdf
+see \cite{GesslerRalf2014EES}\\
 
-%Matlab Simulink System Generator\\
-%Matlab Simulink HDL Coder\\
+Matlab Simulink System Generator\\
+Matlab Simulink HDL Coder\\
 
@@
 
-%direct, or 'manual' implementation\\
-%VHDL/Verilog\\
+direct, or 'manual' implementation\\
+VHDL/Verilog\\

### 2021-07-20 doc/thesis/chapters/system-design-process.tex
similar diagram (wip)
v-model

@@
 
-\includepicture [0.7] [0] {The SIMILAR process in a graphical representation.} {similar_process} {img/similar_process}
+\includepicture [0.95] [0] {The SIMILAR process in a graphical representation \cite{BahillA.Terry2017TDiS}.} {similar_process} {img/draw.io/similar_process}
 
@@
 \subsection{Validation and Verification}
+\label{sec:ValidationAndVerification}
 
@@
 
-applies to general software development, but can directly be mapped to FPGA development.
+The development of a product requires an organized approach, in order to successfully achieve the requirements.
+The common term for such an organized approach is the process model, which represents a guideline on how to organize and structure the development cycle.\\
 
-\includepicture [1.0] [0] [H] {The V-Modell \cite[Abbildung 3.5]{GesslerRalf2014EES}.} {v_modell} {img/draw.io/v_modell}
+A well-known process model is the so called V-Model.
+It is based on the similarly well-known Waterfall Model, and has its name because of the distinct V-shape of its process diagram, as it is shown in Fig.\ref{fig:v_modell}.
+One of the main characteristics of the V-Model is, that one phase needs to be completed before the following phase can start.
+This classifies the model as a static process model and is one of the main disadvantages at the same time, because it makes the model inflexible to adaptions of the requirements.
+The advantages however are that it is easy to install and apply, each phase has its specific deliverables, good results can be achieved for large projects with simple requirements and the early test phases.\\
+
+The disadvantages of inflexibility towards changes of requirements and the fact, that it is difficult to consider all all eventualities in the creation of the V-model, led to the innovation of modern process models, which are iterative and thus more flexible.\\
+
+The V-model applies to general software development, or project development, but can directly be mapped to FPGA development.
+The model is covered in this thesis section, because it shows the clear testing structure of module tests (unit tests), integration tests and system tests, which is especially important in projects that target hardware, such as FPGA projects.\\
+
+Furthermore, the V-model diagram in Fig.\ref{fig:v_modell} shows the difference between validation and verification, as it is explained in the previous section \ref{sec:ValidationAndVerification}.
+
+\includepicture [1.0] [0] {The V-Modell process diagram \cite[Abbildung 3.5]{GesslerRalf2014EES}.} {v_modell} {img/draw.io/v_modell}
 

### 2021-07-20 doc/thesis/chapters/system-architecture-and-concept.tex
structuring
direct impl

@@
       This includes a range of toolboxes for signal processing, such as filter designers or FFT functions, as well as convenient ways to visualize data \cite{SoftwareMatlab}.
-  \item \textbf{FPGA: \cplusplus High-Level Synthesis}\\
+  \item \textbf{FPGA: \cplusplus\ High-Level Synthesis}\\
       In order to use an FPGA, a hardware design needs to be developed for it.
@@
 A good system structure tries to share as many parts as possible in the system.
-This is to minimize the points of failure, as well as to reduce the effort of implementation.
+This is to minimize the points of failure, as well as to reduce the effort of implementation.\\
 
@@
 
-%%%%%%%%%%%%%%%%%%%%
-\section{Test Environment}
+%%%%%%%%%%%%%%%%%%%%%
+%\section{Test Environment}

### 2021-07-20 doc/thesis/chapters/system-design-process.tex
direct impl

@@
 
-A wide range of design approaches are available that are targeting FPGAs.
+A wide range of system-level design approaches are available that are targeting FPGAs.
 In this section, some selected approaches are described.
@@
 \subsection{High-Level Synthesis}
-\cplusplus\\
-SystemC (deprecated for synthesis by Xilinx)\\
 
-A comprehensive list of HLS tools can be found in Section 3.3 of \cite{2016FfSP}
+High-Level Synthes (HLS) is an FPGA design methodology that provides a high level of abstraction, and thus can be seen as a system level design method.
+Different languages are used for HLS - two well-known languages are SystemC, which is a library for \cplusplus\ to provide modelling and simulation features, and plain \cplusplus.
+A comprehensive list of available HLS tools and languages can be found in Section 3.3 of \cite{2016FfSP}.\\
 
-HLS is described into more detail in section \ref{cha:HLS}.
+The main idea of HLS is, that the hardware design can be written in abstract modules, that represent the actual functional meaning of a feature or algorithm, rather than the complex, low-level hardware details.
+HLS achieves this by allowing to develop the code in a high-level language, such as \cplusplus.
+This high-level code is then automatically transformed into hardware description language (HDL) by the respective HLS tool.\\
+
+In this thesis, only the HLS of plain \cplusplus\ is covered.
+SystemC is not discussed, since it may potentially be less important in the future.
+In the latest Xilinx HLS tools 'Vitis HLS' and 'Vivado HLS', the language SystemC is already deprecated \cite{VivadoHlsSystemC}.\\
+
+\noindent
+HLS is described into more detail in Chapter \ref{cha:HLS}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Direct Implementation}
+
+The implementation of low-level hardware description language (HDL) in the register transfer level (RTL) is denoted as 'direct', or 'manual' implementation in this thesis.\\
+
+Two commonly used languages therefor are VHDL and Verilog.
+Both have similar features to describe hardware in the signal level, with parallel structures and timing-dependent architectures.
+The languages can describe a design in a hierarchical order, so that the separate modules can be constructed in a way that an overview of the functionality of the system remains given.
+Synthesis, as well as simulation is supported, which allows efficient development of RTL designs.\\
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Direct Implementation}
+In Fig.\ref{fig:fpga_design_levels}, multiple implementation methods are compared, with their respective levels of abstraction.
 
-direct, or 'manual' implementation\\
-VHDL/Verilog\\
+\includepicture [0.5] [0] {Abstraction levels of the different implementation languages \cite[Abbildung 5.1]{GesslerRalf2014EES}.} {fpga_design_levels} {img/fpga_design_levels}

### 2021-07-20 doc/thesis/chapters/conclusion-and-prospect.tex
level of abstraction

%%%%%%%%%%%%%%%%%%%%
\chapter{Conclusion and Prospect}
\label{cha:ConclusionAndProspect}

% NOTES:
%  - use Matlab Simulink HDL Coder and System Generator
%    as additional implementation variants
%
%
%
%
%
%

### 2021-07-20 doc/thesis/chapters/system-design-process.tex
level of abstraction

@@
 \subsection{Implementation with Tool Support}
-see \cite{GesslerRalf2014EES}\\
 
-Matlab Simulink System Generator\\
-Matlab Simulink HDL Coder\\
+The implementation of FPGA designs with tool support focuses on the development as a model-based approach, that is often featured by a graphical programmers interface.
+This is often used for prototyping in an early phase of product development.
+A well-known example for such a model-based approach is Matlab, and its graphical tool Simulink.\\
 
-In Fig.\ref{fig:fpga_design_levels}, multiple implementation methods are compared, with their respective levels of abstraction.
+Matlab can be used to describe and develop hardware-oriented algorithms, since it supports various library functions, such as the fixed point representation of numbers.
+A major advantage is, that fast prototyping of an algorithm, as well as a fast simulation of the system is achievable.
+Simulink is a graphical editor, that allows to connect functional blocks in a block diagram and create a system in that way.
+Multiple toolboxes support additional blocks, that abstract complex functionalities, so that a fast prototyping is possible here as well.\\
 
-\includepicture [0.5] [0] {Abstraction levels of the different implementation languages \cite[Abbildung 5.1]{GesslerRalf2014EES}.} {fpga_design_levels} {img/fpga_design_levels}
+A powerful feature of Matlab and Simulink is, that it can be used similar to an HLS tool.
+Matlab code, as well as Simulink block diagrams can be transformed into low-level RTL code for an FPGA design.
+
+%%%%%
+\subsubsection{Simulink System Generator}
+
+A full system-level design approach can be covered with the Simulink System Generator.
+This includes the high-level development of the system as a block diagram in Simulink, the automated generation of low-level RTL code for the FPGA hardware, and the verification in simulation and in hardware, in the form of hardware-in-the-loop.
+This workflow is supported for Xilinx FPGAs and is specifically targeting system design engineers, who work in rapid prototyping to generate hardware utilization estimates for the final products' hardware.\\
+
+Fig. \ref{fig:simulink_system_generator} shows this entire design approach in a block diagram.
+The abbreviation ISE stands for Integrated Software Environment, and is Xilinx' development environment that includes the FPGA synthesis tool.
+
+\includepicture [0.8] [0] {Overview of the capabilities of the Simulink System Generator \cite[Abbildung 5.11]{GesslerRalf2014EES}.} {simulink_system_generator} {img/simulink_system_generator}
+
+%%%%%
+\subsubsection{Simulink HDL Coder}
+
+The Simulink HDL Coder tool can generate VHDL or Verilog code from Matlab functions or Simulink block diagrams and state diagrams.
+The generated code can be used in Xilinx, as well as Intel/Altera FPGA hardware.
+This allows prototyping of algorithms, while continuously having detailed information of the target implementation, such as an estimate of the hardware logic utilization, or even the critical path in the logic.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Level of Abstraction}
+
+Multiple implementation methods are introduced in above sections, which all have different levels of abstraction.
+VHDL and Verilog require, or allow, a very detailed description of the RTL code, while SystemC, \cplusplus, or Matlab and Simulink provide a high level of abstraction.
+In Fig.\ref{fig:fpga_design_levels}, the various implementation methods are compared, with their respective levels of abstraction.\\
+
+In conclusion, it is an engineers' choice which implementation method to use in order to develop a product.
+A trade-off between optimization and the level of abstraction is likely to be made.
+The decision may not always be obvious, but should be based on the fundamental requirements of the final product.
+
+\includepicture [0.5] [0] [H] {Abstraction levels of the different implementation languages \cite[Abbildung 5.1]{GesslerRalf2014EES}.} {fpga_design_levels} {img/fpga_design_levels}

### 2021-07-20 doc/thesis/chapters/system-design-process.tex
new diagrams

@@
 The languages can describe a design in a hierarchical order, so that the separate modules can be constructed in a way that an overview of the functionality of the system remains given.
-Synthesis, as well as simulation is supported, which allows efficient development of RTL designs.\\
+Synthesis, as well as simulation is supported, which allows efficient development of RTL designs.
 
@@
 The abbreviation ISE stands for Integrated Software Environment, and is Xilinx' development environment that includes the FPGA synthesis tool.
+ModelSim is a HDL simulator software, that is developed by Mentor Graphics.
 
-\includepicture [0.8] [0] {Overview of the capabilities of the Simulink System Generator \cite[Abbildung 5.11]{GesslerRalf2014EES}.} {simulink_system_generator} {img/simulink_system_generator}
+\includepicture [0.9] [0] {Capabilities of the Simulink System Generator \cite[Abbildung 5.11]{GesslerRalf2014EES}.} {simulink_system_generator} {img/draw.io/simulink_system_generator}
 
@@
 
-\includepicture [0.5] [0] [H] {Abstraction levels of the different implementation languages \cite[Abbildung 5.1]{GesslerRalf2014EES}.} {fpga_design_levels} {img/fpga_design_levels}
+\includepicture [0.65] [0] [H] {Abstraction levels of the implementation methods \cite[Abbildung 5.1]{GesslerRalf2014EES}.} {fpga_design_levels} {img/draw.io/fpga_design_levels}
+
+%TODO: try to get this figure up into the last page...

### 2021-07-20 doc/thesis/chapters/dsp.tex
impl details
new diagram (matlab transmitter)
matlab transmitter
transmitrer
fix all newlines

@@
 The frequency needs to be high, to radiate the signal by an antenna.
-
 On the receiver side, once the signal reaches the antenna of a receiver, there are further similarities.
@@
 ; HF to IF to BB --> real signal ) ??
-
 These three steps - amplification, filtering and down-conversion to an IF - are often done in the analog domain, since digital circuits cannot work with such high frequencies.
@@
 This is done by an analog-digital-converter (ADC).
-
 After the signal is processed to digital baseband, the received signal can finally be decoded to get the actual data content.
@@
 A disadvantage is that the signal noise around the image frequency band is translated into the IF output band.
-Thus, the filter requirements are relatively high, since they need to deliver a strong attenuation in order to suppress noise outside the desired band.
+Thus, the filter requirements are relatively high, since they need to deliver a strong attenuation in order to suppress noise outside the desired band.\\
 
@@
 A direct-conversion receiver, as its name suggests, converts an RF signal to baseband directly.
-It can also be called homodyne, synchrodyne, or zero-IF receiver.
+It can also be called homodyne, synchrodyne, or zero-IF receiver.\\
 
 In order to achieve direct down-conversion, the local oscillator of the mixer needs to be set exactly to the center frequency of the desired signal.
-
 The main advantage of direct-conversion receivers is, that their output is an image-free signal.
@@
 This lower frequency can either be zero, i.e. baseband, or an intermediate frequency, such as in the application of a superheterodyne receiver.
-
 In order to achieve this, the incoming signal needs to be multiplicated with a carrier signal, that is generated locally in the receiver, by a local oscillator (LO).
 Depending on the frequency $f_{LO}$ of this local carrier, the resulting signal will be shifted to baseband, or an IF.
-Note, that in the case of down-conversion, the local oscillator frequency $f_{LO}$ is interchangeably denoted as carrier frequency $f_c$.
+Note, that in the case of down-conversion, the local oscillator frequency $f_{LO}$ is interchangeably denoted as carrier frequency $f_c$.\\
 
@@
 This equation is depicted as a block diagram and signal spectrum in Fig.\ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.
-
 The signal $m(t)$ is a baseband signal.
@@
 The received signal is a real signal, and thus its negative frequency spectrum is an exact mirror of the positive frequencies.
-This fact can be explained in two simple equations.
+This fact can be explained in two simple equations.\\
 
@@
 \section{Sampling Theorem}
+\label{sec:SamplingTheorem}
 
@@
 A digital signal is obtained by sampling an analog signal in a periodic interval, the so-called sampling frequency $f_s$.
-
 The sampling frequency needs to meet the sampling theorem, in order to correctly represent the analog signal.
@@
 Any digital signal processing technique must hold the specification of the sampling theorem \cite[chpt. 4.2]{AlessioSilviaMaria2016DSPa}.
-
 In case the theorem is violated, the effect of aliasing occurs on the generated signal.
@@
 The probably best known usecase therefor is commercial broadcast radio, where an audio stream is transmitted.
-Devices to receive these streams are available for low prices to the public.
+Devices to receive these streams are available for low prices to the public.\\
 
@@
 Simply looking at this equation, the frequency deviation, which is sometimes also called the swing, varies in the range between $f_c \pm \Delta f$.
-
 Considering the relationship between frequency and angle, as described above, and after some simple substitutions which will not be described into detail here, the equation for a generic frequency modulated wave is the following
@@
 % More info on modulation index: https://fas.org/man/dod-101/navy/docs/es310/FM.htm
-
 Formula \eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
@@
 \subsection{Frequency Band}
+\label{sec:FrequencyBand}
 
@@
 It must not be fixed to a single center frequency.
-Through this specification, a transmitter is capable of switching its transmission center frequency, in order to avoid interference with another transmitters signal.
+Through this specification, a transmitter is capable of switching its transmission center frequency, in order to avoid interference with another transmitters signal.\\
 
@@
 \subsection{Frequency Spectrum of a Broadcast FM Channel}
+\label{sec:FrequencySpectrumOfABroadcastFmChannel}
 
@@
 The upper limit at 15 kHz is chosen to maintain a sufficient spacing to the first subcarrier.
-
 For audio streaming, as in FM broadcasting, this is not a limitation, since it is unlikely to have frequencies higher than 15 kHz in an audio stream.
-Also, this already reaches the upper limit of the human ears' bandwith, so higher spectral parts will not be heard anyway.
+Also, this already reaches the upper limit of the human ears' bandwidth, so higher spectral parts will not be heard anyway.
 
@@
 The modulation technique that is used for this spectral part is called dual-sideband suppressed-carrier (DSB-SC).
-Even though the carrier is suppressed, it can still be recovered, because it is phase coherent with the 19 kHz pilot tone per definition.
+Even though the carrier is suppressed, it can still be recovered, because it is phase coherent with the 19 kHz pilot tone per definition.\\
 
@@
 Considering all these parts in the spectrum, there is still free bandwidth available to use up to the maximum channel bandwitdh of 100 kHz in one sideband.
-Because of that, additional services were added to the pure audio transmission, to provide additional data services and information.
+Because of that, additional services were added to the pure audio transmission, to provide additional data services and information.\\
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Emphasis and De-Emphasis}
-
-TODO
+%\subsection{Pre-Emphasis and De-Emphasis}
+%\label{sec:PreEmphasisAndDeEmphasis}
+%TODO
 
@@
 The FM signal that is evaluated here is assumed to be a quadrature-mixed signal in baseband, which means that inphase and quadrature (I/Q) signals are available.
-
+%
 %TODO: show downconversion path (maybe as a figure?)
 %https://www.veron.nl/wp-content/uploads/2014/01/FmDemodulator.pdf
-
+%
 In the following sections, two digital FM demodulator variants are described.
@@
 % TODO: add img/discrimination_method_bd2.png here (?)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Baseband Delay Demodulator}
+\label{sec:BasebandDelayDemodulator}
+TODO\\
+see /home/mike/Dropbox/Studium/ma\_semester\_4\_SS21/ma/literature/FmDemodulator.pdf
 
@@
 
-\includepicture [0.75] {Frequency discriminator block diagram \cite{FigBdFreqDiscriminator}.} {bd_freq_discriminator} {img/draw.io/bd_freq_discriminator}
+\includepicture [0.75] [0] [H] {Frequency discriminator block diagram \cite{FigBdFreqDiscriminator}.} {bd_freq_discriminator} {img/draw.io/bd_freq_discriminator}
 
@@
 In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.
-A block diagram is shown in Fig.\ref{fig:bd_freq_discriminator}.
+A block diagram is shown in Fig.\ref{fig:bd_freq_discriminator}.\\
 
@@
 In the case of FM demodulation, the loop filter in the feedback branch is configured to be able to follow the frequency variations on the input, which is directly related to the encoded information.
-The output of the PLL directly delivers an AM signal, which corresponds to the transmitted information. \cite{SchnyderHaller2002}
+The output of the PLL directly delivers an AM signal, which corresponds to the transmitted information. \cite{SchnyderHaller2002}\\
 
@@
 The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly, including the de-emphasis of the signal.
-
 So at this point the FM signal is converted to an AM signal.
-Thus, traditional AM DSP techniques can be employed to decode the FM broadcast channel.
+Thus, traditional AM DSP techniques can be employed to decode the FM broadcast channel.\\
 
@@
 Afterwards, the pilot tone frequency is doubled to achieve a 38 kHz subcarrier.
-This guarantees a phase-coherency between pilot and subcarrier, which is required to correctly shift the difference signal down to baseband.
+This guarantees a phase-coherency between pilot and subcarrier, which is required to correctly shift the difference signal down to baseband.\\
 

### 2021-07-20 doc/thesis/chapters/introduction.tex
v-model
direct impl
new diagrams
remove unused/old pictures in img/
fix all newlines

@@
 Instead, it is often treated as a standalone part that fulfills its own, specific functionality.
-
-Therefore, one of the main motivation points of this thesis is to combine several disciplines and multiple implementation methods in a common context of a single project.\\
+Therefore, one of the main motivation points of this thesis is to combine several disciplines and multiple implementation methods into a single system design, and treat it in this common project context.\\
 
@@
 The main objective of this thesis is to elaborate a system design for an embedded system, that combines all the knowledge of different disciplines and topics which are accumulated over the time of the university program, in a single project.
-
 This includes the development of a system model in a tool like Matlab, the practical usage and implementation of DSP in an FPGA, the development of a digital receiver and the deployment on actual hardware.\\
@@
 The first variant is traditional, manually written hardware description language (HDL) in the programming language VHDL, as it has been done for years.
-For the second variant, high-level synthesis (HLS) is employed, to generate the low-level HDL from high-level \cplusplus code.
-These two methods are implemented and compared on the basis of usability and some metrics.
+For the second variant, high-level synthesis (HLS) is employed, to generate the low-level HDL from high-level \cplusplus\ code.
+These two methods are implemented and compared on the basis of usability and some metrics.\\
 
@@
 %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO General
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% ## unify all graphs and diagrams
+%       - all gray; use exact same gray tone; eliminate the yellow?
+%
+%

### 2021-07-20 doc/thesis/chapters/system-design-process.tex
fix all newlines

@@
 
+\noindent
 The acronym SIMILAR consists of the following components.
@@
   \item \textbf{A}ssess Performance
-  \item \textbf{R}e-evaluate
+  \item \textbf{R}e-evaluate\\
 \end{itemize}
@@
 VHDL and Verilog require, or allow, a very detailed description of the RTL code, while SystemC, \cplusplus, or Matlab and Simulink provide a high level of abstraction.
-In Fig.\ref{fig:fpga_design_levels}, the various implementation methods are compared, with their respective levels of abstraction.\\
+In Fig.\ref{fig:fpga_design_levels}, the various implementation methods are compared.
+The respective level of abstraction is higher towards the upper end, while the level of detail increases towards the bottom of the figure.\\
 

### 2021-07-20 doc/thesis/chapters/implementation.tex
add a new chapter
fix all newlines
adaptions; white/black box; move figures

@@
 
-Fig.\ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram.
+\noindent
+Fig.\ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram.\\
 
-\includepicture [1.02] [0] [H] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/draw.io/bd_dsp_detailed}
+%\includepicture [1.02] [0] [H] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/draw.io/bd_dsp_detailed}
 
+\noindent
 The following list explains implementation details about the respective DSP blocks.
@@
       It is important that the recovered carrier is phase-coherent with the received input, as described in Section \ref{sec:demodulation-of-broadcast-fm}.
-      This can be achieved, by simply multiplying the pilot tone with itself, which is the way it is implemented here.
+      This can be achieved, by simply multiplying the pilot tone with itself, which is the way it is implemented here.\\
 \end{itemize}
+
+\includepicture [1.02] [0] [H] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/draw.io/bd_dsp_detailed}
 
@@
 Matlab is used to create a model of the receiver, which is used as a reference for the other implementation variants.
-Here, DSP algorithms are tested and evaluated, to find a method that is suitable to implement in hardware, specifically in an FPGA.
+Here, DSP algorithms are tested and evaluated, to find a method that is suitable to implement in hardware, specifically in an FPGA.\\
 
@@
 The audio file contains data in a sample rate of 48 kHz.
-Next, the sample rate is increased by an oversampling factor of 20, to achieve a sufficient sample rate for the subsequent DSP procedures.
+Next, the sample rate is increased by an oversampling factor of 20, to achieve a sufficient sample rate for the subsequent DSP procedures.\\
 
@@
 However, the higher rate was chosen, mainly because a higher number of signal samples beautify any plots that are created during analysis.
-Since a thorough optimization of the DSP chain is not the focus of this thesis, this was an acceptable trade-off.
+Since a thorough optimization of the DSP chain is not the focus of this thesis, this was an acceptable trade-off.\\
 
@@
 This is done to reduce some complexity in the hardware implementation.
-The following stage generates the Mono and LR Diff signals, by simply calculating the summation and diffence of the left and right channels, respectively.
+The following stage generates the Mono and LR Diff signals, by simply calculating the summation and diffence of the left and right channels, respectively.\\
 
@@
 Thus, the Mono signal, as well as the Pilot Tone need to be delayed by the amount of samples defined by the BP filters' group delay.
-The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone \cite{FmMultiplexingForStereo}.
+The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone \cite{FmMultiplexingForStereo}.\\
 
@@
 However, the Hinz Triller is still transmitted in many radio stations and can often be heard before and after the traffic announcements \cite{HinzTriller}.
-In the implemented transmitter, the Hinz Triller is implemented, but is disabled, since it only represents a temporarily transmitted signal.
+In the implemented transmitter, the Hinz Triller is implemented, but is disabled, since it only represents a temporarily transmitted signal.\\
 
 The Radio Data System (RDS) signal is not implemented in the transmitter, since it consists of a comprehensive protocol and multiple steps of modulation, such as differentially-coded BPSK \cite{IntroFmStereoRdsModulation} and is therefor too time-consuming to implement.
-Furthermore, it would not contribute any additional value to this project, since this is not the main focus.
+Furthermore, it would not contribute any additional value to this project, since this is not the main focus.\\
 
-At this point in the DSP chain, the signal resembles the frequency spectrum of Fig.\ref{fig:fm_channel_baseband_freqs}.
+At this point in the DSP chain, the signal resembles the frequency spectrum that is displayed in Fig.\ref{fig:fm_channel_baseband_freqs}.\\
 
@@
 As a final step, the sample rate is decreased again, back to the chosen sampling rate of 960 kHz.
-
 This signal can now be processed in a digital FM receiver.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Pre-Emphasis and De-Emphasis}
-TODO: show effects in spectrum, show how to design the filters
+%\subsection{Pre-Emphasis and De-Emphasis}
+%TODO: show effects in spectrum, show how to design the filters
+%see \ref{PreEmphasisAndDeEmphasis}
 
@@
 Further analysis is done in the frequency domain, especially in order to better understand the modulation of the different spectral parts in the FM channel.
-An example thereof is shown in Fig.\ref{fig:matlab_analysis_freq_domain}.
+An example thereof is shown in Fig.\ref{fig:matlab_analysis_freq_domain}.\\
 
@@
 A signal range from 70 MHz to 6 GHz can be covered by the front end.
-The board further features a Xilinx Spartan-6 FPGA that is user-programmable, which makes the device very flexible for various applications with the received signal \cite{USRPb200Mini}.
+The board further features a Xilinx Spartan-6 FPGA that is user-programmable, which makes the device very flexible for various applications with the received signal \cite{USRPb200Mini}.\\
 
@@
 It is based on the RTL2832U chipset, which includes an 8-bit ADC and a digital signal processor.
-The device delivers IQ data via the USB interface \cite{RTLSDR}.
+The device delivers IQ data via the USB interface \cite{RTLSDR}.\\
 
+\noindent
 Based on its hardware, the RTL-SDR can only be used as a receiver.
@@
 Multiple blocks are connected to a block diagram, that is very similar to the one in Fig.\ref{fig:bd_matlab_transmitter}.
-The USRP device is used to broadcast the RF signal over the air.
+The USRP device is used to broadcast the RF signal over the air.\\
 
@@
 
-\noindent
 Special care needs to be taken when transmitting signals over the air.
@@
 Please make sure to adapt your setup to the respective local regulations, to avoid any legal issues.
-Some of the regulations are mentioned in \ref{sec:FrequencyBand}, but may differ in various countries, however.
+Some of the regulations are mentioned in \ref{sec:FrequencyBand}, but may differ in various countries, however.\\
 
+\noindent
 Fig.\ref{fig:bd_gnuradio_setup_transmitter} shows a block diagram of the transmitter application.
@@
 This simple receiver was intended as a proof-of-concept and a first example of the cooperation of the external hardware and the GNU Radio software.
-Thus, the mono receiver was sufficient, as the focus of the GNU Radio application is put on the transmitter side.
+Thus, the mono receiver was sufficient, as the focus of the GNU Radio application is put on the transmitter side.\\
 
 This receiver works with both devices, the USRP b200mini and the RTL-SDR.
-
 Fig.\ref{fig:bd_gnuradio_setup_receiver} shows a block diagram of the receiver application.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{VHDL (TODO: no shortcut?)}
+\section{VHDL}
 
@@
 Various verification and testing libraries exist, in order to create unit tests, for example.
-The advantage is, that the code that is developed there, can be compiled and run on the target CPU directly, while maintaining the direct step-by-step representation of the code.
+The advantage is, that the code that is developed there, can be compiled and run on the target CPU directly, while maintaining the direct step-by-step representation of the code.\\
 
@@
 Each time the code is adapted, the entire synthesis process needs to be started over again.
-
-Because of all these reasons, debugging HDL code step-by-step as in software is not possible.\\
+Because of all of these reasons, debugging HDL code step-by-step as in software is not possible.\\
 
@@
 Using Python provides a large factor of flexibility and possibilities, since it is a comprehensive language, considering all its extension libraries.
-Additionally, Python may be more familiar to engineers in comparison to VHDL and thus, writing testbenches is made simpler with cocotb.
+Additionally, Python may be more familiar to engineers in comparison to VHDL and thus, writing testbenches is made simpler with cocotb.\\
 
 The test cases in Python are implemented as coroutines.
-By the way, the name cocotb is constructed by the words \textit{CO}routine, \textit{CO}simulation and \textit{T}est\textit{B}ench \cite{cocotbDoc}.\\
+By the way, this is where cocotb gets its name from.
+It is constructed by the words \textit{CO}routine, \textit{CO}simulation and \textit{T}est\textit{B}ench \cite{cocotbDoc}.\\
 
@@
 
-The architecture for this testbench is created under the assumptions and according to the guidelines that are developed in Chapter \ref{cha:SystemArchitectureAndConcept}.
+The architecture for this testbench is created under the assumptions and according to the guidelines that are developed in Chapter \ref{cha:SystemArchitectureAndConcept}.\\
 
@@
 In that way, the interface signals are abstracted to simple read- and write-function calls in the Python code, while the actual signal handling is done in the mentioned cocotb library classes.
-
-A clock generator class is also available from cocotb, which is used to provide the system clock to the DUT.
+A clock generator class is also available from cocotb, which is used to provide the system clock to the DUT.\\
 
@@
 This could be implemented differently, since Python can be used to actually implement an FM receiver including the entire DSP functionality, just like in Matlab.
-Because the Matlab model already exists, this is not done here - loading the Matlab data is sufficient as it is been seen as 'the single source of truth' in this project.
+Because the Matlab model already exists, this is not done here - loading the Matlab data is sufficient as it is been seen as 'the single source of truth' in this project.\\
 
@@
 However, their combination and interaction may still be subject to error - this is where integration tests unfold their potential.
-Implementation errors in the glue logic between the units, as well as the interaction of the respective interfaces can be found in these integration tests.
+Implementation errors in the glue logic between the units, as well as the interaction of the respective interfaces can be found in these integration tests.\\
 
@@
 However, in case the DUT fails to produce a correct output in the integration test, it is not always visible where exactly the error is located.
-Thus, in addition to the DUTs' interfaces, additional signals are logged from the inside of the DUT.
-that can be accessed through the testbench.
-white box: disadvantage: test needs to change as well, when implementation changes
---> white box, black box
+The source of error may be hidden anywhere within the nested and interconnected units.
+Thus, in addition to the DUTs' interfaces, further signals are logged from the inside of the DUT.
+These internal signals can be accessed through the Python testbench.\\
+
+In this context, there are further keywords that need to be mentioned - black box testing and white box testing.
+These are testing strategies, and describe how the DUT is treated.
+In the case of black box, the DUTs internals, such as implementation details, are not known.
+Thus, only the exposed interfaces can be tested.
+In the white box as a contrary, every detail of the internal structure is known.
+Thus, even single signals can be accessed, which is what is done in this Python testbench.
+The advantage of the white box is, that these internals can be tested.
+However, any time the implementation changes, even if it is just the naming of a signal, the testbench code needs to be adapted as well.
+In black box testing, the testbench code is independent of the detailed implementation of the DUT, as long as it maintains its required behaviour at the exposed interfaces.
+However, the disadvantage is that internal states cannot be observed.\\
+
+The Python testbench that is implemented combines the advantages of both worlds and uses black box testing, as well as white box testing.
 
@@
 \subsection{Unit Tests}
-FIR the only unit that is worth the effort (others are only delays, and additions/substractions respectively)
 
-A unit test only targets a single unit of the system, as the name already describes.
-This is in the contrary to an integration test, which tests the combination of several units into an integrated system.
-The testbench architecture that is developed in \ref{sub:VHDL:TestbenchArchitecture} represents such an integration test.
-However, if an integration test fails, it is not always
+A unit test only targets a single unit, or module, of the system, as the name already describes.
+This is in the contrary to an integration test, which tests the combination of several units into an integrated system, as explained above.\\
+
+In this project, the only module that has its dedicated unit test is the FIR filter.
+The reason therefor is, that it is the most complex unit within the entire design and is used in multiple instances.
+All the other design units mostly consist of simple delay lines, additions or substractions and thus a special unit test is not implemented.
 

### 2021-07-20 doc/thesis/chapters/deployment-on-hardware.tex
deployment on hardware

--- 2021-07-09 doc/thesis/chapters/deployment-on-hardware.tex
@@
 \label{cha:DeploymentOnHardware}
-  \section{Hardware Platform}
-    \subsection{RTL2832u Dongle}
-    \subsection{ZedBoard}
+
+This chapter describes the additional parts of implementation that are necessary in order to run the implemented FM radio on actual hardware.
+The hardware that is used is introduced as well.
+
+%%%%%%%%%%%%%%%%%%%%
+\section{GNU Radio}
+
+The GNU Radio implementations, both receiver and transmitter, are used in combination with external hardware.
+The specific hardware that is used is briefly described here.\\
+
+In order to interface with the real world, two different hardware devices are used.
+Both devices are supported by GNU Radio and are briefly described here.
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Ettus Research USRP b200mini}
+
+The USRP b200mini is a device that can be used to create a software-defined radio.
+It resembles the size of a business card and is developed by Ettus Research.
+A signal range from 70 MHz to 6 GHz can be covered by the front end.
+The board further features a Xilinx Spartan-6 FPGA that is user-programmable, which makes the device very flexible for various applications with the received signal \cite{USRPb200Mini}.\\
+
+GNU Radio supports this device with the blocks \textit{UHD: USRP Sink} and \textit{UHD: USRP Source}.
+Therefore, it can be used as both a receiver and a transmitter.
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{RTL-SDR}
+
+The RTL-SDR is a cheap USB dongle, that was originally developed as a DVB-T TV tuner.
+It is based on the RTL2832U chipset, which includes an 8-bit ADC and a digital signal processor.
+The device delivers IQ data via the USB interface \cite{RTLSDR}.\\
+
+\noindent
+Based on its hardware, the RTL-SDR can only be used as a receiver.
+
+%%%%%%%%%%%%%%%%%%%%
+\section{FPGA Hardware Platform}
+
+The ZedBoard is a development board that is developed by Xilinx.
+It is specifically designed around the Xilinx Zynq-7000 SoC, which contains a dual-core ARM Cortex A9 processor core and a programmable logic, the FPGA.
+The board further offers a range of features from external DDR3 memory, over various connectors, such as RJ45 Ethernet, USB, 3.5 mm audio, HDMI, Pmod, buttons to LEDs and a display.\\
+
+The most important features for this project are the audio line-out connector and the SD card port.
+Additionally, the LEDs are used to display various status information during runtime of the FM receiver.
+
+\includepicture [0.75] [0] [H] {The ZedBoard PCB \cite{ZedBoard}.} {zedboard_photo} {img/zedboard_photo}
+
+%%%%%%%%%%%%%%%%%%%%
+\section{System-On-Chip Design}
+
+Chapter \ref{cha:Implementation} already described the development of the main IP core, which is the FM radio receiver in the VHDL and HLS variants.
+In order to embed this IP on actual FPGA, or rather SoC hardware, a significant additional amount of implementation is neccessary.
+In this section, these additional parts of the implementation, which specifically enable and support the deployment on hardware, are described.
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Architecture}
+
+The block diagram in Fig.\ref{fig:bd_impl_vivado} shows the implemented architecture that is used in the SoC.
+The various design domains in and around the SoC are depicted in different visual representations.
+The FPGA domain blocks are highlighted in gray.
+The processor system domain is shown as white blocks.
+Blocks that are external to the SoC are drawn with dashed lines.
+
+\includepicture [1.0] {Block diagram of the Vivado implementation.} {bd_impl_vivado} {img/draw.io/bd_impl_vivado}
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Functionality}
+%TODO: re-work this section for better wording
+%TODO: refer to above theory chapters, with verification and HIL, etc..
+The FM receiver IP is the main functionality of the design.
+However, the design around it needs to function properly in order for the IP to work correctly, which is explained in the following passage.\\
+
+The CPU is the main actor that determines any actions on the system.
+It reads a file from the external SD card and stores this data in a dedicated memory area.
+Next, the DMA is configured and a continuous transfer mode is initiated, so that the data in the mentioned memory area is transferred into the FPGA in a loop.
+Any data connection in the following IP block chain is implemented as an AXI stream bus.
+The first block after the DMA is an AXI Stream Switch.
+It can switch its single input to be forwarded to either of the outputs.
+The switch is configured by the CPU, to enable one output at a time.
+Consequently, one of the FM radio receiver IPs is provided with data, which is processed to produce an audio output.
+This audio output data is delivered to another AXI Stream Switch, that fulfills the function to either select the output of above or below IP, in coordination with the first switch, obviously.
+The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input.
+One of these duplicates is sourced to an AXI-Stream-To-I2S converter, which generates the I2S signals for the external audio codec chip.
+The second duplicate is connected to a FIFO, which simply stores the generated output data.
+The FIFO asserts an interrupt to the CPU, on which it reads out the entire FIFO buffer.
+The CPU then stores this data on the external SD card again.
+In that way, the loop of data is closed, which enables the verification of the respective IPs correct functionality.
+Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.\\
+
+It is to note, that Xilinx provides all the AXI infrastructure IPs, and thus the only self-implemented IPs are the FM radio receivers, as well as the I2S interface.
+Nevertheless, the correct design and implementation of the architecture still needs to be formulated by the user.
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Software}
+
+drivers for IPs, libraries to interface with e.g. SD card, etc\\
+--> look at software which parts may be interesting to highlight

### 2021-07-20 doc/thesis/chapters/implementation.tex
deployment on hardware

@@
 Further analysis is done in the frequency domain, especially in order to better understand the modulation of the different spectral parts in the FM channel.
-An example thereof is shown in Fig.\ref{fig:matlab_analysis_freq_domain}.\\
-
-The spectrum of an FM channel, as it is generated by a transmitter before the modulation, is shown in the upper diagram.
+An example thereof is shown in Fig.\ref{fig:matlab_analysis_freq_domain}.
+The spectrum of an FM channel multiplex signal, as it is generated by a transmitter before the modulation, is shown in this diagram.
 The spectrum after the demodulation through a receiver is shown in the lower diagram.
@@
 GNU Radio supports interfaces to external RF hardware \cite{SoftwareGnuRadio}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{External Hardware}
-
-In order to interface with the real world, two different hardware devices are used.
-Both devices are supported by GNU Radio and are briefly described here.
-
-\subsubsection{USRP b200mini}
-
-The USRP b200mini is a device that can be used to create a software-defined radio.
-It resembles the size of a business card and is developed by Ettus Research.
-A signal range from 70 MHz to 6 GHz can be covered by the front end.
-The board further features a Xilinx Spartan-6 FPGA that is user-programmable, which makes the device very flexible for various applications with the received signal \cite{USRPb200Mini}.\\
-
-GNU Radio supports this device with the blocks \textit{UHD: USRP Sink} and \textit{UHD: USRP Source}.
-Therefore, it can be used as both a receiver and a transmitter.
-
-\subsubsection{RTL-SDR}
-
-The RTL-SDR is a cheap USB dongle, that was originally developed as a DVB-T TV tuner.
-It is based on the RTL2832U chipset, which includes an 8-bit ADC and a digital signal processor.
-The device delivers IQ data via the USB interface \cite{RTLSDR}.\\
-
-\noindent
-Based on its hardware, the RTL-SDR can only be used as a receiver.
 

### 2021-07-21 doc/thesis/chapters/deployment-on-hardware.tex
update

@@
 This chapter describes the additional parts of implementation that are necessary in order to run the implemented FM radio on actual hardware.
-The hardware that is used is introduced as well.
 
@@
 The ZedBoard is a development board that is developed by Xilinx.
-It is specifically designed around the Xilinx Zynq-7000 SoC, which contains a dual-core ARM Cortex A9 processor core and a programmable logic, the FPGA.
-The board further offers a range of features from external DDR3 memory, over various connectors, such as RJ45 Ethernet, USB, 3.5 mm audio, HDMI, Pmod, buttons to LEDs and a display.\\
+It is specifically designed around the Xilinx Zynq-7000 SoC, which contains a dual-core ARM Cortex A9 processor and a programmable logic fabric, the FPGA.
+The board further offers a range of features from external DDR3 memory, over various connectors, such as RJ45 Ethernet, USB, 3.5 mm audio, HDMI, Pmod, to buttons, LEDs and a display.\\
 
@@
 Chapter \ref{cha:Implementation} already described the development of the main IP core, which is the FM radio receiver in the VHDL and HLS variants.
-In order to embed this IP on actual FPGA, or rather SoC hardware, a significant additional amount of implementation is neccessary.
-In this section, these additional parts of the implementation, which specifically enable and support the deployment on hardware, are described.
+In order to embed these IPs on the actual FPGA, or rather SoC hardware, a significant additional amount of implementation is neccessary.
+This section describes these additional parts of the implementation, which specifically enable and support the deployment on the SoC.
 
@@
 The various design domains in and around the SoC are depicted in different visual representations.
-The FPGA domain blocks are highlighted in gray.
+The FPGA domain blocks are highlighted in gray, with the main IPs being surrounded by emphasized lines.
 The processor system domain is shown as white blocks.
@@
 \subsection{Functionality}
-%TODO: re-work this section for better wording
-%TODO: refer to above theory chapters, with verification and HIL, etc..
+
 The FM receiver IP is the main functionality of the design.
@@
 The CPU is the main actor that determines any actions on the system.
-It reads a file from the external SD card and stores this data in a dedicated memory area.
-Next, the DMA is configured and a continuous transfer mode is initiated, so that the data in the mentioned memory area is transferred into the FPGA in a loop.
-Any data connection in the following IP block chain is implemented as an AXI stream bus.
+First, it reads a file from the external SD card and stores this data in a dedicated area in the memory.
+Next, the DMA is configured and its data transfer mode is initiated, so that the DMA starts to transfer data from the mentioned memory area towards the FPGA.
+The DMA is configured in a mode that creates a continuous transfer, by looping through the data - once the end is reached, the transfer starts at the beginning of the memory area.
+Any data connection in the following processing chain is implemented as an AXI stream bus.
 The first block after the DMA is an AXI Stream Switch.
@@
 The switch is configured by the CPU, to enable one output at a time.
-Consequently, one of the FM radio receiver IPs is provided with data, which is processed to produce an audio output.
-This audio output data is delivered to another AXI Stream Switch, that fulfills the function to either select the output of above or below IP, in coordination with the first switch, obviously.
-The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input.
+Consequently, one of the FM radio receiver IPs is provided with data.
+The IP processes this data to produce an audio output, which is delivered to another AXI Stream Switch.
+It fulfills the function to either select the output of above or below IP, in coordination with the first switch, obviously.
+The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input stream.
 One of these duplicates is sourced to an AXI-Stream-To-I2S converter, which generates the I2S signals for the external audio codec chip.
 The second duplicate is connected to a FIFO, which simply stores the generated output data.
-The FIFO asserts an interrupt to the CPU, on which it reads out the entire FIFO buffer.
+Once the FIFO is almost full, it asserts an interrupt to the CPU, which triggers it to read out the entire FIFO buffer.
 The CPU then stores this data on the external SD card again.
-In that way, the loop of data is closed, which enables the verification of the respective IPs correct functionality.
-Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.\\
+In that way, the loop of data is closed, which enables the verification of the respective IPs' correct functionality.
+Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.
+This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section \ref{sec:SimilarProcess}.
+Specifically, this implemented system design enables the application of the steps 'Assess Performance' and 'Re-evaluate', since the IPs' performance can be measured and their design can be re-evaluated.\\
 

### 2021-07-21 doc/thesis/chapters/implementation.tex
update

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TODO: Other Stuff...}
+%\subsection{TODO: Other Stuff...}
 
@@
 
-A testbench is equally important in the HLS implementation, as it is for the VHDL implementation, explained in \ref{sub:VHDL:Testbench:General}.
-The testbench simulation is
+A testbench is equally important in the HLS implementation, as it is for the VHDL implementation.
+Just like the IPs' HDL code, the testbench is also written in the language \cplusplus.
+The execution of the testbench is done in the Xilinx HLS software.
+
+A major difference between a VHDL testbench and a HLS testbench is its execution speed.
+HLS faster because it does not simulate any timing info; its just plain c++ code thats run
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Design for Reuse}
+
+testbench is using Firmware code! --> enables reuse and early testing on the code, before a hardware prototype exists
 

### 2021-07-21 doc/thesis/chapters/hls.tex
add list of abbrev
add a new chapter
hls

@@
-%%%%%%%%%%%%%%%%%%%%
-\chapter[HLS]{High-Level Synthesis}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{High-Level Synthesis}
 \label{cha:HLS}
 
-\section{Introduction / State of the Art}
+The main sources of information for this chapter are found in \cite{VivadoUgHLS}, TODO.
 
-\section{Functionality (transform high-level code to HDL)}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{State Of The Art}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Functionality}
+transform high-level code to HDL
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Language Support}
 
-  \subsection{C++}
+%%%%%%%%%%%%%%%%%%%%
+\subsection{\cplusplus}
 
+%%%%%%%%%%%%%%%%%%%%
 \subsection{SystemC}
@@
 
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Compiler Directives (\#pragma's)}
 
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Data types}
 
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Functions}
 
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Loops}
 
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Conditional statements}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Testbench}
+
+An HLS testbench can be executed in different levels of abstraction, which is a major difference between a VHDL testbench and an HLS testbench.
+The most important impact thereof is the execution speed - the time it takes to simulate a number of test cases on the DUT.
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{C Simulation}
+
+In C Simulation, the HLS compiler is used to create an executable binary that includes the testbench code, as well as the IP design of the DUT.
+This binary can be run just like any other program.
+The most important note here is, that the HLS code is \textit{not} transformed into RTL at this point.
+Instead, the HLS code is treated as a regular \cplusplus\ code, that uses some libraries, like the fixed-point library.
+This means, that the HLS IP cores' top-level function is used as a regular function call in the testbench.\\
+
+The main advantage of this type of simulation is the major increase of execution speed.
+The reason for this speed-up is the abstraction of timing, such as clock cycles or similar constraints.
+In fact, there is absolutely no timing information that needs to be taken into account in this simulation.
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{C/RTL Co-Simulation}
+
+This type of simulation actually transforms the HLS IP code into RTL code, such as Verilog or VHDL, for the testbench.
+This is done in three steps.\\
+
+\begin{itemize}
+  \item \textbf{Create input vectors}\\
+  In this first step, the above explained, regular C simulation is executed.
+  Hereby, the inputs that are sent to the DUT are stored as 'input vectors'.
+
+  \item \textbf{Perform RTL simulation}\\
+  Next, the previously generated input vectors are applied to the DUT.
+  At this point, the DUT is already translated into RTL code, so the code that is used here is either VHDL or Verilog code.
+  The default simulator is the Vivado Simulator (XSim).
+  However, various other simulators like ModelSim are supported.
+  The generated outputs of this step are stored as 'output vectors'.
+  \item \textbf{Verify the results}\\
+  The final verification step sources the output vectors back to the C testbench, which compares the outputs with the expected results.\\
+\end{itemize}
+
+\noindent
+This type of simulation takes a significantly longer time to execute, because timing information now actually needs to be considered for the RTL code.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Advantages / Disadvantages}

### 2021-07-21 doc/thesis/chapters/implementation.tex
hls

@@
 A testbench is equally important in the HLS implementation, as it is for the VHDL implementation.
-Just like the IPs' HDL code, the testbench is also written in the language \cplusplus.
-The execution of the testbench is done in the Xilinx HLS software.
-
-A major difference between a VHDL testbench and a HLS testbench is its execution speed.
-HLS faster because it does not simulate any timing info; its just plain c++ code thats run
+The HDL code for the IP is written in \cplusplus and thus the testbench is also written in this language.
+The execution of the testbench is done in the Xilinx HLS software.\\
 

### 2021-07-21 doc/thesis/chapters/hls.tex
comment dividers

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter{High-Level Synthesis}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{State Of The Art}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Functionality}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Language Support}
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{\cplusplus}
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{SystemC}
 
+In the latest Xilinx HLS tools 'Vitis HLS' and 'Vivado HLS', the language SystemC is already deprecated \cite{VivadoHlsSystemC}.\\
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Coding}
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Compiler Directives (\#pragma's)}
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Data types}
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Functions}
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Loops}
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Conditional statements}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Testbench}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{C Simulation}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{C/RTL Co-Simulation}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Advantages / Disadvantages}
+
 easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV.

### 2021-07-21 doc/thesis/chapters/hls.tex
hls intro

@@
 
-The main sources of information for this chapter are found in \cite{VivadoUgHLS}, TODO.
+This chapter talks about High-Level Synthesis in general, but also specifically describes the details of the Xilinx HLS.
+The main sources of information for this chapter are found in \cite{VivadoUgHLSIntro}and \cite{VivadoUgHLS}.
 
@@
 \section{Introduction}
+
+In software development, engineers are used to writing code in a high-level language, such as C or \cplusplus.
+With these languages, specific hardware details are abstracted and thus, very little knowledge of the target hardware is required.
+The compiler takes over these details as it transforms the code by using the underlying, supported instruction set of the target hardware.
+This has been the standard for many years by now.
+In the earlier times however, it was common practise to develop programs in the low-level assembly language.
+The supported instruction set of the target CPU had to be implemented directly by the software engineer.\\
+
+A direct comparison can be drawn to FPGA design.
+An FPGA engineer historically develops code in the register-transfer level (RTL), which is common practise until today.
+This can be compared to developing software in the assembly language.
+However, the recent development towards automation and code generation brought up High-Level Synthesis (HLS).
+There, the FPGA design can be described in a high-level language, such as C or \cplusplus, which is convenient and allows to focus on the algorithm, rather than hardware details.
+This high-level code is then transformed into the low-level RTL automatically, by the HLS tool.\\
+
+The usage of HLS tries to bring software- and hardware development closer together and tries to make these two worlds more similar.
+In software development, there are different compilers for multiple processor architectures, which can be used to port a program to another target hardware.
+The idea of HLS is to create the same level of abstraction - write code in a high-level language, which is agnostic of its target hardware architecture.
 
@@
 easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV.
-However, it is less optimized, and a developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.
+However, it may be less optimized.
+A developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.

### 2021-07-21 doc/thesis/chapters/hls.tex
structuring of HLS..

@@
 
-This chapter talks about High-Level Synthesis in general, but also specifically describes the details of the Xilinx HLS.
+This chapter covers High-Level Synthesis in general, but also describes details that specifically apply to Xilinx.
 The main sources of information for this chapter are found in \cite{VivadoUgHLSIntro}and \cite{VivadoUgHLS}.
@@
 In software development, there are different compilers for multiple processor architectures, which can be used to port a program to another target hardware.
-The idea of HLS is to create the same level of abstraction - write code in a high-level language, which is agnostic of its target hardware architecture.
+The idea of HLS is to create the same level of abstraction - write code in a high-level language, which is agnostic of its target hardware architecture.\\
+
+A major achievement in the usage of HLS is the shorter development time.
+The time that it takes to develop a first working version of the intended design can be significantly reduced, compared to the development in RTL code directly.
+This is especially important when an FPGA design is combined with a software part, like in a System On Chip, since software development is comparably fast.
+Fig.\ref{fig:timeline_performance_first_prototype} schematically displays the difference in development time, over performance and degree of optimization, when the same functionality of a prototype is implemented in different methods for various targets.
+The compared targets are an x86 processor CPU, a digital signal processor (DSP), a graphics processing unit (GPU) and the FPGA design in RTL or HLS implementation.
+It is clearly visible, that the FPGA design with the HLS design approach achieves a first working prototype much earlier than the direct RTL approach.
+
+\includepicture [0.8] [0] {Design Time vs.
+Application Performance, if the same functionality is implemented in different methods for various targets \cite[Figure 1-1]{VivadoUgHLSIntro}.} {timeline_performance_first_prototype} {img/timeline_performance_first_prototype}
 
@@
 
+- usage in percentage (market share), compared to direct RTL coding\\
+
+-  Language Support\\
+\cplusplus and SystemC\\
+In the latest Xilinx HLS tools 'Vitis HLS' and 'Vivado HLS', the language SystemC is already deprecated \cite{VivadoHlsSystemC}.\\
+
+- efficiency (see \cite[II.C.: State of the Art]{CompareHlsVHDLArticle})\\
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Functionality}
-transform high-level code to HDL
+
+transform high-level code to HDL.\\
+-analysis\\
+-scheduling\\
+-pipelining\\
+
+- Limitations\\
+dynamic memory allocation, etc.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Language Support}
+\section{Workflow}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\cplusplus}
+see Chapter "Software Verification and Vivado HLS" of \cite{VivadoUgHLSIntro}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{SystemC}
+- recommended to use software verification tools, like valgrind or Coverity (dynamic code checker)\\
+- use code coverage with e.g. gcov for testbench
 
-In the latest Xilinx HLS tools 'Vitis HLS' and 'Vivado HLS', the language SystemC is already deprecated \cite{VivadoHlsSystemC}.\\
+general workflow:
+- develop c/c++\\
+- verify it using above mentioned software tools \\
+- HLS testbench it (c-sim)\\
+- transform to RTL\\
+- HLS testbench it (co-sim)\\
+- integrate the generated RTL to a larger FPGA design\\
+- synthesize FPGA bitstream\\
+- use on FPGA\\
 
@@
 This type of simulation takes a significantly longer time to execute, because timing information now actually needs to be considered for the RTL code.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Advantages / Disadvantages}
-
-easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV.
-However, it may be less optimized.
-A developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.

### 2021-07-21 doc/thesis/chapters/results.tex
add list of abbrev
structuring
emphasis
impl details
adaptions; white/black box; move figures
structuring of HLS..

@@
 %%%%%%%%%%%%%%%%%%%%
-\chapter{Test Results}
-\label{cha:Results}
+\chapter{Comparison and Results}
+\label{cha:ComparisonAndResults}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Demodulated Signal}
+% show plots of the common analyzer python script
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Functionality}
 % Functionality same?
+% DSP yes, AXI no (FSM of axi-stream)
+% unit tests
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Code Development}
+% lines of code
+% Implementation effort/time
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Simulation}
 % Sim speed of x min of recorded FM signal
-  % lines of code
-  \subsection{Implementation effort/time}
-  \subsection{Hardware Utilization}
-  \subsection{Others}
 
-  % TODO Überlegungen in diesem Kapitel
-  % kann Demodulator geändert werden? Bsp. von FM zu DAB+, QAM, etc..
-  % - von matlab heraus VHDL generieren, etc
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Useability}
+% IP generation and integration into Vivado
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Applicability}
+% of implementation variant (GNU Radio for fast prototyping, Matlab for research, HLS/VHDL deployment on HW, etc.)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Advantages / Disadvantages}
+
+easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV.\\
+However, it may be less optimized.\\
+A developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Hardware Utilization}
+% especially VHDL<->HLS
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Latency}
+% especially VHDL<->HLS
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \section{Lines of Code}
+  % of the entire project
+
+  % NOTES:
+  % - The system level architecture that was chosen, can be applied to any DSP project. Simply replace the main IP, e.g. from FM to DAB+, QAM, etc. and the environment can run in the exact same way.
+  %
+  % Further steps:
+  % - generate VHDL from Matlab receiver code
+  % - use the common input file and feed it to GNU Radio; also produce an output file with GNU Radio for comparison
+  %- Develop HLS compiler independent code (optional)}
+  %   -- Different compilers use different #pragmas, etc.
+  %   -- These #pragmas need to be within the code, at the appropriate positions, which makes it difficult to write compiler independent code....

### 2021-07-21 doc/thesis/chapters/hls.tex
skip the Functionality section
state of the art HLS

@@
 
-- usage in percentage (market share), compared to direct RTL coding\\
+Information for this section is found in \cite{EvolutionOfHLS}, \cite{XilinxVivisHLSOpenSource} and \cite{CompareHlsVHDLArticle}.\\
 
--  Language Support\\
-\cplusplus and SystemC\\
-In the latest Xilinx HLS tools 'Vitis HLS' and 'Vivado HLS', the language SystemC is already deprecated \cite{VivadoHlsSystemC}.\\
+HLS has been developed over 20 years ago, but has not been able to replace the manual, traditional implementation of RTL, using VHDL and Verilog, so far.
+Unfortunately, no published numbers about the market share of HLS in relation to traditional implementations could be found during the research in this thesis.
+However, according to the given sources it is becoming more and more popular in recent years.
+This is mainly based on the fact that the HLS tools keep producing results of better quality as they are developed further.
+This especially regards the efficiency of the produced RTL code, the useability of the tools, as well as the better understandable correlation between the high-level code and the generated RTL code.\\
 
-- efficiency (see \cite[II.C.: State of the Art]{CompareHlsVHDLArticle})\\
+Xilinx is the largest manufacturer of FPGAs globally and therefore its tool 'Xilinx HLS' is widely used.
+Intel, the second largest FPGA manufacturer, also has its HLS tool called 'Intel oneAPI' \cite{IntelHLSOneAPI}.
+Both companies are conducting serious effort to push the HLS implementation method.
+Xilinx recently open-sourced the code for the front-end of their HLS compiler.
+The front-end processes the HLS code, specifically \cplusplus, and creates an intermediate representation from it.
+The company expects to expand the community engagement and the encouragement of innovation, driven by users, with this step.
+Also Intel with its oneAPI toolkit is trying to introduce a wider usage of HLS to the community.\\
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Functionality}
+In terms of language support, several tools settle for \cplusplus.
+Intel oneAPI uses library APIs and DP\cplusplus\ (Data-Parallel \cplusplus), which is an extension to the ISO \cplusplus\ standard.
+Xilinx also decided in favor of the \cplusplus\ language for their tools 'Vivado HLS' and 'Vitis HLS'.
+In older versions, SystemC was supported for synthesis as well, but is now deprecated in the respective latest versions \cite{VivadoHlsDeprecatesSystemC}.\\
 
-transform high-level code to HDL.\\
--analysis\\
--scheduling\\
--pipelining\\
+Summing up, HLS is gaining more and more attention as it becomes more efficient.
+Its usage is being promoted and pushed forward with serious effort by the major FPGA chip manufacturers and tool vendors.
 
-- Limitations\\
-dynamic memory allocation, etc.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Functionality}
+%
+%transform high-level code to HDL.\\
+%-analysis\\
+%-scheduling\\
+%-pipelining\\
+%
+%- Limitations\\
+%dynamic memory allocation, etc.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Compiler Directives (\#pragma's)}
+\subsection{Compiler Directives}
+
+e.g. \#pragma's
 
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Data types}
+\subsection{Data Types}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Conditional statements}
+\subsection{Conditional Statements}
 

### 2021-07-21 doc/thesis/chapters/system-architecture-and-concept.tex
state of the art HLS

@@
       This can be done using high-level synthesis.
-      High-level synthesis (HLS) is explained in detail in Chapter \ref{cha:HLS}.
+      This thesis focuses in \cplusplus and does not implement a SystemC variant, because Xilinx has deprecated SystemC support for synthesis in their latest HLS tools.
+      This, and the details of High-level synthesis (HLS) are explained in Chapter \ref{cha:HLS}.
+
   \item \textbf{FPGA: VHDL}\\

### 2021-07-21 doc/thesis/chapters/hls.tex
workflow wip

@@
 
-see Chapter "Software Verification and Vivado HLS" of \cite{VivadoUgHLSIntro}
+This section describes the workflow for Xilinx HLS, as it is recommended in the user guide \textit{Introduction to FPGA Design with Vivado High-Level Synthesis} \cite{VivadoUgHLSIntro}.\\
 
-- recommended to use software verification tools, like valgrind or Coverity (dynamic code checker)\\
-- use code coverage with e.g. gcov for testbench
+Generally, as the user guide states, the quality and the correctness of the generated output can only be as good as its input.
+Consequently, it is very important to follow the design recommendations and adhere to specific rules in order to produce a high-quality input software.\\
 
-general workflow:
-- develop c/c++\\
-- verify it using above mentioned software tools \\
-- HLS testbench it (c-sim)\\
-- transform to RTL\\
-- HLS testbench it (co-sim)\\
-- integrate the generated RTL to a larger FPGA design\\
-- synthesize FPGA bitstream\\
-- use on FPGA\\
+The following list presents the main steps that are required in the implementation workflow of an HLS IP core project.\\
+
+\begin{itemize}
+  \item \textbf{Software Testbench}\\
+      The developed HLS IP code needs to be verified and checked for correct functionality in a software testbench, before it gets processed to RTL code.
+      This can be done by using using conventional software development tools.
+      Xilinx specifically recommends dynamic code checker tools like \textit{valgrind} or \textit{Coverity}.
+      They support a developer in finding memory leaks, out-of-bounds memory access, uninitialized variables and similar issues.
+      Furthermore, the application of code coverage tools like \textit{gcov} is recommended as well.
+      In that way, the percentage of executed code lines of testbench and HLS code can be measured, to ensure a sufficient level of testing.
+      The software testbench is also called 'C Simulation'.\\
+
+      The implementation of the HLS IP code is subject to restrictions.
+      As an example, dynamic memory allocation is not supported, since the entire code must be analyzable at runtime to be processed into RTL code.
+      %For further details, please read the user guides regarding Xilinx HLS.
+      In contrast, the testbench code does not have any restrictions.
+      Anything that is supported in C/\cplusplus\ may be used here.\\
+
+  \item \textbf{Co-Simulation}\\
+      Co-Simulation serves the purpose verify the correct functionality of the generated RTL code.
+      However, the main goal is not to verify the algorithm here.
+      This is already done by above described C simulation.
+
+  \item \textbf{integrate the generated RTL to a larger FPGA design}\\
+  \item \textbf{synthesize FPGA bitstream}\\
+  \item \textbf{use on FPGA}\\
+\end{itemize}
 

### 2021-07-21 doc/thesis/chapters/hls.tex
workflow done

@@
 \section{Workflow}
+\label{sec:hls:workflow}
 
@@
   \item \textbf{Co-Simulation}\\
-      Co-Simulation serves the purpose verify the correct functionality of the generated RTL code.
-      However, the main goal is not to verify the algorithm here.
-      This is already done by above described C simulation.
+      Co-Simulation serves the purpose to verify the correct functionality of the generated RTL code.
+      It is not the goal to verify the algorithm, though.
+      This is already done in the previous C simulation step.
+      However, the issue is that the C simulation is executed like a regular program, which means it is executed sequentially, instruction by instruction and thus parallelism cannot be represented.\\
 
-  \item \textbf{integrate the generated RTL to a larger FPGA design}\\
-  \item \textbf{synthesize FPGA bitstream}\\
-  \item \textbf{use on FPGA}\\
+      This is where the Co-Simulation becomes neccessary.
+      The RTL code is simulated with an HDL simulator software, which can actually simulate the parallelism as it exists in the FPGA hardware.
+      In that way it can be ensured, that the user did not break the functional correctness of the algorithm, by giving wrong parallelization guidance to the HLS tool.\\
+
+      The external simulator software has to run the complex HDL simulation and return data results, which is a time consuming task.
+      Therefore, the execution time of this simulation type is expected to be approximately 10.000 times slower, compared to the C simulation.\\
+
+  \item \textbf{Integration of the generated IP}\\
+      At this point, the developed HLS code is fully verified and can successfully be transformed into working RTL code.
+      Xilinx HLS exports the generated RTL code in an IP structure that is supported by its IP integrator tool, that is used in Vivado or Vitis.
+      Therefore, the developed IP core can be imported to the block design as another processing block, get connected with other IPs, and in that way get integrated into a larger design.\\
+
+  \item \textbf{Synthesize FPGA bitstream}\\
+      Like in any other FPGA development cycle, the created block design can now be synthesized.
+      This generates a bitstream file, which can be programmed into an actual FPGA hardware.
+      Depending on the architecture of the system, the developed IP may now interface with the on-chip CPU, in the example of a System-On-Chip.\\
 \end{itemize}
+
+These steps represent the standard workflow for the development of an integrated HLS IP.
+A successful deployment on hardware can be achieved if they are completed accurately.
+There may be more and deeper knowledge necessary, in order to use this design approach.
+This thesis however focuses on system design, not on HLS specifically, and thus a comprehensive study of the referenced literature is recommended.
 
@@
 An HLS testbench can be executed in different levels of abstraction, which is a major difference between a VHDL testbench and an HLS testbench.
-The most important impact thereof is the execution speed - the time it takes to simulate a number of test cases on the DUT.
+The most important impact thereof is the execution speed - the time it takes to simulate a number of test cases on the DUT.\\
+
+Section \ref{sec:hls:workflow} above already mentioned the topic of testbenches.
+The following descriptions are a comprehensive extension around this topic and the simulation types.
 

### 2021-07-21 doc/thesis/chapters/hls.tex
comments
move the entire section vhdl:testbench:general to hls:testbench:general

@@
 
+This section gives a brief overview of the \cplusplus\ HLS code style that is supported with Xilinx tools.
+More detailed information can be found in the respective user guide \cite{VivadoUgHLS}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Data Types}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Interfaces}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Data Types}
+\subsection{Optimization}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Functions}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Loops}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Conditional Statements}
+%\subsection{Functions}
+%\subsection{Loops}
+%\subsection{Conditional Statements}
 
@@
 
-An HLS testbench can be executed in different levels of abstraction, which is a major difference between a VHDL testbench and an HLS testbench.
+The general significance of the simulation testbench for a HDL design is explained here, as well as the different levels of abstraction in which a HLS testbench can be executed.
+These different levels of abstraction in the testbench execution are a major difference between a VHDL testbench and an HLS testbench.
 The most important impact thereof is the execution speed - the time it takes to simulate a number of test cases on the DUT.\\
 
-Section \ref{sec:hls:workflow} above already mentioned the topic of testbenches.
-The following descriptions are a comprehensive extension around this topic and the simulation types.
+Section \ref{sec:hls:workflow} above already mentioned the topics of testbenches and simulation types.
+The following descriptions shall be seen as a comprehensive extension around these topics.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{General}
+\label{sub:hls:testbench:general}
+
+In the development of software that runs on a CPU, the tool of step-by-step debugging is commonly used during the development process.
+Various verification and testing libraries exist, in order to create unit tests, for example.
+The advantage is, that the code that is developed there, can be compiled and run on the target CPU directly, while maintaining the direct instruction-by-instruction representation of the code.\\
+
+However, in the development of hardware with any HDL, this is not possible.
+HDL languages provide a syntax and semantic representation that allow to create parallel structures and a dependency on timing, such as a system clock.
+The design is based on signals that perform logic operations and transfer values into registers.
+Thus, HDL code is often called register-transfer level, or short, RTL code.
+Before this code can be run on the target hardware, i.e. an FPGA, it needs to run through a process called synthesis.
+In synthesis, the HDL code is translated into a netlist and is consequently mapped into hardware logic cells that are available in the target device.
+Various optimizations are done during this process, so that the direct connection to the code representation is lost.
+Additionally, it is impossible to access every single signal within the design during debugging on the hardware, which is probably also one of the main issues.
+Synthesis tools like Xilinx Vivado provide support to make debugging possible in hardware, but require a large effort to implement them, especially in terms of synthesis time.
+Another big factor that speaks against debugging in hardware generally, is that the process of synthesis takes a comparably large amount of time to transform code into the target technology.
+Each time the code is adapted, the entire synthesis process needs to be started over again.
+Because of all of these reasons, debugging HDL code step-by-step as in software is not possible.\\
+
+Thus, in order to allow debugging and verification of HDL code, the technology of simulation was invented.
+Here, the HDL is also translated into a netlist that represents all the parallel processes just like in hardware, but the subsequent step of synthesis is not taken.
+Therefore, simulation comes to a result much faster.
+Since HDL designs actually only consist of electrical wires and some basic block primitives, like flipflops or lookup tables, it is difficult to verify correct values.
+They would need to be compared in their binary bit-pattern format, at the correct point in time.
+Of course the HDL language provides support with different datatypes, that abstract the binary number representation and make the numbers human-readable.
+However, it remains an enormous effort to manually write testing code to verify a bus interface, for example.
+In order to simplify such complex operations, various testbench frameworks and libraries are available to developers.
 

### 2021-07-21 doc/thesis/chapters/implementation.tex
move the entire section vhdl:testbench:general to hls:testbench:general

@@
 
-The general significance of a simulation testbench in hardware description language (HDL) design is explained here, as well as an introduction to the used testbench framework that is used in this project.
-
-%%%%%%%%%%%%%
-\subsubsection{General}
-\label{sub:VHDL:Testbench:General}
-
-In the development of software that runs on a CPU, the tool of step-by-step debugging is commonly used during the development process.
-Various verification and testing libraries exist, in order to create unit tests, for example.
-The advantage is, that the code that is developed there, can be compiled and run on the target CPU directly, while maintaining the direct step-by-step representation of the code.\\
-
-However, in the development of hardware with any HDL, this is not possible.
-HDL languages provide a syntax and semantic representation that allow to create parallel structures and a dependency on timing, such as a system clock.
-The design is based on signals that perform logic operations and transfer values into registers.
-Thus, HDL code is often called register-transfer level, or short, RTL code.
-Before this code can be run on the target hardware, i.e. an FPGA, it needs to run through a process called synthesis.
-In synthesis, the HDL code is translated into a netlist and is consequently mapped into hardware logic cells that are available in the target device.
-Various optimizations are done during this process, so that the direct connection to the code representation is lost.
-Additionally, it is impossible to access every single signal within the design during debugging on the hardware, which is probably also one of the main issues.
-Synthesis tools like Xilinx Vivado provide support to make debugging possible in hardware, but require a large effort to implement them, especially in terms of synthesis time.
-Another big factor that speaks against debugging in hardware generally, is that the process of synthesis takes a comparably large amount of time to transform code into the target technology.
-Each time the code is adapted, the entire synthesis process needs to be started over again.
-Because of all of these reasons, debugging HDL code step-by-step as in software is not possible.\\
-
-Thus, in order to allow debugging and verification of HDL code, the technology of simulation was invented.
-Here, the HDL is also translated into a netlist that represents all the parallel processes just like in hardware, but the subsequent step of synthesis is not taken.
-Therefore, simulation comes to a result much faster.
-Since HDL designs actually only consist of electrical wires and some basic block primitives, like Flipflops or lookup tables, it is difficult to verify correct values.
-They would need to be compared in their binary bit-pattern format, at the correct point in time.
-Of course the HDL language provides support with different datatypes, that abstract the binary number representation and make the numbers human-readable.
-However, it remains an enormous effort to manually write testing code to verify a bus interface, for example.
-In order to simplify such complex operations, various testbench frameworks are available to developers.
+The HDL compiler and the testbench framework that are used in this project are introduced here.
 
@@
 %%%%%%%%%%%%%
-\subsubsection{Framework cocotb}
+\subsubsection{Testbench framework cocotb}
 \label{sub:VHDL:FrameworkCocotb}

### 2021-07-22 doc/thesis/chapters/hls.tex
data types

@@
 
-This section gives a brief overview of the \cplusplus\ HLS code style that is supported with Xilinx tools.
+This section gives a brief overview of some HLS-specific \cplusplus\ features that are supported with Xilinx tools.
 More detailed information can be found in the respective user guide \cite{VivadoUgHLS}.
@@
 \subsection{Data Types}
+
+HLS supports special data types, which are not available in standard \cplusplus.
+The focus of these data types is mainly to support an optimized translation into RTL, which requires the least amount of logic cells in the FPGA.\\
+
+Standard \cplusplus\ only supports variables in the bitwidth of byte-boundaries, so in multiples of 8 bit.
+However, this may be inefficient to fit into available FPGA hardware.
+One example therefor is the DSP48 macro that Xilinx FPGAs have built-in.
+They support the multiplication of only up to 18 bit.
+In an example application, the reachable number range of a variable might be limited to 17 bit for the multiplication.
+However, in order to represent that, using a 32 bit data type would be necessary in standard \cplusplus.
+The HLS compiler does not have this prior knowledge of the number range limit and would thus implement a 32 bit multiplication, which is a waste of resources in this case.
+Because of this reason, a major feature of HLS is the possibility to define data types with arbitrary bitwidths.
+In that way, the above example can be optimized to use a data type that has a bitwidth of 17 bit, which is wide enough for the application.
+
+todo: fixed point
 

### 2021-07-22 doc/thesis/chapters/hls.tex
data types

@@
 Standard \cplusplus\ only supports variables in the bitwidth of byte-boundaries, so in multiples of 8 bit.
-However, this may be inefficient to fit into available FPGA hardware.
+However, this may be inefficient to fit into the available FPGA target hardware.
 One example therefor is the DSP48 macro that Xilinx FPGAs have built-in.
@@
 Because of this reason, a major feature of HLS is the possibility to define data types with arbitrary bitwidths.
-In that way, the above example can be optimized to use a data type that has a bitwidth of 17 bit, which is wide enough for the application.
+In that way, the above example can be optimized to use a data type that has a bitwidth of 17 bit, which is exactly the bitwidth that is required for the application.
+This results in a lower logic utilization, which in turn results in a higher maximum clock frequency that can be used.
+Consequently, more functionality an be implemented in the FPGA device.
+The referenced user guide shows a design example where an area reduction of 75\% is achieved by just reducing the bitwidths.\\
 
-todo: fixed point
+The argumentation is equally valid for integer data types, as well as fixed point data types.
+In calculations it is often necessary to represent fractional numbers.
+In standard \cplusplus, floating point data types are used therefor.
+However, floating point operations require a large amount of logic to implement and thus, HLS provides fixed point data types.
+These can be instantiated in arbitrary bitwidths, where integer and fractional bitwidths can be specified separately.
+The fixed point format is often described in the format 'integer.fractional', like '2.14', which represents a 16 bit fixed point data type that has 2 integer bits and 14 fractional bits.\\
+
+An example for the usage of integer and fixed point data types is given in the following code section.
+
+\begin{CppCode}
+  #include <ap_int.h>
+  #include <ap_fixed.h>
+
+  ap_int<9>   custom_integer_var1;         //  9 bit integer, signed
+  ap_uint<17> custom_integer_var2;         // 17 bit integer, unsigned
+
+  ap_fixed<16,2>  custom_fixedpoint_var1;  // 16 bit fixed point, 2.14 format, signed
+  ap_ufixed<14,4> custom_fixedpoint_var2;  // 12 bit fixed point, 4.10 format, unsigned
+\end{CppCode}
 

### 2021-07-22 doc/thesis/chapters/hls.tex
directives

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interfaces}
+\subsection{Optimization Directives}
+
+The HLS code can be instumented with additional optimization directives.
+This is used to provide the HLS compiler with additional information and instruct it on how to implement certain parts of the code in RTL.\\
+
+Directives can be applied to interfaces, functions, loops, arrays or code regions.
+The directives can be written in a Tcl file that is provided to the HLS project, or in the design code directly, in the form of a pre-processor pragma.
+Both variants have their advantages and disadvantages.\\
+
+Examples for such directives are optimization strategies for thoughput and dataflow optimization, such as pipelining or loop unrolling.
+Also the implementation of the required interface types, such as various AXI interfaces, is controlled with directives.\\
+
+Since this topic is very comprehensive, it cannot be covered into more detail here.
+For further information, refer to the user guide \cite{VivadoUgHLS}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Compiler Directives}
-
-e.g. \#pragma's
+\subsection{Interfaces}
 

### 2021-07-22 doc/thesis/chapters/hls.tex
interfaces

--- 2021-07-22 doc/thesis/chapters/hls.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interfaces}
+\subsection{Optimization Directives}
+
+The HLS code can be instumented with optimization directives.
+This is used to provide the HLS compiler with additional information and to instruct it on how to implement certain parts of the code in RTL.\\
+
+Directives can be applied to interfaces, functions, loops, arrays or code regions.
+They can be written in a Tcl file that is provided to the HLS project, or in the design code directly, in the form of a pre-processor pragma.
+Both variants have their advantages and disadvantages.\\
+
+Examples for such directives are optimization strategies for throughput and dataflow optimization, such as pipelining or loop unrolling.
+Also the implementation of the required interface types, such as various AXI interfaces, is controlled with directives.\\
+
+Since this topic is very comprehensive, it cannot be covered into more detail here.
+For further information, refer to the user guide \cite{VivadoUgHLS}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Compiler Directives}
-
-e.g. \#pragma's
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Optimization}
-
+%\subsection{Optimization}
 %\subsection{Functions}
@@
 %\subsection{Conditional Statements}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Interfaces}
+
+Interfaces are an important asset to any IP design, since this is where they communicate with their surroundings.
+This may be another IP that processes data, but may also be a CPU that reads status information or configures the IPs' settings.\\
+
+Xilinx HLS implements two different types of interfaces: block-level, and port-level interfaces.
+The block-level interface is a general interface to the IP, that can provide a handshaking functionality for when it is idle and ready to start a new operation, or when the operation is completed.
+The port-level interfaces on the other hand are specific to input and output ports.
+They can also provide handshaking functionality, such as data-valid indicators, acknowledge signals, or implement a FIFO for the IOs.
+However, port-level interfaces can also be implemented as AXI interfaces, such as AXI4-Stream, AXI4-Lite Slave or an AXI4 Master.
+This is an especially useful function, since the implementation of the actual interface logic is abstracted from the user.
+Additionally, software drivers are automatically generated along with the respective interfaces, which provides a fast and simple integration into a larger system.\\
+
+The following code example shows the specification of an AXI4-Lite memory-mapped slave interface.
+The provided registers are separated into config registers and status registers.
+The config is readable and writeable, while the status registers are implemented as read-only, which requires less logic in the RTL implementation.\\
+
+\begin{program}
+  \caption{Implementation of an AXI4-Lite memory-mapped slave with read-only and read-write registers.}
+  \label{prog:axi4_lite}
+  \begin{CppCode}
+    typedef struct {
+      ap_int<NUM_LEDS> led_ctrl;
+      uint8_t code;
+    } config_t;
+
+    typedef struct {
+      uint32_t uptime;
+      uint32_t result;
+    } status_t;
+
+    void hls_ip_top_level(config_t& config,
+                          status_t* status) {
+        #pragma HLS INTERFACE s_axilite port = status bundle = API
+        #pragma HLS INTERFACE s_axilite port = config bundle = API
+
+        if (config.code == 0xFF) {
+          status->result = 0;
+          ...
+        }
+        ...
+    }
+  \end{CppCode}
+\end{program}
+
+Another code example demonstrates the equally simple definition of an AXI4-Stream interface.
+
+\begin{CppCode}
+  #include <hls_stream.h>
+
+  void hls_ip_top_level(hls::stream<uint32_t>& sample_in,
+                        hls::stream<uint32_t>& audio_out) {
+      #pragma HLS INTERFACE axis port = sample_in
+      #pragma HLS INTERFACE axis port = audio_out
+  }
+\end{CppCode}
 

### 2021-07-22 doc/thesis/chapters/hls.tex
code formatting

@@
 
-The following code example shows the specification of an AXI4-Lite memory-mapped slave interface.
-The provided registers are separated into config registers and status registers.
-The config is readable and writeable, while the status registers are implemented as read-only, which requires less logic in the RTL implementation.\\
+The program code example \ref{prog:axi4_lite} shows the specification of an AXI4-Lite memory-mapped slave interface.
+The provided registers are separated into configuration registers and status registers.
+The configuration is readable and writeable, while the status registers are implemented as read-only, which requires less logic in the RTL implementation.\\
 
 \begin{program}
-  \caption{Implementation of an AXI4-Lite memory-mapped slave with read-only and read-write registers.}
+  \caption{Implementation of an AXI4-Lite memory-mapped slave with a separation of read-only status registers and read-write configuration registers.}
   \label{prog:axi4_lite}
@@
 
-Another code example demonstrates the equally simple definition of an AXI4-Stream interface.
+The second program code example \ref{prog:axi4_stream} demonstrates the equally simple definition of an AXI4-Stream interface.
 
+\begin{program}
+  \caption{Implementation of an AXI4-Stream interface.}
+  \label{prog:axi4_stream}
 \begin{CppCode}
@@
      #pragma HLS INTERFACE axis port = audio_out
+
+     ...
  }
 \end{CppCode}
+\end{program}
 

### 2021-07-22 doc/thesis/chapters/implementation.tex
axi lite register engine

@@
 
+%see duplicate below
 %\includepicture [1.02] [0] [H] {Detailed block diagram of the digital signal processing chain.} {bd_dsp_detailed} {img/draw.io/bd_dsp_detailed}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{TODO: Other Stuff...}
+\subsection{Interfaces}
+
+The main interfaces of the FM radio receiver IP are two AXI4-Stream interfaces and one AXI4-Lite memory-mapped slave interface.
+The streaming interfaces are used for the input of IQ samples and output of the processed audio samples, while the memory-mapped interface provides configuration options and status information.
+
+%%%%%
+\subsubsection{AXI4-Stream}
+
+An AXI4-Stream interface in the minimum configuration includes signals for the ready-indicator, and the data with its valid-indicator.
+The ready-indicator is used as a flag to signalize whether the interface can process data or not at this time, which allows handshaking between subsequent processing blocks.
+An IP may indicate that it is not ready, while its current operation is still running, for example.\\
+
+In the VHDL version of the FM radio receiver IP, this handshaking is not implemented according to the AXI stream specification.
+Specifically, the ready signal handshake is implemented with a workaround, to simplify the development.
+The DSP chain of the IP is not fully pipelined and thus cannot consume an input sample on every clock cycle.
+This is mainly because of the FIR filter implementation - it takes multiple clock cycles to compute an output sample, before it can take a new sample.
+Ideally, this situation should be handled by the ready signal, since this is exactly what it is intended for.
+However, in order to simplify the implementation a strobe generator is used, which is a counter that generates a repetitive pulse with a defined period.
+This pulse, in combination with some additional logic, is used as the ready signal, to guarantee enough spacing between two consecutive input samples.
+
+%%%%%
+\subsubsection{AXI4-Lite Slave}
+
+The IP provides configuration registers and status registers on this memory-mapped bus interface.
+It is connected with the CPU, which uses these registers in the application software.\\
+
+The VHDL code for this register interface is generated by a 'Register Engine' Python script.
+The registers can be defined in a YAML-file.
+Additionally, documentation in the form of a Markdown language file is generated, and - more importantly - a C/\cplusplus\ code header, which defines all the register addresses and can be used in the application software.\\
+
+Fig.\ref{fig:register_engine} gives an overview of how the Register Engine script works.
+The Python script takes the register definitions from the YAML file, and uses templates to generate the respecitive output files.\\
+
+\includepicture [1.0] [0] [H] {Overview of the Register Engine script.} {register_engine} {img/draw.io/register_engine}
 

### 2021-07-22 doc/thesis/chapters/introduction.tex
comments
move the entire section vhdl:testbench:general to hls:testbench:general
axi lite register engine

@@
 In the view of a system design approach, another focus is put on a sustainable testing and development strategy for the entire system.
-Above mentioned implementation variants should optimally share as much as possible of testing and verification code, in order to save time and effort in implementation.
+Above mentioned implementation variants should ideally share as much as possible of testing and verification code, in order to save time and effort in implementation.
 Also, as many tasks as possible should be automated by scripts, to save time in manual labor for the engineers working on the development.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%Thesis Design Decisions: \\
-%\\
-%Main focus on system design. \\
-%\\
-%How to achieve the same result in 3 (4?) different levels of abstraction?
-%
-%\begin{enumerate}
-%  \item GnuRadio (?)
-%  \item Matlab/Python
-%  \item HLS\\
-%      Implementation targeting Xilinx' ZedBoard.\\
-%      Optionally: \\
-%      Intel as a comparison (synth only, no HW) - "How platform-independent is HLS?"
-%  \item VHDL
-%\end{enumerate}
-%\vspace{.5 cm}
-%
-%Compare:
-%- Implementation effort (difficulty)
-%- Implementation time
-%- Usefulness on Target Systems (uC, SoC, RasPi, etc)
-%- others?
 
@@
 %
+% ## check all abbreviation and where they are introduced the first time
+%       - introduce only once, like "device under test (DUT)",
+%         then use the abbreviation only
 %

### 2021-07-22 doc/thesis/chapters/system-design-process.tex
update
move the entire section vhdl:testbench:general to hls:testbench:general
axi lite register engine

@@
 \subsection{The SIMILAR Process}
+\label{sec:SimilarProcess}
 %TODO: maybe include additional facts from https://slideplayer.com/slide/14327503/
@@
 In order to be able to do that, interfaces between the components and subsystems need to be defined.
-Optimally, these interfaces are designed in a way that the amount of data that needs to be exchanged is kept to a minimum.
+Ideally, these interfaces are designed in a way that the amount of data that needs to be exchanged is kept to a minimum.
 It is of advantage, if a subsystem only transfers finished products that it produced to another subsystem.
@@
 SystemC is not discussed, since it may potentially be less important in the future.
-In the latest Xilinx HLS tools 'Vitis HLS' and 'Vivado HLS', the language SystemC is already deprecated \cite{VivadoHlsSystemC}.\\
+In the latest Xilinx HLS tools 'Vitis HLS' and 'Vivado HLS', the language SystemC is already deprecated \cite{VivadoHlsDeprecatesSystemC}.\\
 

### 2021-07-22 doc/thesis/chapters/implementation.tex
tb stuff

@@
 
+%TODO: use mathematical typesetting in this diagram
 \includepicture [1.0] [0]  {Block diagram of the interface between cocotb and a simulator \cite{cocotbDoc}.} {bd_cocotb_interface_simulator} {img/draw.io/bd_cocotb_interface_simulator}
@@
 
+This section describes the HLS version of the FM radio receiver implementation.
 The implemented testbench structure, as well as some specific implementation details are explained here.
@@
 A testbench is equally important in the HLS implementation, as it is for the VHDL implementation.
-The HDL code for the IP is written in \cplusplus and thus the testbench is also written in this language.
-The execution of the testbench is done in the Xilinx HLS software.\\
+The HDL code for the IP is written in \cplusplus\ and thus the testbench is also written in this language.
+The execution of the testbench is done in the Xilinx HLS software.
+A detailed explanation of how an HLS testbench works in general is given in Chapter \ref{cha:HLS}.\\
+
+The testbench is developed in a similar manner as the VHDL testbench.
+It applies input data to the DUT and generates output data of multiple signals, which need to be observed.
+This output data is in turn saved to output files, which are analyzed in a subsequent step.\\
+
+The actual implementation to write the HLS signal data to a file, is implemented as a file-writer \cplusplus\ class within the HLS design code.
+These code parts are only evaluated during simulation, which is achieved by using pre-processor defines.
+The following code shows an example of this file-writer class.\\
+
+\begin{CppCode}
+audio_sample_t fm_receiver( ... ) {
+  ...
+#ifndef __SYNTHESIS__
+  static DataWriter writer("data_out_audio.txt");
+  writer.write(audio);
+#endif
+  ...
+  return audio;
+}
+\end{CppCode}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Design for Reuse}
+\subsection{Design For Reuse}
 
-testbench is using Firmware code! --> enables reuse and early testing on the code, before a hardware prototype exists
+The testbench is especially designed around the 'Design For Reuse' segment of the SIMILAR concept, which is introduced in Section \ref{sec:SimilarProcess}.
+
+testbench is using Firmware code to load the data! --> enables reuse and early testing on the code, before a hardware prototype exists\\
+
+also, the same Python analyzer code is used as in the VHDL tb
 
@@
 \subsection{TODO: Other Stuff...}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Common Testbench}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Architecture (same tb for VHDL and HLS-generated HDL)}

### 2021-07-22 doc/thesis/chapters/implementation.tex
reuse of application code

@@
 \subsection{Testbench}
+\label{sec:impl:hls:testbench}
 
@@
 The testbench is especially designed around the 'Design For Reuse' segment of the SIMILAR concept, which is introduced in Section \ref{sec:SimilarProcess}.
+This is the case because of two major parts.
+%It is not only using the exact same Python code for analysis, but is also using parts of the application software code directly, to load the testbench input data.\\
 
-testbench is using Firmware code to load the data! --> enables reuse and early testing on the code, before a hardware prototype exists\\
+%%%%%
+\subsubsection{Testbench}
 
-also, the same Python analyzer code is used as in the VHDL tb
+In the previous Section \ref{sec:impl:hls:testbench} it is explained, that the HLS testbench generates output files during its execution.
+These output files are analyzed with the exact same Python analysis code, that is already used in the VHDL testbench analysis.
+
+%%%%%
+\subsubsection{Application Software Code}
+
+The testbench is developed in the language \cplusplus\ and does not have any restrictions in the usage thereof, as Section \ref{sec:hls:workflow} described.
+Thus, the \cplusplus\ code that is later used in the application software to interface with the IP, can already be used in the testbench here, for the same purpose.
+This is especially useful in the development process of such a combined system, consisting of software and hardware, that relies on these two parts working together successfully.
+The fact that the application software can directly be used in the testbench, enables the development of software that interfaces with the hardware in a very early stage of the project, before any actual hardware exists.
+It also allows that the software part is already tested before the deployment on hardware.\\
+
+Fig.\ref{fig:testbench_hls_overview_reuse} gives an overview of the testbench and specifically points out the reuse of the application software.
+The diagram displays the HLS IP in the testbench and in the hardware with dotted lines, to emphasize that they are actually the same.\\
+
+\includepicture [0.8] [0] [H] {Overview of the testbench and the reuse of the application software.
+The HLS IP is displayed with dotted lines, to point out that it is the same in testbench and hardware.} {testbench_hls_overview_reuse} {img/draw.io/testbench_hls_overview_reuse}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TODO: Other Stuff...}
+%\subsection{TODO: Other Stuff...}

### 2021-07-22 doc/thesis/chapters/deployment-on-hardware.tex
update
product design

@@
 \includepicture [0.75] [0] [H] {The ZedBoard PCB \cite{ZedBoard}.} {zedboard_photo} {img/zedboard_photo}
+
+%%%%%%%%%%%%%%%%%%%%
+\section{Product Design}
+
+The implemented version of the FM Radio Receiver on the ZedBoard is a working prototype at this point.
+The implementation does not support live data streaming.
+Instead, antenna data is recorded to a file with the RTL-SDR hardware, using a regular PC.
+The antenna data is recorded as IQ data in the baseband.
+The file is then stored onto an SD card, that is plugged into the ZedBoard.
+The CPU on the SoC then reads it and replays the antenna data to the FPGA.
+In the future, this design may be extended to use an embedded Linux operating system.
+In that way, the RTL-SDR could be connected to the ZedBoard PCB directly via USB, which would enable a live data stream from the antenna to the FPGA.
+However, the chosen solution is sufficient for the aims of this thesis.
 
@@
 
-drivers for IPs, libraries to interface with e.g. SD card, etc\\
---> look at software which parts may be interesting to highlight
+The application software is running on a single core of the dual-core ARM Cortex A9 CPU and is built upon a real-time operating system, namely FreeRTOS.
+Xilinx provides a ported version of this RTOS with the installation of their software development kit, the Xilinx SDK.
+The SDK also includes drivers to support various functionalities, such as support for accessing a FAT file system on the external SD card.
+The respective software drivers for the IPs that are used in the FPGA design are automatically imported into the SDK project.
+They are embedded in the so-called hardware definition file (HDF) that is generated by Vivado after synthesis, and provided together with the bitstream file.
+It is a file with the extension \textit{*.hdf} and can be opened like a regular archive file, with any ZIP archive viewer.
+
+\texttt{--->} look at software which parts may be interesting to highlight

### 2021-07-22 doc/thesis/chapters/deployment-on-hardware.tex
software

@@
 Xilinx provides a ported version of this RTOS with the installation of their software development kit, the Xilinx SDK.
-The SDK also includes drivers to support various functionalities, such as support for accessing a FAT file system on the external SD card.
-The respective software drivers for the IPs that are used in the FPGA design are automatically imported into the SDK project.
-They are embedded in the so-called hardware definition file (HDF) that is generated by Vivado after synthesis, and provided together with the bitstream file.
-It is a file with the extension \textit{*.hdf} and can be opened like a regular archive file, with any ZIP archive viewer.
+The SDK also includes drivers to support various functionalities, such as support for accessing a FAT file system on the external SD card.\\
 
-\texttt{--->} look at software which parts may be interesting to highlight
+The respective software drivers for several IPs that are used in the FPGA design are automatically imported into the SDK project.
+They are embedded in the so-called hardware definition file (HDF) that is generated by Vivado after the synthesis process, and provided together with the bitstream file.
+It is a file with the extension \textit{*.hdf} and can be opened like a regular archive file, with any ZIP archive viewer.\\
+
+Generally, the application software for this project is developed as an object-oriented \cplusplus\ program.
+It uses common software development patterns and a class structure that fits the purpose.
+The main tasks of the application are to configure several IPs in the FPGA, so that they can fulfill their respective purpose in the system, and to read the recorded data from the SD card, which is then streamed to the FPGA.
+It also provides a basic user interface over the serial console, with which a user can interact with the FM Radio Receiver product.\\
+
+However, software design is not the main focus of this thesis, and thus no further details are discussed here.

### 2021-07-23 doc/thesis/chapters/hls.tex
one newline :)
new diagram timeline_performance_first_prototype

@@
 
-\includepicture [0.8] [0] {Design Time vs.
-Application Performance, if the same functionality is implemented in different methods for various targets \cite[Figure 1-1]{VivadoUgHLSIntro}.} {timeline_performance_first_prototype} {img/timeline_performance_first_prototype}
+\includepicture [1.0] [0] {Design Time vs.
+Application Performance, if the same functionality is implemented in different methods for various targets \cite[Figure 1-1]{VivadoUgHLSIntro}.} {timeline_performance_first_prototype} {img/draw.io/timeline_performance_first_prototype}
 
@@
 
-An example for the usage of integer and fixed point data types is given in the following code section.
+An example for the usage of integer and fixed point data types is given in the following code section.\\
 

### 2021-07-28 doc/thesis/chapters/hls.tex
hls interface

@@
 
-This chapter covers High-Level Synthesis in general, but also describes details that specifically apply to Xilinx.
+This chapter covers the High-Level Synthesis of \cplusplus\ in general, but also describes details that specifically apply to HLS tools that are provided by Xilinx.
+An overview of the most important steps in the development workflow, as well as main features and possibilities of the code implementation and testbench are given.
 The main sources of information for this chapter are found in \cite{VivadoUgHLSIntro}and \cite{VivadoUgHLS}.
@@
 \subsection{Interfaces}
+\label{sec:hls:coding:interfaces}
 

### 2021-07-28 doc/thesis/chapters/implementation.tex
math. typesetting for diagram
hls interface

@@
 
-%TODO: use mathematical typesetting in this diagram
 \includepicture [1.0] [0]  {Block diagram of the interface between cocotb and a simulator \cite{cocotbDoc}.} {bd_cocotb_interface_simulator} {img/draw.io/bd_cocotb_interface_simulator}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{TODO: Other Stuff...}
+\subsection{Interfaces}
+
+HLS can automatically implement interfaces, as described in Section \ref{sec:hls:coding:interfaces}.
+This is a major feature that is used in the implementation of the HLS IP.\\
+
+The design uses an AXI4-Lite memory mapped interface, as well as multiple AXI4-Streams.
+Additionally, a data-only interface is used for the connected LEDs.
+The actual implementation is demonstrated in the following lines of code.\\
+
+\begin{CppCode}
+void fm_receiver_hls(hls::stream<iq_sample_t>& iq_in,
+                     hls::stream<audio_sample_t>& audio_out,
+                     config_t& config,
+                     status_t* status,
+                     ap_int<NUM_LEDS>* led_out) {
+
+  /*----------- HLS interface settings ------------*/
+  #pragma HLS INTERFACE ap_ctrl_hs port = return
+
+  #pragma HLS INTERFACE axis port = iq_in
+  #pragma HLS DATA_PACK variable  = iq_in
+
+  #pragma HLS INTERFACE axis port = audio_out
+  #pragma HLS DATA_PACK variable  = audio_out
+
+  #pragma HLS INTERFACE s_axilite port = status bundle = API
+  #pragma HLS INTERFACE s_axilite port = config bundle = API
+
+  #pragma HLS INTERFACE ap_none port = led_out
+
+  ...
+}
+\end{CppCode}
+
+These few lines of code already implement the entire functionality of each interface type.
+The streams use the types \texttt{iq\_sample\_t} and \texttt{audio\_sample\_t}, which are structs of two 16-bit fixed-point variables and thus represent a width of 32 bit.
+The configuration and status registers are also implemented as structs, that contain multiple variables to represent the separate registers.
+For the registers, a special detail is the distinction between the register types read-write and read-only.
+This difference is made by using a value-by-reference or a pointer value, for configuration and status interface, respectively.
+Consequently, the configuration registers become read-writeable and the status registers remain read-only.
+The \texttt{bundle} keyword in the pragma-directives combine the two register groups into a single AXI4-Lite bus interface, instead of creating one bus each.
+This may be of advantage in the application software later, since it only needs to handle a single base address.
+The interface for the LEDs that can be connected is implemented as a data-only type, which does not have any protocol assigned.
+It simply represents the data value that is written to the \cplusplus\ variable.
+The usage of a data type with variable bitwidth, such as \texttt{ap\_int} in the example, is of advantage since only the actually required number of connections is implemented in the interface.

### 2021-07-28 doc/thesis/chapters/hls.tex
sample rate reduction hls

@@
 
+     auto input = sample_in.read();
      ...
+     audio_out.write(audio);
  }
@@
 \end{program}
+
+HLS stream interface variables can be accessed via read- and write-functions.
+One very important property of these functions is, that they are blocking calls.
+The HLS streams are implemented with an underlying FIFO.
+Thus, values can only be written into it when there is an empty spot, and can only be read if there is a at least one value available.
+If the respective condition is not satisfied, the functions block the execution, until the FIFO state changes.
+During simulation in the testbench, this may result in an endlessly blocking behaviour.
 

### 2021-07-28 doc/thesis/chapters/implementation.tex
sample rate reduction hls

@@
 \section{Digital Signal Processing Chain}
+\label{sec:impl:dsp-chain}
 
@@
 Additionally, a data-only interface is used for the connected LEDs.
-The actual implementation is demonstrated in the following lines of code.\\
+The actual implementation is demonstrated in Program \ref{prog:hls_impl_interfaces}.
+These few lines of code already implement the entire functionality of each interface type.\\
 
+\begin{program}
+  \caption{Implementation of sample rate reduction, using for-loops and AXI stream variables with their underlying FIFO behaviour.}
+  \label{prog:hls_impl_interfaces}
 \begin{CppCode}
@@
 \end{CppCode}
+\end{program}
 
-These few lines of code already implement the entire functionality of each interface type.
 The streams use the types \texttt{iq\_sample\_t} and \texttt{audio\_sample\_t}, which are structs of two 16-bit fixed-point variables and thus represent a width of 32 bit.
@@
 This difference is made by using a value-by-reference or a pointer value, for configuration and status interface, respectively.
-Consequently, the configuration registers become read-writeable and the status registers remain read-only.
+Consequently, the configuration registers become read-writeable and the status registers remain read-only.\\
+
 The \texttt{bundle} keyword in the pragma-directives combine the two register groups into a single AXI4-Lite bus interface, instead of creating one bus each.
@@
 The usage of a data type with variable bitwidth, such as \texttt{ap\_int} in the example, is of advantage since only the actually required number of connections is implemented in the interface.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sample Rate Reduction}
+
+Multiple steps in the DSP chain require the operation of sample rate reduction.
+This is explained in section \ref{sec:impl:dsp-chain}.\\
+
+In the VHDL implementation, this is implemented with a counter that counts the number of incoming samples.
+If the counter reaches the number that matches the downsampling rate, is forwards the current sample.
+Otherwise, the incoming samples are discarded.\\
+
+In the HLS implementation however, this is done in another way.
+Here the previous block, which still runs in the oversampled domain, is executed N times, where N is the oversampling rate, before a single output sample is forwarded.
+This behaviour is possible because of the FIFO that is implemented in an AXI stream interface.\\
+
+The implementation of the sample rate reduction can be shown in a short code example in Program \ref{prog:sample_rate_reduction_fifo_behaviour}.
+The FM Demodulator runs in the oversampled domain and is executed \texttt{OSR\_RX} times in the inner for-loop.
+Only the last sample is written to the \texttt{fm\_channel\_data} variable, which itself is an AXI stream type with an underlying FIFO depth of \texttt{OSR\_AUDIO}, as defined with the accompanying pragma.
+Thus, the inner for-loop is executed \texttt{OSR\_AUDIO} times by the outer loop, to fill this FIFO until it is full.
+The subsequent Channel Decoder then reads the entire FIFO and also implements downsampling with a similar behaviour.
+Several samples are processed, where only the last one is forwarded to the subsequent processing step.
+The FIFO behaviour of AXI stream variables is described in Section \ref{sec:hls:coding:interfaces}.
+
+\begin{program}
+  \caption{Implementation of sample rate reduction, using for-loops and AXI stream variables with their underlying FIFO behaviour.}
+  \label{prog:sample_rate_reduction_fifo_behaviour}
+\begin{CppCode}
+audio_sample_t fm_receiver(hls::stream<iq_sample_t>& iq_in) {
+
+  hls::stream<sample_t> fm_channel_data;
+  #pragma HLS STREAM depth = OSR_AUDIO variable = fm_channel_data
+
+  sample_t fm_demod;
+  for (uint8_t i = 0; i < OSR_AUDIO; i++) {
+    for (uint8_t k = 0; k < OSR_RX; k++) {
+      iq_sample_t iq = iq_in.read();
+      fm_demod       = fm_demodulator(iq);
+
+      if (k == OSR_RX - 1) {
+        fm_channel_data.write(fm_demod);
+      }
+    }
+  }
+
+  audio_sample_t audio_sample = channel_decoder(fm_channel_data);
+
+  return audio_sample;
+}
+
+audio_sample_t channel_decoder(hls::stream<sample_t>& in_sample) {
+  sample_t audio_mono;
+
+  for (uint8_t i = 0; i < OSR_AUDIO; i++) {
+    sample_t sample = in_sample.read();
+
+    audio_mono = recover_mono(sample);
+    ...
+  }
+  ...
+  return audio;
+}
+\end{CppCode}
+\end{program}

### 2021-07-28 doc/thesis/chapters/hls.tex
add audio output graphic

@@
 
-The program code example \ref{prog:axi4_lite} shows the specification of an AXI4-Lite memory-mapped slave interface.
+The code example in Program \ref{prog:axi4_lite} shows the specification of an AXI4-Lite memory-mapped slave interface.
 The provided registers are separated into configuration registers and status registers.
@@
 
-The second program code example \ref{prog:axi4_stream} demonstrates the equally simple definition of an AXI4-Stream interface.
+The second code example, Program \ref{prog:axi4_stream}, demonstrates the equally simple definition of an AXI4-Stream interface.
 

### 2021-07-28 hardware/common/analysis/analyze_ip_output_data.m

@@ -61,7 +61,9 @@ audioDataLeft_VHDL_FPGA  = circshift(audioDataLeft_VHDL_FPGA, 3);
 audioDataRight_VHDL_FPGA = circshift(audioDataRight_VHDL_FPGA,3);
 
 % Plot
-fig_title = 'Audio Output';
+set(0,'defaulttextinterpreter','latex')
+
+fig_title = 'Audio Output Comparison Of All Implementation Types';
 fig_audio_time = figure('Name',fig_title);
 sgtitle(fig_title);
 
@@ -100,4 +102,9 @@ if EnableZoomToBegin
     xlim([0,length(audioDataLeft_VHDL_SIM)]);
     ylim([-0.035,0.035]);
 end
-saveas(fig_audio_time, "./audio_output.png");
+
+% Adapt figure size and save it
+fig_audio_time.Position(3:4) = [900 600];
+%saveas(fig_audio_time, "./audio_output_compare_all_ips.png");
+dir_output_doc = "../../../doc/thesis/img/matlab";
+exportgraphics(fig_audio_time,sprintf("%s/%s",dir_output_doc, "audio_output_compare_all_ips.pdf"),'ContentType','vector')

### 2021-07-29 doc/thesis/chapters/conclusion-and-prospect.tex
hls
comments
product design
re-structure the results

@@
 % NOTES:
+%
+% Further steps:
+%  - generate VHDL from Matlab receiver code
+%  - use the common input file and feed it to GNU Radio;
+%    also produce an output file with GNU Radio for comparison
+%  - Develop HLS compiler independent code (optional)}
+%    -- Different compilers use different #pragmas, etc.
+%    -- These #pragmas need to be within the code, at the appropriate positions,
+%       which makes it difficult to write compiler independent code....
 %  - use Matlab Simulink HDL Coder and System Generator
 %     as additional implementation variants
+%  - use Intel oneAPI HLS compiler
+%      -- try to develop compiler-independent code
+%      -- how platform-independent is HLS?
+%  - explore HLS optimizations
+%  - explore HLS libraries (e.g. DSP library --> pre-made FIR filter)
+%  - create another IP in the same system
+%      -- due to the modular system design this is easy to do
+%  - use Linux OS on the SoC, to enable live-data streaming with the RTL-SDR
 %
@@
 %
-%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Advantages / Disadvantages}
+
+easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV.\\
+However, it may be less optimized.\\
+A developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.

### 2021-07-29 doc/thesis/chapters/results.tex
comments
data types
interfaces
code formatting
axi lite register engine
tb stuff
add audio output graphic
re-structure the results

--- 2021-07-09 doc/thesis/chapters/results.tex
@@
 %%%%%%%%%%%%%%%%%%%%
-\chapter{Test Results}
-\label{cha:Results}
+\chapter{Comparison and Results}
+\label{cha:ComparisonAndResults}
+
+This chapter presents the results that are achieved in the development of the FM Radio Receiver project, with the chosen different methods.
+However, the main focus is to compare the results that are achieved with HLS and VHDL, since these are the major implementation variants of the project.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{General}
+
+The main goal of this thesis is to implement an FM radio receiver in multiple different methods.
+This includes GNU Radio, Matlab, VHDL and HLS.
+The target functionality is to listen to the audio broadcast signal of an FM radio station.\\
+
+All of these methods are implemented and the target functionality is achieved with all of them.
+However, the level of detail that is required in the implementation, and the resulting effort and time it takes to implement the respective variant, differs by a large factor.\\
+
+In the following sections, especially the HLS and VHDL implementations are compared on the basis of metrics, but also based on the experience during development.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Functionality}
+
+Generally, both tools - HLS and VHDL - provide the capabilities to implement any functionality in one or the other method.
+However,
+
 % Functionality same?
+% DSP yes, AXI no (FSM of axi-stream)
+% unit tests
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Audio Output}
+% show plots of the common analyzer python script
+Demodulated Signal
+
+\includepicture [1.0] [0] {Comparison of the audio signal.} {audio_output_compare_all_ips} {img/matlab/audio_output_compare_all_ips}
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Interfaces}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Code Development}
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Lines of Code}
+% of the entire project
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Implementation Time}
+
+% Implementation effort/time
+%    -- see logged times; is subjective; had no previous knowledge
+%    -- AXI interfaces --> huge effort in VHDL...
+%        (mention Johannes Walter for generously providing/allowing the Register Engine script, adapted for this thesis)
+%    -- drivers are auto-generated with HLS! huge time-saving..
+%       (maybe show the code of AXI4-Lite interface and its generated driver
+%        --> difference between R/W and RO!;
+%        and show screenshots of the IP in the block design)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Simulation}
 % Sim speed of x min of recorded FM signal
-  % lines of code
-  \subsection{Implementation effort/time}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{System Design}
+
+% - The system level architecture that was chosen, can be applied to any DSP project. Simply replace the main IP, e.g. from FM to DAB+, QAM, etc. and the environment can run in the exact same way.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Deployment On Hardware}
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Integration}
+% IP generation and integration into Vivado
+
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Hardware Utilization}
-  \subsection{Others}
+% --> utilization reports
 
-  % TODO Überlegungen in diesem Kapitel
-  % kann Demodulator geändert werden? Bsp. von FM zu DAB+, QAM, etc..
-  % - von matlab heraus VHDL generieren, etc
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Latency}
+% --> utilization reports
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Applicability}
+% of all implementation variants
+%   GNU Radio for fast prototyping,
+%   Matlab for research
+%   HLS/VHDL for deployment on HW, product development

### 2021-07-29 doc/thesis/chapters/results.tex
comments
results

@@
 
-The main goal of this thesis is to implement an FM radio receiver in multiple different methods.
+A main objective of this thesis is to implement an FM radio receiver in multiple different methods.
 This includes GNU Radio, Matlab, VHDL and HLS.
@@
 
-All of these methods are implemented and the target functionality is achieved with all of them.
-However, the level of detail that is required in the implementation, and the resulting effort and time it takes to implement the respective variant, differs by a large factor.\\
+All of the listed methods are implemented and the target functionality is achieved with all of them.
+However, the level of detail that is required in the implementation, and the resulting effort and time it takes to implement the respective variant, differs by a large factor.
+This is mainly due to the different levels of abstraction, so that the low-level algorithms do not need to be known.\\
 
@@
 
-Generally, both tools - HLS and VHDL - provide the capabilities to implement any functionality in one or the other method.
-However,
+Generally, both tools --- HLS and VHDL --- provide the capabilities to implement any functionality in one or the other method.
+However, the implementation is done in a different level.
+VHDL uses an approach that is very close to the hardware, such as clock cycles and flipflops, while HLS describes the logic on an algorithmic level.\\
+
+The implementation is split into two main parts, the communication interfaces and the DSP.
+The results are presented in the following sections.
 
@@
 %%%%%%%%%%%%%%%%%%%%
+\subsection{Interfaces}
+
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Audio Output}
+
+Regarding the DSP, both implementation variants achieve a very similar result.
+
 % show plots of the common analyzer python script
@@
 \includepicture [1.0] [0] {Comparison of the audio signal.} {audio_output_compare_all_ips} {img/matlab/audio_output_compare_all_ips}
-
-%%%%%%%%%%%%%%%%%%%%
-\subsection{Interfaces}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Deployment On Hardware}
+\section{Deployment on Hardware}
 

### 2021-07-29 doc/thesis/chapters/results.tex
results wip..

@@
 
-Generally, both tools --- HLS and VHDL --- provide the capabilities to implement any functionality in one or the other method.
+Generally, both tools -- HLS and VHDL -- provide the capabilities to implement any functionality in one or the other method.
 However, the implementation is done in a different level.
@@
 
+The AXI4-Lite memory-mapped bus is implemented to have an exactly equal behaviour in both variants.
+There are read-only and read-writeable registers, which are all mapped onto a single base address.\\
+
+The AXI stream interface however does not show the exact same behaviour.
+Here, the HLS variant is using the AXI stream according to its protocol, while the VHDL variant is implemented differently.
+It uses a simplified logic for the ready-flag, which reduces the effort in implementation.
+However, from the perspective of the communicating blocks, the interface is usable like a regular AXI stream.
+The implementation details are explained in Section \ref{sec:impl:vhdl:interfaces}.\\
+
+In the final, integrated system on the SoC hardware, the CPU is able to communicate with both IPs via their interfaces successfully.
+Status and configuration data can be read and written through the AXI4-Lite interface, and the streaming data for the DSP chain is successfully sent through the AXI streams as well.
+
 %%%%%%%%%%%%%%%%%%%%
@@
 
-Regarding the DSP, both implementation variants achieve a very similar result.
+The audio output of the HLS and VHDL variants is compared with the Matlab model, which serves as the reference.
+Additionally, the results of the testbench are compared with their respective result in the actual hardware.
+In summary, it can be stated that the DSP chain produces a very similar audio output in all the compared data sets.
+However, differences remain, which are presented in the following paragraphs and diagrams.
 
-% show plots of the common analyzer python script
-Demodulated Signal
+\subsubsection{Comparing against the Matlab reference}
 
-\includepicture [1.0] [0] {Comparison of the audio signal.} {audio_output_compare_all_ips} {img/matlab/audio_output_compare_all_ips}
+In the comparison against the Matlab model, the HLS variant achieves a very exact match of the output signal.
+However, the VHDL implementation differs by a certain amount.
+There seems to be an issue in the signal separation between the left and right audio channels.
+This is visible in the analysis of the audio signal, as shown is Fig. \ref{fig:audio_output_compare_ips_vs_matlab}.
+The signal strength of the left channel in the upper diagram is weaker than expected, while the right channel in the lower diagram is stronger.
+This imbalance can also be observed by listening to the audio output via speakers.\\
+
+The cause for this issue may be located in several components in the design.
+This includes the FM demodulator, the carrier recovery, i.e. the 38~kHz carrier, as well as a sample timing shift in the final summation and substraction to recover the left and right channels.
+The FIR filter has a successfully passing unit test and is thus assumed to be correct.
+Also the fixed-point data type with its overflow- and rounding behaviour is suspected to be a potential issue.
+Due to the time limitation in the elaboration of this thesis, this issue can not be traced down to the root cause and thus still persists in the current implementation.
+However, from the system-level point of view of this thesis, the IP works and can be integrated into the final system.
+
+%TODO: find a good time-range for this picture and format it correctly
+\includepicture [1.0] [0] {Comparison of the IPs audio signals against the Matlab model.
+The HLS variant matches the model, while the VHDL output diverts by a certain amount.} {audio_output_compare_ips_vs_matlab} {img/matlab/audio_output_compare_ips_vs_matlab}
+
+\subsubsection{Comparing simulation- against hardware results}
+
+Here, the results of the simulation testbenches are compared against the values that are read from the FPGA directly.
+Ideally, the values should be exactly the same.
+However, Figure \ref{fig:audio_output_compare_tb_vs_hw} shows that this is not always the case.
+Again, the HLS implementation does have a matching result, whereas the VHDL variant has deviating values.
+The explanation here may be linked with above suspicions, but may also be linked to the initial reset values that exist in the hardware.
+In the HLS implementation, the reset logic of several registers is implemented automatically, whereas in VHDL these conditions need to be implemented manually.
+Therefore, the IPs internal state, i.e. the register values, may be different at the beginning, which leads to an error propagation throughout the design.
+The entire DSP chain is built like a pipeline and therefor it takes a number of samples to process, before the 'wrong' intermediate values are flushed out.
+All that may be the cause for the deviating values in the VHDL design.\\
+
+Again, the time limitation in the elaboration of this thesis does not allow a deeper analysis of this issue.
+However, the end-to-end system design enables the developer to analyze an issue like this in a convenient way, in quick iterations.
+Any adaptions in the VHDL design can be simulated and integrated into the FPGA with automated scripts.
+Both results can then be analyzed and compared, which may lead to further adaptions.\\
+
+%TODO: find a good time-range for this picture and format it correctly
+\includepicture [1.0] [0] [H] {Comparison of the audio signal in simulation testbench and hardware.} {audio_output_compare_tb_vs_hw} {img/matlab/audio_output_compare_tb_vs_hw}
 

### 2021-07-29 hardware/common/analysis/analyze_ip_output_data.m

@@ -12,7 +12,11 @@ close all;
 clc;
 
 %% Settings
-EnableZoomToBegin = false;
+dir_output_doc = "../../../doc/thesis/img/matlab";
+
+EnableSaveToFile                = true;
+EnableZoomToBegin               = true;
+EnableZoomToComparisonTimeRange = false;
 
 %=========================================================================
 %% Read data
@@ -80,8 +84,10 @@ h4 = plot(audioDataLeft_HLS_SIM,  'r--','DisplayName', 'HLS SIM');
 xline(length(audioDataLeft_HLS_FPGA),'k--','FPGA');
 xline(length(audioDataLeft_HLS_SIM), 'k--','HLS SIM');
 xline(length(audioDataLeft_VHDL_SIM),'k--','VHDL SIM');
-grid on; legend([h0,h1,h2,h3,h4],'Location','east');
+grid on; legend([h0,h1,h2,h3,h4],'Location','southeast');
 ylim([ymin,ymax]);
+xlabel('time [sample]');
+ylabel('amplitude');
 
 ax2 = subplot(2,1,2); hold on;
 title('Right channel');
@@ -93,18 +99,27 @@ h4 = plot(audioDataRight_HLS_SIM,  'r--','DisplayName', 'HLS SIM');
 xline(length(audioDataRight_HLS_FPGA),'k--','FPGA');
 xline(length(audioDataRight_HLS_SIM), 'k--','HLS SIM');
 xline(length(audioDataRight_VHDL_SIM),'k--','VHDL SIM');
-grid on; legend([h0,h1,h2,h3,h4],'Location','east');
+grid on; legend([h0,h1,h2,h3,h4],'Location','southeast');
 ylim([ymin,ymax]);
 
 xlabel('time [sample]');
+ylabel('amplitude');
 linkaxes([ax1,ax2],'xy');
+
+% Adapt figure size
+fig_audio_time.Position(3:4) = [900 450];
+
+% Zoom area of interest, and save
 if EnableZoomToBegin
     xlim([0,length(audioDataLeft_VHDL_SIM)]);
     ylim([-0.035,0.035]);
 end
+if EnableZoomToComparisonTimeRange
+    xlim([3450,4250]);
+    ylim([-0.45,0.65]);
+end
 
-% Adapt figure size and save it
-fig_audio_time.Position(3:4) = [900 600];
-%saveas(fig_audio_time, "./audio_output_compare_all_ips.png");
[…]

### 2021-07-29 hardware/vhdl/ip/tb/integration_test/tests.py

@@ -92,7 +92,7 @@ async def axi_stream_dsp_test(dut):
     # --------------------------------------------------------------------------
 
     # Number of seconds to process
-    n_sec = 0.0025
+    n_sec = 0.1
 
     # --------------------------------------------------------------------------
     # Prepare environment
@@ -143,12 +143,12 @@ async def axi_stream_dsp_test(dut):
     #  This speedup factor needs to be chosen carefully. The IP needs to complete the calculation for one sample,
     #  before the next sample can be taken from the input. Therefore, set the factor small, to ensure enough
     #  low-cycles, to allow the IP to complete one entire calculation.
-    output_speedup_factor = 50
+    output_speedup_factor = 120
     strobe_num_cycles_high = 1
     strobe_num_cycles_low = tb.CLOCK_FREQ_MHZ * 1e6 // fm_global.fs_audio_c // output_speedup_factor - strobe_num_cycles_high
     ratio = strobe_num_cycles_low // strobe_num_cycles_high
     tb.backpressure_i2s.start(bit_toggler(repeat(strobe_num_cycles_high), repeat(strobe_num_cycles_low)))
-    assert ratio >= 9, "output_speedup_factor is set too high! --> IP won't have enough time to calculate a sample, before the next one arrives"
+    assert ratio >= 9, f"output_speedup_factor is set too high ({ratio}<9)! --> IP won't have enough time to calculate a sample, before the next one arrives"
 
     # Fork the 'receiving parts'
     fm_demod_output_fork = cocotb.fork(tb.read_fm_demod_output())

### 2021-07-29 doc/thesis/chapters/results.tex
figure

@@
 
+\includepicture [1.0] [0] {Comparison of the IP audio output signals, in simulation and hardware, against the Matlab model.
+The HLS variant matches the model, while the VHDL output diverts by a certain amount.
+Also, VHDL differs between simulation- and hardware results.} {audio_output_compare_tb_vs_hw} {img/matlab/audio_output_compare_tb_vs_hw}
+
 \subsubsection{Comparing against the Matlab reference}
@@
 There seems to be an issue in the signal separation between the left and right audio channels.
-This is visible in the analysis of the audio signal, as shown is Fig. \ref{fig:audio_output_compare_ips_vs_matlab}.
+This is visible in the analysis of the audio signal, as shown is Fig. \ref{fig:audio_output_compare_tb_vs_hw}.
 The signal strength of the left channel in the upper diagram is weaker than expected, while the right channel in the lower diagram is stronger.
@@
 
-%TODO: find a good time-range for this picture and format it correctly
-\includepicture [1.0] [0] {Comparison of the IPs audio signals against the Matlab model.
+%\includepicture [1.0] [0] {Comparison of the IP audio output signals against the Matlab model.
 The HLS variant matches the model, while the VHDL output diverts by a certain amount.} {audio_output_compare_ips_vs_matlab} {img/matlab/audio_output_compare_ips_vs_matlab}
@@
 Both results can then be analyzed and compared, which may lead to further adaptions.\\
-
-%TODO: find a good time-range for this picture and format it correctly
-\includepicture [1.0] [0] [H] {Comparison of the audio signal in simulation testbench and hardware.} {audio_output_compare_tb_vs_hw} {img/matlab/audio_output_compare_tb_vs_hw}
 

### 2021-07-29 hardware/common/analysis/analyze_ip_output_data.m

@@ -15,7 +15,7 @@ clc;
 dir_output_doc = "../../../doc/thesis/img/matlab";
 
 EnableSaveToFile                = true;
-EnableZoomToBegin               = true;
+EnableZoomToVhdlTimeRange       = true;
 EnableZoomToComparisonTimeRange = false;
 
 %=========================================================================
@@ -67,7 +67,7 @@ audioDataRight_VHDL_FPGA = circshift(audioDataRight_VHDL_FPGA,3);
 % Plot
 set(0,'defaulttextinterpreter','latex')
 
-fig_title = 'Audio Output Comparison Of All Implementation Types';
+fig_title = 'Audio Output Comparison';
 fig_audio_time = figure('Name',fig_title);
 sgtitle(fig_title);
 
@@ -107,12 +107,14 @@ ylabel('amplitude');
 linkaxes([ax1,ax2],'xy');
 
 % Adapt figure size
-fig_audio_time.Position(3:4) = [900 450];
+fig_audio_time.Position(3:4) = [900 600];
 
 % Zoom area of interest, and save
-if EnableZoomToBegin
+if EnableZoomToVhdlTimeRange
     xlim([0,length(audioDataLeft_VHDL_SIM)]);
-    ylim([-0.035,0.035]);
+    xlim([2800,3950]);
+    exportgraphics(fig_audio_time,sprintf("%s/%s",dir_output_doc, "audio_output_compare_tb_vs_hw.pdf"),'ContentType','vector')
+    ylim([-0.45,0.6]);
 end
 if EnableZoomToComparisonTimeRange
     xlim([3450,4250]);
@@ -120,6 +122,9 @@ if EnableZoomToComparisonTimeRange
 end
 
 if EnableSaveToFile
-    %saveas(fig_audio_time, "./audio_output_compare_all_ips.png");
-    exportgraphics(fig_audio_time,sprintf("%s/%s",dir_output_doc, "audio_output_compare_all_ips.pdf"),'ContentType','vector')
+    %exportgraphics(fig_audio_time,sprintf("%s/%s",dir_output_doc, "audio_output_compare_ips_vs_matlab.pdf"),'ContentType','vector')
+    
+%    xlim auto
+%    ylim auto
+%    exportgraphics(fig_audio_time,sprintf("%s/%s",dir_output_doc, "audio_output_compare_all_ips.pdf"),'ContentType','vector')
 end

### 2021-07-29 hardware/common/analysis/analyze_ip_output_data.m
bit smaller figure

@@ -107,7 +107,7 @@ ylabel('amplitude');
 linkaxes([ax1,ax2],'xy');
 
 % Adapt figure size
-fig_audio_time.Position(3:4) = [900 600];
+fig_audio_time.Position(3:4) = [900 550];
 
 % Zoom area of interest, and save
 if EnableZoomToVhdlTimeRange

### 2021-07-29 doc/thesis/chapters/dsp.tex
hls interface
results
Fig. -> Figure

@@
 \noindent
-The block diagram in Fig.\ref{fig:bd_dsp_overview} gives a high-level overview of a receiver structure.
+The block diagram in Figure \ref{fig:bd_dsp_overview} gives a high-level overview of a receiver structure.
 
@@
 
-Fig.\ref{fig:bd_superhet} shows a block diagram of a superhet receiver.
+Figure \ref{fig:bd_superhet} shows a block diagram of a superhet receiver.
 The antenna signal first runs through a band selection filter.
@@
 
-\includepicture [1.0] {Block design of a superhetorodyne receiver \cite[Fig.2.1]{PassosFábio2020AHSo}} {bd_superhet} {img/draw.io/superheterodyne-receiver-block-diagram}
+\includepicture [1.0] {Block design of a superhetorodyne receiver \cite{PassosFábio2020AHSo}} {bd_superhet} {img/draw.io/superheterodyne-receiver-block-diagram}
 
@@
 
-\includepicture [1.0] {Block design of a direct-conversion receiver \cite[Fig.2.3]{PassosFábio2020AHSo}.} {bd_direct_conversion} {img/draw.io/direct-conversion-receiver-block-diagram}
+\includepicture [1.0] {Block design of a direct-conversion receiver \cite{PassosFábio2020AHSo}.} {bd_direct_conversion} {img/draw.io/direct-conversion-receiver-block-diagram}
 
@@
 
-A block diagram, as well as the signal spectrum of the process is shown in Fig.\ref{fig:bd_down_conversion_spectrum}.
+A block diagram, as well as the signal spectrum of the process is shown in Figure \ref{fig:bd_down_conversion_spectrum}.
 In this example, the frequency of the LO is set to shift the signal from $f_m$ to baseband.
@@
 Their respective amplitudes are $A_m$ and $A_c$.
-This equation is depicted as a block diagram and signal spectrum in Fig.\ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.
+This equation is depicted as a block diagram and signal spectrum in Figure \ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.
 The signal $m(t)$ is a baseband signal.
@@
 
-The graphical representation in the frequency spectrum is shown in Fig.\ref{fig:down_conversion_spectrum_steps}.
+The graphical representation in the frequency spectrum is shown in Figure \ref{fig:down_conversion_spectrum_steps}.
 As mentioned above, the assumption is a direct conversion to baseband, so $w_m = w_c$.
@@
 These replicas overlap, if $f_s$ does not meet the theorem, which results in signal distortion.
-Fig.\ref{fig:aliasing_freq_spectrum} illustrates the effect of aliasing in the frequency spectrum \cite{ThyagarajanK.S2019ItDS}.
+Figure \ref{fig:aliasing_freq_spectrum} illustrates the effect of aliasing in the frequency spectrum \cite{ThyagarajanK.S2019ItDS}.
 
@@
 
-Fig.\ref{fig:fm_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
+Figure \ref{fig:fm_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
 For a better overview, only the positive spectum part is shown.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Baseband Delay Demodulator}
-\label{sec:BasebandDelayDemodulator}
-TODO\\
-see /home/mike/Dropbox/Studium/ma\_semester\_4\_SS21/ma/literature/FmDemodulator.pdf
+%\subsection{Baseband Delay Demodulator}
+%\label{sec:BasebandDelayDemodulator}
+%TODO\\
+%see /home/mike/Dropbox/Studium/ma\_semester\_4\_SS21/ma/literature/%FmDemodulator.pdf
 
@@
 Different strategies can be implemented for both parts.
-A block diagram for a frequency discriminator with its two main parts is shown in Fig.\ref{fig:bd_freq_discriminator}.
+A block diagram for a frequency discriminator with its two main parts is shown in Figure \ref{fig:bd_freq_discriminator}.
 
@@
 
-A time-domain example for this signal is shown in the top left diagram in Fig.\ref{fig:time_domain_envelope_detect}.\\
+A time-domain example for this signal is shown in the top left diagram in Figure \ref{fig:time_domain_envelope_detect}.\\
 
@@
 In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.
-A block diagram is shown in Fig.\ref{fig:bd_freq_discriminator}.\\
+A block diagram is shown in Figure \ref{fig:bd_freq_discriminator}.\\
 
-The method implemented in Fig.\ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as `Asynchronous Half-Wave Envelope Detector' \cite{DSPRelatedEnvelopeDetection}.
+The method implemented in Figure \ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as `Asynchronous Half-Wave Envelope Detector' \cite{DSPRelatedEnvelopeDetection}.
 It consists of a thresholding unit to remove negative values and a conventional lowpass filter.
@@
 
-A frequency spectrum of the multiplex signal of an FM channel is shown in Fig.\ref{fig:fm_channel_baseband_freqs}.
+A frequency spectrum of the multiplex signal of an FM channel is shown in Figure \ref{fig:fm_channel_baseband_freqs}.
 
@@
 
-The block diagram in Fig.\ref{fig_bd_stereo_demod} illustrates the signal processing chain for stereo audio.
+The block diagram in Figure \ref{fig_bd_stereo_demod} illustrates the signal processing chain for stereo audio.
 In the block diagram, $x(t)$ represents the multiplex signal from the FM demodulator.

### 2021-07-29 doc/thesis/chapters/hls.tex
Fig. -> Figure

@@
 This is especially important when an FPGA design is combined with a software part, like in a System On Chip, since software development is comparably fast.
-Fig.\ref{fig:timeline_performance_first_prototype} schematically displays the difference in development time, over performance and degree of optimization, when the same functionality of a prototype is implemented in different methods for various targets.
+Figure \ref{fig:timeline_performance_first_prototype} schematically displays the difference in development time, over performance and degree of optimization, when the same functionality of a prototype is implemented in different methods for various targets.
 The compared targets are an x86 processor CPU, a digital signal processor (DSP), a graphics processing unit (GPU) and the FPGA design in RTL or HLS implementation.

### 2021-07-29 doc/thesis/chapters/implementation.tex
results wip..
Fig. -> Figure

@@
 \noindent
-Fig.\ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram.\\
+Figure \ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram.\\
 
@@
   \item \textbf{FM Demodulator}\\
+       TODO\\
+       % TODO
       The FM demodulator is implemented as a Baseband Delay Modulator, as described in Section \ref{sec:BasebandDelayDemodulator}.
@@
 
-The transmitter is implemented according to the structure that is depicted as a block diagram in Fig.\ref{fig:bd_matlab_transmitter}.
-To facilitate an easy understanding of the block diagram, it is of advantage to keep in mind that it exactly resembles the frequency spectrum that is previously shown in Fig.\ref{fig:fm_channel_baseband_freqs}.\\
+The transmitter is implemented according to the structure that is depicted as a block diagram in Figure \ref{fig:bd_matlab_transmitter}.
+To facilitate an easy understanding of the block diagram, it is of advantage to keep in mind that it exactly resembles the frequency spectrum that is previously shown in Figure \ref{fig:fm_channel_baseband_freqs}.\\
 
@@
 
-At this point in the DSP chain, the signal resembles the frequency spectrum that is displayed in Fig.\ref{fig:fm_channel_baseband_freqs}.\\
+At this point in the DSP chain, the signal resembles the frequency spectrum that is displayed in Figure \ref{fig:fm_channel_baseband_freqs}.\\
 
@@
 
-The receiver is implemented, following the block diagram in Fig.\ref{fig:bd_dsp_detailed}.
+The receiver is implemented, following the block diagram in Figure \ref{fig:bd_dsp_detailed}.
 Further assumptions and simplifications, such as the fixed point arithmetic are described into detail above.
@@
 
-An example of a plot that is created by the analysis is depicted in Fig.\ref{fig:matlab_analysis_time_domain}.
+An example of a plot that is created by the analysis is depicted in Figure \ref{fig:matlab_analysis_time_domain}.
 The diagram shows the respective signals in the time domain.
@@
 Further analysis is done in the frequency domain, especially in order to better understand the modulation of the different spectral parts in the FM channel.
-An example thereof is shown in Fig.\ref{fig:matlab_analysis_freq_domain}.
+An example thereof is shown in Figure \ref{fig:matlab_analysis_freq_domain}.
 The spectrum of an FM channel multiplex signal, as it is generated by a transmitter before the modulation, is shown in this diagram.
@@
 GNU Radio provides a comprehensive set of DSP blocks, which are available to build various functionalities.
-Multiple blocks are connected to a block diagram, that is very similar to the one in Fig.\ref{fig:bd_matlab_transmitter}.
+Multiple blocks are connected to a block diagram, that is very similar to the one in Figure \ref{fig:bd_matlab_transmitter}.
 The USRP device is used to broadcast the RF signal over the air.\\
@@
 \noindent
-Fig.\ref{fig:bd_gnuradio_setup_transmitter} shows a block diagram of the transmitter application.
+Figure \ref{fig:bd_gnuradio_setup_transmitter} shows a block diagram of the transmitter application.
 
@@
 This receiver works with both devices, the USRP b200mini and the RTL-SDR.
-Fig.\ref{fig:bd_gnuradio_setup_receiver} shows a block diagram of the receiver application.
+Figure \ref{fig:bd_gnuradio_setup_receiver} shows a block diagram of the receiver application.
 
@@
 
-Fig.\ref{fig:bd_cocotb_interface_simulator} gives an overview of how cocotb interfaces with a simulator.
+Figure \ref{fig:bd_cocotb_interface_simulator} gives an overview of how cocotb interfaces with a simulator.
 The Python test cases are shown as a list of coroutines, which are managed by a scheduler.
@@
 
-Fig.\ref{fig:bd_testbench_architecture_vhdl} shows a block diagram of the implemented architecture.
+Figure \ref{fig:bd_testbench_architecture_vhdl} shows a block diagram of the implemented architecture.
 The split between the Python environment and the VHDL implementation is clearly visible but is represented from the developers perspective.
@@
 \subsection{Interfaces}
+\label{sec:impl:vhdl:interfaces}
 
@@
 
-Fig.\ref{fig:register_engine} gives an overview of how the Register Engine script works.
+Figure \ref{fig:register_engine} gives an overview of how the Register Engine script works.
 The Python script takes the register definitions from the YAML file, and uses templates to generate the respecitive output files.\\
@@
 
-Fig.\ref{fig:testbench_hls_overview_reuse} gives an overview of the testbench and specifically points out the reuse of the application software.
+Figure \ref{fig:testbench_hls_overview_reuse} gives an overview of the testbench and specifically points out the reuse of the application software.
 The diagram displays the HLS IP in the testbench and in the hardware with dotted lines, to emphasize that they are actually the same.\\
@@
 \subsection{Interfaces}
+\label{sec:impl:hls:interfaces}
 

### 2021-07-29 doc/thesis/chapters/system-architecture-and-concept.tex
spelling
Fig. -> Figure

@@
 %%%%%%%%%%%%%%%%%%%%
-\section{The project}
+\section{The Project}
 
@@
 \noindent
-A very high-level overview of the entire system is given by the block diagram in Fig.\ref{fig:bd_system_overview_highlevel}.
+A very high-level overview of the entire system is given by the block diagram in Figure \ref{fig:bd_system_overview_highlevel}.
 It shows the front end, which processes the received antenna signal, so that it can be used in the FM radio receiver block.
@@
 %%%%%%%%%%%%%%%%%%%%
-\section{System concept}
+\section{System Concept}
 
@@
 
-An example of a system concept and its architecture is developed and shown in Fig.\ref{fig:bd_system_overview_detailed}.
+An example of a system concept and its architecture is developed and shown in Figure \ref{fig:bd_system_overview_detailed}.
 Here, a common data source is provided by the block called \textit{Front End}.

### 2021-07-29 doc/thesis/chapters/system-design-process.tex
Fig. -> Figure

@@
 
-Each of these terms represents a separate step in the process diagram shown in Fig.\ref{fig:similar_process}.
+Each of these terms represents a separate step in the process diagram shown in Figure \ref{fig:similar_process}.
 The steps can be seen as iterative and parallel, as the process runs during the product development cycle.
@@
 A well-known process model is the so called V-Model.
-It is based on the similarly well-known Waterfall Model, and has its name because of the distinct V-shape of its process diagram, as it is shown in Fig.\ref{fig:v_modell}.
+It is based on the similarly well-known Waterfall Model, and has its name because of the distinct V-shape of its process diagram, as it is shown in Figure \ref{fig:v_modell}.
 One of the main characteristics of the V-Model is, that one phase needs to be completed before the following phase can start.
@@
 
-Furthermore, the V-model diagram in Fig.\ref{fig:v_modell} shows the difference between validation and verification, as it is explained in the previous section \ref{sec:ValidationAndVerification}.
+Furthermore, the V-model diagram in Figure \ref{fig:v_modell} shows the difference between validation and verification, as it is explained in the previous section \ref{sec:ValidationAndVerification}.
 
@@
 
-Fig. \ref{fig:simulink_system_generator} shows this entire design approach in a block diagram.
+Figure \ref{fig:simulink_system_generator} shows this entire design approach in a block diagram.
 The abbreviation ISE stands for Integrated Software Environment, and is Xilinx' development environment that includes the FPGA synthesis tool.
@@
 VHDL and Verilog require, or allow, a very detailed description of the RTL code, while SystemC, \cplusplus, or Matlab and Simulink provide a high level of abstraction.
-In Fig.\ref{fig:fpga_design_levels}, the various implementation methods are compared.
+In Figure \ref{fig:fpga_design_levels}, the various implementation methods are compared.
 The respective level of abstraction is higher towards the upper end, while the level of detail increases towards the bottom of the figure.\\

### 2021-07-29 hardware/vhdl/ip/tb/integration_test/tests.py

@@ -92,7 +92,7 @@ async def axi_stream_dsp_test(dut):
     # --------------------------------------------------------------------------
 
     # Number of seconds to process
-    n_sec = 0.1
+    n_sec = 0.0025
 
     # --------------------------------------------------------------------------
     # Prepare environment

### 2021-07-29 doc/thesis/chapters/deployment-on-hardware.tex
Fig. -> Figure
incorporate some of the feedback (wip)

@@
 It resembles the size of a business card and is developed by Ettus Research.
-A signal range from 70 MHz to 6 GHz can be covered by the front end.
+A signal range from 70~MHz to 6~GHz can be covered by the front end.
 The board further features a Xilinx Spartan-6 FPGA that is user-programmable, which makes the device very flexible for various applications with the received signal \cite{USRPb200Mini}.\\
@@
 
-The block diagram in Fig.\ref{fig:bd_impl_vivado} shows the implemented architecture that is used in the SoC.
+The block diagram in Figure \ref{fig:bd_impl_vivado} shows the implemented architecture that is used in the SoC.
 The various design domains in and around the SoC are depicted in different visual representations.

### 2021-07-29 doc/thesis/chapters/dsp.tex
incorporate some of the feedback (wip)

@@
 Another advantage is the ability to adapt the ongoing transmission to the changing characteristics of the channel.
-This is especially useful for mobile receivers, since the signals' path is inherently changing over time there, which leads to challenging conditions for the receiver.
+This is especially useful for mobile receivers, since the signals' path is inherently changing over time, which leads to challenging conditions for the receiver.
 Additionally, varying signal strength, or multipath propagation may occur.
-Multipath propagation is caused by reflections of the signal, which - for example - can occur on a house facade, or a large vehicle that is passing by.
-The receiver then gets a signal, that is a summation of the direct path and the reflected path - a signal from multiple paths.
+Multipath propagation is caused by reflections of the signal, which, for example, can occur on a house facade, or a large vehicle that is passing by.
+The receiver then gets a signal that is a summation of the direct path and the reflected path -- a signal from multiple paths.
 Besides that, environmental impacts, such as rain or fog have an effect on the transmission channel and thus the received signal.\\
@@
 ; HF to IF to BB --> real signal ) ??
-These three steps - amplification, filtering and down-conversion to an IF - are often done in the analog domain, since digital circuits cannot work with such high frequencies.
+These three steps -- amplification, filtering and down-conversion to an IF -- are often done in the analog domain, since digital circuits cannot work with such high frequencies.
 After the analog signal is down-converted to baseband, it needs to be digitized.
@@
 Two common examples are explained in the following sections.
+The main source of information for this section is \cite{PassosFábio2020AHSo}.
 
@@
 Thus, it is also called an image rejection filter.
-The demodulator stage can now perform any further signal processing on the signal, until the required signal is retrieved \cite{PassosFábio2020AHSo}.\\
+The demodulator stage can now perform any further signal processing on the signal, until the required signal is retrieved.\\
 
-The depicted block diagram may represent an FM broadcast receiver, where an audio output is generated from the antenna input.
+The depicted block diagram represents a superhet receiver structure, for example an FM broadcast receiver, where an audio output is generated from the antenna input.
 
@@
 
-A direct-conversion receiver, as its name suggests, converts an RF signal to baseband directly.
+A direct-conversion receiver, as its name suggests, converts an RF signal directly to baseband.
 It can also be called homodyne, synchrodyne, or zero-IF receiver.\\
@@
 This is due to the properties of the mixer.
-The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer output, and the flicker noise, also called \textit{1/f}-noise, which induces strong noise at low frequencies \cite{PassosFábio2020AHSo}.
+The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer's output, and the flicker noise, also called \textit{1/f}-noise, which induces strong noise at low frequencies.
 These effects are not explained into more detail here, since this is not the focus of this thesis.
@@
 It may be antenna signal that was filtered by a bandpass filter.
-The received signal is a real signal, and thus its negative frequency spectrum is an exact mirror of the positive frequencies.
+The received signal is a real signal, and thus, its negative frequency spectrum is an exact mirror of the positive frequencies.
 This fact can be explained in two simple equations.\\
@@
 It is well-known that the real signal $cos(wt) = \frac{1}{2} \Big( e^{jw} + e^{-jw} \Big)$, which follows from $e^{jw} = cos(wt)+jsin(wt)$.
-Looking at the frequencies of the first equation, the negative and positive parts can be seen.\\
+Looking at the frequencies $\omega$ of the first equation, the negative and positive parts can be seen.\\
 
@@
 
+\noindent
 Assuming that $m(t) = cos(\omega_mt)$, $c(t) = cos(\omega_ct)$ and $A_m = A_c = 1$,
@@
 
+\noindent
 follows.
@@
 
+\noindent
 the final formula of the conversion can be represented as
@@
 
-In words: The sampling frequency must be higher than twice the frequency of the signals' highest frequency content.
-Any digital signal processing technique must hold the specification of the sampling theorem \cite[chpt. 4.2]{AlessioSilviaMaria2016DSPa}.
+In words: The sampling frequency must be higher than twice the frequency of the signal's highest frequency content.
+Any digital signal processing technique must hold the specification of the sampling theorem \cite[Ch. 4.2]{AlessioSilviaMaria2016DSPa}.
 In case the theorem is violated, the effect of aliasing occurs on the generated signal.
@@
   Frequency spectrum of the sampled baseband signal with its replicas at integer multiples of $f_s$ \cite{ThyagarajanK.S2019ItDS}.\\
-  A) $f_s > 2f_{signal}$: $f_s$ is high enough and thus meets the theorem.\\
+  A) $f_s > 2f_{signal}$: $f_s$ is high enough, and thus, meets the theorem.\\
   B) $f_s = 2f_{signal}$: $f_s$ exactly meets the theorem.\\
-  C) $f_s < 2f_{signal}$: $f_s$ is too low and thus creates overlaps of the replicas - aliasing.
+  C) $f_s < 2f_{signal}$: $f_s$ is too low, and thus, creates overlaps of the replicas -- aliasing.
 } {aliasing_freq_spectrum} {img/draw.io/sampling-aliasing}
@@
 
-As an example, assuming the signal of interest has a bandwidth of 20 kHz.
-The Nyquist theorem thus requires a remaining sampling rate of at least twice the signals' frequency, so 40 kHz.
-The signal is sampled with an oversampling factor of 10 - meaning a sample rate of 200 kHz.
-A downsampling factor of 4 is now applied, which results in a sample rate of 50 kHz.
+As an example, assuming the signal of interest has a bandwidth of 20~kHz.
+The Nyquist theorem thus requires a remaining sampling rate of at least twice the signals' frequency, so 40~kHz.
+The signal is sampled with an oversampling factor of 10 -- meaning a sample rate of 200~kHz.
+A downsampling factor of 4 is now applied, which results in a sample rate of 50~kHz.
 The resulting signal is now downsampled and can be processed at this lower data rate, while none of the original data content is lost.\\
@@
 
-More detailed information can be found in \cite[chpt. 6.9]{AlessioSilviaMaria2016DSPa}, \cite[chpt. 4.1]{GopiE.S2018MDSP} and \cite[chpt. 10.2.2]{ThyagarajanK.S2019ItDS}.
+More detailed information can be found in \cite[Ch.~6.9]{AlessioSilviaMaria2016DSPa}, \cite[Ch.~4.1]{GopiE.S2018MDSP} and \cite[Ch.~10.2.2]{ThyagarajanK.S2019ItDS}.
 
@@
 \begin{equation}
-  s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int m(t) dt \Big)
+  s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int m(t) dt \Big),
   \label{equ:fm_func}
@@
 \begin{equation}
-  \beta = \frac{\Delta f}{f_m} = \frac{k_f A_m}{f_m}
+  \beta = \frac{\Delta f}{f_m} = \frac{k_f A_m}{f_m},
 \end{equation}
@@
 \end{equation}
-By inserting this message signal into the generic FM formula \eqref{equ:fm_func}, the final equation for an FM signal transforms into the following form.
+By inserting this message signal into the generic FM Equation \eqref{equ:fm_func}, the final formula for an FM signal transforms into the following form.
 \begin{equation}
@@
 
-The frequency band that is used for FM broadcasting is defined worldwide and spans from 87.5 to 108 MHz.
+The frequency band that is used for FM broadcasting is defined worldwide and spans from 87.5 to 108~MHz.
 However, some countries only partially use the band \cite[RR5-39]{ItuRadioRegulations}.
@@
 
-The entire FM broadcasting band from 87.5 to 108 MHz divides into multiple channels which can be used.
+The entire FM broadcasting band from 87.5 to 108~MHz divides into multiple channels which can be used.
 In Europe, the European Telecommunications Standards Institute (ETSI) sets the respective standards for the usage of these frequencies.
@@
 
-Each channel may allocate a bandwidth of 200 kHz.
-Within one of these channels, the maximum deviation from the center frequency shall not exceed 75 kHz.
-This leaves a guard band of 25 kHz on either side, to minimize interference with adjacent channels.
-Center frequencies may be allocated on multiples of 100 kHz.
+Each channel may allocate a bandwidth of 200~kHz.
+Within one of these channels, the maximum deviation from the center frequency shall not exceed 75~kHz.
+This leaves a guard band of 25~kHz on either side, to minimize interference with adjacent channels.
+Center frequencies may be allocated on multiples of 100~kHz.
 However, neighboring channels need to be considered therefore, since this may result in an overlap in the spectrum.\\
@@
 Therefore, it is also referred to as `multiplex signal' in the literature.
-Several parts are described in the following paragraphs \cite{EbuFmRegulators}\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard}\cite{ElectronicsNotesFmTutorial}.
+Several parts are described in the following paragraphs \cite{EbuFmRegulators},\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard},\cite{ElectronicsNotesFmTutorial}.
 
@@
 
-The mono audio stream is located between 30 Hz and 15 kHz.
+The mono audio stream is located between 30~Hz and 15~kHz.
 This signal is built by the sum of the left and right audio channels.
-%The lower limit at 30 Hz is to prevent the transmission of a direct current (DC) part, which would require a large amount of power in transmission.
+%The lower limit at 30~Hz is to prevent the transmission of a direct current (DC) part, which would require a large amount of power in transmission.
 %This effect is one of the downsides of amplitude modulation (AM), where the carrier is located at the channels' center frequeny and requires a high transmission power. % TODO: check this fact
-The upper limit at 15 kHz is chosen to maintain a sufficient spacing to the first subcarrier.
-For audio streaming, as in FM broadcasting, this is not a limitation, since it is unlikely to have frequencies higher than 15 kHz in an audio stream.
+The upper limit at 15~kHz is chosen to maintain a sufficient spacing to the first subcarrier.
+For audio streaming, as in FM broadcasting, this is not a limitation, since it is unlikely to have frequencies higher than 15~kHz in an audio stream.
 Also, this already reaches the upper limit of the human ears' bandwidth, so higher spectral parts will not be heard anyway.
@@
 
-The first subcarrier is allocated at an offset of 19 kHz from the center frequency.
-This subcarrier is also called the 'pilot tone', since it is a continuous signal.
+The first subcarrier is allocated at an offset of 19~kHz from the center frequency.
+This subcarrier is also called the pilot tone, since it is a continuous signal.
 It is independent of any varying message signal content.
@@
 
-A signal that is constructed by the difference between the left and right audio channel is modulated on a 38 kHz subcarrier.
-This subcarrier is an integer multiple of the pilot tone at 19 kHz, for practical reasons.
-However, the 38 kHz carrier is suppressed and thus not visible in the received spectrum.
+A signal that is constructed by the difference between the left and right audio channel is modulated on a 38~kHz subcarrier.
+This subcarrier is an integer multiple of the pilot tone at 19~kHz, for practical reasons.
+However, the 38~kHz carrier is suppressed and thus not visible in the received spectrum.
 The modulation technique that is used for this spectral part is called dual-sideband suppressed-carrier (DSB-SC).
-Even though the carrier is suppressed, it can still be recovered, because it is phase coherent with the 19 kHz pilot tone per definition.\\
+Even though the carrier is suppressed, it can still be recovered, because it is phase coherent with the 19~kHz pilot tone per definition.\\
 
-The bandwidth for this difference-signal spans 15 kHz on either side of the subcarrier.
+The bandwidth for this difference-signal spans 15~kHz on either side of the subcarrier.
 It is used to generate a stereo audio signal, in combination with the mono signal.
-This process is explained into more detail in chapter \ref{sec:demodulation-of-broadcast-fm}.
+This process is explained into more detail in Chapter \ref{sec:demodulation-of-broadcast-fm}.
 
@@
 
-Considering all these parts in the spectrum, there is still free bandwidth available to use up to the maximum channel bandwitdh of 100 kHz in one sideband.
+Considering all these parts in the spectrum, there is still free bandwidth available to use up to the maximum channel bandwitdh of 100~kHz in one sideband.
 Because of that, additional services were added to the pure audio transmission, to provide additional data services and information.\\
@@
 Out of these, RDS is the most significant service in Europe.
-It is used to transmit additional information about the channel, such as the radio stations' name, the currently playing songs' title or traffic information.
+It is used to transmit additional information about the channel, such as the radio stations' names, the currently playing song's title or traffic information.
 
@@
 The differentiator already converts the FM signal into an AM signal, but still leaves an FM content.
-The transformation that happens here can be expressed in a formula, by simply differentiating Eq.\eqref{equ:fm_func}.
+The transformation that happens here can be expressed in a formula, by simply differentiating Equation \eqref{equ:fm_func}.
 This operation, with some additional simple algebraic transformations, results in
@@
 
-This formula shows the AM and an FM contents clearly.
+This formula shows AM and FM contents clearly.
 The FM part is described by the sine function, with its argument.
@@
 The transmitted signal is subject to random additive noise on the transmission channel.
-The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Eq.\eqref{equ:differentiator_two_samples}.
+The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Equation \eqref{equ:differentiator_two_samples}.
 To achieve a constant amplitude, the received complex baseband signal $s(t)$ can be normalized to a value of one.
 \begin{equation}
-  s_{norm} = \frac{s}{|s|} = \frac{A(n)\ e^{j\phi_{FM}(n)}}{|{A(n)\ e^{j\phi_{FM}(n)}|}} = e^{j\phi_{FM}(n)}
+  s_{norm} = \frac{s}{|s|} = \frac{A(n)\ e^{j\phi_{\text{FM}}(n)}}{|{A(n)\ e^{j\phi_{\text{FM}}(n)}|}} = e^{j\phi_{\text{FM}}(n)}
 \end{equation}
 \begin{equation}
-  e^{j\phi_{FM}(n)} = |1|
+  |e^{j\phi_{FM}(n)}| = 1
 \end{equation}
@@
 
-The differentiated signal still consists of a FM part, as explained above, using Eq.\eqref{equ:fm_demod_discriminator}.
+The differentiated signal still consists of a FM part, as explained above, using Equation \eqref{equ:fm_demod_discriminator}.
 In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.
@@
 A mono receiver can thus simply replay the so-called multiplex signal that is produced by the FM demodulator, without any further filtering or processing.
-The 19 kHz pilot tone, and any higher frequency components, will not be audible by most people, because it is outside of the range of a human ear.
+The 19~kHz pilot tone, and any higher frequency components, will not be audible by most people, because it is outside of the range of a human ear.
 Besides that, any used speaker may also be unable to produce a frequency in this range.\\
@@
 Consequently, in order to demodulate this spectral part, a lowpass filter is required.
-The filter needs to have a cutoff frequency of 15 kHz, to allow the entire mono audio spectrum to pass.
-Sufficient attenutation must be available before 19 kHz, in order to suppress the pilot tone.
+The filter needs to have a cutoff frequency of 15~kHz, to allow the entire mono audio spectrum to pass.
+Sufficient attenutation must be available before 19~kHz, in order to suppress the pilot tone.
 
@@
 In the block diagram, $x(t)$ represents the multiplex signal from the FM demodulator.
-The signals $x_l(t)$ and $x_r(t)$ describe the left and right audio signal, respectively.\\
+The signals $x_L(t)$ and $x_R(t)$ describe the left and right audio signal, respectively.\\
 
@@
 
-The central branch selects the pilot tone at 19 kHz with a bandpass filter.
-This bandpass needs to have a frequency response that is sharp enough to have a sufficient attenuation at $\pm$4 kHz, since this is where the mono and difference signals' frequency spectra are allocated.
-Afterwards, the pilot tone frequency is doubled to achieve a 38 kHz subcarrier.
+The central branch selects the pilot tone at 19~kHz with a bandpass filter.
+This bandpass needs to have a frequency response that is sharp enough to have a sufficient attenuation at $\pm$4~kHz, since this is where the mono and difference signals' frequency spectra are allocated.
+Afterwards, the pilot tone frequency is doubled to achieve a 38~kHz subcarrier.
 This guarantees a phase-coherency between pilot and subcarrier, which is required to correctly shift the difference signal down to baseband.\\
@@
 Another option is to use a PLL and configure it to generate the double frequency at its output.
-The PLL lock indicator can then be re-used as an indicator, whether a pilot tone is existent.\\
+The PLL lock indicator can then be re-used as an indicator, whether a pilot tone exists.\\
 
-In the upper branch, a bandpass filter selects the difference signal with ranges from 23 to 53 kHz.
-It is then modulated to baseband by the previously generated 38 kHz subcarrier.
-Subsequently, a lowpass limits signal bandwidth to 15 kHz, which removes the generated modulation artifacts.\\
+In the upper branch, a bandpass filter selects the difference signal with ranges from 23 to 53~kHz.
+It is then modulated to baseband by the previously generated 38~kHz subcarrier.
+Subsequently, a lowpass limits signal bandwidth to 15~kHz, which removes the generated modulation artifacts.\\
 
@@
 
-The rightmost part in the diagram performs the combination of the mono and the difference signal, according to Eq.\eqref{equ_stereo_from_sum_diff}.
+The rightmost part in the diagram performs the combination of the mono and the difference signal, according to Equation \eqref{equ_stereo_from_sum_diff}.

### 2021-07-29 doc/thesis/chapters/implementation.tex
incorporate some of the feedback (wip)

@@
 Again, an audio file is used as an input, which is modulated as an FM channel signal.
-The entire multiplex signal is then modulated onto an actual carrier frequency of 99 MHz, that is within the FM radio broadcast band.
+The entire multiplex signal is then modulated onto an actual carrier frequency of 99~MHz, that is within the FM radio broadcast band.
 This entire DSP chain is being run in the GNU Radio software, on a host PC.
-Only the final signal - FM modulated, and on a carrier of 99 MHz - is then sent to the USRP device in the form of digital signal samples.
+Only the final signal - FM modulated, and on a carrier of 99~MHz - is then sent to the USRP device in the form of digital signal samples.
 The USRP hardware then transmits the signal via its antenna.\\

### 2021-07-29 doc/thesis/chapters/dsp.tex
check all \ref{} for a unified format

@@
 This section describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
-The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly, including the de-emphasis of the signal.
+The assumption here is, that the actual FM demodulation, as described in Chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly, including the de-emphasis of the signal.
 So at this point the FM signal is converted to an AM signal.
@@
 It is then modulated to baseband by the previously generated 38~kHz subcarrier.
-Subsequently, a lowpass limits signal bandwidth to 15~kHz, which removes the generated modulation artifacts.\\
-
-The lower branch lowpass-filters the summation, or mono signal, as described in paragraph \ref{subsec:demod_mono}.\\
-
+Subsequently, a lowpass limits signal bandwidth to 15~kHz, which removes the generated modulation artifacts.
+The lower branch lowpass-filters the summation, or mono signal, as described in Section \ref{subsec:demod_mono}.
 The rightmost part in the diagram performs the combination of the mono and the difference signal, according to Equation \eqref{equ_stereo_from_sum_diff}.

### 2021-07-29 doc/thesis/chapters/implementation.tex
check all \ref{} for a unified format

@@
       It is not required, because of the following explanation.
-      At this point the sampling frequency is set to 960 kHz.
-      The input is band-limited at 75 kHz, which is the maximum frequency in an FM channel, as explained in Section \ref{sec:FrequencySpectrumOfABroadcastFmChannel}.
+      At this point the sampling frequency is set to 960~kHz.
+      The input is band-limited at 75~kHz, which is the maximum frequency in an FM channel, as explained in Section \ref{sec:FrequencySpectrumOfABroadcastFmChannel}.
       Thus, the input has an oversampling rate of 12.8, while downsampling is only performed with a factor of 8.
@@
       Several filters are implemented as FIR filters.
-      The 15 kHz lowpass (LP) filter is instantiated twice, with the exact same coefficients.
+      The 15~kHz lowpass (LP) filter is instantiated twice, with the exact same coefficients.
 
@@
   \item \textbf{Carrier Recovery}\\
-      The 38 kHz carrier can be recovered from the 19 kHz pilot tone.
+      The 38~kHz carrier can be recovered from the 19~kHz pilot tone.
       It is important that the recovered carrier is phase-coherent with the received input, as described in Section \ref{sec:demodulation-of-broadcast-fm}.
@@
 The transmitter reads an audio file, including a left and a right channel in the first step.
-The audio file contains data in a sample rate of 48 kHz.
+The audio file contains data in a sample rate of 48~kHz.
 Next, the sample rate is increased by an oversampling factor of 20, to achieve a sufficient sample rate for the subsequent DSP procedures.\\
 
-The achieved sample rate of 960 kHz could actually be chosen much lower, since an FM channels' maximum frequency is 75 kHz (see \ref{sec:FrequencySpectrumOfABroadcastFmChannel}), and according to the Nyquist theorem the sample rate thus only needs to have 150 kHz (see \ref{sec:SamplingTheorem}).
+The achieved sample rate of 960~kHz could actually be chosen much lower, since an FM channels' maximum frequency is 75~kHz (see Section \ref{sec:FrequencySpectrumOfABroadcastFmChannel}), and according to the Nyquist theorem the sample rate thus only needs to have 150~kHz (see Section \ref{sec:SamplingTheorem}).
 However, the higher rate was chosen, mainly because a higher number of signal samples beautify any plots that are created during analysis.
@@
 The next section in the block diagram represents multiple processes in parallel.
-The LR Diff part is shifted to 38 kHz in the spectrum.
-This is done by modulating the signal with its sub-carrier of 38 kHz.
+The LR Diff part is shifted to 38~kHz in the spectrum.
+This is done by modulating the signal with its sub-carrier of 38~kHz.
 A subsequent BP filter serves to remove the artifacts of the modulation, such as the image replica.
-It band-limits the spectrum of the LR Diff part between 23 and 53 kHz.
+It band-limits the spectrum of the LR Diff part between 23 and 53~kHz.
 The BP filter introduces a constant group delay of the signal.
@@
 An LP filter band-limits the signal and removes artifacts of the mixer.
-As a final step, the sample rate is decreased again, back to the chosen sampling rate of 960 kHz.
+As a final step, the sample rate is decreased again, back to the chosen sampling rate of 960~kHz.
 This signal can now be processed in a digital FM receiver.
@@
 The spectrum after the demodulation through a receiver is shown in the lower diagram.
-Spectral parts, such as the pilot tone at 19 kHz, or the LR Diff part around 38 kHz can be seen clearly.
-The demodulated spectrum shows an artifact at 57 kHz, which does not exist in the original transmitter spectrum.
+Spectral parts, such as the pilot tone at 19~kHz, or the LR Diff part around 38~kHz can be seen clearly.
+The demodulated spectrum shows an artifact at 57~kHz, which does not exist in the original transmitter spectrum.
 This is an artifact of the demodulation process, specifically the frequency-shift of the LR Diff part to baseband.
@@
 Please make sure to adapt your setup to the respective local regulations, to avoid any legal issues.
-Some of the regulations are mentioned in \ref{sec:FrequencyBand}, but may differ in various countries, however.\\
+Some of the regulations are mentioned in Section \ref{sec:FrequencyBand}, but may differ in various countries, however.\\
 
@@
 The split between the Python environment and the VHDL implementation is clearly visible but is represented from the developers perspective.
-The actual simulator interfaces, as described in \ref{sub:VHDL:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.\\
+The actual simulator interfaces, as described in Section \ref{sub:VHDL:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.\\
 
@@
 Multiple steps in the DSP chain require the operation of sample rate reduction.
-This is explained in section \ref{sec:impl:dsp-chain}.\\
+This is explained in Section \ref{sec:impl:dsp-chain}.\\
 

### 2021-07-29 doc/thesis/chapters/system-design-process.tex
check all \ref{} for a unified format

@@
 
-Furthermore, the V-model diagram in Figure \ref{fig:v_modell} shows the difference between validation and verification, as it is explained in the previous section \ref{sec:ValidationAndVerification}.
+Furthermore, the V-model diagram in Figure \ref{fig:v_modell} shows the difference between validation and verification, as it is explained in the previous Section \ref{sec:ValidationAndVerification}.
 

### 2021-07-29 doc/thesis/chapters/deployment-on-hardware.tex
fix all the 'Gedankenstriche' in sentences

@@
 Next, the DMA is configured and its data transfer mode is initiated, so that the DMA starts to transfer data from the mentioned memory area towards the FPGA.
-The DMA is configured in a mode that creates a continuous transfer, by looping through the data - once the end is reached, the transfer starts at the beginning of the memory area.
+The DMA is configured in a mode that creates a continuous transfer, by looping through the data -- once the end is reached, the transfer starts at the beginning of the memory area.
 Any data connection in the following processing chain is implemented as an AXI stream bus.

### 2021-07-29 doc/thesis/chapters/dsp.tex
fix all the 'Gedankenstriche' in sentences

@@
 For a better overview, only the positive spectum part is shown.
-The negative part is an exact mirror, because the transmitted and received signals from the antenna are real signals - they do not have an imaginary part.
+The negative part is an exact mirror, because the transmitted and received signals from the antenna are real signals -- they do not have an imaginary part.
 The spectrum consists of multiple parts.
@@
 In other words, it converts FM to AM.
-The discriminator consists of two main parts - a differentiator and an envelope detector.
+The discriminator consists of two main parts -- a differentiator and an envelope detector.
 Different strategies can be implemented for both parts.
@@
 Two parts need to be combined in order to achieve stereo.
-The sum and the difference of the left and right audio channel signals - the mono and the stereo difference part, respectively.
+The sum and the difference of the left and right audio channel signals -- the mono and the stereo difference part, respectively.
 The following equations need to be applied, to compute a left and a right channel.

### 2021-07-29 doc/thesis/chapters/hls.tex
check all \ref{} for a unified format
fix all the 'Gedankenstriche' in sentences

@@
 An overview of the most important steps in the development workflow, as well as main features and possibilities of the code implementation and testbench are given.
-The main sources of information for this chapter are found in \cite{VivadoUgHLSIntro}and \cite{VivadoUgHLS}.
+The main sources of information for this chapter are found in \cite{VivadoUgHLSIntro} and \cite{VivadoUgHLS}.
 
@@
 In software development, there are different compilers for multiple processor architectures, which can be used to port a program to another target hardware.
-The idea of HLS is to create the same level of abstraction - write code in a high-level language, which is agnostic of its target hardware architecture.\\
+The idea of HLS is to create the same level of abstraction -- write code in a high-level language, which is agnostic of its target hardware architecture.\\
 
@@
 These different levels of abstraction in the testbench execution are a major difference between a VHDL testbench and an HLS testbench.
-The most important impact thereof is the execution speed - the time it takes to simulate a number of test cases on the DUT.\\
+The most important impact thereof is the execution speed -- the time it takes to simulate a number of test cases on the DUT.\\
 

### 2021-07-29 doc/thesis/chapters/implementation.tex
fix all the 'Gedankenstriche' in sentences

@@
 The audio file contains data that represents speech with the words \textit{'Left channel... Right channel...'}.
-In terms of audio channels, the words exactly match the respective channels' output - the first part is modulated on the left audio channel only, while the second part only targets the right audio channel.
+In terms of audio channels, the words exactly match the respective channels' output -- the first part is modulated on the left audio channel only, while the second part only targets the right audio channel.
 This is particulary useful for verification, especially to verify the functionality of channel separation.
@@
 In the final chain of DSP blocks, the sample rate is increased again, to be able to frequency-modulate the signal.
-The resulting frequency modulated signal represents the signal that is received by an antenna - the RF signal of an FM channel.
+The resulting frequency modulated signal represents the signal that is received by an antenna -- the RF signal of an FM channel.
 This signal is then processed by a complex baseband mixer, which results in inphase-, and quadrature components in the baseband.
@@
 This entire DSP chain is being run in the GNU Radio software, on a host PC.
-Only the final signal - FM modulated, and on a carrier of 99~MHz - is then sent to the USRP device in the form of digital signal samples.
+Only the final signal -- FM modulated, and on a carrier of 99~MHz -- is then sent to the USRP device in the form of digital signal samples.
 The USRP hardware then transmits the signal via its antenna.\\
@@
 An FM receiver is to be implemented in multiple different versions.
-All of them need to be verified somehow - this is where the SDR transmitter created with GNU Radio and external hardware opens up an entire new possibility from the viewpoint of system level design integration.\\
+All of them need to be verified somehow -- this is where the SDR transmitter created with GNU Radio and external hardware opens up an entire new possibility from the viewpoint of system level design integration.\\
 
@@
 It can be used to verify VHDL and SystemVerilog RTL code, in combination with an external simulator.
-Various different simulators can be used with cocotb - amongst others, GHDL is supported.
+Various different simulators can be used with cocotb -- amongst others, GHDL is supported.
 The interface that is used to the simulator is the industry-standard VPI.\\
@@
 This could be implemented differently, since Python can be used to actually implement an FM receiver including the entire DSP functionality, just like in Matlab.
-Because the Matlab model already exists, this is not done here - loading the Matlab data is sufficient as it is been seen as 'the single source of truth' in this project.\\
+Because the Matlab model already exists, this is not done here -- loading the Matlab data is sufficient as it is been seen as 'the single source of truth' in this project.\\
 
@@
 Given all unit tests being successful, it is ensured that all these single units are implemented correctly.
-However, their combination and interaction may still be subject to error - this is where integration tests unfold their potential.
+However, their combination and interaction may still be subject to error -- this is where integration tests unfold their potential.
 Implementation errors in the glue logic between the units, as well as the interaction of the respective interfaces can be found in these integration tests.\\
@@
 
-In this context, there are further keywords that need to be mentioned - black box testing and white box testing.
+In this context, there are further keywords that need to be mentioned -- black box testing and white box testing.
 These are testing strategies, and describe how the DUT is treated.
@@
 The DSP chain of the IP is not fully pipelined and thus cannot consume an input sample on every clock cycle.
-This is mainly because of the FIR filter implementation - it takes multiple clock cycles to compute an output sample, before it can take a new sample.
+This is mainly because of the FIR filter implementation -- it takes multiple clock cycles to compute an output sample, before it can take a new sample.
 Ideally, this situation should be handled by the ready signal, since this is exactly what it is intended for.
@@
 The registers can be defined in a YAML-file.
-Additionally, documentation in the form of a Markdown language file is generated, and - more importantly - a C/\cplusplus\ code header, which defines all the register addresses and can be used in the application software.\\
+Additionally, documentation in the form of a Markdown language file is generated, and -- more importantly -- a C/\cplusplus\ code header, which defines all the register addresses and can be used in the application software.\\
 

### 2021-07-29 doc/thesis/chapters/introduction.tex
sample rate reduction hls
comments
results
results wip..
Fig. -> Figure
incorporate some of the feedback (wip)
fix all the 'Gedankenstriche' in sentences

@@
 
-Those major reasons lead to the choice of topic for this thesis - understanding all the single parts by themselves, while maintaining the understanding over an integrated embedded system that fulfills a task as a whole.
+Those major reasons lead to the choice of topic for this thesis -- understanding all the single parts by themselves, while maintaining the understanding over an integrated embedded system that fulfills a task as a whole.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
 \section{Objective}
@@
 
-%develop an FM receiver and explain it, so that it can be followed in a tutorial with an affordable budget in hardware.\\
+%%%%%%%%%%%%%%%%%%%%%%%
+\section{Outline}
+
+TODO
 
@@
 %                   a single project
+%                  --> design space exploration
 %
 %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FEEDBACK, Meeting 07-29-2021
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% ## Absätze
+%   -- hab jetzt alle MIT Leerzeile
+%
+% ## Outline?
+%   -- 1-2 Sätze Zusammenfassung zu jedem Chapter
+%
+% ## TITEL ändern
+%   -- design of an fm receiver at system level using different methods
+%
+%   -- different approaches designing an fm receiver at system level
+%
+%   -- without the "fm receiver" part?
+%
+% ## maximale Länge der Arbeit?
+%
+% ## PRÜFUNG
+%   -- H.G. nicht da --> 2 Möglichkeiten:  1. Oktober
+%                                          2. Zweitfach PFA (AMV?)
 %
@@
 %
+% ## unify all references (use format like in caption)
+%       - Figure 7.10
+%       - Chapter 7.3
+%       - Section 7.8
+%       - Program 3.5
+%       - etc.
+%
 % ## check all abbreviation and where they are introduced the first time
@@
 %
+% ## unify highlighted text ("Betonungen")
+%       - use 'asdf' or \textit{asdf}
+%
+%
+%
+%

### 2021-07-29 doc/thesis/chapters/system-design-process.tex
fix all the 'Gedankenstriche' in sentences

@@
 High-Level Synthes (HLS) is an FPGA design methodology that provides a high level of abstraction, and thus can be seen as a system level design method.
-Different languages are used for HLS - two well-known languages are SystemC, which is a library for \cplusplus\ to provide modelling and simulation features, and plain \cplusplus.
+Different languages are used for HLS -- two well-known languages are SystemC, which is a library for \cplusplus\ to provide modelling and simulation features, and plain \cplusplus.
 A comprehensive list of available HLS tools and languages can be found in Section 3.3 of \cite{2016FfSP}.\\

### 2021-07-30 doc/thesis/chapters/results.tex
bit smaller figure
Fig. -> Figure
sysdesign

@@
 
+%%%%%
 \subsubsection{Comparing against the Matlab reference}
@@
 There seems to be an issue in the signal separation between the left and right audio channels.
-This is visible in the analysis of the audio signal, as shown is Fig. \ref{fig:audio_output_compare_tb_vs_hw}.
+This is visible in the analysis of the audio signal, as shown is Figure \ref{fig:audio_output_compare_tb_vs_hw}.
 The signal strength of the left channel in the upper diagram is weaker than expected, while the right channel in the lower diagram is stronger.
@@
 
+%%%%%
 \subsubsection{Comparing simulation- against hardware results}
@@
 
-Again, the time limitation in the elaboration of this thesis does not allow a deeper analysis of this issue.
+Again, the time limitation in the elaboration of this thesis does not allow a deeper analysis and investigation of this issue.
 However, the end-to-end system design enables the developer to analyze an issue like this in a convenient way, in quick iterations.
@@
 
+asdf
+
 %%%%%%%%%%%%%%%%%%%%
 \subsection{Lines of Code}
+
 % of the entire project
+asdf
 
@@
 % Sim speed of x min of recorded FM signal
+%
+% VHDL:
+%          time                samples
+%   audio         sim        in      out
+%  0.1000 s    73:44 min  (96000 -> 4000)
+%  0.0100 s    09:37 min  ( 9600 ->  400)
+%  0.0025 s    02:36 min  ( 2400 ->  100)
+%
+% HLS:
 
@@
 
-% - The system level architecture that was chosen, can be applied to any DSP project. Simply replace the main IP, e.g. from FM to DAB+, QAM, etc. and the environment can run in the exact same way.
+The system design architecture that is chosen in this project covers several parts of the system.
+This includes a reference model, the development of the IP itself, including the simulation testbench, as well as the final integration into the productive system.
+Because of this structured approach, that is given by this system design, it can be applied to almost any DSP project.
+The main processing IP can simply be replaced with another implementation, while the environment does not need to be modified.
+The reason therefor is, that DSP projects almost always use streaming interfaces for their inputs and outputs.
+Also a configuration interface is commonly used, to adapt certain parameters of the IP during runtime.
 

### 2021-07-30 doc/thesis/chapters/results.tex
pie chart lines of code

@@
 A main objective of this thesis is to implement an FM radio receiver in multiple different methods.
-This includes GNU Radio, Matlab, VHDL and HLS.
+This includes GNU Radio, Matlab, VHDL and HLS, which all have their advantages and disadvantages and surface different challenges to the developer.
 The target functionality is to listen to the audio broadcast signal of an FM radio station.\\
@@
 
-asdf
+In the development of a software project, the metric for the lines of code is sometimes used to give an idea of the complexity of the project.
+The time it takes to implement a certain set of features can also give such an insight, even though it is a subjective metric, since it is influenced by the respective delopers experience.
+However, both metrics are applied to the FM Radio Receiver project.
 
@@
 
-% of the entire project
-asdf
+Several parts of the system design are taken into account to analyze the lines of code, to give an impression over the largest code parts in the project.
+
+\includepicture [0.75] [0] {Lines of code of the respective system design parts, presented in a pie chart.} {lines_of_code_pie_chart} {img/matlab/lines_of_code_pie_chart}
+
+\begin{GenericCode}
+Matlab:
+  Language     files blank comment code TotalWithoutBlank Total
+  ____________ _____ _____ _______ ____ _________________ _____
+  MATLAB          34   551     728 1451              2179  2730
+
+VHDL IP Design:
+  Language     files blank comment code TotalWithoutBlank Total
+  ____________ _____ _____ _______ ____ _________________ _____
+  VHDL            20   525     787 1431              2218  2743
+  Python           1    45      22  141               163   208
+
+VHDL Testbench:
+  Language     files blank comment code TotalWithoutBlank Total
+  ____________ _____ _____ _______ ____ _________________ _____
+  Python           5   141     141  426               567   708
+  make             2    19      18   70                88   107
+  Bourne Shell     1     5       6   16                22    27
+
+HLS IP Design:
+  Language     files blank comment code TotalWithoutBlank Total
+  ____________ _____ _____ _______ ____ _________________ _____
+  C/C++ Header    14    90     119  221               340   430
+  C++              9    97     197  195               392   489
+
+HLS Testbench:
+  Language     files blank comment code TotalWithoutBlank Total
+  ____________ _____ _____ _______ ____ _________________ _____
+  C++              1    39      44  141               185   224
+  Tcl/Tk           5    33      37  112               149   182
+  C/C++ Header     2    24      18   87               105   129
+  make             1    24      26   69                95   119
+  Python           1    19      21   45                66    85
+\end{GenericCode}
 

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.m

@@ -1,0 +1,41 @@
+%-------------------------------------------------------------------------
+% File        : lines_of_code_diagram.m
+% Author      : Michael Wurm <wurm.michael95@gmail.com>
+% Description : Create a diagram to represent the lines of code.
+%-------------------------------------------------------------------------
+
+%% Prepare environment
+clear;
+close all;
+clc;
+
+%% Settings
+dir_output_doc = "../thesis/img/matlab";
+
+
+%% Plot
+
+%values    = [2730, 2743, 208, 708, 107, 27, 430+489, 224+129, 182, 119, 85];
+%value_txt = {'Matlab','VHDL IP Design (VHDL)', 'VHDL IP Design (Python)', 'VHDL Tb (Py)','VHDL Tb (make)', 'VHDL Tb (shell)', 'HLS IP Design (C++)', 'HLS Tb (C++)','HLS Tb (Tcl)','HLS Tb (make)','HLS Tb (Py)'};
+
+values    = [2730, 2743+208, 708+107+27, 430+489, 224+129+182+119+85];
+value_txt = {'Matlab','VHDL IP Design', 'VHDL Testbench', 'HLS IP Design', 'HLS Testbench'};
+
+set(0,'defaulttextinterpreter','latex');
+
+p = pie(values);
+title('\textbf{Lines Of Code}');
+
+pText = findobj(p,'Type','text');
+percentValues = get(pText,'String');
+combinedtxt = strcat(strcat(value_txt,'~(~'),percentValues'); 
+
+for i=1:length(value_txt)
+    tmp = char(combinedtxt(i));
+    pText(i).String = [tmp(1:end-1) '\%~)'];
+end
+
+fig = gcf;
+fig.Position(3:4) = [700 500];
+
+exportgraphics(fig,sprintf("%s/%s",dir_output_doc, "lines_of_code_pie_chart.pdf"),'ContentType','vector')

### 2021-07-30 doc/thesis/chapters/results.tex
create piechart with python

@@
 
-\includepicture [0.75] [0] {Lines of code of the respective system design parts, presented in a pie chart.} {lines_of_code_pie_chart} {img/matlab/lines_of_code_pie_chart}
+\includepicture [0.75] [0] {Lines of code of the respective system design parts, presented in a pie chart.} {lines_of_code_pie_chart} {img/matlab/lines_of_code_pie_chart_py}
 

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.m

@@ -38,4 +38,4 @@ end
 fig = gcf;
 fig.Position(3:4) = [700 500];
 
-exportgraphics(fig,sprintf("%s/%s",dir_output_doc, "lines_of_code_pie_chart.pdf"),'ContentType','vector')
+exportgraphics(fig,sprintf("%s/%s",dir_output_doc, "lines_of_code_pie_chart_matlab.pdf"),'ContentType','vector')

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.py

@@ -1,0 +1,73 @@
+# -------------------------------------------------------------------------
+# File        : lines_of_code_diagram.py
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Create a diagram to represent the lines of code.
+# -------------------------------------------------------------------------
+
+import matplotlib.pyplot as plt
+
+# plt.rcParams.update({
+#    "text.usetex": True,
+#    "font.family": "sans-serif",
+#    "font.sans-serif": ["Helvetica"]
+# })
+# for Palatino and other serif fonts use:
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    #    "font.serif": ["Palatino"],
+    "font.sans-serif": ["Helvetica"]
+})
+
+
+# Make data: I have 3 groups and 7 subgroups
+group_names = ['\\textbf{Matlab}', '\\textbf{VHDL}', '\\textbf{HLS}']
+group_size = [2730,
+              2743 + 208 + 708 + 107 + 27,
+              430 + 489 + 224 + 129 + 182 + 119 + 85]
+subgroup_names = ['', 'IP Design', 'Testbench', 'IP Design', 'Testbench']
+subgroup_size = [2730, 2743 + 208, 708 + 107 + 27, 430 + 489, 224 + 129 + 182 + 119 + 85]
+# subgroup_size = [2730,
+#                 2743, 208,
+#                 708, 107, 27,
+#                 430, 489,
+#                 224, 129, 182, 119, 85]
+
+# Create colors
+color_matlab, color_vhdl, color_hls = [plt.cm.Blues, plt.cm.Reds, plt.cm.Greens]
+
+# First Ring (outside)
+fig, ax = plt.subplots()
+fig.set_size_inches(6, 4)
+ax.axis('equal')
+mypie, _ = ax.pie(group_size, radius=1.45,
+                  labels=group_names,
+                  labeldistance=0.75,
+                  colors=[
+                      color_matlab(0.6), color_vhdl(0.6), color_hls(0.6)])
+plt.setp(mypie, width=0.6, edgecolor='white')
+
+# Second Ring (Inside)
+mypie2, _ = ax.pie(subgroup_size, radius=1.3 - 0.3,
+                   labels=subgroup_names,
+                   labeldistance=0.5,
+                   colors=[
+                       color_matlab(0.5),
[…]

### 2021-07-30 doc/thesis/chapters/results.tex
re-generate

--- 2021-07-30 doc/thesis/chapters/results.tex
@@
 
-% of the entire project
-asdf
+Several parts of the system design are taken into account to analyze the lines of code, to give an impression over the largest code parts in the project.
+
+\includepicture [0.9] [0] {Lines of code of the respective system design parts, presented in a pie chart.} {lines_of_code_pie_chart} {img/matlab/lines_of_code_pie_chart_py}
+
+%\begin{GenericCode}
+%Matlab:
+%  Language     files blank comment code TotalWithoutBlank Total
+%  ____________ _____ _____ _______ ____ _________________ _____
+%  MATLAB          34   551     728 1451              2179  2730
+%
+%VHDL IP Design:
+%  Language     files blank comment code TotalWithoutBlank Total
+%  ____________ _____ _____ _______ ____ _________________ _____
+%  VHDL            20   525     787 1431              2218  2743
+%  Python           1    45      22  141               163   208
+%
+%VHDL Testbench:
+%  Language     files blank comment code TotalWithoutBlank Total
+%  ____________ _____ _____ _______ ____ _________________ _____
+%  Python           5   141     141  426               567   708
+%  make             2    19      18   70                88   107
+%  Bourne Shell     1     5       6   16                22    27
+%
+%HLS IP Design:
+%  Language     files blank comment code TotalWithoutBlank Total
+%  ____________ _____ _____ _______ ____ _________________ _____
+%  C/C++ Header    14    90     119  221               340   430
+%  C++              9    97     197  195               392   489
+%
+%HLS Testbench:
+%  Language     files blank comment code TotalWithoutBlank Total
+%  ____________ _____ _____ _______ ____ _________________ _____
+%  C++              1    39      44  141               185   224
+%  Tcl/Tk           5    33      37  112               149   182
+%  C/C++ Header     2    24      18   87               105   129
+%  make             1    24      26   69                95   119
+%  Python           1    19      21   45                66    85
+%\end{GenericCode}
 

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.py

@@ -6,26 +6,22 @@
 
 import matplotlib.pyplot as plt
 
-# plt.rcParams.update({
-#    "text.usetex": True,
-#    "font.family": "sans-serif",
-#    "font.sans-serif": ["Helvetica"]
-# })
-# for Palatino and other serif fonts use:
 plt.rcParams.update({
     "text.usetex": True,
     "font.family": "serif",
-    #    "font.serif": ["Palatino"],
     "font.sans-serif": ["Helvetica"]
 })
 
 
-# Make data: I have 3 groups and 7 subgroups
+# Create data
 group_names = ['\\textbf{Matlab}', '\\textbf{VHDL}', '\\textbf{HLS}']
 group_size = [2730,
               2743 + 208 + 708 + 107 + 27,
               430 + 489 + 224 + 129 + 182 + 119 + 85]
-subgroup_names = ['', 'IP Design', 'Testbench', 'IP Design', 'Testbench']
+subgroup_names = [' ', ' ', ' ', ' ', ' ']
+# subgroup_names = ['Matlab',
+#                       'IP Design', 'Testbench',
+#                       'IP Design', 'Testbench']
 subgroup_size = [2730, 2743 + 208, 708 + 107 + 27, 430 + 489, 224 + 129 + 182 + 119 + 85]
 # subgroup_size = [2730,
 #                 2743, 208,
@@ -42,7 +38,7 @@ fig.set_size_inches(6, 4)
 ax.axis('equal')
 mypie, _ = ax.pie(group_size, radius=1.45,
                   labels=group_names,
-                  labeldistance=0.75,
+                  labeldistance=0.72,
                   colors=[
                       color_matlab(0.6), color_vhdl(0.6), color_hls(0.6)])
 plt.setp(mypie, width=0.6, edgecolor='white')

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.py
re-order

@@ -14,15 +14,33 @@ plt.rcParams.update({
 
 
 # Create data
-group_names = ['\\textbf{Matlab}', '\\textbf{VHDL}', '\\textbf{HLS}']
-group_size = [2730,
-              2743 + 208 + 708 + 107 + 27,
-              430 + 489 + 224 + 129 + 182 + 119 + 85]
-subgroup_names = [' ', ' ', ' ', ' ', ' ']
-# subgroup_names = ['Matlab',
-#                       'IP Design', 'Testbench',
-#                       'IP Design', 'Testbench']
-subgroup_size = [2730, 2743 + 208, 708 + 107 + 27, 430 + 489, 224 + 129 + 182 + 119 + 85]
+group_names = [
+    '\\textbf{HLS}',
+    '\\textbf{Matlab}',
+    '\\textbf{VHDL}',
+]
+group_size = [
+    430 + 489 + 224 + 129 + 182 + 119 + 85,
+    2730,
+    2743 + 208 + 708 + 107 + 27,
+]
+#subgroup_names = [' ', ' ', ' ', ' ', ' ']
+subgroup_names = [
+    'IP Design', 'Testbench',
+    'Matlab',
+    'IP Design', 'Testbench',
+]
+subgroup_names_legs = [
+    'IP Design', 'Testbench',
+    'Matlab',
+    'IP Design', 'Testbench',
+]
+
+subgroup_size = [
+    430 + 489, 224 + 129 + 182 + 119 + 85,
+    2730,
+    2743 + 208, 708 + 107 + 27,
+]
 # subgroup_size = [2730,
 #                 2743, 208,
 #                 708, 107, 27,
@@ -40,7 +58,10 @@ mypie, _ = ax.pie(group_size, radius=1.45,
                   labels=group_names,
                   labeldistance=0.72,
                   colors=[
-                      color_matlab(0.6), color_vhdl(0.6), color_hls(0.6)])
+                      color_hls(0.6),
+                      color_matlab(0.6),
+                      color_vhdl(0.6),
+                  ])
 plt.setp(mypie, width=0.6, edgecolor='white')
 
 # Second Ring (Inside)
@@ -48,16 +69,14 @@ mypie2, _ = ax.pie(subgroup_size, radius=1.3 - 0.3,
                    labels=subgroup_names,
                    labeldistance=0.5,
                    colors=[
+                       color_hls(0.5), color_hls(0.4),
                        color_matlab(0.5),
[…]

### 2021-07-30 doc/thesis/chapters/results.tex
wip.....

@@
 
+\begin{figure}%
+  \centering
+  \subfloat[\centering label 1]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py} }}%
+  \\
+  %\hspace{\fill}
+  \subfloat[\centering label 2]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py} }}%
+  %\hspace{\fill}
+  \vspace{3cm}
+  \subfloat[\centering label 3]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py} }}
+
+  \caption{Multiple Figures side by side}%
+  \label{fig:example}%
+\end{figure}
+
 %\begin{GenericCode}

### 2021-07-30 doc/lines-of-code-analysis/asdf.m

@@ -1,0 +1,10 @@
+close all
+
+L = {'A','B','C','D','E'};
+X = [  1,  3,0.5,2.5,  2];
+H = pie(X);
+%
+T = H(strcmpi(get(H,'Type'),'text'));
+P = cell2mat(get(T,'Position'));
+set(T,{'Position'},num2cell(P*0.6,2))
+text(P(:,1),P(:,2),L(:))

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.m

@@ -15,16 +15,19 @@ dir_output_doc = "../thesis/img/matlab";
 
 %% Plot
 
-%values    = [2730, 2743, 208, 708, 107, 27, 430+489, 224+129, 182, 119, 85];
-%value_txt = {'Matlab','VHDL IP Design (VHDL)', 'VHDL IP Design (Python)', 'VHDL Tb (Py)','VHDL Tb (make)', 'VHDL Tb (shell)', 'HLS IP Design (C++)', 'HLS Tb (C++)','HLS Tb (Tcl)','HLS Tb (make)','HLS Tb (Py)'};
+values    = [2730, 2743, 208, 708, 107, 27, 430+489, 224+129, 182, 119, 85];
+value_txt = {'Matlab','VHDL IP Design (VHDL)', 'VHDL IP Design (Python)', 'VHDL Tb (Py)','VHDL Tb (make)', 'VHDL Tb (shell)', 'HLS IP Design (C++)', 'HLS Tb (C++)','HLS Tb (Tcl)','HLS Tb (make)','HLS Tb (Py)'};
 
 values    = [2730, 2743+208, 708+107+27, 430+489, 224+129+182+119+85];
 value_txt = {'Matlab','VHDL IP Design', 'VHDL Testbench', 'HLS IP Design', 'HLS Testbench'};
 
+%values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85];
+%value_txt = {'Matlab','VHDL', 'HLS'};
+
 set(0,'defaulttextinterpreter','latex');
 
 p = pie(values);
-title('\textbf{Lines Of Code}');
+%title('\textbf{Lines Of Code}');
 
 pText = findobj(p,'Type','text');
 percentValues = get(pText,'String');

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.py

@@ -9,36 +9,39 @@ import matplotlib.pyplot as plt
 plt.rcParams.update({
     "text.usetex": True,
     "font.family": "serif",
-    "font.sans-serif": ["Helvetica"]
+    "font.sans-serif": ["Helvetica"],
+    "font.size": 20,
+    #    "legend.fontsize": 10,
+    #    "axes.labelsize": 10,
 })
 
 
 # Create data
 group_names = [
-    '\\textbf{HLS}',
     '\\textbf{Matlab}',
+    '\\textbf{HLS}',
     '\\textbf{VHDL}',
 ]
 group_size = [
-    430 + 489 + 224 + 129 + 182 + 119 + 85,
     2730,
+    430 + 489 + 224 + 129 + 182 + 119 + 85,
     2743 + 208 + 708 + 107 + 27,
 ]
 #subgroup_names = [' ', ' ', ' ', ' ', ' ']
 subgroup_names = [
-    'IP Design', 'Testbench',
     'Matlab',
+    'IP Design', 'Testbench',
     'IP Design', 'Testbench',
 ]
 subgroup_names_legs = [
-    'IP Design', 'Testbench',
     'Matlab',
+    'IP Design', 'Testbench',
     'IP Design', 'Testbench',
 ]
 
 subgroup_size = [
-    430 + 489, 224 + 129 + 182 + 119 + 85,
     2730,
+    430 + 489, 224 + 129 + 182 + 119 + 85,
     2743 + 208, 708 + 107 + 27,
 ]
 # subgroup_size = [2730,
@@ -48,7 +51,7 @@ subgroup_size = [
 #                 224, 129, 182, 119, 85]
 
 # Create colors
-color_matlab, color_vhdl, color_hls = [plt.cm.Blues, plt.cm.Reds, plt.cm.Greens]
+color_blue, color_red, color_green = [plt.cm.Blues, plt.cm.Reds, plt.cm.Greens]
 
 # First Ring (outside)
 fig, ax = plt.subplots()
@@ -56,22 +59,22 @@ fig.set_size_inches(6, 4)
 ax.axis('equal')
 mypie, _ = ax.pie(group_size, radius=1.45,
                   labels=group_names,
-                  labeldistance=0.72,
+                  labeldistance=0.78,
                   colors=[
-                      color_hls(0.6),
-                      color_matlab(0.6),
-                      color_vhdl(0.6),
+                      color_blue(0.6),
+                      color_green(0.6),
+                      color_red(0.6),
                   ])
[…]

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.py
do it like this

@@ -10,7 +10,7 @@ plt.rcParams.update({
     "text.usetex": True,
     "font.family": "serif",
     "font.sans-serif": ["Helvetica"],
-    "font.size": 20,
+    "font.size": 16,
     #    "legend.fontsize": 10,
     #    "axes.labelsize": 10,
 })
@@ -18,9 +18,9 @@ plt.rcParams.update({
 
 # Create data
 group_names = [
-    '\\textbf{Matlab}',
-    '\\textbf{HLS}',
-    '\\textbf{VHDL}',
+    '33\%',
+    '25\%',
+    '25\%',
 ]
 group_size = [
     2730,
@@ -29,9 +29,9 @@ group_size = [
 ]
 #subgroup_names = [' ', ' ', ' ', ' ', ' ']
 subgroup_names = [
-    'Matlab',
-    'IP Design', 'Testbench',
-    'IP Design', 'Testbench',
+    '33\%',
+    '25\%', 'xx\%',
+    '25\%', 'xx\%',
 ]
 subgroup_names_legs = [
     'Matlab',
@@ -59,7 +59,7 @@ fig.set_size_inches(6, 4)
 ax.axis('equal')
 mypie, _ = ax.pie(group_size, radius=1.45,
                   labels=group_names,
-                  labeldistance=0.78,
+                  labeldistance=0.75,
                   colors=[
                       color_blue(0.6),
                       color_green(0.6),
@@ -70,7 +70,7 @@ plt.setp(mypie, width=0.6, edgecolor='white')
 # Second Ring (Inside)
 mypie2, _ = ax.pie(subgroup_size, radius=1.3 - 0.3,
                    labels=subgroup_names,
-                   labeldistance=0.45,
+                   labeldistance=0.5,
                    colors=[
                        color_blue(0.5),
                        color_green(0.5), color_green(0.4),
@@ -82,7 +82,7 @@ plt.margins(0, 0)
 plt.legend()
 handles, labels = ax.get_legend_handles_labels()
 
-ax.legend(handles[len(group_names):], subgroup_names_legs, loc=(0.9, 0.35))
+ax.legend(handles[len(group_names):], subgroup_names_legs, loc=(0.95, 0.2))
 #plt.suptitle('\\textbf{Lines Of Code}', fontsize=14)
 plt.tight_layout()
 

### 2021-07-30 doc/thesis/chapters/results.tex
do it like this
all

--- 2021-07-30 doc/thesis/chapters/results.tex
@@
 
+\begin{figure}%
+  \centering
+  \subfloat[\centering label 1]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py_all} }}%
+  \\
+  %\hspace{\fill}
+  \subfloat[\centering label 2]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py_hls} }}%
+  %\hspace{\fill}
+  %\vspace{3cm}
+  \subfloat[\centering label 3]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py_hls} }}%
+
+  \caption{Multiple Figures side by side}%
+  \label{fig:example}%
+\end{figure}
+
 %\begin{GenericCode}

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.m

@@ -21,8 +21,8 @@ value_txt = {'Matlab','VHDL IP Design (VHDL)', 'VHDL IP Design (Python)', 'VHDL 
 values    = [2730, 2743+208, 708+107+27, 430+489, 224+129+182+119+85];
 value_txt = {'Matlab','VHDL IP Design', 'VHDL Testbench', 'HLS IP Design', 'HLS Testbench'};
 
-%values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85];
-%value_txt = {'Matlab','VHDL', 'HLS'};
+values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85];
+value_txt = {'Matlab','VHDL', 'HLS'};
 
 set(0,'defaulttextinterpreter','latex');
 

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.py

@@ -16,6 +16,66 @@ plt.rcParams.update({
 })
 
 
+# -------------------------------------------------------------------------
+# Create data
+group_names = [
+    '33\%',
+    '20\%',
+    '47\%',
+]
+group_size = [
+    2730,
+    430 + 489 + 224 + 129 + 182 + 119 + 85,
+    2743 + 208 + 708 + 107 + 27,
+]
+subgroup_names = [
+    '33\%',
+    '25\%',
+    '25\%',
+]
+subgroup_names_legs = [
+    'Matlab',
+    'HLS',
+    'VHDL',
+]
+
+subgroup_size = [
+    2730,
+    430 + 489 + 224 + 129 + 182 + 119 + 85,
+    2743 + 208 + 708 + 107 + 27,
+]
+
+# Create colors
+color_blue, color_red, color_green = [plt.cm.Blues, plt.cm.Reds, plt.cm.Greens]
+
+# First Ring (outside)
+fig, ax = plt.subplots()
+fig.set_size_inches(6, 4)
+ax.axis('equal')
+mypie, _ = ax.pie(group_size, radius=1.45,
+                  labels=group_names,
+                  labeldistance=0.6,
+                  colors=[
+                      color_blue(0.6),
+                      color_green(0.6),
+                      color_red(0.6),
+                  ])
+plt.setp(mypie, width=0.9, edgecolor='white')
+
+
+plt.margins(0, 0)
+plt.legend()
+handles, labels = ax.get_legend_handles_labels()
+
+ax.legend(handles[:], subgroup_names_legs, loc=(0.95, 0.3))
+#plt.suptitle('\\textbf{Lines Of Code}', fontsize=14)
+plt.tight_layout()
+
+# show and save
+plt.savefig('../thesis/img/matlab/lines_of_code_pie_chart_py_all.pdf')
+plt.show()
+
+# -------------------------------------------------------------------------
 # Create data
 group_names = [
     '33\%',
@@ -87,5 +147,81 @@ ax.legend(handles[len(group_names):], subgroup_names_legs, loc=(0.95, 0.2))
 plt.tight_layout()
 
 # show and save
-plt.savefig('../thesis/img/matlab/lines_of_code_pie_chart_py.pdf')
+plt.savefig('../thesis/img/matlab/lines_of_code_pie_chart_py_hls.pdf')
+plt.show()
+
+# -------------------------------------------------------------------------
+
+# Create data
+group_names = [
[…]

### 2021-07-30 doc/thesis/chapters/results.tex
work on LOC plots.. (still wip...)

--- 2021-07-30 doc/thesis/chapters/results.tex
@@
 
-\includepicture [0.9] [0] {Lines of code of the respective system design parts, presented in a pie chart.} {lines_of_code_pie_chart} {img/matlab/lines_of_code_pie_chart_py}
+\begin{figure}%
+  \centering
+  \subfloat[\centering Overview of ]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py_all} }}%
+  \\
+  %\hspace{\fill}
+  \subfloat[\centering HLS]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py_hls} }}%
+  %\hspace{\fill}
+  %\vspace{3cm}
+  \subfloat[\centering VHDL]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py_vhdl} }}%
+
+  \caption{Lines of code of the respective system design parts, presented in multiple pie charts.}%
+  \label{fig:lines_of_code_pie_chart}%
+\end{figure}
 

### 2021-07-30 doc/lines-of-code-analysis/asdf.m

@@ -1,10 +1,0 @@
-close all
-
-L = {'A','B','C','D','E'};
-X = [  1,  3,0.5,2.5,  2];
-H = pie(X);
-%
-T = H(strcmpi(get(H,'Type'),'text'));
-P = cell2mat(get(T,'Position'));
-set(T,{'Position'},num2cell(P*0.6,2))
-text(P(:,1),P(:,2),L(:))

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.m

@@ -15,14 +15,25 @@ dir_output_doc = "../thesis/img/matlab";
 
 %% Plot
 
-values    = [2730, 2743, 208, 708, 107, 27, 430+489, 224+129, 182, 119, 85];
-value_txt = {'Matlab','VHDL IP Design (VHDL)', 'VHDL IP Design (Python)', 'VHDL Tb (Py)','VHDL Tb (make)', 'VHDL Tb (shell)', 'HLS IP Design (C++)', 'HLS Tb (C++)','HLS Tb (Tcl)','HLS Tb (make)','HLS Tb (Py)'};
+% details
+%values    = [2730, 2743, 208, 708, 107, 27, 430+489, 224+129, 182, 119, 85];
+%value_txt = {'Matlab','VHDL IP Design (VHDL)', 'VHDL IP Design (Python)', 'VHDL Tb (Py)','VHDL Tb (make)', 'VHDL Tb (shell)', 'HLS IP Design (C++)', 'HLS Tb (C++)','HLS Tb (Tcl)','HLS Tb (make)','HLS Tb (Py)'};
 
-values    = [2730, 2743+208, 708+107+27, 430+489, 224+129+182+119+85];
-value_txt = {'Matlab','VHDL IP Design', 'VHDL Testbench', 'HLS IP Design', 'HLS Testbench'};
+% hls, tb/design
+%values    = [430+489, 224+129+182+119+85];
+%value_txt = {'HLS IP Design', 'HLS Testbench'};
 
-values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85];
-value_txt = {'Matlab','VHDL', 'HLS'};
+% vhdl, tb/design
+%values    = [2743+208, 708+107+27];
+%value_txt = {'VHDL IP Design', 'VHDL Testbench'};
+
+% all, tb/design
+%values    = [2730, 2743+208, 708+107+27, 430+489, 224+129+182+119+85];
+%value_txt = {'Matlab','VHDL IP Design', 'VHDL Testbench', 'HLS IP Design', 'HLS Testbench'};
+
+% all
+%values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85];
+%value_txt = {'Matlab','VHDL', 'HLS'};
 
 set(0,'defaulttextinterpreter','latex');
 

### 2021-07-30 doc/lines-of-code-analysis/lines_of_code_diagram.py

@@ -16,7 +16,7 @@ plt.rcParams.update({
 })
 
 
-# -------------------------------------------------------------------------
+# ------------------------- ALL -----------------------------------------------
 # Create data
 group_names = [
     '33\%',
@@ -28,21 +28,10 @@ group_size = [
     430 + 489 + 224 + 129 + 182 + 119 + 85,
     2743 + 208 + 708 + 107 + 27,
 ]
-subgroup_names = [
-    '33\%',
-    '25\%',
-    '25\%',
-]
-subgroup_names_legs = [
+group_names_legs = [
     'Matlab',
     'HLS',
     'VHDL',
-]
-
-subgroup_size = [
-    2730,
-    430 + 489 + 224 + 129 + 182 + 119 + 85,
-    2743 + 208 + 708 + 107 + 27,
 ]
 
 # Create colors
@@ -62,12 +51,11 @@ mypie, _ = ax.pie(group_size, radius=1.45,
                   ])
 plt.setp(mypie, width=0.9, edgecolor='white')
 
-
 plt.margins(0, 0)
 plt.legend()
 handles, labels = ax.get_legend_handles_labels()
 
-ax.legend(handles[:], subgroup_names_legs, loc=(0.95, 0.3))
+ax.legend(handles[:], group_names_legs, loc=(0.95, 0.3))
 #plt.suptitle('\\textbf{Lines Of Code}', fontsize=14)
 plt.tight_layout()
 
@@ -75,40 +63,42 @@ plt.tight_layout()
 plt.savefig('../thesis/img/matlab/lines_of_code_pie_chart_py_all.pdf')
 plt.show()
 
-# -------------------------------------------------------------------------
+# ---------------------- HLS --------------------------------------------------
 # Create data
 group_names = [
-    '33\%',
-    '25\%',
-    '25\%',
+    'IP Design',
+    'Testbench',
+]
+group_names_legs = [
+    'IP Design',
+    'Testbench',
 ]
 group_size = [
-    2730,
-    430 + 489 + 224 + 129 + 182 + 119 + 85,
-    2743 + 208 + 708 + 107 + 27,
-]
-#subgroup_names = [' ', ' ', ' ', ' ', ' ']
-subgroup_names = [
-    '33\%',
-    '25\%', 'xx\%',
-    '25\%', 'xx\%',
-]
-subgroup_names_legs = [
-    'Matlab',
-    'IP Design', 'Testbench',
[…]

### 2021-07-31 doc/lines-of-code-analysis/lines_of_code_diagram.m
vhdl plot

@@ -23,6 +23,10 @@ dir_output_doc = "../thesis/img/matlab";
 %values    = [430+489, 224+129+182+119+85];
 %value_txt = {'HLS IP Design', 'HLS Testbench'};
 
+% hls, tb
+values    = [224+129,182,119,85];
+value_txt = {'cpp','Tcl','make','Python'};
+
 % vhdl, tb/design
 %values    = [2743+208, 708+107+27];
 %value_txt = {'VHDL IP Design', 'VHDL Testbench'};

### 2021-07-31 doc/lines-of-code-analysis/lines_of_code_diagram.py

@@ -79,11 +79,11 @@ group_size = [
 ]
 
 subgroup_names = [
-    '33\%',
-    '33\%',
+    '100\%',
+    '48\%',
     '25\%',
-    'xx\%',
-    'xx\%',
+    '16\%',
+    '12\%',
 ]
 subgroup_names_legs = [
     'C\\texttt{++}',
@@ -109,21 +109,27 @@ fig.set_size_inches(6, 4)
 ax.axis('equal')
 mypie, _ = ax.pie(group_size, radius=1.45,
                   labels=group_names,
-                  labeldistance=0.75,
+                  labeldistance=0.82,
                   colors=[
                       color_blue(0.6),
                       color_red(0.6),
-                  ])
+                  ],
+                  rotatelabels=False,
+                  textprops=dict(rotation_mode='anchor', va='center', ha='center')
+                  )
 plt.setp(mypie, width=0.6, edgecolor='white')
 
 # Second Ring (Inside)
 mypie2, _ = ax.pie(subgroup_size, radius=1.3 - 0.3,
                    labels=subgroup_names,
-                   labeldistance=0.5,
+                   labeldistance=0.7,
                    colors=[
                        color_blue(0.5),
                        color_red(0.5), color_red(0.4), color_red(0.3), color_red(0.2),
-                   ])
+                   ],
+                   rotatelabels=True,
+                   textprops=dict(rotation_mode='anchor', va='center', ha='center')
+                   )
 plt.setp(mypie2, width=0.6, edgecolor='white')
 plt.margins(0, 0)
 

### 2021-07-31 doc/thesis/chapters/results.tex
vhdl now actually (prev. was HLS)

@@
   \centering
-  \subfloat[\centering Overview of ]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py_all} }}%
+  \subfloat[\centering Overview of ]{{\includegraphics[trim=10 10 10 10,clip, width=0.5\textwidth]{img/matlab/lines_of_code_pie_chart_py_all} }}%
   \\
   %\hspace{\fill}
-  \subfloat[\centering HLS]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py_hls} }}%
+  \subfloat[\centering HLS]{{\includegraphics[trim=10 10 10 10,clip, width=0.5\textwidth]{img/matlab/lines_of_code_pie_chart_py_hls} }}%
   %\hspace{\fill}
   %\vspace{3cm}
-  \subfloat[\centering VHDL]{{\includegraphics[trim=10 10 10 10,clip, width=0.45\textwidth]{img/matlab/lines_of_code_pie_chart_py_vhdl} }}%
+  \subfloat[\centering VHDL]{{\includegraphics[trim=10 10 10 10,clip, width=0.5\textwidth]{img/matlab/lines_of_code_pie_chart_py_vhdl} }}%
 
@@
 %\end{GenericCode}
+
+Also note, that the common Python testbench consists of XXX lines of Python code.
+Therefor, the testbench code of the respective IPs is relatively small, because the entire analysis part is shared.
 

### 2021-07-31 doc/lines-of-code-analysis/lines_of_code_diagram.m

@@ -24,12 +24,20 @@ dir_output_doc = "../thesis/img/matlab";
 %value_txt = {'HLS IP Design', 'HLS Testbench'};
 
 % hls, tb
-values    = [224+129,182,119,85];
-value_txt = {'cpp','Tcl','make','Python'};
+%values    = [224+129,182,119,85];
+%value_txt = {'cpp','Tcl','make','Python'};
 
 % vhdl, tb/design
 %values    = [2743+208, 708+107+27];
 %value_txt = {'VHDL IP Design', 'VHDL Testbench'};
+
+% vhdl, tb
+values    = [708, 107, 27];
+value_txt = {'Python','make','shell'};
+
+% vhdl, design
+%values    = [2743,208];
+%value_txt = {'vhdl','Python'};
 
 % all, tb/design
 %values    = [2730, 2743+208, 708+107+27, 430+489, 224+129+182+119+85];

### 2021-07-31 doc/lines-of-code-analysis/lines_of_code_diagram.py

@@ -115,7 +115,8 @@ mypie, _ = ax.pie(group_size, radius=1.45,
                       color_red(0.6),
                   ],
                   rotatelabels=False,
-                  textprops=dict(rotation_mode='anchor', va='center', ha='center')
+                  textprops=dict(rotation_mode='anchor', va='center', ha='center'),
+                  startangle=-10
                   )
 plt.setp(mypie, width=0.6, edgecolor='white')
 
@@ -128,7 +129,8 @@ mypie2, _ = ax.pie(subgroup_size, radius=1.3 - 0.3,
                        color_red(0.5), color_red(0.4), color_red(0.3), color_red(0.2),
                    ],
                    rotatelabels=True,
-                   textprops=dict(rotation_mode='anchor', va='center', ha='center')
+                   textprops=dict(rotation_mode='anchor', va='center', ha='center'),
+                   startangle=-10
                    )
 plt.setp(mypie2, width=0.6, edgecolor='white')
 plt.margins(0, 0)
@@ -156,31 +158,30 @@ group_names_legs = [
     'Testbench',
 ]
 group_size = [
-    # TODO HERE
-    430 + 489,
-    224 + 129 + 182 + 119 + 85,
+    2743 + 208,
+    708 + 107 + 27,
 ]
 
 subgroup_names = [
-    '33\%',
-    '33\%',
-    '25\%',
-    'xx\%',
-    'xx\%',
+    '93\%',
+    '7\%',
+    '84\%',
+    '13\%',
+    '',
 ]
 subgroup_names_legs = [
-    'C\\texttt{++}',
-    'C\\texttt{++}',
-    'Tcl',
-    'make',
+    'VHDL',
     'Python',
+    'Python',
+    'make',
+    'shell',
 ]
 subgroup_size = [
-    430 + 489,  # C++ IP
-    224 + 129,  # C++ Tb
-    182,        # Tcl Tb
-    119,        # make Tb
-    85,         # python Tb
+    2743,  # Tb:     VHDL
+    208,   # Tb:     Python
+    708,   # Design: Python
+    107,   # Design: make
+    27,    # Design: shell
 ]
 
 # Create colors
@@ -192,21 +193,29 @@ fig.set_size_inches(6, 4)
 ax.axis('equal')
 mypie, _ = ax.pie(group_size, radius=1.45,
                   labels=group_names,
-                  labeldistance=0.75,
+                  labeldistance=0.82,
                   colors=[
                       color_blue(0.6),
[…]

### 2021-07-31 doc/thesis/chapters/results.tex
table

@@
   \centering
-  \subfloat[\centering Overview of ]{{\includegraphics[trim=10 10 10 10,clip, width=0.5\textwidth]{img/matlab/lines_of_code_pie_chart_py_all} }}%
+  \subfloat[\centering Overview of the proportions between the Matlab model, the HLS and VHDL IPs.]{{\includegraphics[trim=10 10 10 10,clip, width=0.5\textwidth]{img/matlab/lines_of_code_pie_chart_py_all} }}%
   \\
   %\hspace{\fill}
-  \subfloat[\centering HLS]{{\includegraphics[trim=10 10 10 10,clip, width=0.5\textwidth]{img/matlab/lines_of_code_pie_chart_py_hls} }}%
+  \subfloat[\centering HLS in detail.]{{\includegraphics[trim=10 10 10 10,clip, width=0.49\textwidth]{img/matlab/lines_of_code_pie_chart_py_hls} }}%
   %\hspace{\fill}
   %\vspace{3cm}
-  \subfloat[\centering VHDL]{{\includegraphics[trim=10 10 10 10,clip, width=0.5\textwidth]{img/matlab/lines_of_code_pie_chart_py_vhdl} }}%
+  \subfloat[\centering VHDL in detail.]{{\includegraphics[trim=10 10 10 10,clip, width=0.49\textwidth]{img/matlab/lines_of_code_pie_chart_py_vhdl} }}%
 
-  \caption{Lines of code of the respective system design parts, presented in multiple pie charts.}%
+  \caption{Lines of code of the respective system design parts, presented in multiple pie charts.
+The percentages of code in IP design and testbench for the HLS and VHDL implementations are shown in detail.}%
   \label{fig:lines_of_code_pie_chart}%
 \end{figure}
+
+\begin{table}[]
+  \centering
+  \begin{tabular}{l|c|c|}
+  \cline{2-3}
+  \textbf{}                                     & \multicolumn{1}{r|}{\textbf{IP Design}} & \multicolumn{1}{l|}{\textbf{Testbench}} \\ \hline
+  \multicolumn{1}{|l|}{VHDL IP}                 & 3000                                    & 850                                     \\ \hline
+  \multicolumn{1}{|l|}{HLS IP}                  & 920                                     & 740                                     \\ \hline
+  \multicolumn{1}{|l|}{Common Testbench}        & \multicolumn{2}{c|}{630}                                                          \\ \hline
+  \multicolumn{1}{|l|}{Matlab Model}            & \multicolumn{2}{c|}{2730}                                                         \\ \hline
+  \multicolumn{1}{|l|}{Vivado Scripts}          & \multicolumn{2}{c|}{490}                                                          \\ \hline
+  \multicolumn{1}{|l|}{\cplusplus\ Application} & \multicolumn{2}{c|}{2800}                                                         \\ \hline
+  \end{tabular}
+  \caption{asdf}
+  \label{tab:lines_of_code}
+\end{table}
 

### 2021-07-31 doc/lines-of-code-analysis/count_lines_of_code.sh

@@ -136,7 +136,7 @@ REPORT_FILE=$SCRIPT_PATH_RELATIVE/vivado_sdk_firmware
 OPTIONS="hardware/vivado/                        \
         --by-file-by-lang                        \
         --match-d=(sdk|scripts)                  \
-        --not-match-f=(upgrade_ips|wav_to_txt|synth|create_project) \
+        --not-match-f=(upgrade_ips|wav_to_txt|synth|create_project|.cproject|.project) \
         --ignored=/tmp/${REPORT_FILE}_ignored.txt"
 
 create_report;

### 2021-07-31 doc/lines-of-code-analysis/lines_of_code_diagram.m

@@ -32,8 +32,8 @@ dir_output_doc = "../thesis/img/matlab";
 %value_txt = {'VHDL IP Design', 'VHDL Testbench'};
 
 % vhdl, tb
-values    = [708, 107, 27];
-value_txt = {'Python','make','shell'};
+%values    = [708, 107, 27];
+%value_txt = {'Python','make','shell'};
 
 % vhdl, design
 %values    = [2743,208];
@@ -44,8 +44,8 @@ value_txt = {'Python','make','shell'};
 %value_txt = {'Matlab','VHDL IP Design', 'VHDL Testbench', 'HLS IP Design', 'HLS Testbench'};
 
 % all
-%values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85];
-%value_txt = {'Matlab','VHDL', 'HLS'};
+%values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85, 459];
+%value_txt = {'Matlab','VHDL', 'HLS', 'Common Tb'};
 
 set(0,'defaulttextinterpreter','latex');
 

### 2021-07-31 doc/thesis/chapters/results.tex
scale text better to fit in pie

@@
 
-In the development of a software project, the metric for the lines of code is sometimes used to give an idea of the complexity of the project.
-The time it takes to implement a certain set of features can also give such an insight, even though it is a subjective metric, since it is influenced by the respective delopers experience.
+In the development of a software project, the metric for the lines of code is sometimes used to give an idea of the scale, or volume of the project.
+The time it takes to implement a certain set of features can also give such an insight and a link to the project's complexity.
+It is a subjective metric though, since it is influenced by the respective developers experience in the specific application.
 However, both metrics are applied to the FM Radio Receiver project.
@@
 Several parts of the system design are taken into account to analyze the lines of code, to give an impression over the largest code parts in the project.
+Figure \ref{fig:lines_of_code_pie_charts} shows multiple pie charts, which represent the proportions of the lines of code in the respective system design parts.
 
@@
 The percentages of code in IP design and testbench for the HLS and VHDL implementations are shown in detail.}%
-  \label{fig:lines_of_code_pie_chart}%
+  \label{fig:lines_of_code_pie_charts}%
 \end{figure}

### 2021-07-31 doc/lines-of-code-analysis/lines_of_code_diagram.m

@@ -28,8 +28,8 @@ dir_output_doc = "../thesis/img/matlab";
 %value_txt = {'cpp','Tcl','make','Python'};
 
 % vhdl, tb/design
-%values    = [2743+208, 708+107+27];
-%value_txt = {'VHDL IP Design', 'VHDL Testbench'};
+values    = [2743+208, 708+107+27];
+value_txt = {'VHDL IP Design', 'VHDL Testbench'};
 
 % vhdl, tb
 %values    = [708, 107, 27];

### 2021-07-31 doc/lines-of-code-analysis/lines_of_code_diagram.py

@@ -66,12 +66,8 @@ plt.show()
 # ---------------------- HLS --------------------------------------------------
 # Create data
 group_names = [
-    'IP Design',
-    'Testbench',
-]
-group_names_legs = [
-    'IP Design',
-    'Testbench',
+    'IP Design (55\%)',
+    'Testbench (45\%)',
 ]
 group_size = [
     430 + 489,
@@ -109,7 +105,7 @@ fig.set_size_inches(6, 4)
 ax.axis('equal')
 mypie, _ = ax.pie(group_size, radius=1.45,
                   labels=group_names,
-                  labeldistance=0.82,
+                  labeldistance=0.78,
                   colors=[
                       color_blue(0.6),
                       color_red(0.6),
@@ -150,12 +146,8 @@ plt.show()
 # ---------------------- VHDL --------------------------------------------------
 # Create data
 group_names = [
-    'IP Design',
-    'Testbench',
-]
-group_names_legs = [
-    'IP Design',
-    'Testbench',
+    'IP Design (78\%)',
+    'Testbench (22\%)',
 ]
 group_size = [
     2743 + 208,
@@ -193,14 +185,14 @@ fig.set_size_inches(6, 4)
 ax.axis('equal')
 mypie, _ = ax.pie(group_size, radius=1.45,
                   labels=group_names,
-                  labeldistance=0.82,
+                  labeldistance=0.78,
                   colors=[
                       color_blue(0.6),
                       color_red(0.6),
                   ],
                   rotatelabels=False,
                   textprops=dict(rotation_mode='anchor', va='center', ha='center'),
-                  startangle=-45
+                  startangle=-49
                   )
 plt.setp(mypie, width=0.6, edgecolor='white')
 
@@ -214,7 +206,7 @@ mypie2, _ = ax.pie(subgroup_size, radius=1.3 - 0.3,
                    ],
                    rotatelabels=True,
                    textprops=dict(rotation_mode='anchor', va='center', ha='center'),
-                   startangle=-45
+                   startangle=-49
                    )
 plt.setp(mypie2, width=0.6, edgecolor='white')
 plt.margins(0, 0)

### 2021-07-31 doc/thesis/chapters/results.tex
text

@@
 The time it takes to implement a certain set of features can also give such an insight and a link to the project's complexity.
-It is a subjective metric though, since it is influenced by the respective developers experience in the specific application.
-However, both metrics are applied to the FM Radio Receiver project.
+It is a subjective metric though, since it is influenced by the respective developer's experience in the specific application.
+However, both metrics are applied to the FM Radio Receiver project.\\
+
+At this point, it is important to note that the results that presented in this section are very specific to the chosen architecture, testbench framework, and the implementation style generally.
+Therefore, the numbers that are presented here may vary widely by modifying any of these factors, i.e. using a different testbench framework for VHDL.
+Thus, a generalization in the art of \textit{HLS always requires less code than VHDL} can not be stated.
+However, the results in this section can give an idea and overview of this topic and provide a detailed comparison for this specific project.
 
@@
 Several parts of the system design are taken into account to analyze the lines of code, to give an impression over the largest code parts in the project.
-Figure \ref{fig:lines_of_code_pie_charts} shows multiple pie charts, which represent the proportions of the lines of code in the respective system design parts.
+Figure \ref{fig:lines_of_code_pie_charts} shows multiple pie charts, which represent the proportions of the lines of code in the respective system design parts.\\
+
+Interesting details can be found by analyzing the charts.
+The upper chart reveals that the VHDL implementation with its accompanying testbench requires more than twice the amount of code, compared to the HLS version.
+The lower charts also highlight several differences between HLS and VHDL.
+The code proportions between IP design and testbench show, that HLS requires about the same amount for these parts, while VHDL requires the largest part of code in the IP design.
 
@@
   \textbf{}                                     & \multicolumn{1}{r|}{\textbf{IP Design}} & \multicolumn{1}{l|}{\textbf{Testbench}} \\ \hline
-  \multicolumn{1}{|l|}{VHDL IP}                 & 3000                                    & 850                                     \\ \hline
-  \multicolumn{1}{|l|}{HLS IP}                  & 920                                     & 740                                     \\ \hline
+  \multicolumn{1}{|l|}{VHDL}                 & 3000                                    & 850                                     \\ \hline
+  \multicolumn{1}{|l|}{HLS}                  & 920                                     & 740                                     \\ \hline
   \multicolumn{1}{|l|}{Common Testbench}        & \multicolumn{2}{c|}{630}                                                          \\ \hline

### 2021-08-03 doc/thesis/chapters/results.tex
better pie chart, again...

@@
 
-In the development of a software project, the metric for the lines of code is sometimes used to give an idea of the scale, or volume of the project.
-The time it takes to implement a certain set of features can also give such an insight and a link to the project's complexity.
-It is a subjective metric though, since it is influenced by the respective developer's experience in the specific application.
+In the development of a software project, the metric for the lines of code can be used to give an idea of the scale, or volume of the project.
+The time it takes to implement a certain set of features can also give such an insight and a hint of the project's complexity.
+It is a subjective metric though, since it is heavily influenced by the respective developer's experience in the specific application.
 However, both metrics are applied to the FM Radio Receiver project.\\
 
-At this point, it is important to note that the results that presented in this section are very specific to the chosen architecture, testbench framework, and the implementation style generally.
+At this point, it is important to note that the results that are presented in this section are very specific to the chosen architecture, testbench framework, and the implementation style generally.
 Therefore, the numbers that are presented here may vary widely by modifying any of these factors, i.e. using a different testbench framework for VHDL.
@@
   \centering
-  \subfloat[\centering Overview of the proportions between the Matlab model, the HLS and VHDL IPs.]{{\includegraphics[trim=10 10 10 10,clip, width=0.5\textwidth]{img/matlab/lines_of_code_pie_chart_py_all} }}%
+  \subfloat[\centering Overview of the proportions between all the implemented code parts of the entire project.]{{\includegraphics[trim=10 40 10 10,clip, width=0.65\textwidth]{img/matlab/lines_of_code_pie_chart_py_all} }}%
   \\

### 2021-08-03 doc/lines-of-code-analysis/lines_of_code_diagram.m

@@ -28,8 +28,8 @@ dir_output_doc = "../thesis/img/matlab";
 %value_txt = {'cpp','Tcl','make','Python'};
 
 % vhdl, tb/design
-values    = [2743+208, 708+107+27];
-value_txt = {'VHDL IP Design', 'VHDL Testbench'};
+%values    = [2743+208, 708+107+27];
+%value_txt = {'VHDL IP Design', 'VHDL Testbench'};
 
 % vhdl, tb
 %values    = [708, 107, 27];
@@ -44,8 +44,11 @@ value_txt = {'VHDL IP Design', 'VHDL Testbench'};
 %value_txt = {'Matlab','VHDL IP Design', 'VHDL Testbench', 'HLS IP Design', 'HLS Testbench'};
 
 % all
-%values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85, 459];
-%value_txt = {'Matlab','VHDL', 'HLS', 'Common Tb'};
+%values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85, 490, 2800, 460];
+%value_txt = {'Matlab','VHDL', 'HLS', 'Vivado Scripts', 'C++ Application', 'Common Tb'};
+
+values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85, 2800, 490, 460];
+value_txt = {'Matlab','VHDL', 'HLS', 'C++ Application', 'Vivado Scripts', 'Common Tb'};
 
 set(0,'defaulttextinterpreter','latex');
 

### 2021-08-03 doc/lines-of-code-analysis/lines_of_code_diagram.py

@@ -19,23 +19,37 @@ plt.rcParams.update({
 # ------------------------- ALL -----------------------------------------------
 # Create data
 group_names = [
-    '33\%',
-    '20\%',
-    '47\%',
+    '32\%',
+    '23\%',
+    '23\%',
+    '14\%',
+    '4\%',
+    '4\%',
 ]
 group_size = [
+    2743 + 208 + 708 + 107 + 27,
     2730,
     430 + 489 + 224 + 129 + 182 + 119 + 85,
-    2743 + 208 + 708 + 107 + 27,
+    2800,
+    490,
+    460
 ]
 group_names_legs = [
+    'VHDL',
     'Matlab',
     'HLS',
-    'VHDL',
+    'C++ Application',
+    'Common Testbench',
+    'Vivado Scripts',
 ]
 
 # Create colors
-color_blue, color_red, color_green = [plt.cm.Blues, plt.cm.Reds, plt.cm.Greens]
+color_blue = plt.cm.Blues
+color_red = plt.cm.Reds
+color_green = plt.cm.Greens
+color_purples = plt.cm.Purples
+color_oranges = plt.cm.Oranges
+color_greys = plt.cm.Greys
 
 # First Ring (outside)
 fig, ax = plt.subplots()
@@ -43,19 +57,27 @@ fig.set_size_inches(6, 4)
 ax.axis('equal')
 mypie, _ = ax.pie(group_size, radius=1.45,
                   labels=group_names,
-                  labeldistance=0.6,
+                  labeldistance=0.65,
                   colors=[
                       color_blue(0.6),
                       color_green(0.6),
                       color_red(0.6),
-                  ])
-plt.setp(mypie, width=0.9, edgecolor='white')
+                      color_purples(0.6),
+                      color_oranges(0.6),
+                      color_greys(0.4),
+                  ],
+                  textprops=dict(rotation_mode='anchor', va='center', ha='center', size=13),
+                  startangle=10,
+                  # frame=True,
+                  )
+plt.setp(mypie, width=1.45 - 0.45, edgecolor='white')
+# fig.subplots_adjust(0.15, 0.0, 0.6, 1.0)  # left, bottom, right, top
 
 plt.margins(0, 0)
 plt.legend()
 handles, labels = ax.get_legend_handles_labels()
 
-ax.legend(handles[:], group_names_legs, loc=(0.95, 0.3))
[…]

### 2021-08-03 doc/thesis/chapters/results.tex
text

@@
 
-Interesting details can be found by analyzing the charts.
-The upper chart reveals that the VHDL implementation with its accompanying testbench requires more than twice the amount of code, compared to the HLS version.
-The lower charts also highlight several differences between HLS and VHDL.
-The code proportions between IP design and testbench show, that HLS requires about the same amount for these parts, while VHDL requires the largest part of code in the IP design.
+Interesting details can be found by analyzing these charts.
+The upper chart shows that the implementation of the hardware support, in the form of Vivado project scripts and the \cplusplus\ application software, takes a large part of the code with 27\%.
+The remaining majority of the code belongs to the development of the IPs.
+This includes the implementation of the Matlab model, the VHDL and HLS code, as well as the common testbench.
+The chart also reveals that the VHDL implementation with its accompanying testbench requires more than twice the amount of code, compared to the HLS version, i.e. 32\% versus 14\%.\\
+
+The lower charts highlight several differences between HLS and VHDL.
+Specifically, the code proportions between IP design and testbench are pointed out.
+The charts show that HLS requires about the same amount of code for IP design and testbench.
+VHDL requires the largest part of code in the IP design, while the testbench code amount is relatively small.
 
@@
 
-  \caption{Lines of code of the respective system design parts, presented in multiple pie charts.
-The percentages of code in IP design and testbench for the HLS and VHDL implementations are shown in detail.}%
+  \caption{Lines of code of the respective code parts in the entire project, presented in multiple pie charts.
+The percentages of code in IP design and testbench for the HLS and VHDL implementations are shown in detail in the lower diagrams.}%
   \label{fig:lines_of_code_pie_charts}%

### 2021-08-03 doc/lines-of-code-analysis/lines_of_code_diagram.py

@@ -29,16 +29,16 @@ group_names = [
 group_size = [
     2743 + 208 + 708 + 107 + 27,
     2730,
-    430 + 489 + 224 + 129 + 182 + 119 + 85,
     2800,
+    430 + 489 + 224 + 129 + 182 + 119 + 85,
     490,
     460
 ]
 group_names_legs = [
     'VHDL',
     'Matlab',
-    'HLS',
     'C++ Application',
+    'HLS',
     'Common Testbench',
     'Vivado Scripts',
 ]

### 2021-08-03 doc/thesis/chapters/results.tex
loc

@@
 Several parts of the system design are taken into account to analyze the lines of code, to give an impression over the largest code parts in the project.
-Figure \ref{fig:lines_of_code_pie_charts} shows multiple pie charts, which represent the proportions of the lines of code in the respective system design parts.\\
+Figure \ref{fig:lines_of_code_pie_charts} shows multiple pie charts, which represent the proportions of the lines of code in the respective system design parts.
+Table \ref{tab:lines_of_code} displays the actual numbers of code lines, to quantify the size of the respective parts.\\
 
@@
 This includes the implementation of the Matlab model, the VHDL and HLS code, as well as the common testbench.
-The chart also reveals that the VHDL implementation with its accompanying testbench requires more than twice the amount of code, compared to the HLS version, i.e. 32\% versus 14\%.\\
+The chart also reveals that the VHDL implementation with its accompanying testbench requires more than twice the amount of code compared to the HLS version, i.e. 32\% versus 14\%.\\
 
@@
 VHDL requires the largest part of code in the IP design, while the testbench code amount is relatively small.
+It is to be noted, that the common testbench consists of 4\%, which equals 630 lines of Python code.
+Therefor, the testbench code of both IPs is relatively small, because the entire analysis part is shared.
 
@@
   \end{tabular}
-  \caption{asdf}
+  \caption{The actual number of code lines, split in IP design and testbench, where applicable. All numbers are rounded up to tens.}
   \label{tab:lines_of_code}
@@
 %\end{GenericCode}
-
-Also note, that the common Python testbench consists of XXX lines of Python code.
-Therefor, the testbench code of the respective IPs is relatively small, because the entire analysis part is shared.
 

### 2021-08-03 sim/matlab/fm_transceiver.m
measure matlab sim speed

@@ -84,7 +84,12 @@ fm_sender();
 %% Receiver
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+tic;
+
 fm_receiver();
+
+elapsed_time = toc;
+fprintf('fm_receiver() took %f seconds\n',elapsed_time);
 
 % TODO: implement a version that is actually using 16 bit fixed_point like hardware
 % fm_receiver_fixed_point();

### 2021-08-04 doc/thesis/chapters/results.tex
remove commented table for detailed LOC
measure matlab sim speed
time logging table

@@
 
-%\begin{GenericCode}
-%Matlab:
-%  Language     files blank comment code TotalWithoutBlank Total
-%  ____________ _____ _____ _______ ____ _________________ _____
-%  MATLAB          34   551     728 1451              2179  2730
-%
-%VHDL IP Design:
-%  Language     files blank comment code TotalWithoutBlank Total
-%  ____________ _____ _____ _______ ____ _________________ _____
-%  VHDL            20   525     787 1431              2218  2743
-%  Python           1    45      22  141               163   208
-%
-%VHDL Testbench:
-%  Language     files blank comment code TotalWithoutBlank Total
-%  ____________ _____ _____ _______ ____ _________________ _____
-%  Python           5   141     141  426               567   708
-%  make             2    19      18   70                88   107
-%  Bourne Shell     1     5       6   16                22    27
-%
-%HLS IP Design:
-%  Language     files blank comment code TotalWithoutBlank Total
-%  ____________ _____ _____ _______ ____ _________________ _____
-%  C/C++ Header    14    90     119  221               340   430
-%  C++              9    97     197  195               392   489
-%
-%HLS Testbench:
-%  Language     files blank comment code TotalWithoutBlank Total
-%  ____________ _____ _____ _______ ____ _________________ _____
-%  C++              1    39      44  141               185   224
-%  Tcl/Tk           5    33      37  112               149   182
-%  C/C++ Header     2    24      18   87               105   129
-%  make             1    24      26   69                95   119
-%  Python           1    19      21   45                66    85
-%\end{GenericCode}
-
 %%%%%%%%%%%%%%%%%%%%
 \subsection{Implementation Time}
+
+The metric of implementation time -- the time it takes to implement a specific feature -- is a subjective measure, since it heavily depends on the respective developer's experience, as already noted above.
+However, this metric is used in this section to specifically compare HLS with VHDL.
+Certain features are chosen for this comparison and their implementation times were logged during the actual implementation process.
+Table \ref{tab:implementation_time_logging} shows the direct comparison between HLS and VHDL implementation times.
+
+\begin{table}[]
+  \centering
+  \begin{tabular}{l|c|c|}
+  \cline{2-3}
+                                                   & \multicolumn{2}{c|}{\textbf{Time [h]}} \\ \hline
+  \multicolumn{1}{|l|}{\textbf{Feature}}           & \textbf{HLS}      & \textbf{VHDL}      \\ \hline
+  \multicolumn{1}{|l|}{AXI Stream Interface}       & 8                 & 12                 \\ \hline
+  \multicolumn{1}{|l|}{AXI4-Lite Interface}        & 2                 & 10                 \\ \hline
+  \multicolumn{1}{|l|}{LED Control Register}       & 1                 & 2                  \\ \hline
+  \multicolumn{1}{|l|}{Build Information Register} & 8                 & 5                  \\ \hline
+  \multicolumn{1}{|l|}{Pass-Through Mode Register} & 1                 & 1                  \\ \hline
+  \multicolumn{1}{|l|}{FIR Filter}                 & 13                & 7                  \\ \hline
+  \multicolumn{1}{|l|}{Sample Decimation}          & 15                & 1                  \\ \hline
+  \multicolumn{1}{|l|}{Receiver Structure}         & 13                & 33                 \\ \hline
+  \multicolumn{1}{|l|}{\textbf{Sum}}               & \textbf{61}       & \textbf{71}        \\ \hline
+  \end{tabular}
+  \caption{Implementation time logging for a specific set of features, for the direct comparison between HLS and VHDL.}
+  \label{tab:implementation_time_logging}
+\end{table}
 
@@
 
+% Issues
+% -- HLS Debug Error with Bitwidth:
+%       Not sure what the error was.
+%       Parallel delay for I/Q didn't work.
+%       The order of every 2nd data value was wrong.
+%       Using a single delay fixed it.
+% -- HLS driver generation
+%       Fixing issue with auto-generated driver.
+%       It did not get imported into the SDK.
+%       Problem was that HLS project and top-level
+%       did not have the same name....
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Simulation}
-% Sim speed of x min of recorded FM signal
+
+Sim speed of x min of recorded FM signal\\
+bla bla
 %
@@
 % HLS:
+%
+%
+% Matlab:
+%          time                samples
+%   audio         sim        in      out
+%  1.7 s    0.45 sec     1632000   68000
+%
 
@@
 % IP generation and integration into Vivado
-
+% Issue: HLS driver import did not work - Vivado SDK is a mess (bad UI based on Eclipse, has bugs, no error output, etc...)
 %%%%%%%%%%%%%%%%%%%%

### 2021-08-04 doc/thesis/chapters/results.tex
axi stream

@@
   \end{tabular}
-  \caption{The actual number of code lines, split in IP design and testbench, where applicable. All numbers are rounded up to tens.}
+  \caption{The actual number of code lines for each code part, split in IP design and testbench, where applicable. All numbers are rounded up to tens.}
   \label{tab:lines_of_code}
@@
 The metric of implementation time -- the time it takes to implement a specific feature -- is a subjective measure, since it heavily depends on the respective developer's experience, as already noted above.
-However, this metric is used in this section to specifically compare HLS with VHDL.
+In the present case, there is a strong background of knowledge and practical experience in VHDL design, while there is only very little of such in HLS design.
+However, this metric is used in this section to specifically compare HLS with VHDL in this project.
 Certain features are chosen for this comparison and their implementation times were logged during the actual implementation process.
-Table \ref{tab:implementation_time_logging} shows the direct comparison between HLS and VHDL implementation times.
+The implementation is always considered as the HDL description, including the corresponding testbench code, to verify the respective correct functionality.
+Table \ref{tab:implementation_time_logging} shows the direct comparison between HLS and VHDL implementation times for these features.\\
+
+The following paragraphs present an insight into how the implementation times of the listed code parts arise.
+
+\subsubsection{AXI Stream Inferface}
+
+This is the main interface for the data input and -output of the DSP chain.
+In HLS it can be automatically inferred by adding pragma directives to the respective function parameter variables.
+Also the testbench code is implemented like a regular function call.
+Therefore, the actual interface logic is completely hidden from the developer.
+In VHDL on the other hand, this logic has to be implemented manually.
+Also the testbench has to be hooked up with the correct signals and have a verificator logic, that supports the functional verification of an AXI stream interface.
+Otherwise, this logic would have to be implemented manually as well.
+In the time comparison, the larger amount of manual implementation in VHDL is clearly visible.
+It is to note that this comparison does not compare functionally equivalent implementations, because the actually implemented logic in VHDL does not use the ready-flag correctly, as explained in the previous chapters.
+According to this fact, an increase in implementation hours should actually be considered for VHDL.
+However, also the HLS implementation has its particular characteristics that need to be understood for a fully functional implementation.
+Following the respective user guides and implementation examples provides a good starting point for that.
+
+\subsubsection{AXI4-Lite Interface}
+
+\subsubsection{Registers}
+
+LED Control, Build Information Register, Pass-Through
+\subsubsection{FIR Filter}
+
+\subsubsection{Sample Decimation}
+
+\subsubsection{Receiver Structure}
+
+In the final summation, the implementation time of the two variants only differ by around 15\%.
+However,  ... \dots
 

### 2021-08-04 doc/thesis/chapters/results.tex
axi-lite regs

@@
 
+This interface is used to control IP settings and to read status information.
+Similar to the previous AXI Stream interface, also this AXI4-Lite memory-mapped interface is automatically generated by HLS, while this has to be done manually in VHDL.
+To create this process more efficiently and extendable, a Python script is used, which can also automatically generate VHDL code for the registers' memory map.
+This turns out to be a multiplicator in productivity for the implementation of registers.
+The register generator script was previously used in projects and is therefor already well-known and tested before it is used in this project.
+However, the VHDL code only represents one part of the interface -- the second part is the firmware driver that is used in the application software.
+Even though the register generator script also generates a \cplusplus\ header file that includes several address offset definitions of the registers, the actual driver functions need to be implemented manually.
+Several files need to be integrated into a pre-defined folder structure, in order to enable a correct integration of the IP in a Vivado block diagram.\\
+
+Again, the amount of manual steps in the VHDL implementation are clearly represented in the number of hours that are spent.
+
+% TODO: mention figure out of read-only and read-write types (pointer/reference)
+% TODO: mention testbench (vhdl: support block, hls: simple access to variable)
+
 \subsubsection{Registers}
 
-LED Control, Build Information Register, Pass-Through
+Once the AXI4-Lite memory-mapped interface's main functionality is implemented, further registers can be added.
+In the specific implementation for this project, registers for LED control, the DSP mode, as well as a build information register are added.
+The latter holds information about the build date and -time of the IPs, and is particularly time-consuming because of the implementation of the respective drivers to convert the date format in both variants.
+In contrast, the addition of the LED control and DSP mode registers is simple and fast, because of the auto-generated code in HLS and VHDL with usage of the register engine.
+The single bits in the LED control register directly represent the LEDs' values on the hardware.
+The DSP mode register can select the regular FM radio mode and a pass-through mode, which simply forwards the streaming data without any processing.
+Summing up, the addition of registers to the design can be time-consuming in both variants, depending on the required driver functionality.
+However, the general tendency is that VHDL requires more time, because of this driver issue.
+
 \subsubsection{FIR Filter}
 
+with TB
+
 \subsubsection{Sample Decimation}
+
+VHDL: counter that only forwards one when valid, every n samples\\
+HLS: consume and process n, before one output (understand blocking AXI stream)
 
@@
 \subsection{Integration}
-% IP generation and integration into Vivado
-% Issue: HLS driver import did not work - Vivado SDK is a mess (bad UI based on Eclipse, has bugs, no error output, etc...)
+% - IP generation and integration into Vivado
+% - Another manual step in VHDL is the addition of the AXI bus interfaces to the IP.
+% - Issue: HLS driver import did not work - Vivado SDK is a mess (bad UI based on Eclipse, has bugs, no error output, etc...)
+
 %%%%%%%%%%%%%%%%%%%%

### 2021-08-13 doc/thesis/chapters/results.tex
todo
fir filter

@@
 
-with TB
+The FIR filter is one of the main DSP blocks of the design, as it is used in multiple instances throughout the design.
+Therefore, its implementation was done with special care, in order to guarantee its correct functionality.
+Consequently, a unit test is implemented in the testbench, which is also included in the listed number of implementation hours.\\
+
+For the \cplusplus\ version, a code example for an FIR filter, which is provided by Xilinx HLS, is taken and modified for the present purpose.
+The VHDL variant is taken from an older DSP project, which was previously developed during a university course.
+However, the integration and testbench code is developed for this project, specifically.
+Here, the HLS implementation took significantly longer, because of the limited knowledge regarding the usage of an AXI stream interface.
+It is the first time in the project, that this type of interface is applied in the IP design and used in the testbench.
+The VHDL implementation was straightforward in comparison.
+The testbenches compare the FIR filters' output with the Matlab model's pilot tone output.
+Both variants match the output results of the Matlab model successfully.
 
@@
 \subsubsection{Receiver Structure}
+
+todo \\
 

### 2021-08-13 doc/thesis/chapters/results.tex
sample dec

@@
 
-VHDL: counter that only forwards one when valid, every n samples\\
-HLS: consume and process n, before one output (understand blocking AXI stream)
+The implementation of the sample decimation functionality differs slightly between the two variants, HLS and VHDL.
+This is explained in Section \ref{sec:impl:hls:sample_rate_reduction}.
+The implementation time of the HLS version is severely larger than the VHDL version.
+This is mainly due to the limited knowledge of the functionality of the AXI stream interface at this stage in the project's implementation, just as previously mentioned with the FIR filter.
+The initial version was implemented like the VHDL version -- a counter that only triggers the subsequent DSP blocks upon a certain number of samples was received.
+However, this led to an endlessly blocking AXI stream in the testbench execution.
+The reason was, that the read-function can only read a sample when the respective stream contains a sample, and that it is a blocking function.
+Because the sample decimator did not write a sample every time, the subsequent read-function blocked indefinitely.
+A thorough investigation in several user guides, in combination with step-by-step debugging, the issue was found and led to the current, working implementation, which processes a number of samples, before it forwards one sample.
+The process of finding this issue explains the largely varying implementation times, as listed in Table \ref{tab:implementation_time_logging}.
 

### 2021-08-13 doc/thesis/chapters/results.tex
summarize impl time; move table

@@
 
+\begin{table}[]
+  \centering
+  \begin{tabular}{l|c|c|}
+  \cline{2-3}
+                                                   & \multicolumn{2}{c|}{\textbf{Time [h]}} \\ \hline
+  \multicolumn{1}{|l|}{\textbf{Feature}}           & \textbf{HLS}      & \textbf{VHDL}      \\ \hline
+  \multicolumn{1}{|l|}{AXI Stream Interface}       & 8                 & 12                 \\ \hline
+  \multicolumn{1}{|l|}{AXI4-Lite Interface}        & 2                 & 10                 \\ \hline
+  \multicolumn{1}{|l|}{LED Control Register}       & 1                 & 2                  \\ \hline
+  \multicolumn{1}{|l|}{Build Information Register} & 8                 & 5                  \\ \hline
+  \multicolumn{1}{|l|}{Pass-Through Mode Register} & 1                 & 1                  \\ \hline
+  \multicolumn{1}{|l|}{FIR Filter}                 & 13                & 7                  \\ \hline
+  \multicolumn{1}{|l|}{Sample Decimation}          & 15                & 1                  \\ \hline
+  \multicolumn{1}{|l|}{Receiver Structure}         & 13                & 33                 \\ \hline
+  \multicolumn{1}{|l|}{\textbf{Sum}}               & \textbf{61}       & \textbf{71}        \\ \hline
+  \end{tabular}
+  \caption{Implementation time logging for a specific set of features, for the direct comparison between HLS and VHDL.}
+  \label{tab:implementation_time_logging}
+\end{table}
+
 The following paragraphs present an insight into how the implementation times of the listed code parts arise.
@@
 It is to note that this comparison does not compare functionally equivalent implementations, because the actually implemented logic in VHDL does not use the ready-flag correctly, as explained in the previous chapters.
-According to this fact, an increase in implementation hours should actually be considered for VHDL.
+According to this fact, a further increase in implementation hours should actually be considered for VHDL.
 However, also the HLS implementation has its particular characteristics that need to be understood for a fully functional implementation.
@@
 
-todo \\
+The implementation of the receiver structure requires the combination of several previously implemented blocks, such as the FIR filter and any interfaces.
+This is done according to the Matlab model, to resemble its functionality.
+The VHDL version is significantly more time consuming in comparison to HLS.
+However, it is to note at this point that the VHDL version is implemented first, and thus some issues are already solved by the time of the HLS implementation.
+Nevertheless, the VHDL implementation takes longer, as it requires more detailed care in the signal level.
+This specifically includes the clock-alignment of any signals that need to be used in a combination.
+For example, if two parallel FIR filters produce output signals that are to be added, they need to be aligned so that they are valid at the time of the summation.
+This often requires analysis and debugging in a signal wave form.
+In HLS this is not required, since the validity of a signal is determined internally by the HLS compiler, according to the usage of the respective data variables.
+Another factor that adds up to the implementation time is the number of lines of code, as presented in Table \ref{tab:lines_of_code}.
+The VHDL IP design requires 3000 lines, while the HLS IP design only takes 920 for the same functionality.
+Again, this is a subjective measure -- several factors such as code style, or the developer's experience play a role -- but still shows a good representation of the implementation effort for this project.\\
 
-In the final summation, the implementation time of the two variants only differ by around 15\%.
-However,  ... \dots
-
-\begin{table}[]
-  \centering
-  \begin{tabular}{l|c|c|}
-  \cline{2-3}
-                                                   & \multicolumn{2}{c|}{\textbf{Time [h]}} \\ \hline
-  \multicolumn{1}{|l|}{\textbf{Feature}}           & \textbf{HLS}      & \textbf{VHDL}      \\ \hline
-  \multicolumn{1}{|l|}{AXI Stream Interface}       & 8                 & 12                 \\ \hline
-  \multicolumn{1}{|l|}{AXI4-Lite Interface}        & 2                 & 10                 \\ \hline
-  \multicolumn{1}{|l|}{LED Control Register}       & 1                 & 2                  \\ \hline
-  \multicolumn{1}{|l|}{Build Information Register} & 8                 & 5                  \\ \hline
-  \multicolumn{1}{|l|}{Pass-Through Mode Register} & 1                 & 1                  \\ \hline
-  \multicolumn{1}{|l|}{FIR Filter}                 & 13                & 7                  \\ \hline
-  \multicolumn{1}{|l|}{Sample Decimation}          & 15                & 1                  \\ \hline
-  \multicolumn{1}{|l|}{Receiver Structure}         & 13                & 33                 \\ \hline
-  \multicolumn{1}{|l|}{\textbf{Sum}}               & \textbf{61}       & \textbf{71}        \\ \hline
-  \end{tabular}
-  \caption{Implementation time logging for a specific set of features, for the direct comparison between HLS and VHDL.}
-  \label{tab:implementation_time_logging}
-\end{table}
+In the final summation over several implementation times of all the listed features, the two variants HLS and VHDL only differ by 10 hours, which is around 15\%.
+However, as the previous paragraphs explain, prior knowledge regarding HLS was limited and had to be gathered during the ongoing implementation.
+Nevertheless, even consiering this fact, the HLS implementation is faster than the VHDL equivalent.
+Thus, with an equal amount of prior knowledge and experience, it is assumed that an HLS implementation can lead to a successfully functioning result significantly faster.
 
@@
 Sim speed of x min of recorded FM signal\\
-bla bla
+Sim speed of x min of recorded FM signal\\
+Sim speed of x min of recorded FM signal\\
+Sim speed of x min of recorded FM signal\\
+Sim speed of x min of recorded FM signal\\
+Sim speed of x min of recorded FM signal\\
+Sim speed of x min of recorded FM signal\\
+Sim speed of x min of recorded FM signal\\
+Sim speed of x min of recorded FM signal\\
+
 %

### 2021-08-13 doc/thesis/chapters/results.tex
impl_time_analysis plot

@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Interfaces}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Audio Output}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Lines of Code}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Implementation Time}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{AXI Stream Inferface}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{AXI4-Lite Interface}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Registers}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{FIR Filter}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Sample Decimation}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Receiver Structure}
@@
 Sim speed of x min of recorded FM signal\\
-Sim speed of x min of recorded FM signal\\
-Sim speed of x min of recorded FM signal\\
-Sim speed of x min of recorded FM signal\\
-Sim speed of x min of recorded FM signal\\
-Sim speed of x min of recorded FM signal\\
-Sim speed of x min of recorded FM signal\\
-Sim speed of x min of recorded FM signal\\
-Sim speed of x min of recorded FM signal\\
 
+measurements are always including loading of data, running through DUT and storing the results.
+exclusively result analysis, since this is shared and thus taking the same amount of time.
+several measurements are done multiple times, to achieve a good average result, that is independent of the current processor load of the host PC.
+
+Figure \ref{fig:impl_time_analysis} shows something
 %
@@
 %   audio         sim           in      out
-%  0.1000 s    73:44 min  (96000 -> 4000)
-%  0.0100 s    09:37 min  ( 9600 ->  400)
-%  0.0025 s    02:36 min  ( 2400 ->  100)
+%  0.1000 s    73:44 min       96000 -> 4000
+%  0.0100 s    09:37 min        9600 ->  400
+%  0.0025 s    02:36 min        2400 ->  100
 %
 % HLS:
+%          time                   samples
+%   audio         sim           in      out
+%     0.8 s    00:17 min       768000 -> 32000
+%     1.7 s    00:36 min      1632000 -> 68000
 %
@@
 %   audio         sim           in      out
-%  1.7 s    0.45 sec     1632000   68000
+%     1.7 s  00:00:00.52 min   1632000   68000
 %
+
+\includepicture [0.55] [0, trim=10 20 10 10, clip] {The implementation time of the implementation variants.
+Note the logarithmic scale of the y-axis.} {impl_time_analysis} {img/matlab/impl_time_analysis}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Integration}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Hardware Utilization}
@@
 
-%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Latency}

### 2021-08-13 doc/thesis/img/matlab/impl_time_analysis.py

@@ -1,0 +1,36 @@
+# -------------------------------------------------------------------------
+# File        : impl_time_analysis.py
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Create a diagram to represent the implementation time.
+# -------------------------------------------------------------------------
+
+import matplotlib.pyplot as plt
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.sans-serif": ["Helvetica"],
+    "font.size": 18,
+})
+
+x = ['VHDL', 'HLS', 'Matlab']
+energy = [18, 45000, 3100000]
+
+x_pos = [i for i, _ in enumerate(x)]
+
+fig = plt.figure()
+fig.set_size_inches(6, 5)
+ax = fig.add_subplot(1, 1, 1)
+ax.set_yscale('log')
+
+plt.bar(x_pos, energy, width=0.5)
+plt.xlabel("Implementation Variant")
+plt.ylabel("Samples per second")
+plt.title("Processed Samples Per Second")
+
+plt.xticks(x_pos, x)
+
+# show and save
+plt.tight_layout()
+plt.savefig('impl_time_analysis.pdf')
+plt.show()

### 2021-08-13 sim/matlab/fm_transceiver.m

@@ -34,7 +34,7 @@ dir_output_doc = "../../doc/thesis/img/matlab/";
 EnableWriteDataFiles = true;
 EnablePlots          = true;
 EnableSavePlotsToPng = true;
-EnableSavePlotsToPDF = true;
+EnableSavePlotsToPDF = false;
 
 EnableSenderSourceRecordedFile = false;
 EnableSenderSourceCreateSim    = true;

### 2021-08-13 doc/thesis/img/matlab/impl_time_analysis.py
beautify plot

@@ -10,7 +10,7 @@ plt.rcParams.update({
     "text.usetex": True,
     "font.family": "serif",
     "font.sans-serif": ["Helvetica"],
-    "font.size": 18,
+    "font.size": 16,
 })
 
 x = ['VHDL', 'HLS', 'Matlab']
@@ -21,11 +21,15 @@ x_pos = [i for i, _ in enumerate(x)]
 fig = plt.figure()
 fig.set_size_inches(6, 5)
 ax = fig.add_subplot(1, 1, 1)
+
 ax.set_yscale('log')
 
-plt.bar(x_pos, energy, width=0.5)
+ax.spines['top'].set_visible(False)
+ax.spines['right'].set_visible(False)
+
+plt.bar(x_pos, energy, width=0.5, edgecolor='black')
 plt.xlabel("Implementation Variant")
-plt.ylabel("Samples per second")
+#plt.ylabel("Samples per second")
 plt.title("Processed Samples Per Second")
 
 plt.xticks(x_pos, x)

### 2021-08-14 doc/thesis/img/matlab/impl_time_analysis.py
beautify plots (factor arrows)

@@ -13,10 +13,10 @@ plt.rcParams.update({
     "font.size": 16,
 })
 
-x = ['VHDL', 'HLS', 'Matlab']
-energy = [18, 45000, 3100000]
+variant = ['VHDL', 'HLS', 'Matlab']
+samples_per_second = [18, 45000, 3100000]
 
-x_pos = [i for i, _ in enumerate(x)]
+x_pos = [i for i, _ in enumerate(variant)]
 
 fig = plt.figure()
 fig.set_size_inches(6, 5)
@@ -27,12 +27,32 @@ ax.set_yscale('log')
 ax.spines['top'].set_visible(False)
 ax.spines['right'].set_visible(False)
 
-plt.bar(x_pos, energy, width=0.5, edgecolor='black')
+plt.bar(x_pos, samples_per_second, width=0.5, edgecolor='black')
 plt.xlabel("Implementation Variant")
 #plt.ylabel("Samples per second")
 plt.title("Processed Samples Per Second")
 
-plt.xticks(x_pos, x)
+plt.xticks(x_pos, variant)
+
+plt.annotate('x2500',
+             xy=(0.5, 10000),
+             xycoords='data',
+             xytext=(-70, -40),
+             textcoords='offset points',
+             fontsize=16,
+             horizontalalignment='left',
+             arrowprops=dict(arrowstyle='->', connectionstyle='arc3,rad=-0.2')
+             )
+plt.annotate('x70',
+             xy=(1.6, 1000000),
+             xycoords='data',
+             xytext=(-70, -40),
+             textcoords='offset points',
+             fontsize=16,
+             horizontalalignment='left',
+             arrowprops=dict(arrowstyle='->', connectionstyle='arc3,rad=-0.2')
+             )
+
 
 # show and save
 plt.tight_layout()

### 2021-08-14 doc/thesis/chapters/results.tex
beautify plot
sim speed done

@@
 
-Sim speed of x min of recorded FM signal\\
+In this section, the simulation speed of the multiple implementation variants in their various abstraction levels are compared.
+Simulation speed in this context refers to the number of input samples that the DUT can process in a time unit, like one second.
+In the current application, the antenna signal is recorded and replayed to the DUT in the simulation.
+This ensures a reproducable result, so that the different variants can be compared against each other reliably.\\
 
-measurements are always including loading of data, running through DUT and storing the results.
-exclusively result analysis, since this is shared and thus taking the same amount of time.
-several measurements are done multiple times, to achieve a good average result, that is independent of the current processor load of the host PC.
+Figure \ref{fig:impl_time_analysis} shows this direct comparison of processed samples per second for the implementation variants.
+The measurements always include the testbench setup-phase, and the run-phase.
+The setup-phase mainly consists of loading the recorded data and its preparation, so that it can be fed into the DUT.
+The run-phase then actually applies this test data to the DUT and stores the respective output results.
+All time measurements are taken without the result analysis, since this is the exact same, shared code between all variants and thus takes the same amount of time everywhere.
+Several measurements are done multiple times, to achieve an average result, that is independent of the current processor load of the host PC.\\
 
-Figure \ref{fig:impl_time_analysis} shows something
+The diagram shows an immense difference of simulation speed between the variants.
+It is important to note that the y-axis has a logarithmic scale, as a method to enable the presentation of the values in multiple orders of magnitude.
+VHDL only achieves around 20 samples per second.
+This is by far the slowest method, which is expected because of the way of how VHDL is simulated.
+This can not be explained in detail here, but one main feature of a VHDL simulator is its capability to simulate parallel processes, as they occur in actual parallel hardware.
+The VHDL simulator therefor needs to evaluate several concurrent processes and statements, whenever any of the interacting signals change their value, and this is time-consuming.
+In the current case, there is an additional interface between the VHDL simulator and the Python testbench framework, which likely additionally slows down the simulation.
+The HLS testbench in comparison achieves around 50000 samples per second, which is a speed-up factor of 2500.
+This is specifically talking about the HLS C/\cplusplus-simulation.
+The main reason for the speed-up therefor is, that in this type of simulation no timing behaviour is simulated -- any processes are treated as C/\cplusplus\ functions that runs like a regular executable on the host CPU.
+Also, the HLS compiler is using native \cplusplus\ data types to represent the signals, where possible, which results in the faster execution of the code.
+Finally, the simulation in Matlab achieves a further speed-up by a factor of 70.
+Matlab is specifically designed for applications which require matrix calculations and utilizes highly optimized algorithms in its underlying functions therefor.
+Thus, Matlab runs very efficiently and achieves the highest simulation speed of around 3 million samples per second.
+
 %
@@
 
-\includepicture [0.55] [0, trim=10 20 10 10, clip] {The implementation time of the implementation variants.
+\includepicture [0.55] [0, trim=10 10 10 10, clip] {The measured simulation speed of the different implementation variants, in samples per second.
+The arrows show the speed-up factors.
 Note the logarithmic scale of the y-axis.} {impl_time_analysis} {img/matlab/impl_time_analysis}
@@
 This includes a reference model, the development of the IP itself, including the simulation testbench, as well as the final integration into the productive system.
-Because of this structured approach, that is given by this system design, it can be applied to almost any DSP project.
+Because of this modular and structured approach, that is given by this system design, it can be applied to almost any DSP project.
 The main processing IP can simply be replaced with another implementation, while the environment does not need to be modified.

### 2021-08-14 doc/thesis/chapters/results.tex
integration

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Simulation}
+\section{Simulation Speed}
 
@@
 
+The VHDL and HLS implementation variants are implemented, simulated in their testbenches to verify the correct functionality, as well as deployed on hardware, to test their performance in an actual application.
+Important parts of this end-to-end integration are presented here.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Integration}
-% - IP generation and integration into Vivado
-% - Another manual step in VHDL is the addition of the AXI bus interfaces to the IP.
-% - Issue: HLS driver import did not work - Vivado SDK is a mess (bad UI based on Eclipse, has bugs, no error output, etc...)
+
+The aim is to create IP blocks, that can be integrated into a block design in the Xilinx Vivado software.
+The IP blocks should be able to be connected with their AXI interfaces, in order to successfully communicate and interact with the embedded hard-processor, the ARM CPU.
+In the case of HLS, several files that are required for this integration are generated automatically.
+The files are already placed into the specific folder structure, that Vivado requires to recognize it as an IP block.
+In the case of VHDL on the other hand, this needs to be done manually.
+There is an assistant, the IP Packager, which helps the developer to create the IP.
+However, at this point the folder structure already needs to be in place, so that the assistant works correctly, which again requires manual action.
+Regarding the top-level interfaces, the IP Packager supports an auto-detection of the signals that belong to the respective interface types, as long as they follow the expected naming guidelines.
+However, the interface itself needs to be defined and updated manually in the IP Packager.
+This is especially important, if the interfaces change in the VHDL code.\\
+
+However, also the auto-generation of the HLS IP has characteristics that require special attention.
+One issue during the implementation was, that the auto-generated software driver did not get imported into the Vivado SDK.
+The reason therefor was, that the top-level function of the IP did not have the exact same naming as the Vivado HLS project.
+For issues like that, the Eclipse-based SDK does not give any error output or any kind of information.
+This however is a general issue with the SDK, where error handling is poor and error log outputs are not existent, or at least not useful to the developer, so that the SDK sometimes crashes without an obvious reason.
+It is to note, that version v2018.2 is used for the entire Vivado toolchain, which is not the latest version that is currently available.
+Thus, some of the occuring issues may already be fixed in the newer versions.\\
+
+However, both IPs could successfully be integrated into the Vivado block design, and the tests on the hardware show that they function correctly within the system.
 

### 2021-08-14 doc/analysis/lines-of-code/count_lines_of_code.sh
move lines-of-code analysis

@@ -8,7 +8,7 @@
 set -euo pipefail
 
 SCRIPT_PATH=$(dirname $(readlink -f $0))
-REPO_ROOT=$SCRIPT_PATH/../../
+REPO_ROOT=$SCRIPT_PATH/../../../
 
 cd $REPO_ROOT
 
@@ -25,7 +25,7 @@ cd $REPO_ROOT
 CLOC_CMD="docker run --rm -v $REPO_ROOT:/tmp/ aldanial/cloc:latest"
 
 # Clean previous output products
-SCRIPT_PATH_RELATIVE=doc/lines-of-code-analysis/output
+SCRIPT_PATH_RELATIVE=doc/analysis/lines-of-code/output
 mkdir -p $SCRIPT_PATH_RELATIVE
 rm -f $SCRIPT_PATH_RELATIVE/*.txt
 

### 2021-08-14 doc/analysis/lines-of-code/get_total_columns.sh

@@ -11,7 +11,7 @@
 ###
 
 SCRIPT_PATH=$(dirname $(readlink -f $0))
-REPO_ROOT=$SCRIPT_PATH/../../
+REPO_ROOT=$SCRIPT_PATH/../../../
 cd $REPO_ROOT
 
 if test $# -eq 0 ; then

### 2021-08-14 doc/analysis/lines-of-code/lines_of_code_diagram.m

@@ -10,7 +10,7 @@ close all;
 clc;
 
 %% Settings
-dir_output_doc = "../thesis/img/matlab";
+dir_output_doc = "../../thesis/img/matlab";
 
 
 %% Plot

### 2021-08-14 doc/analysis/lines-of-code/lines_of_code_diagram.py

@@ -82,7 +82,7 @@ ax.legend(handles[:], group_names_legs, loc=(1.0, 0.25), prop={'size': 13})
 plt.tight_layout()
 
 # show and save
-plt.savefig('../thesis/img/matlab/lines_of_code_pie_chart_py_all.pdf')
+plt.savefig('../../thesis/img/matlab/lines_of_code_pie_chart_py_all.pdf')
 plt.show()
 
 # ---------------------- HLS --------------------------------------------------
@@ -161,7 +161,7 @@ ax.legend(handles[len(group_names):], subgroup_names_legs, loc=(0.95, 0.2))
 plt.tight_layout()
 
 # show and save
-plt.savefig('../thesis/img/matlab/lines_of_code_pie_chart_py_hls.pdf')
+plt.savefig('../../thesis/img/matlab/lines_of_code_pie_chart_py_hls.pdf')
 plt.show()
 
 
@@ -241,7 +241,7 @@ ax.legend(handles[len(group_names):], subgroup_names_legs, loc=(0.95, 0.2))
 plt.tight_layout()
 
 # show and save
-plt.savefig('../thesis/img/matlab/lines_of_code_pie_chart_py_vhdl.pdf')
+plt.savefig('../../thesis/img/matlab/lines_of_code_pie_chart_py_vhdl.pdf')
 plt.show()
 
 # -------------------------------------------------------------------------

### 2021-08-14 doc/analysis/implementation-time/impl_time_analysis.py
move impl time analysis

@@ -56,5 +56,5 @@ plt.annotate('x70',
 
 # show and save
 plt.tight_layout()
-plt.savefig('impl_time_analysis.pdf')
+plt.savefig('../../thesis/img/matlab/impl_time_analysis.pdf')
 plt.show()

### 2021-08-14 doc/thesis/chapters/results.tex
add data for plot

@@
 \subsection{Hardware Utilization}
-% --> utilization reports
+
+The hardware utilization of the HLS and VHDL implementations are compared in detail in this section.
+The numbers are taken from the final utilization report, that is created by Vivado at the end of the bitstream creation, since this is assumed to be the most accurate report that shows the actual utilization.
+
+Figure \ref{fig:hardware_utilization} shows
+
+\includepicture [0.75] [0, trim=10 10 10 10, clip] {hardware utilization} {hardware_utilization} {img/matlab/hardware_utilization}
 

### 2021-08-14 doc/analysis/hardware-utilization/hardware_utilization.py

@@ -1,0 +1,67 @@
+# -------------------------------------------------------------------------
+# File        : hardware_utilization.py
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Create a diagram to represent the hardware utilization.
+# -------------------------------------------------------------------------
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.sans-serif": ["Helvetica"],
+    "font.size": 16,
+})
+
+# Data
+labels = ['VHDL', 'HLS']
+hw_util_lut = [1203, 650]
+hw_util_ff = [1361, 494]
+hw_util_lutram = [176, 0]
+hw_util_srl = [0, 56]
+hw_util_ramb18 = [0, 7]
+hw_util_dsp48 = [8, 6]
+
+barwidth = 0.15
+
+# the label locations
+x1 = np.arange(len(labels))
+x2 = [x + barwidth for x in x1]
+x3 = [x + barwidth for x in x2]
+x4 = [x + barwidth for x in x3]
+x5 = [x + barwidth for x in x4]
+
+# Plot
+fig, ax = plt.subplots()
+fig.set_size_inches(8, 5)
+ax.spines['top'].set_visible(False)
+ax.spines['right'].set_visible(False)
+
+rects1 = ax.bar(x2, hw_util_lut, barwidth, label='LUT', edgecolor='black')
+rects2 = ax.bar(x1, hw_util_ff, barwidth, label='FF', edgecolor='black')
+rects3 = ax.bar(x3, hw_util_lutram, barwidth, label='LUTRAM', edgecolor='black')
+rects4 = ax.bar(x3, hw_util_srl, barwidth, label='SRL', edgecolor='black')
+rects5 = ax.bar(x4, hw_util_ramb18, barwidth, label='RAMB18', edgecolor='black')
+rects6 = ax.bar(x5, hw_util_dsp48, barwidth, label='DSP48', edgecolor='black')
+
+# Add some text for labels, title and custom x-axis tick labels, etc.
+# ax.set_ylabel('Scores')
+ax.set_title('Hardware Utilization')
+#ax.set_xticks([r + barwidth for r in range(len(labels))], labels)
+plt.xticks([r + 2 * barwidth for r in range(len(labels))], labels)
+ax.set_xticklabels(labels)
[…]

### 2021-08-14 doc/thesis/chapters/results.tex
re-structure and beautify plot

--- 2021-08-14 doc/thesis/chapters/results.tex
@@
 \subsection{Hardware Utilization}
-% --> utilization reports
+
+The hardware utilization of the HLS and VHDL implementations are compared in detail in this section.
+The numbers are taken from the final utilization report, that is created by Vivado at the end of the bitstream creation, since this is assumed to be the most accurate report that shows the actual utilization.\\
+
+Figure \ref{fig:hardware_utilization} shows the utilization numbers of different FPGA resources.
+
+\includepicture [0.75] [0, trim=10 10 10 10, clip] {hardware utilization} {hardware_utilization} {img/matlab/hardware_utilization}
 

### 2021-08-14 doc/analysis/hardware-utilization/hardware_utilization.py

@@ -15,22 +15,15 @@ plt.rcParams.update({
 })
 
 # Data
-labels = ['VHDL', 'HLS']
-hw_util_lut = [1203, 650]
-hw_util_ff = [1361, 494]
-hw_util_lutram = [176, 0]
-hw_util_srl = [0, 56]
-hw_util_ramb18 = [0, 7]
-hw_util_dsp48 = [8, 6]
+labels = ['LUT', 'FF', 'LUTRAM', 'SRL', 'RAMB18', 'DSP48']
+values_vhdl = [1203, 1361, 176, 0, 0, 8]
+values_hls = [650, 494, 0, 56, 7, 6]
 
-barwidth = 0.15
+barwidth = 0.4
 
 # the label locations
-x1 = np.arange(len(labels))
+x1 = np.arange(len(labels)) * 1.5
 x2 = [x + barwidth for x in x1]
-x3 = [x + barwidth for x in x2]
-x4 = [x + barwidth for x in x3]
-x5 = [x + barwidth for x in x4]
 
 # Plot
 fig, ax = plt.subplots()
@@ -38,28 +31,20 @@ fig.set_size_inches(8, 5)
 ax.spines['top'].set_visible(False)
 ax.spines['right'].set_visible(False)
 
-rects1 = ax.bar(x2, hw_util_lut, barwidth, label='LUT', edgecolor='black')
-rects2 = ax.bar(x1, hw_util_ff, barwidth, label='FF', edgecolor='black')
-rects3 = ax.bar(x3, hw_util_lutram, barwidth, label='LUTRAM', edgecolor='black')
-rects4 = ax.bar(x3, hw_util_srl, barwidth, label='SRL', edgecolor='black')
-rects5 = ax.bar(x4, hw_util_ramb18, barwidth, label='RAMB18', edgecolor='black')
-rects6 = ax.bar(x5, hw_util_dsp48, barwidth, label='DSP48', edgecolor='black')
+rects1 = ax.bar(x1, values_vhdl, barwidth, label='VHDL', color=plt.cm.Greens(0.6), edgecolor='black')
+rects2 = ax.bar(x2, values_hls, barwidth, label='HLS', color=plt.cm.Blues(0.6), edgecolor='black')
 
 # Add some text for labels, title and custom x-axis tick labels, etc.
 # ax.set_ylabel('Scores')
 ax.set_title('Hardware Utilization')
[…]

### 2021-08-14 doc/thesis/chapters/results.tex
hardware util

@@
 
-Figure \ref{fig:hardware_utilization} shows the utilization numbers of different FPGA resources.
+Figure \ref{fig:hardware_utilization} shows the utilization numbers of the different FPGA resources.
+It reveals that the manual VHDL implementation clearly requires more resources in the FPGA.
+The study of the detailed utilization report however does not enable a deeper direct comparison of the respective single blocks, such as VHDL FIR filter against HLS FIR filter.
+This is because the HLS compiler automatically applies optimizations, so that the direct connection between code and resource report is partially lost.
+Examples for optimizations are inlining of functions, re-ordering of operations and time-sharing.
+The latter has an especially large impact on the resource utilization.
+Generally, in hardware design there is a trade-off between resource utilization and throughput, which leads to time-sharing.
+For example, an entirely parallel-duplicated pipeline can process the double amount of data, but also requires the double amount of resources.
+On the other hand, time-sharing a resource may take the double amount of time, thus only processes half of the data, but also only requires half of the resource utilization.
+In HLS, this trade-off can be influenced and determined by the developer.
+Depending on the requirements of the design, the optimization may either be set towards a lower resource utilization, or towards a high throughput.
+In the current HLS implementation, no measures are taken towards any optimization and thus, the compiler is free to decide.
+However, the only requirement is the correct functionality within the given situation, of course.
+This specifically means that the throughput needs to be high enough, in order to process the audio signal output accordingly.\\
 
-\includepicture [0.75] [0, trim=10 10 10 10, clip] {hardware utilization} {hardware_utilization} {img/matlab/hardware_utilization}
+Nevertheless, the resource report allows the comparison of the configuration interface, which is the AXI4-Lite interface.
+In VHDL, it requires 44 LUTs and 54 FFs, while HLS consumes only 36 LUTs and 37 FFs.
+So it seems that the manual VHDL implementation requires more resources in all the comparable parts.
+By further investigating the report, the VHDL version apparently utilizes a large amount of LUTs for the delays, which are used to clock-align the FIR output signals.
+For an optimization of the VHDL design, this would be a good point to start at.
+
+\includepicture [0.75] [0, trim=10 10 10 10, clip] {A comparison of the resource utilization of the VHDL and HLS implementations.
+The VHDL variant requires significantly more resources.} {hardware_utilization} {img/matlab/hardware_utilization}
 

### 2021-08-14 doc/analysis/hardware-utilization/hardware_utilization.py

@@ -39,7 +39,7 @@ rects2 = ax.bar(x2, values_hls, barwidth, label='HLS', color=plt.cm.Blues(0.6), 
 ax.set_title('Hardware Utilization')
 plt.xticks([r + barwidth / 2 for r in x1], labels)
 ax.set_xticklabels(labels)
-ax.legend(loc='best')
+ax.legend(loc='right')
 #ax.legend(loc=(1.0, 0.2))
 #ax.legend(loc='center left', bbox_to_anchor=(1.0, 0.5))
 

### 2021-08-14 doc/thesis/chapters/results.tex
latency

@@
 For an optimization of the VHDL design, this would be a good point to start at.
+However, for this project, both implementation variants are compared without any optimizations from the developer's side.
 
@@
 \subsection{Latency}
-% --> utilization reports
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Applicability}
-% of all implementation variants
-%   GNU Radio for fast prototyping,
-%   Matlab for research
-%   HLS/VHDL for deployment on HW, product development
+Latency is the time it takes for the design to produce an output sample.
+The latency of the HLS version is taken from the report that is produced after the C-synthesis of the HLS compiler.
+The report shows an estimated resource utilization, as well as the latency as a number of clock cycles.
+The latency is given with 2712 clock cycles.
+For the VHDL version, a similar report is not available.
+Instead, the testbench of the integration test is executed, which produces a signal wave form.
+The latency is measured from this wave form and results in a number of around 2470 clock cycles.
+With regards to latency, the two implementation variants appear to have a similar behaviour.

### 2021-08-14 doc/thesis/chapters/conclusion-and-prospect.tex
comments
results wip..
latency
conclusion begin

@@
 \label{cha:ConclusionAndProspect}
+
+This thesis covers a comprehensive variety of topics, that ranges from digital signal processing to system modeling and -design, IP design in different methods, verification, to hardware-software interaction in hardware.
+The implementation of the FM radio receiver in multiple different methods, namely GNU Radio, Matlab, VHDL and HLS so that they achieve the same functionality, is a challenge, especially with the latter methods.
+However, this challenge is solved within an end-to-end system design, that strongly supports the development process.\\
+
+GNU Radio is a software, that allows an efficient development of a software-defined radio as a prototype, which can run on actual hardware in a live environment.
+It is definitely worth to use it during the development of a project like the one in this thesis, since it allows to achieve a functioning result with relatively low effort.\\
+
+The usage of Matlab as a tool to research DSP algorithms and strategies to demodulate a signal, in order to ultimately find a system model that can be implemented in hardware, turns out to be a powerful asset.
+It allows the development in fast iterations, through short simulation times, to achieve a result quickly.
+Based on this system model, the implementation of actual hardware IPs, with VHDL and HLS can be done more efficiently and in a direct way, because issues in the algorithm level are already fixed in the model at this point.
+In combination with the simulation testbenches that share major parts, such as the input source, as well as a the analysis tool, the IP development works smoothly and efficiently.
+The verification as a comparison against the Matlab system model gives instant feedback of whether the IP functions correctly.
+Even if the system model changes, the testbench automatically adapts, since the input source files are re-generated by the model and thus picked up at the next iteration of the testbench simulation, so the hardware engineer can also adapt the IP code.
+Once the IP is deployed on actual hardware, it is tested by the firmware, in order to check its correct functionality again, which closes the loop from the system model and the IP simulation to the hardware deployment.
+This entire process proved itself to be a successful approach.\\
+
+Regarding the implementation of the VHDL and HLS IPs and the accompanying comparison, the result turns out to be unexpected.
+The initial expectation, that HLS is inefficient in terms of resource utilization, is contradicted.
+Also the amount of time spent in the implementation and the metric of lines of code create a positive picture of HLS.
+Further, the automatic generation of several files that are required to integrate the IP into a system design, such as interface definitions or software drivers, support this attitude.
+VHDL may have its advantages in specific areas of application, but HLS is definitely a powerful tool that can achieve great results in less time.
+This is especially true, when it comes to the development of complex functionalities, such as image- or video processing, with the support of built-in libraries like OpenCV.
 
@@
 %
-%
-%
+% ...the list shows how comprehensive the topic of this thesis is.
+% Based on the results that were made during the elaboration, many more ideas came into play.
+% Many ideas to extend the project, ...
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Advantages / Disadvantages}
+%\section{Advantages / Disadvantages}
 
 easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV.\\
-However, it may be less optimized.\\
 A developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.

### 2021-08-14 doc/thesis/chapters/hls.tex
conclusion begin

@@
 In older versions, SystemC was supported for synthesis as well, but is now deprecated in the respective latest versions \cite{VivadoHlsDeprecatesSystemC}.\\
+
+Another feature of HLS is the support of libraries for complex functionalities, such as video processing.
+Xilinx HLS for example provides a port of the OpenCV image- and video processing library, that enables the development of complex functions in this area, inclusively the verification in the testbench.\\
 

### 2021-08-14 doc/thesis/chapters/conclusion-and-prospect.tex
conclusion done

@@
 VHDL may have its advantages in specific areas of application, but HLS is definitely a powerful tool that can achieve great results in less time.
-This is especially true, when it comes to the development of complex functionalities, such as image- or video processing, with the support of built-in libraries like OpenCV.
+This is especially true, when it comes to the development of complex functionalities, such as image- or video processing, with the support of built-in libraries like OpenCV.\\
+
+However, the often heard statement that HLS enables any software developer to develop hardware IPs, can not be supported.
+A developer that writes HLS code, still needs to have a strong hardware background, to be able to understand what is being generated in hardware.
+It is important to know about the limitations of the target hardware, i.e. FPGA, in order to write an efficient design.
+Therefore, the connection between specific HLS \cplusplus\ code lines, as well as the general structure and algorithm, and the hardware target need to be understood.
 
@@
 % Many ideas to extend the project, ...
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Advantages / Disadvantages}
-
-easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV.\\
-A developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.

### 2021-08-14 doc/thesis/chapters/conclusion-and-prospect.tex
prospect

@@
 It is important to know about the limitations of the target hardware, i.e. FPGA, in order to write an efficient design.
-Therefore, the connection between specific HLS \cplusplus\ code lines, as well as the general structure and algorithm, and the hardware target need to be understood.
+Therefore, the connection between specific HLS \cplusplus\ code lines, as well as the general structure and algorithm, and the hardware target need to be understood.\\
+
+In prospect, there are many ideas on how the project can potentially be extended.
+Additional implementation methods could be used to develop the FM radio receiver IP.
+For example, Matlab can be used to create VHDL code, in a high-level synthesis fashion, which can be used to create the IP.
+Regarding the verification environment, also the GNU Radio implementation could be included.
+The common input that is produced by the Matlab model, and the common output analysis, which are both used for the VHDL and HLS testbenches, can be used in GNU Radio as well.
+This would allow another comparison.
+Looking at the HLS code, it could be evaluated how compiler-independent it is.
+The question could be, if it is possible to write code that can be compiled with multiple compilers for different target platforms and vendors, e.g. Xilinx and Intel FPGAs, at the same time.
+Furthermore, the various optimization strategies that Xilinx HLS provides could be evaluated on the existing HLS code.
+Also, currently manually implemented functions, such as the FIR filter, could be replaced with build-in library functions.
+Finally, the application software could be upgraded from the current bare-metal, real-time OS, to an embedded Linux OS which runs on the ARM CPU.
+This would additionally allow the support of the RTL-SDR device to be used for live data streaming, instead of the current replay of the recorded antenna data from the SD card.\\
+
+The amount of ideas for potential extensions to the project shows how comprehensive the project is and how many topics it covers.
+Summing up, it can be said that the project revealed a range of interesting and partially unexpected insights into multiple topics, and that a lot of knowledge was gained during the elaboration of this thesis.
 
@@
 %  - generate VHDL from Matlab receiver code
+%  - use Matlab Simulink HDL Coder and System Generator
+%     as additional implementation variants
+
 %  - use the common input file and feed it to GNU Radio;
 %    also produce an output file with GNU Radio for comparison
+
 %  - Develop HLS compiler independent code (optional)}
@@
 %       which makes it difficult to write compiler independent code....
-%  - use Matlab Simulink HDL Coder and System Generator
-%     as additional implementation variants
 %  - use Intel oneAPI HLS compiler
@@
 %      -- how platform-independent is HLS?
+
 %  - explore HLS optimizations
 %  - explore HLS libraries (e.g. DSP library --> pre-made FIR filter)
+
 %  - create another IP in the same system

### 2021-08-14 doc/thesis/chapters/hls.tex
prospect

--- 2021-07-29 doc/thesis/chapters/hls.tex
@@
 In older versions, SystemC was supported for synthesis as well, but is now deprecated in the respective latest versions \cite{VivadoHlsDeprecatesSystemC}.\\
+
+Another feature of HLS is the support of libraries for complex functionalities, such as video processing.
+Xilinx HLS for example provides a port of the OpenCV image- and video processing library, that enables the development of complex functions in this area, inclusively the verification in the testbench.
+Also, other libraries are built-in, which feature commonly used functions, such as an FIR filter, for example.\\
 

### 2021-08-14 doc/thesis/chapters/introduction.tex
text
outline

@@
 
-TODO
+Chapter \ref{cha:SignalProcessingTheory} describes the signal processing theory that is required to demodulate an FM broadcast radio signal.
+This includes an introduction to receiver types, the description of an FM signal generally and also specifically as it is used in FM audio broadcasting, as well as various general topics of signal processing.
+Chapter \ref{cha:SystemDesignProcess} covers the topic of a certain system design process, as well as system design tools that specifically target FPGAs.
+Chapter \ref{cha:HLS} is about high-level synthesis, which includes a general introduction to the state of the art, as well as a description of the workflow, inclusively the coding and testbench features and characteristics.
+Chapter \ref{cha:SystemArchitectureAndConcept} describes the chosen system architecture and concept for the project.
+The actual implementation, with a detailed explanation of some selected components, is covered in Chapter \ref{cha:Implementation}.
+Chapter \ref{cha:DeploymentOnHardware} gives an overview of the required parts of development that are necessary to deploy the developed IPs on actual hardware.
+The comparison and achieved results are covered in Chapter \ref{cha:ComparisonAndResults}.
+Chapter \ref{cha:ConclusionAndProspect} concludes this thesis with a retrospective and gives an outlook of potential future extensions to the project.
 
@@
 %
-% ## check all abbreviation and where they are introduced the first time
+% ## check all abbreviations and where they are introduced the first time
 %       - introduce only once, like "device under test (DUT)",
@@
 %
+% ## fix the usage of apastrophes everywhere...
+%       - see https://englishlive.ef.com/de-de/blog/grammatik-und-vokabeln/englischen-apostroph/
 %
 %
-%

### 2021-08-15 doc/thesis/chapters/introduction.tex
minor adaptions Ch.1

@@
 
-In the master's degree program Embedded Systems Design a wide range of topics relate to digital signal processing, chip design and software development.
+In the master's degree program Embedded Systems Design, a wide range of topics relate to digital signal processing, chip design and software development.
 All of these disciplines find their application in an embedded system.
@@
 Instead, it is often treated as a standalone part that fulfills its own, specific functionality.
-Therefore, one of the main motivation points of this thesis is to combine several disciplines and multiple implementation methods into a single system design, and treat it in this common project context.\\
+Therefore, one of the main motivation points for this thesis is to combine several disciplines and multiple implementation methods into a single system design, and treat it in this common project context.\\
 
 In a more general perspective, the current electronics market situation is a huge motivational factor as well.
-Smart devices, smart sensors, Internet of Things (IoT) and many similar terms are well-known and are becoming more and more present in the publics' daily lives.
-In some degree, all of these devices represent embedded systems, since they all consist of processing units, combined with sensors and interfaces to communicate with their outside world.
-In the future, the number of devices and the range of applications in this area is going to grow from todays' point of view.
-This will require a lot of engineering work to advance the state of technology we have today and explore new areas of application.\\
+Smart devices, smart sensors, Internet of Things (IoT) and many similar terms are well-known and are becoming more and more present in the public's daily lives.
+To some degree, all of these devices represent embedded systems, since they all consist of processing units, combined with sensors and interfaces to communicate with their surrounding world.
+In the future, the number of devices and the range of applications in this area is going to grow from today's point of view.
+This will require a lot of engineering work to advance the state of technology we have today, to push the limits and explore new areas of applications.\\
 
@@
 
-Those major reasons lead to the choice of topic for this thesis -- understanding all the single parts by themselves, while maintaining the understanding over an integrated embedded system that fulfills a task as a whole.
+Those major reasons lead to the choice of topic for this thesis -- having the competence for all the single parts by themselves, while maintaining the understanding over the integrated embedded system that fulfills a task as a whole.
 
@@
 The project of choice is a digital FM broadcast radio receiver.
-It combines several disciplines that appear in the master's program Embedded Systems Design, such as digital signal processing (DSP), software development and system design and thus is a perfect candidate for this project.\\
+It combines several disciplines that appear in the master's program Embedded Systems Design, such as digital signal processing, software and hardware development, as well as system design and therefore is a perfect candidate for this project.\\
 
@@
 This is subjectively more attractive than a generic data stream.
-Additionally, it is a comparably simple signal to demodulate, but still requires several techniques of signal processing to decode it.
+Additionally, it is a comparably simple signal to demodulate, but still requires several techniques of signal processing to achieve that.
 
@@
 The main objective of this thesis is to elaborate a system design for an embedded system, that combines all the knowledge of different disciplines and topics which are accumulated over the time of the university program, in a single project.
-This includes the development of a system model in a tool like Matlab, the practical usage and implementation of DSP in an FPGA, the development of a digital receiver and the deployment on actual hardware.\\
+This includes the development of a system model in a tool like Matlab, the practical usage and implementation of a digital signal processing (DSP) chain in an FPGA, the development of a digital receiver and the deployment on actual hardware.\\
 
 The receiver implementation on the FPGA is to be done in two variants.
-The first variant is traditional, manually written hardware description language (HDL) in the programming language VHDL, as it has been done for years.
-For the second variant, high-level synthesis (HLS) is employed, to generate the low-level HDL from high-level \cplusplus\ code.
+The first variant is traditional, manually written hardware description language (HDL) in the programming language VHDL.
+For the second variant, high-level synthesis (HLS) is employed, which generates the low-level HDL from high-level \cplusplus\ code.
 These two methods are implemented and compared on the basis of usability and some metrics.\\
@@
 By the end of this thesis, the implementation of an FM radio receiver in multiple methods and abstraction levels should be achieved, as well as a good understanding of the overall system level design.
-Knowledge of each methods' preferred field of application, efficiency and application should be gained.
+Knowledge of each method's advantages and disadvantages, as well as efficiency and useability should be gained.
 
@@
 The actual implementation, with a detailed explanation of some selected components, is covered in Chapter \ref{cha:Implementation}.
-Chapter \ref{cha:DeploymentOnHardware} gives an overview of the required parts of development that are necessary to deploy the developed IPs on actual hardware.
+Chapter \ref{cha:DeploymentOnHardware} gives an overview of the required parts of development that are necessary to deploy the deeloped system on actual hardware.
 The comparison and achieved results are covered in Chapter \ref{cha:ComparisonAndResults}.

### 2021-08-15 doc/thesis/chapters/deployment-on-hardware.tex
all /ref to ~/ref (to avoid newline separations)

@@
 
-Chapter \ref{cha:Implementation} already described the development of the main IP core, which is the FM radio receiver in the VHDL and HLS variants.
+Chapter~\ref{cha:Implementation} already described the development of the main IP core, which is the FM radio receiver in the VHDL and HLS variants.
 In order to embed these IPs on the actual FPGA, or rather SoC hardware, a significant additional amount of implementation is neccessary.
@@
 
-The block diagram in Figure \ref{fig:bd_impl_vivado} shows the implemented architecture that is used in the SoC.
+The block diagram in Figure~\ref{fig:bd_impl_vivado} shows the implemented architecture that is used in the SoC.
 The various design domains in and around the SoC are depicted in different visual representations.
@@
 Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.
-This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section \ref{sec:SimilarProcess}.
+This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section~\ref{sec:SimilarProcess}.
 Specifically, this implemented system design enables the application of the steps 'Assess Performance' and 'Re-evaluate', since the IPs' performance can be measured and their design can be re-evaluated.\\

### 2021-08-15 doc/thesis/chapters/dsp.tex
all /ref to ~/ref (to avoid newline separations)

@@
 \noindent
-The block diagram in Figure \ref{fig:bd_dsp_overview} gives a high-level overview of a receiver structure.
+The block diagram in Figure~\ref{fig:bd_dsp_overview} gives a high-level overview of a receiver structure.
 
@@
 
-Figure \ref{fig:bd_superhet} shows a block diagram of a superhet receiver.
+Figure~\ref{fig:bd_superhet} shows a block diagram of a superhet receiver.
 The antenna signal first runs through a band selection filter.
@@
 
-A block diagram, as well as the signal spectrum of the process is shown in Figure \ref{fig:bd_down_conversion_spectrum}.
+A block diagram, as well as the signal spectrum of the process is shown in Figure~\ref{fig:bd_down_conversion_spectrum}.
 In this example, the frequency of the LO is set to shift the signal from $f_m$ to baseband.
@@
 Their respective amplitudes are $A_m$ and $A_c$.
-This equation is depicted as a block diagram and signal spectrum in Figure \ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.
+This equation is depicted as a block diagram and signal spectrum in Figure~\ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.
 The signal $m(t)$ is a baseband signal.
@@
 
-The graphical representation in the frequency spectrum is shown in Figure \ref{fig:down_conversion_spectrum_steps}.
+The graphical representation in the frequency spectrum is shown in Figure~\ref{fig:down_conversion_spectrum_steps}.
 As mentioned above, the assumption is a direct conversion to baseband, so $w_m = w_c$.
@@
 These replicas overlap, if $f_s$ does not meet the theorem, which results in signal distortion.
-Figure \ref{fig:aliasing_freq_spectrum} illustrates the effect of aliasing in the frequency spectrum \cite{ThyagarajanK.S2019ItDS}.
+Figure~\ref{fig:aliasing_freq_spectrum} illustrates the effect of aliasing in the frequency spectrum \cite{ThyagarajanK.S2019ItDS}.
 
@@
 % More info on modulation index: https://fas.org/man/dod-101/navy/docs/es310/FM.htm
-Formula \eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
+Formula~\eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
 A widely used example application of FM is broadcast radio, where audio streams are transmitted.
@@
 \end{equation}
-By inserting this message signal into the generic FM Equation \eqref{equ:fm_func}, the final formula for an FM signal transforms into the following form.
+By inserting this message signal into the generic FM Equation~\eqref{equ:fm_func}, the final formula for an FM signal transforms into the following form.
 \begin{equation}
@@
 
-Figure \ref{fig:fm_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
+Figure~\ref{fig:fm_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
 For a better overview, only the positive spectum part is shown.
@@
 It is used to generate a stereo audio signal, in combination with the mono signal.
-This process is explained into more detail in Chapter \ref{sec:demodulation-of-broadcast-fm}.
+This process is explained into more detail in Chapter~\ref{sec:demodulation-of-broadcast-fm}.
 
@@
 Different strategies can be implemented for both parts.
-A block diagram for a frequency discriminator with its two main parts is shown in Figure \ref{fig:bd_freq_discriminator}.
+A block diagram for a frequency discriminator with its two main parts is shown in Figure~\ref{fig:bd_freq_discriminator}.
 
@@
 The differentiator already converts the FM signal into an AM signal, but still leaves an FM content.
-The transformation that happens here can be expressed in a formula, by simply differentiating Equation \eqref{equ:fm_func}.
+The transformation that happens here can be expressed in a formula, by simply differentiating Equation~\eqref{equ:fm_func}.
 This operation, with some additional simple algebraic transformations, results in
@@
 
-A time-domain example for this signal is shown in the top left diagram in Figure \ref{fig:time_domain_envelope_detect}.\\
+A time-domain example for this signal is shown in the top left diagram in Figure~\ref{fig:time_domain_envelope_detect}.\\
 
@@
 The transmitted signal is subject to random additive noise on the transmission channel.
-The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Equation \eqref{equ:differentiator_two_samples}.
+The differentiator would deliver wrong results because of this noise, when substracting two consecutive samples, as described in Equation~\eqref{equ:differentiator_two_samples}.
 To achieve a constant amplitude, the received complex baseband signal $s(t)$ can be normalized to a value of one.
@@
 
-The differentiated signal still consists of a FM part, as explained above, using Equation \eqref{equ:fm_demod_discriminator}.
+The differentiated signal still consists of a FM part, as explained above, using Equation~\eqref{equ:fm_demod_discriminator}.
 In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.
-A block diagram is shown in Figure \ref{fig:bd_freq_discriminator}.\\
+A block diagram is shown in Figure~\ref{fig:bd_freq_discriminator}.\\
 
-The method implemented in Figure \ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as `Asynchronous Half-Wave Envelope Detector' \cite{DSPRelatedEnvelopeDetection}.
+The method implemented in Figure~\ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as `Asynchronous Half-Wave Envelope Detector' \cite{DSPRelatedEnvelopeDetection}.
 It consists of a thresholding unit to remove negative values and a conventional lowpass filter.
@@
 This section describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
-The assumption here is, that the actual FM demodulation, as described in Chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly, including the de-emphasis of the signal.
+The assumption here is, that the actual FM demodulation, as described in Chapter~\ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly, including the de-emphasis of the signal.
 So at this point the FM signal is converted to an AM signal.
@@
 
-A frequency spectrum of the multiplex signal of an FM channel is shown in Figure \ref{fig:fm_channel_baseband_freqs}.
+A frequency spectrum of the multiplex signal of an FM channel is shown in Figure~\ref{fig:fm_channel_baseband_freqs}.
 
@@
 
-The block diagram in Figure \ref{fig_bd_stereo_demod} illustrates the signal processing chain for stereo audio.
+The block diagram in Figure~\ref{fig_bd_stereo_demod} illustrates the signal processing chain for stereo audio.
 In the block diagram, $x(t)$ represents the multiplex signal from the FM demodulator.
@@
 Subsequently, a lowpass limits signal bandwidth to 15~kHz, which removes the generated modulation artifacts.
-The lower branch lowpass-filters the summation, or mono signal, as described in Section \ref{subsec:demod_mono}.
-The rightmost part in the diagram performs the combination of the mono and the difference signal, according to Equation \eqref{equ_stereo_from_sum_diff}.
+The lower branch lowpass-filters the summation, or mono signal, as described in Section~\ref{subsec:demod_mono}.
+The rightmost part in the diagram performs the combination of the mono and the difference signal, according to Equation~\eqref{equ_stereo_from_sum_diff}.

### 2021-08-15 doc/thesis/chapters/hls.tex
all /ref to ~/ref (to avoid newline separations)

@@
 This is especially important when an FPGA design is combined with a software part, like in a System On Chip, since software development is comparably fast.
-Figure \ref{fig:timeline_performance_first_prototype} schematically displays the difference in development time, over performance and degree of optimization, when the same functionality of a prototype is implemented in different methods for various targets.
+Figure~\ref{fig:timeline_performance_first_prototype} schematically displays the difference in development time, over performance and degree of optimization, when the same functionality of a prototype is implemented in different methods for various targets.
 The compared targets are an x86 processor CPU, a digital signal processor (DSP), a graphics processing unit (GPU) and the FPGA design in RTL or HLS implementation.
@@
 
-The code example in Program \ref{prog:axi4_lite} shows the specification of an AXI4-Lite memory-mapped slave interface.
+The code example in Program~\ref{prog:axi4_lite} shows the specification of an AXI4-Lite memory-mapped slave interface.
 The provided registers are separated into configuration registers and status registers.
@@
 
-The second code example, Program \ref{prog:axi4_stream}, demonstrates the equally simple definition of an AXI4-Stream interface.
+The second code example, Program~\ref{prog:axi4_stream}, demonstrates the equally simple definition of an AXI4-Stream interface.
 
@@
 
-Section \ref{sec:hls:workflow} above already mentioned the topics of testbenches and simulation types.
+Section~\ref{sec:hls:workflow} above already mentioned the topics of testbenches and simulation types.
 The following descriptions shall be seen as a comprehensive extension around these topics.

### 2021-08-15 doc/thesis/chapters/implementation.tex
sample dec
all /ref to ~/ref (to avoid newline separations)

@@
 
-The theory behind the demodulation of an FM signal is already explained in Chapter \ref{cha:SignalProcessingTheory}.
-The demodulation of an FM broadcast channel specifically, is described in Section \ref{sec:demodulation-of-broadcast-fm}.
+The theory behind the demodulation of an FM signal is already explained in Chapter~\ref{cha:SignalProcessingTheory}.
+The demodulation of an FM broadcast channel specifically, is described in Section~\ref{sec:demodulation-of-broadcast-fm}.
 In this section however, the DSP chain is described into more detail and with more information, as it is implemented in the various methods.\\
@@
 \noindent
-Figure \ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram.\\
+Figure~\ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram.\\
 
@@
        % TODO
-      The FM demodulator is implemented as a Baseband Delay Modulator, as described in Section \ref{sec:BasebandDelayDemodulator}.
+      The FM demodulator is implemented as a Baseband Delay Modulator, as described in Section~\ref{sec:BasebandDelayDemodulator}.
 
@@
           Thus, their filter order may be different as well, which leads to a non-matching group delay.
-          Since the two signals are multiplicated in the subsequent mixer, and an important factor is the phase-coherency of the recovered carrier (as explained in Section \ref{sec:demodulation-of-broadcast-fm}), the group delay compensation becomes necessary here.\\
+          Since the two signals are multiplicated in the subsequent mixer, and an important factor is the phase-coherency of the recovered carrier (as explained in Section~\ref{sec:demodulation-of-broadcast-fm}), the group delay compensation becomes necessary here.\\
 
@@
       The 38~kHz carrier can be recovered from the 19~kHz pilot tone.
-      It is important that the recovered carrier is phase-coherent with the received input, as described in Section \ref{sec:demodulation-of-broadcast-fm}.
+      It is important that the recovered carrier is phase-coherent with the received input, as described in Section~\ref{sec:demodulation-of-broadcast-fm}.
       This can be achieved, by simply multiplying the pilot tone with itself, which is the way it is implemented here.\\
@@
 
-The transmitter is implemented according to the structure that is depicted as a block diagram in Figure \ref{fig:bd_matlab_transmitter}.
-To facilitate an easy understanding of the block diagram, it is of advantage to keep in mind that it exactly resembles the frequency spectrum that is previously shown in Figure \ref{fig:fm_channel_baseband_freqs}.\\
+The transmitter is implemented according to the structure that is depicted as a block diagram in Figure~\ref{fig:bd_matlab_transmitter}.
+To facilitate an easy understanding of the block diagram, it is of advantage to keep in mind that it exactly resembles the frequency spectrum that is previously shown in Figure~\ref{fig:fm_channel_baseband_freqs}.\\
 
@@
 
-The achieved sample rate of 960~kHz could actually be chosen much lower, since an FM channels' maximum frequency is 75~kHz (see Section \ref{sec:FrequencySpectrumOfABroadcastFmChannel}), and according to the Nyquist theorem the sample rate thus only needs to have 150~kHz (see Section \ref{sec:SamplingTheorem}).
+The achieved sample rate of 960~kHz could actually be chosen much lower, since an FM channels' maximum frequency is 75~kHz (see Section~\ref{sec:FrequencySpectrumOfABroadcastFmChannel}), and according to the Nyquist theorem the sample rate thus only needs to have 150~kHz (see Section~\ref{sec:SamplingTheorem}).
 However, the higher rate was chosen, mainly because a higher number of signal samples beautify any plots that are created during analysis.
@@
 
-At this point in the DSP chain, the signal resembles the frequency spectrum that is displayed in Figure \ref{fig:fm_channel_baseband_freqs}.\\
+At this point in the DSP chain, the signal resembles the frequency spectrum that is displayed in Figure~\ref{fig:fm_channel_baseband_freqs}.\\
 
@@
 
-The receiver is implemented, following the block diagram in Figure \ref{fig:bd_dsp_detailed}.
+The receiver is implemented, following the block diagram in Figure~\ref{fig:bd_dsp_detailed}.
 Further assumptions and simplifications, such as the fixed point arithmetic are described into detail above.
@@
 
-An example of a plot that is created by the analysis is depicted in Figure \ref{fig:matlab_analysis_time_domain}.
+An example of a plot that is created by the analysis is depicted in Figure~\ref{fig:matlab_analysis_time_domain}.
 The diagram shows the respective signals in the time domain.
@@
 Further analysis is done in the frequency domain, especially in order to better understand the modulation of the different spectral parts in the FM channel.
-An example thereof is shown in Figure \ref{fig:matlab_analysis_freq_domain}.
+An example thereof is shown in Figure~\ref{fig:matlab_analysis_freq_domain}.
 The spectrum of an FM channel multiplex signal, as it is generated by a transmitter before the modulation, is shown in this diagram.
@@
 GNU Radio provides a comprehensive set of DSP blocks, which are available to build various functionalities.
-Multiple blocks are connected to a block diagram, that is very similar to the one in Figure \ref{fig:bd_matlab_transmitter}.
+Multiple blocks are connected to a block diagram, that is very similar to the one in Figure~\ref{fig:bd_matlab_transmitter}.
 The USRP device is used to broadcast the RF signal over the air.\\
@@
 Please make sure to adapt your setup to the respective local regulations, to avoid any legal issues.
-Some of the regulations are mentioned in Section \ref{sec:FrequencyBand}, but may differ in various countries, however.\\
+Some of the regulations are mentioned in Section~\ref{sec:FrequencyBand}, but may differ in various countries, however.\\
 
 \noindent
-Figure \ref{fig:bd_gnuradio_setup_transmitter} shows a block diagram of the transmitter application.
+Figure~\ref{fig:bd_gnuradio_setup_transmitter} shows a block diagram of the transmitter application.
 
@@
 This receiver works with both devices, the USRP b200mini and the RTL-SDR.
-Figure \ref{fig:bd_gnuradio_setup_receiver} shows a block diagram of the receiver application.
+Figure~\ref{fig:bd_gnuradio_setup_receiver} shows a block diagram of the receiver application.
 
@@
 
-Figure \ref{fig:bd_cocotb_interface_simulator} gives an overview of how cocotb interfaces with a simulator.
+Figure~\ref{fig:bd_cocotb_interface_simulator} gives an overview of how cocotb interfaces with a simulator.
 The Python test cases are shown as a list of coroutines, which are managed by a scheduler.
@@
 
-The architecture for this testbench is created under the assumptions and according to the guidelines that are developed in Chapter \ref{cha:SystemArchitectureAndConcept}.\\
+The architecture for this testbench is created under the assumptions and according to the guidelines that are developed in Chapter~\ref{cha:SystemArchitectureAndConcept}.\\
 
-Figure \ref{fig:bd_testbench_architecture_vhdl} shows a block diagram of the implemented architecture.
+Figure~\ref{fig:bd_testbench_architecture_vhdl} shows a block diagram of the implemented architecture.
 The split between the Python environment and the VHDL implementation is clearly visible but is represented from the developers perspective.
-The actual simulator interfaces, as described in Section \ref{sub:VHDL:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.\\
+The actual simulator interfaces, as described in Section~\ref{sub:VHDL:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.\\
 
@@
 
-Figure \ref{fig:register_engine} gives an overview of how the Register Engine script works.
+Figure~\ref{fig:register_engine} gives an overview of how the Register Engine script works.
 The Python script takes the register definitions from the YAML file, and uses templates to generate the respecitive output files.\\
@@
 The execution of the testbench is done in the Xilinx HLS software.
-A detailed explanation of how an HLS testbench works in general is given in Chapter \ref{cha:HLS}.\\
+A detailed explanation of how an HLS testbench works in general is given in Chapter~\ref{cha:HLS}.\\
 
@@
 
-The testbench is especially designed around the 'Design For Reuse' segment of the SIMILAR concept, which is introduced in Section \ref{sec:SimilarProcess}.
+The testbench is especially designed around the 'Design For Reuse' segment of the SIMILAR concept, which is introduced in Section~\ref{sec:SimilarProcess}.
 This is the case because of two major parts.
@@
 
-In the previous Section \ref{sec:impl:hls:testbench} it is explained, that the HLS testbench generates output files during its execution.
+In the previous Section~\ref{sec:impl:hls:testbench} it is explained, that the HLS testbench generates output files during its execution.
 These output files are analyzed with the exact same Python analysis code, that is already used in the VHDL testbench analysis.
@@
 
-The testbench is developed in the language \cplusplus\ and does not have any restrictions in the usage thereof, as Section \ref{sec:hls:workflow} described.
+The testbench is developed in the language \cplusplus\ and does not have any restrictions in the usage thereof, as Section~\ref{sec:hls:workflow} described.
 Thus, the \cplusplus\ code that is later used in the application software to interface with the IP, can already be used in the testbench here, for the same purpose.
@@
 
-Figure \ref{fig:testbench_hls_overview_reuse} gives an overview of the testbench and specifically points out the reuse of the application software.
+Figure~\ref{fig:testbench_hls_overview_reuse} gives an overview of the testbench and specifically points out the reuse of the application software.
 The diagram displays the HLS IP in the testbench and in the hardware with dotted lines, to emphasize that they are actually the same.\\
@@
 
-HLS can automatically implement interfaces, as described in Section \ref{sec:hls:coding:interfaces}.
+HLS can automatically implement interfaces, as described in Section~\ref{sec:hls:coding:interfaces}.
 This is a major feature that is used in the implementation of the HLS IP.\\
@@
 Additionally, a data-only interface is used for the connected LEDs.
-The actual implementation is demonstrated in Program \ref{prog:hls_impl_interfaces}.
+The actual implementation is demonstrated in Program~\ref{prog:hls_impl_interfaces}.
 These few lines of code already implement the entire functionality of each interface type.\\
@@
 \subsection{Sample Rate Reduction}
+\label{sec:impl:hls:sample_rate_reduction}
 
 Multiple steps in the DSP chain require the operation of sample rate reduction.
-This is explained in Section \ref{sec:impl:dsp-chain}.\\
+This is explained in Section~\ref{sec:impl:dsp-chain}.\\
 
@@
 
-The implementation of the sample rate reduction can be shown in a short code example in Program \ref{prog:sample_rate_reduction_fifo_behaviour}.
+The implementation of the sample rate reduction can be shown in a short code example in Program~\ref{prog:sample_rate_reduction_fifo_behaviour}.
 The FM Demodulator runs in the oversampled domain and is executed \texttt{OSR\_RX} times in the inner for-loop.
@@
 Several samples are processed, where only the last one is forwarded to the subsequent processing step.
-The FIFO behaviour of AXI stream variables is described in Section \ref{sec:hls:coding:interfaces}.
+The FIFO behaviour of AXI stream variables is described in Section~\ref{sec:hls:coding:interfaces}.
 

### 2021-08-15 doc/thesis/chapters/introduction.tex
all /ref to ~/ref (to avoid newline separations)

@@
 
-Chapter \ref{cha:SignalProcessingTheory} describes the signal processing theory that is required to demodulate an FM broadcast radio signal.
+Chapter~\ref{cha:SignalProcessingTheory} describes the signal processing theory that is required to demodulate an FM broadcast radio signal.
 This includes an introduction to receiver types, the description of an FM signal generally and also specifically as it is used in FM audio broadcasting, as well as various general topics of signal processing.
-Chapter \ref{cha:SystemDesignProcess} covers the topic of a certain system design process, as well as system design tools that specifically target FPGAs.
-Chapter \ref{cha:HLS} is about high-level synthesis, which includes a general introduction to the state of the art, as well as a description of the workflow, inclusively the coding and testbench features and characteristics.
-Chapter \ref{cha:SystemArchitectureAndConcept} describes the chosen system architecture and concept for the project.
-The actual implementation, with a detailed explanation of some selected components, is covered in Chapter \ref{cha:Implementation}.
-Chapter \ref{cha:DeploymentOnHardware} gives an overview of the required parts of development that are necessary to deploy the deeloped system on actual hardware.
-The comparison and achieved results are covered in Chapter \ref{cha:ComparisonAndResults}.
-Chapter \ref{cha:ConclusionAndProspect} concludes this thesis with a retrospective and gives an outlook of potential future extensions to the project.
+Chapter~\ref{cha:SystemDesignProcess} covers the topic of a certain system design process, as well as system design tools that specifically target FPGAs.
+Chapter~\ref{cha:HLS} is about high-level synthesis, which includes a general introduction to the state of the art, as well as a description of the workflow, inclusively the coding and testbench features and characteristics.
+Chapter~\ref{cha:SystemArchitectureAndConcept} describes the chosen system architecture and concept for the project.
+The actual implementation, with a detailed explanation of some selected components, is covered in Chapter~\ref{cha:Implementation}.
+Chapter~\ref{cha:DeploymentOnHardware} gives an overview of the required parts of development that are necessary to deploy the deeloped system on actual hardware.
+The comparison and achieved results are covered in Chapter~\ref{cha:ComparisonAndResults}.
+Chapter~\ref{cha:ConclusionAndProspect} concludes this thesis with a retrospective and gives an outlook of potential future extensions to the project.
 

### 2021-08-15 doc/thesis/chapters/results.tex
remove unnecessary appendix
all /ref to ~/ref (to avoid newline separations)

@@
 However, from the perspective of the communicating blocks, the interface is usable like a regular AXI stream.
-The implementation details are explained in Section \ref{sec:impl:vhdl:interfaces}.\\
+The implementation details are explained in Section~\ref{sec:impl:vhdl:interfaces}.\\
 
@@
 There seems to be an issue in the signal separation between the left and right audio channels.
-This is visible in the analysis of the audio signal, as shown is Figure \ref{fig:audio_output_compare_tb_vs_hw}.
+This is visible in the analysis of the audio signal, as shown is Figure~\ref{fig:audio_output_compare_tb_vs_hw}.
 The signal strength of the left channel in the upper diagram is weaker than expected, while the right channel in the lower diagram is stronger.
@@
 Ideally, the values should be exactly the same.
-However, Figure \ref{fig:audio_output_compare_tb_vs_hw} shows that this is not always the case.
+However, Figure~\ref{fig:audio_output_compare_tb_vs_hw} shows that this is not always the case.
 Again, the HLS implementation does have a matching result, whereas the VHDL variant has deviating values.
@@
 Several parts of the system design are taken into account to analyze the lines of code, to give an impression over the largest code parts in the project.
-Figure \ref{fig:lines_of_code_pie_charts} shows multiple pie charts, which represent the proportions of the lines of code in the respective system design parts.
-Table \ref{tab:lines_of_code} displays the actual numbers of code lines, to quantify the size of the respective parts.\\
+Figure~\ref{fig:lines_of_code_pie_charts} shows multiple pie charts, which represent the proportions of the lines of code in the respective system design parts.
+Table~\ref{tab:lines_of_code} displays the actual numbers of code lines, to quantify the size of the respective parts.\\
 
@@
 The implementation is always considered as the HDL description, including the corresponding testbench code, to verify the respective correct functionality.
-Table \ref{tab:implementation_time_logging} shows the direct comparison between HLS and VHDL implementation times for these features.\\
+Table~\ref{tab:implementation_time_logging} shows the direct comparison between HLS and VHDL implementation times for these features.\\
 
@@
 The implementation of the sample decimation functionality differs slightly between the two variants, HLS and VHDL.
-This is explained in Section \ref{sec:impl:hls:sample_rate_reduction}.
+This is explained in Section~\ref{sec:impl:hls:sample_rate_reduction}.
 The implementation time of the HLS version is severely larger than the VHDL version.
@@
 A thorough investigation in several user guides, in combination with step-by-step debugging, the issue was found and led to the current, working implementation, which processes a number of samples, before it forwards one sample.
-The process of finding this issue explains the largely varying implementation times, as listed in Table \ref{tab:implementation_time_logging}.
+The process of finding this issue explains the largely varying implementation times, as listed in Table~\ref{tab:implementation_time_logging}.
 
@@
 In HLS this is not required, since the validity of a signal is determined internally by the HLS compiler, according to the usage of the respective data variables.
-Another factor that adds up to the implementation time is the number of lines of code, as presented in Table \ref{tab:lines_of_code}.
+Another factor that adds up to the implementation time is the number of lines of code, as presented in Table~\ref{tab:lines_of_code}.
 The VHDL IP design requires 3000 lines, while the HLS IP design only takes 920 for the same functionality.
@@
 
-Figure \ref{fig:impl_time_analysis} shows this direct comparison of processed samples per second for the implementation variants.
+Figure~\ref{fig:impl_time_analysis} shows this direct comparison of processed samples per second for the implementation variants.
 The measurements always include the testbench setup-phase, and the run-phase.
@@
 
-Figure \ref{fig:hardware_utilization} shows the utilization numbers of the different FPGA resources.
+Figure~\ref{fig:hardware_utilization} shows the utilization numbers of the different FPGA resources.
 It reveals that the manual VHDL implementation clearly requires more resources in the FPGA.
@@
 Instead, the testbench of the integration test is executed, which produces a signal wave form.
-The latency is measured from this wave form and results in a number of around 2470 clock cycles.
+The latency is measured from this wave form and results in a number of 2470 clock cycles.
 With regards to latency, the two implementation variants appear to have a similar behaviour.

### 2021-08-15 doc/thesis/chapters/system-architecture-and-concept.tex
all /ref to ~/ref (to avoid newline separations)

@@
 \noindent
-A very high-level overview of the entire system is given by the block diagram in Figure \ref{fig:bd_system_overview_highlevel}.
+A very high-level overview of the entire system is given by the block diagram in Figure~\ref{fig:bd_system_overview_highlevel}.
 It shows the front end, which processes the received antenna signal, so that it can be used in the FM radio receiver block.
@@
 This mainly regards the signal processing theory, to be able to decode the signal that is received by the antenna.
-Chapter \ref{cha:SignalProcessingTheory} describes this part into detail.
+Chapter~\ref{cha:SignalProcessingTheory} describes this part into detail.
 
@@
       This thesis focuses in \cplusplus and does not implement a SystemC variant, because Xilinx has deprecated SystemC support for synthesis in their latest HLS tools.
-      This, and the details of High-level synthesis (HLS) are explained in Chapter \ref{cha:HLS}.
+      This, and the details of High-level synthesis (HLS) are explained in Chapter~\ref{cha:HLS}.
 
@@
 
-An example of a system concept and its architecture is developed and shown in Figure \ref{fig:bd_system_overview_detailed}.
+An example of a system concept and its architecture is developed and shown in Figure~\ref{fig:bd_system_overview_detailed}.
 Here, a common data source is provided by the block called \textit{Front End}.

### 2021-08-15 doc/thesis/chapters/system-design-process.tex
all /ref to ~/ref (to avoid newline separations)

@@
 
-Each of these terms represents a separate step in the process diagram shown in Figure \ref{fig:similar_process}.
+Each of these terms represents a separate step in the process diagram shown in Figure~\ref{fig:similar_process}.
 The steps can be seen as iterative and parallel, as the process runs during the product development cycle.
@@
 A well-known process model is the so called V-Model.
-It is based on the similarly well-known Waterfall Model, and has its name because of the distinct V-shape of its process diagram, as it is shown in Figure \ref{fig:v_modell}.
+It is based on the similarly well-known Waterfall Model, and has its name because of the distinct V-shape of its process diagram, as it is shown in Figure~\ref{fig:v_modell}.
 One of the main characteristics of the V-Model is, that one phase needs to be completed before the following phase can start.
@@
 
-Furthermore, the V-model diagram in Figure \ref{fig:v_modell} shows the difference between validation and verification, as it is explained in the previous Section \ref{sec:ValidationAndVerification}.
+Furthermore, the V-model diagram in Figure~\ref{fig:v_modell} shows the difference between validation and verification, as it is explained in the previous Section~\ref{sec:ValidationAndVerification}.
 
@@
 \noindent
-HLS is described into more detail in Chapter \ref{cha:HLS}.
+HLS is described into more detail in Chapter~\ref{cha:HLS}.
 
@@
 
-Figure \ref{fig:simulink_system_generator} shows this entire design approach in a block diagram.
+Figure~\ref{fig:simulink_system_generator} shows this entire design approach in a block diagram.
 The abbreviation ISE stands for Integrated Software Environment, and is Xilinx' development environment that includes the FPGA synthesis tool.
@@
 VHDL and Verilog require, or allow, a very detailed description of the RTL code, while SystemC, \cplusplus, or Matlab and Simulink provide a high level of abstraction.
-In Figure \ref{fig:fpga_design_levels}, the various implementation methods are compared.
+In Figure~\ref{fig:fpga_design_levels}, the various implementation methods are compared.
 The respective level of abstraction is higher towards the upper end, while the level of detail increases towards the bottom of the figure.\\
@@
 
-%TODO: try to get this figure up into the last page...
+%TODO: try to get this figure into the upper page...

### 2021-08-15 doc/thesis/chapters/introduction.tex
adapt motivation

@@
 Instead, it is often treated as a standalone part that fulfills its own, specific functionality.
-Therefore, one of the main motivation points for this thesis is to combine several disciplines and multiple implementation methods into a single system design, and treat it in this common project context.\\
+Therefore, one of the main motivation points for this thesis is to combine several disciplines and multiple implementation methods into a single system design, to find a universal workflow for design and verification, and treat them in this common project context.\\
 
@@
 However, not only understanding single parts, but to have an overview and understanding of the entire, integrated system and the interaction between those single parts is important in an embedded system.\\
+%This further creates the motivation to find a solution for a universal design flow, that enables to combine the different implementation methods into a single system design.\\
 

### 2021-08-15 doc/thesis/chapters/dsp.tex
adapt all the diagrams
minor modifications

@@
 The superheterodyne receiver is often referred to as \textit{superhet}, or \textit{super} in the literature.
-The main characteristic of a superhet is, that it converts the radio-frequency (RF) antenna signal to an IF in a first stage, instead of going to baseband directly.\\
+The main characteristic of a superhet is that it converts the radio-frequency (RF) antenna signal to an IF in a first stage, instead of going to baseband directly.\\
 
@@
 
-The depicted block diagram represents a superhet receiver structure, for example an FM broadcast receiver, where an audio output is generated from the antenna input.
+The depicted block diagram represents a superhet receiver structure, for example as it can be used in an FM broadcast receiver, where an audio output is generated from the antenna input.
 
@@
 These effects are not explained into more detail here, since this is not the focus of this thesis.
+Figure~\ref{fig:bd_direct_conversion} shows a block diagram of a direct-conversion receiver.
 
@@
 A block diagram, as well as the signal spectrum of the process is shown in Figure~\ref{fig:bd_down_conversion_spectrum}.
-In this example, the frequency of the LO is set to shift the signal from $f_m$ to baseband.
+In this example, the frequency of the LO is set to shift the signal from $f_m$, which is drawn in solid lines, to baseband, which is drawn in dashed lines.
 
@@
 
-The procedure that is used for down-conversion is an amplitude modulation (AM).
 The equation for a generic AM signal is $ S_{AM}(t) = A_m m(t) \cdot A_c c(t) $, where $m(t)$ represents the message signal, and $c(t)$ is the high-frequency carrier.
 Their respective amplitudes are $A_m$ and $A_c$.
-This equation is depicted as a block diagram and signal spectrum in Figure~\ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.
-The signal $m(t)$ is a baseband signal.
-It may be antenna signal that was filtered by a bandpass filter.
-The received signal is a real signal, and thus, its negative frequency spectrum is an exact mirror of the positive frequencies.
+This equation is depicted as a block diagram and signal spectrum in Figure~\ref{fig:bd_down_conversion_spectrum}, in which $S_{AM}$ is denoted as the baseband signal $m_{BB}$, because it is already down-converted.
+The original HF signal $m$, before the down-conversion, is a bandpass signal, such as an antenna signal that was filtered by a bandpass filter.
+This received signal is a real signal, and thus, its negative frequency spectrum is an exact mirror of the positive frequencies.
 This fact can be explained in two simple equations.\\
@@
 Sample rate reduction is also called downsampling.
-Downsampling is a way to reduce the amount of data that needs to be processed, without losing any of the signals' information.
+It is a way to reduce the amount of data that needs to be processed, without losing any of the signals' information.
 In some applications, the digital signal is oversampled by a large factor.
 However, if the frequency spectrum of the desired signal is a narrow band, a high oversampling rate is not necessary and the signal can be downsampled without losing any information content.
-To guarantee this, the Nyquist sampling theorem must still be met after the downsampling process.
+To guarantee this, the Nyquist sampling theorem must still be met after the downsampling process.\\
 
@@
 As an example, assuming the signal of interest has a bandwidth of 20~kHz.
-The Nyquist theorem thus requires a remaining sampling rate of at least twice the signals' frequency, so 40~kHz.
+The Nyquist theorem thus requires a remaining sampling rate of at least twice the signal's frequency, so 40~kHz.
 The signal is sampled with an oversampling factor of 10 -- meaning a sample rate of 200~kHz.
@@
 
-More detailed information can be found in \cite[Ch.~6.9]{AlessioSilviaMaria2016DSPa}, \cite[Ch.~4.1]{GopiE.S2018MDSP} and \cite[Ch.~10.2.2]{ThyagarajanK.S2019ItDS}.
+More detailed information can be found in the literature \cite[Ch.~6.9]{AlessioSilviaMaria2016DSPa}, \cite[Ch.~4.1]{GopiE.S2018MDSP} and \cite[Ch.~10.2.2]{ThyagarajanK.S2019ItDS}.
 
@@
     (L+R) + (L-R) = (2)L \\
-    (L+R) - (L-R) = (2)R
+    (L+R) - (L-R) = (2)R,
     \label{equ_stereo_from_sum_diff}
@@
 
+\noindent
 where $(L+R)$ represents the mono part and $(L-R)$ the stereo difference.\\

### 2021-08-16 doc/thesis/chapters/dsp.tex
adapt dsp chapter

@@
 The probably best known usecase therefor is commercial broadcast radio, where an audio stream is transmitted.
-Devices to receive these streams are available for low prices to the public.\\
-
+Devices to receive these streams are available for low prices to the public.
 This section describes the main properties of broadcast FM, such as the mathematical description, frequency bands that are used, or the specific frequency parts within a channel spectrum.
@@
 The mathematical description of an FM baseband signal can be described with the formulas presented in this section.
-FM encodes the information content in its instantanious frequency.
+FM encodes the information content in its instantaneous frequency.
 This means that the measured frequency at any moment in time represents a specific value of a transmitted message.
-For a general classification, FM belongs to the group of angle, or phase modulated signals (PM).
-The simple reason therefor is, that a frequency has a direct relationship to an angle, if the signal is seen on a unit circle.\\
+For a general classification, FM belongs to the group of angle-, or phase modulated signals (PM).
+The simple reason therefor is that a frequency has a direct relationship to an angle, if the signal is seen on a unit circle.\\
 
-The instantanious frequency $f_i$ of an FM signal can be described as
+\noindent
+The instantaneous frequency $f_i$ of an FM signal can be described as
 \begin{equation}
@@
 where $f_c$ is the carrier- or center frequency, $\Delta f$ is the maximum frequency deviation and $m(t)$ is the information or message signal that is to be transmitted.
-Simply looking at this equation, the frequency deviation, which is sometimes also called the swing, varies in the range between $f_c \pm \Delta f$.
+Simply looking at this equation, the instantaneous frequency varies in the range between $f_c \pm \Delta f$, which is sometimes also called the swing.
 Considering the relationship between frequency and angle, as described above, and after some simple substitutions which will not be described into detail here, the equation for a generic frequency modulated wave is the following
 \begin{equation}
-  s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int m(t) dt \Big),
+  s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int_{0}^{t} m(\tau) d\tau \Big),
   \label{equ:fm_func}
@@
 % More info on modulation index: https://fas.org/man/dod-101/navy/docs/es310/FM.htm
-Formula~\eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
+Equation~\eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
 A widely used example application of FM is broadcast radio, where audio streams are transmitted.
@@
 Therefore, it is also referred to as `multiplex signal' in the literature.
-Several parts are described in the following paragraphs \cite{EbuFmRegulators},\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard},\cite{ElectronicsNotesFmTutorial}.
+Several of these parts are described in the following paragraphs, according to \cite{EbuFmRegulators},\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard} and \cite{ElectronicsNotesFmTutorial}.
 
@@
 \begin{equation}
-  \frac{d s(t)}{dt} = A_c 2 \pi \Big[f_c + k_f m(t) \Big] sin \Big(\omega_c t + 2 \pi k_f \int m(t) dt -pi \Big)
+  \frac{d s(t)}{dt} = A_c 2 \pi \Big[f_c + k_f m(t) \Big] sin \Big(\omega_c t + 2 \pi k_f \int_{0}^{t} m(\tau) d\tau - \pi \Big)
   \label{equ:fm_demod_discriminator}
@@
 It is described by the term within the square brackets.
-
 A time-domain example for this signal is shown in the top left diagram in Figure~\ref{fig:time_domain_envelope_detect}.\\
@@
 
+\noindent
 In a hardware implementation, a differentiation is simply performed by substracting two consecutive samples, like
 \begin{equation}
-  \frac{d s(t)}{dt} = s(t) - s(t-\Delta t)
+  \frac{d s(t)}{dt} = s(t) - s(t-\Delta t),
   \label{equ:differentiator_two_samples}
@@
 This section describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
-The assumption here is, that the actual FM demodulation, as described in Chapter~\ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly, including the de-emphasis of the signal.
-So at this point the FM signal is converted to an AM signal.
+The assumption here is, that the actual FM demodulation, as described in Chapter~\ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly.
+So at this point, the FM signal is converted to an AM signal.
 Thus, traditional AM DSP techniques can be employed to decode the FM broadcast channel.\\
 
+\noindent
 A frequency spectrum of the multiplex signal of an FM channel is shown in Figure~\ref{fig:fm_channel_baseband_freqs}.
@@
     (L+R) + (L-R) = (2)L \\
-    (L+R) - (L-R) = (2)R,
+    (L+R) - (L-R) = (2)R
     \label{equ_stereo_from_sum_diff}

### 2021-08-16 doc/thesis/chapters/system-design-process.tex
adapt system-design-process

@@
 The process of system design is a central topic in the development of any product.
-This chapter describes general techniques for system design.
-However, they can also be directly applied to FPGA design.
+This chapter describes a general technique for system design, which can however also be directly applied to FPGA design.
 Additionally, tools that are specifically used in FPGA design are introduced.
@@
 The SIMILAR process provides a general approach to problem solving.
-It can be applied to any kind of problem that can be of technical, as well as non-technical nature, in order to successfully achieve the specified requirements.
+It can be applied to any kind of problem that can be of technical-, as well as non-technical nature, in order to successfully achieve the specified requirements.
 The concept was developed by A.T. Bahill and B. Gissing, who evaluated and compared multiple existing processes to find their similarities.\\
@@
 
-The very first step of the process is to state the problem that is to be solved.
+\hspace{5mm} The very first step of the process is to state the problem that is to be solved.
 This can be the initial definition of a new product, but can also be the definition of an improvement that is to be made for an existing product.
@@
 
-Generally speaking, the problem statement serves the purpose to build a common understanding of the problem between a stakeholder and the developer.
+Generally speaking, the problem statement serves the purpose to build a common understanding of the problem between the stakeholders and the developers.
 In case there is a mismatch, an engineer may spend valuable time to develop a sophisticated solution to a problem that was not even a requirement.
@@
 
-In almost any kind of problem, there are multiple ways in order to achieve a successful solution.
+\hspace{5mm} In almost any kind of problem, there are multiple ways in order to achieve a successful solution.
 However, some solutions may be preferred over others because of certain criteria that are evaluated.
@@
 Further design steps, such as a first estimation of the intended design implementation, simulation results, or a prototype can provide more data and facts.
-These can be used to re-evaluate the alternatives and consequently optimize the design process.
+These can be used to re-evaluate the alternatives and subsequently optimize the design process.
 
@@
 
-Based on the previous steps, a model for the system is developed here.
-This includes a process model, that is used for stragetic planning towards cost reduction, optimization of the implementation effort, or the schedule of subprojects, and a product model, which supports trade-offs and the identification of risks, or helps to explain the system as a whole.\\
+\hspace{5mm} Based on the previous steps, a model for the system is developed here.
+This includes a process model, that is used for strategic planning towards cost reduction, optimization of the implementation effort, or the schedule of subprojects, and a product model, which supports trade-offs and the identification of risks, or helps to explain the system as a whole.\\
 
-Typically, a coarse model is created for almost all alternative designs that were found in the previous step.
-A more detailed model is developed for the single, selected design.
+Typically, a coarse model is created for almost all alternative designs that are found in the previous step.
+A more detailed model is only developed for the single, selected design.
 The model itself may be represented in many different ways, such as state machines, block diagrams, flow diagrams, simulations or object-oriented diagrams.
@@
 
-This step combines multiple single components into a larger, integrated system, which can then be treated as a whole.
+\hspace{5mm} This step combines multiple single components into a larger, integrated system, which can then be treated as a whole.
 In order to be able to do that, interfaces between the components and subsystems need to be defined.
 Ideally, these interfaces are designed in a way that the amount of data that needs to be exchanged is kept to a minimum.
-It is of advantage, if a subsystem only transfers finished products that it produced to another subsystem.
+Therefor, it is of advantage, if a subsystem only transfers finished products that it produced to another subsystem.
 The aim to develop simple interfaces is of special advantage in systems that include a feedback loop.
@@
 
-The system launch actually runs the system in a defined environment, so that it can perform what it is intended and designed to do.
+\hspace{5mm} The system launch actually runs the system in a defined environment, so that it can perform what it is intended and designed to do.
 The defined environment can be the final, operational environment, or just a simulation which provides an output that is detailed enough for an evaluation.
@@
 
-This step is all about measurements.
+\hspace{5mm} This step is all about measurements.
 The performance of a system can only be evaluated, if it can be measured by using different metrics.
@@
 
-The re-evaluate step is a central building block in the SIMILAR concept.
+\hspace{5mm} The re-evaluate step is a central building block in the SIMILAR concept.
 All of the other steps have a connection to it, in the form of a feedback loop.
@@
 The idea of re-evaluation and the subsequent improvement of a process is a widely applied and well-known concept.
-Its main function is to observe an output and to use this information to improve the systems' input.
+Its main function is to observe an output and to use this information to improve the system's input.
 
@@
 After all these positive arguments, there is a counter argument to it, though: designing a system with reuseable subsystems comes with cost.
-Additional effort is required to design the subsystems in a way so that they can be reused in another project.
-However, this additional effort may potentially already be paid off, when a single subsystem can be reused in a future product and thus, it is worth the effort.
+Additional effort is required to design the subsystems in a way so that they can be reused in another project at all.
+However, this additional effort may potentially already be paid off, when a single subsystem can be reused in a future product and thus, it is usually worth the effort.
 
@@
 Verification tests the properties of the developed product against the previously determined specification and requirements.
-It stives to build the \textit{system right}.
+It strives to build the \textit{system right}.
 Validation on the other hand tries to build the \textit{right system}.
@@
 This is important in order to develop the product according to what the customer actually needs.
-To avoid to divert from these customer needs, validation is done throughout the products' development cycle, from the beginning to the end.
-Various artifacts can be collected during the process, in order to have a traceable documentation of the decision that were made.\\
+To avoid to divert from these customer needs, validation is done throughout the product's development cycle, from the beginning to the end.
+Various artifacts can be collected during the process, in order to have a traceable documentation of the decision that are made.\\
 
@@
 As an example, the system may be very sensitive to variations of a parameter that a customer cannot influence.
-Another example is when a system is more sensitive to certain parameters than to its inputs.
+Another example is, when a system is more sensitive to certain parameters than to its inputs.
 In both cases, the system may be the wrong system for the specific problem and should urgently be re-evaluated.\\
@@
 
-A wide range of system-level design approaches are available that are targeting FPGAs.
+A wide range of system-level design approaches, which are targeting FPGAs, are available.
 In this section, some selected approaches are described.
@@
 
-A well-known process model is the so called V-Model.
+A well-known process model is the so-called V-Model.
 It is based on the similarly well-known Waterfall Model, and has its name because of the distinct V-shape of its process diagram, as it is shown in Figure~\ref{fig:v_modell}.
@@
 This classifies the model as a static process model and is one of the main disadvantages at the same time, because it makes the model inflexible to adaptions of the requirements.
-The advantages however are that it is easy to install and apply, each phase has its specific deliverables, good results can be achieved for large projects with simple requirements and the early test phases.\\
+The advantages however are that it is easy to install and apply, each phase has its specific deliverables, good results can be achieved for large projects with simple requirements, and the early test phases.\\
 
-The disadvantages of inflexibility towards changes of requirements and the fact, that it is difficult to consider all all eventualities in the creation of the V-model, led to the innovation of modern process models, which are iterative and thus more flexible.\\
+The disadvantages of the inflexibility regarding changes of requirements, and the fact that it is difficult to already consider all eventualities that may occur in the creation of the V-model, led to the innovation of modern process models, which are iterative and thus more flexible.\\
 
 The V-model applies to general software development, or project development, but can directly be mapped to FPGA development.
-The model is covered in this thesis section, because it shows the clear testing structure of module tests (unit tests), integration tests and system tests, which is especially important in projects that target hardware, such as FPGA projects.\\
+The model is covered in this thesis section, because it shows the clear testing structure of module tests (unit tests), integration tests and system tests, which are especially important in projects that target hardware, such as FPGA projects.\\
 
@@
 
-High-Level Synthes (HLS) is an FPGA design methodology that provides a high level of abstraction, and thus can be seen as a system level design method.
+High-Level Synthes (HLS) is an FPGA design methodology that provides a high level of abstraction from the low-level hardware, and thus can be seen as a system level design method.
 Different languages are used for HLS -- two well-known languages are SystemC, which is a library for \cplusplus\ to provide modelling and simulation features, and plain \cplusplus.
-A comprehensive list of available HLS tools and languages can be found in Section 3.3 of \cite{2016FfSP}.\\
+A comprehensive list of available HLS tools and languages can be found in Section~3.3 of \cite{2016FfSP}.\\
 
-The main idea of HLS is, that the hardware design can be written in abstract modules, that represent the actual functional meaning of a feature or algorithm, rather than the complex, low-level hardware details.
+The main idea of HLS is, that the hardware design can be written in abstract modules, which represent the actual functional meaning of a feature or algorithm, rather than the complex, low-level hardware details.
 HLS achieves this by allowing to develop the code in a high-level language, such as \cplusplus.
@@
 
-The implementation of low-level hardware description language (HDL) in the register transfer level (RTL) is denoted as 'direct', or 'manual' implementation in this thesis.\\
+The implementation of low-level HDL in the register transfer level (RTL) is denoted as 'direct', or 'manual' implementation in this thesis.\\
 
@@
 The languages can describe a design in a hierarchical order, so that the separate modules can be constructed in a way that an overview of the functionality of the system remains given.
-Synthesis, as well as simulation is supported, which allows efficient development of RTL designs.
+Synthesis, as well as simulation is supported, which allows an efficient development of RTL designs.
 
@@
 This is often used for prototyping in an early phase of product development.
-A well-known example for such a model-based approach is Matlab, and its graphical tool Simulink.\\
+A well-known example for such a model-based approach is Matlab and its graphical tool Simulink.\\
 
@@
 Simulink is a graphical editor, that allows to connect functional blocks in a block diagram and create a system in that way.
-Multiple toolboxes support additional blocks, that abstract complex functionalities, so that a fast prototyping is possible here as well.\\
+Multiple toolboxes support additional blocks, that abstract complex functionalities, so that fast prototyping is possible here as well.\\
 
@@
 This includes the high-level development of the system as a block diagram in Simulink, the automated generation of low-level RTL code for the FPGA hardware, and the verification in simulation and in hardware, in the form of hardware-in-the-loop.
-This workflow is supported for Xilinx FPGAs and is specifically targeting system design engineers, who work in rapid prototyping to generate hardware utilization estimates for the final products' hardware.\\
+This workflow is supported for Xilinx FPGAs and is specifically targeting system design engineers, who work in rapid prototyping to generate hardware utilization estimates for the final product's hardware.\\
 
@@
 
-Multiple implementation methods are introduced in above sections, which all have different levels of abstraction.
+Multiple implementation methods are introduced in the previous sections, which all have different levels of abstraction.
 VHDL and Verilog require, or allow, a very detailed description of the RTL code, while SystemC, \cplusplus, or Matlab and Simulink provide a high level of abstraction.
@@
 
-In conclusion, it is an engineers' choice which implementation method to use in order to develop a product.
-A trade-off between optimization and the level of abstraction is likely to be made.
-The decision may not always be obvious, but should be based on the fundamental requirements of the final product.
+\includepicture [0.65] [0, trim=0 5 0 5, clip]  {Abstraction levels of the implementation methods \cite[Abbildung 5.1]{GesslerRalf2014EES}.} {fpga_design_levels} {img/draw.io/fpga_design_levels}
 
-\includepicture [0.65] [0] [H] {Abstraction levels of the implementation methods \cite[Abbildung 5.1]{GesslerRalf2014EES}.} {fpga_design_levels} {img/draw.io/fpga_design_levels}
-
-%TODO: try to get this figure into the upper page...
+In conclusion, it is an engineer's choice which implementation method to use to develop a product.
+While decisions and trade-offs may not always be obvious, they should consistently be based on the fundamental requirements of the final product.

### 2021-08-20 doc/thesis/chapters/deployment-on-hardware.tex
unify 'highlights' in the text

@@
 This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section~\ref{sec:SimilarProcess}.
-Specifically, this implemented system design enables the application of the steps 'Assess Performance' and 'Re-evaluate', since the IPs' performance can be measured and their design can be re-evaluated.\\
+Specifically, this implemented system design enables the application of the steps \textit{Assess Performance} and \textit{Re-evaluate}, since the IPs' performance can be measured and their design can be re-evaluated.\\
 

### 2021-08-20 doc/thesis/chapters/hls.tex
unify 'highlights' in the text

@@
 
-Xilinx is the largest manufacturer of FPGAs globally and therefore its tool 'Xilinx HLS' is widely used.
-Intel, the second largest FPGA manufacturer, also has its HLS tool called 'Intel oneAPI' \cite{IntelHLSOneAPI}.
+Xilinx is the largest manufacturer of FPGAs globally and therefore its tool \textit{Xilinx HLS} is widely used.
+Intel, the second largest FPGA manufacturer, also has its HLS tool called \textit{Intel oneAPI} \cite{IntelHLSOneAPI}.
 Both companies are conducting serious effort to push the HLS implementation method.
@@
 Intel oneAPI uses library APIs and DP\cplusplus\ (Data-Parallel \cplusplus), which is an extension to the ISO \cplusplus\ standard.
-Xilinx also decided in favor of the \cplusplus\ language for their tools 'Vivado HLS' and 'Vitis HLS'.
+Xilinx also decided in favor of the \cplusplus\ language for their tools \textit{Vivado HLS} and \textit{Vitis HLS}.
 In older versions, SystemC was supported for synthesis as well, but is now deprecated in the respective latest versions \cite{VivadoHlsDeprecatesSystemC}.\\
@@
       In that way, the percentage of executed code lines of testbench and HLS code can be measured, to ensure a sufficient level of testing.
-      The software testbench is also called 'C Simulation'.\\
+      The software testbench is also called \textit{C Simulation}.\\
 
@@
 These can be instantiated in arbitrary bitwidths, where integer and fractional bitwidths can be specified separately.
-The fixed point format is often described in the format 'integer.fractional', like '2.14', which represents a 16 bit fixed point data type that has 2 integer bits and 14 fractional bits.\\
+The fixed point format is often described in the format \textit{integer.fractional}, like \textit{2.14}, which represents a 16 bit fixed point data type that has 2 integer bits and 14 fractional bits.\\
 
@@
   In this first step, the above explained, regular C simulation is executed.
-  Hereby, the inputs that are sent to the DUT are stored as 'input vectors'.
-
+  Hereby, the inputs that are sent to the DUT are stored as \textit{input vectors}.
   \item \textbf{Perform RTL simulation}\\
@@
   However, various other simulators like ModelSim are supported.
-  The generated outputs of this step are stored as 'output vectors'.
+  The generated outputs of this step are stored as \textit{output vectors}.
   \item \textbf{Verify the results}\\

### 2021-08-20 doc/thesis/chapters/implementation.tex
unify 'highlights' in the text

@@
 
-          To overcome the issue at this point in the actual hardware implementation, the filters were manually 'tuned' to have a matching order.
+          To overcome the issue at this point in the actual hardware implementation, the filters are manually adapted to have a matching order.
       \end{itemize}
@@
 In the implemented version, an audio file is modulated into the FM channel.
-The audio file contains data that represents speech with the words \textit{'Left channel... Right channel...'}.
+The audio file contains data that represents speech with the words \textit{Left channel... Right channel...}.
 In terms of audio channels, the words exactly match the respective channels' output -- the first part is modulated on the left audio channel only, while the second part only targets the right audio channel.
@@
 The \textit{tb\_data\_handler} only works as a data storage class, that encapsulates the data that is captured from the DUT.
-The class called \textit{fm\_receiver\_model} serves as the 'golden model reference' that is used to compare the DUT against.
+The class called \textit{fm\_receiver\_model} serves as the golden model reference that is used to compare the DUT against.
 Internally, it loads the Matlab verification data.
 This could be implemented differently, since Python can be used to actually implement an FM receiver including the entire DSP functionality, just like in Matlab.
-Because the Matlab model already exists, this is not done here -- loading the Matlab data is sufficient as it is been seen as 'the single source of truth' in this project.\\
+Because the Matlab model already exists, this is not done here -- loading the Matlab data is sufficient as it is been seen as the single source of truth in this project.\\
 
@@
 
-The VHDL code for this register interface is generated by a 'Register Engine' Python script.
+The VHDL code for this register interface is generated by a Python script which is named \textit{Register Engine}.
 The registers can be defined in a YAML-file.
@@
 
-The testbench is especially designed around the 'Design For Reuse' segment of the SIMILAR concept, which is introduced in Section~\ref{sec:SimilarProcess}.
+The testbench is especially designed around the \textit{Design For Reuse} segment of the SIMILAR concept, which is introduced in Section~\ref{sec:SimilarProcess}.
 This is the case because of two major parts.

### 2021-08-20 doc/thesis/chapters/results.tex
adapt system-design-process
unify 'highlights' in the text

@@
 The FIR filter has a successfully passing unit test and is thus assumed to be correct.
-Also the fixed-point data type with its overflow- and rounding behaviour is suspected to be a potential issue.
+Also the fixed point data type with its overflow- and rounding behaviour is suspected to be a potential issue.
 Due to the time limitation in the elaboration of this thesis, this issue can not be traced down to the root cause and thus still persists in the current implementation.
@@
 Therefore, the IPs internal state, i.e. the register values, may be different at the beginning, which leads to an error propagation throughout the design.
-The entire DSP chain is built like a pipeline and therefor it takes a number of samples to process, before the 'wrong' intermediate values are flushed out.
+The entire DSP chain is built like a pipeline and therefor it takes a number of samples to process, before the wrong, intermediate values are flushed out.
 All that may be the cause for the deviating values in the VHDL design.\\

### 2021-08-20 doc/thesis/chapters/system-architecture-and-concept.tex
unify 'highlights' in the text

@@
 This output data is again written to a file.
-Hereby it is important to note, that the output of the \textit{Matlab} implementation is chosen to be the 'golden reference model' of correct output.
+Hereby it is important to note, that the output of the \textit{Matlab} implementation is chosen to be the \textit{golden reference model} of correct output.
 Its output is used as the verification data later in the process.

### 2021-08-20 doc/thesis/chapters/system-design-process.tex
unify 'highlights' in the text

@@
 SystemC is not discussed, since it may potentially be less important in the future.
-In the latest Xilinx HLS tools 'Vitis HLS' and 'Vivado HLS', the language SystemC is already deprecated \cite{VivadoHlsDeprecatesSystemC}.\\
+In the latest Xilinx HLS tools \textit{Vitis HLS} and \textit{Vivado HLS}, the language SystemC is already deprecated \cite{VivadoHlsDeprecatesSystemC}.\\
 
@@
 
-The implementation of low-level HDL in the register transfer level (RTL) is denoted as 'direct', or 'manual' implementation in this thesis.\\
+The implementation of low-level HDL in the register transfer level (RTL) is denoted as \textit{direct}, or \textit{manual} implementation in this thesis.\\
 

### 2021-08-20 doc/thesis/chapters/hls.tex
adapt hls

@@
 In the earlier times however, it was common practise to develop programs in the low-level assembly language.
-The supported instruction set of the target CPU had to be implemented directly by the software engineer.\\
+The software engineer had to implement the software by directly using the supported instruction set of the target CPU.\\
 
@@
 An FPGA engineer historically develops code in the register-transfer level (RTL), which is common practise until today.
-This can be compared to developing software in the assembly language.
-However, the recent development towards automation and code generation brought up High-Level Synthesis (HLS).
+This can be compared to the development of software in the assembly language.
+However, the recent efforts towards automation and code generation brought up High-Level Synthesis (HLS).
 There, the FPGA design can be described in a high-level language, such as C or \cplusplus, which is convenient and allows to focus on the algorithm, rather than hardware details.
-This high-level code is then transformed into the low-level RTL automatically, by the HLS tool.\\
+This high-level code is then transformed into the low-level RTL automatically by the HLS tool.\\
 
@@
 A major achievement in the usage of HLS is the shorter development time.
-The time that it takes to develop a first working version of the intended design can be significantly reduced, compared to the development in RTL code directly.
-This is especially important when an FPGA design is combined with a software part, like in a System On Chip, since software development is comparably fast.
+The time that it takes to develop a first working version of the intended design can be significantly reduced, compared to the development of RTL code directly.
+This is especially important if an FPGA design is combined with a software part, like in a System On Chip, because software development is comparably fast and thus, it is important to also be fast on the hardware design side, to be able to bring the two parts together as soon as possible.\\
+
 Figure~\ref{fig:timeline_performance_first_prototype} schematically displays the difference in development time, over performance and degree of optimization, when the same functionality of a prototype is implemented in different methods for various targets.
-The compared targets are an x86 processor CPU, a digital signal processor (DSP), a graphics processing unit (GPU) and the FPGA design in RTL or HLS implementation.
+The compared targets are an x86 processor CPU, a digital signal processor (DSP), a graphics processing unit (GPU) and the FPGA design in RTL and HLS implementation.
 It is clearly visible, that the FPGA design with the HLS design approach achieves a first working prototype much earlier than the direct RTL approach.
+Even the optimized HLS version can be achieved within comparably short time, while both RTL versions, the first working version and the optimization, are beyond the typical design time limit in a software project.
 
@@
 
-HLS has been developed over 20 years ago, but has not been able to replace the manual, traditional implementation of RTL, using VHDL and Verilog, so far.
+HLS has been developed over 20 years ago in a first version, but has not yet been able to replace the manual, traditional implementation of RTL, by using the languages VHDL or Verilog.
 Unfortunately, no published numbers about the market share of HLS in relation to traditional implementations could be found during the research in this thesis.
 However, according to the given sources it is becoming more and more popular in recent years.
-This is mainly based on the fact that the HLS tools keep producing results of better quality as they are developed further.
+This is mainly based on the fact that the HLS tools keep producing results of better quality as they evolve.
 This especially regards the efficiency of the produced RTL code, the useability of the tools, as well as the better understandable correlation between the high-level code and the generated RTL code.\\
 
-Xilinx is the largest manufacturer of FPGAs globally and therefore its tool \textit{Xilinx HLS} is widely used.
+Xilinx is the largest manufacturer of FPGAs globally and therefore its tool called \textit{Vivado HLS} is widely used.
 Intel, the second largest FPGA manufacturer, also has its HLS tool called \textit{Intel oneAPI} \cite{IntelHLSOneAPI}.
@@
 In terms of language support, several tools settle for \cplusplus.
-Intel oneAPI uses library APIs and DP\cplusplus\ (Data-Parallel \cplusplus), which is an extension to the ISO \cplusplus\ standard.
-Xilinx also decided in favor of the \cplusplus\ language for their tools \textit{Vivado HLS} and \textit{Vitis HLS}.
+Intel oneAPI uses library APIs and Data-Parallel \cplusplus\ (DP\cplusplus), which is an extension to the ISO \cplusplus\ standard.
+Xilinx also decided in favor of the \cplusplus\ language for their tools \textit{Vivado HLS} and the latest tool \textit{Vitis HLS}.
 In older versions, SystemC was supported for synthesis as well, but is now deprecated in the respective latest versions \cite{VivadoHlsDeprecatesSystemC}.\\
@@
 Generally, as the user guide states, the quality and the correctness of the generated output can only be as good as its input.
-Consequently, it is very important to follow the design recommendations and adhere to specific rules in order to produce a high-quality input software.\\
+Consequently, it is very important to follow the design recommendations and adhere to specific coding rules in order to produce a high-quality input software.\\
 
@@
       The developed HLS IP code needs to be verified and checked for correct functionality in a software testbench, before it gets processed to RTL code.
-      This can be done by using using conventional software development tools.
-      Xilinx specifically recommends dynamic code checker tools like \textit{valgrind} or \textit{Coverity}.
+      This can be done by using conventional software development tools.
+      Xilinx specifically recommends dynamic code checker tools like \textit{valgrind} and \textit{Coverity}.
       They support a developer in finding memory leaks, out-of-bounds memory access, uninitialized variables and similar issues.
@@
 
-      The external simulator software has to run the complex HDL simulation and return data results, which is a time consuming task.
+      The external simulator software has to run the complex HDL simulation and return data results, which is a time-consuming task.
       Therefore, the execution time of this simulation type is expected to be approximately 10.000 times slower, compared to the C simulation.\\
@@
       At this point, the developed HLS code is fully verified and can successfully be transformed into working RTL code.
-      Xilinx HLS exports the generated RTL code in an IP structure that is supported by its IP integrator tool, that is used in Vivado or Vitis.
+      Xilinx HLS exports the generated RTL code in an IP file- and folder structure that is supported by its IP integrator tool, that is used in Vivado or Vitis.
       Therefore, the developed IP core can be imported to the block design as another processing block, get connected with other IPs, and in that way get integrated into a larger design.\\
@@
       Like in any other FPGA development cycle, the created block design can now be synthesized.
-      This generates a bitstream file, which can be programmed into an actual FPGA hardware.
+      This generates a bitstream file, which can be programmed into the actual FPGA hardware.
       Depending on the architecture of the system, the developed IP may now interface with the on-chip CPU, in the example of a System-On-Chip.\\
@@
 These steps represent the standard workflow for the development of an integrated HLS IP.
-A successful deployment on hardware can be achieved if they are completed accurately.
+The successful deployment of hardware can be achieved if they are completed accurately.
 There may be more and deeper knowledge necessary, in order to use this design approach.

### 2021-08-21 doc/thesis/chapters/hls.tex
new title again
hls if and coding

@@
       This is already done in the previous C simulation step.
-      However, the issue is that the C simulation is executed like a regular program, which means it is executed sequentially, instruction by instruction and thus parallelism cannot be represented.\\
+      However, the issue is that the C simulation is executed like a regular program, which means it is executed sequentially, instruction by instruction and thus parallelism can not be represented.\\
 
@@
 \section{Coding}
+%continue here
 
 This section gives a brief overview of some HLS-specific \cplusplus\ features that are supported with Xilinx tools.
-More detailed information can be found in the respective user guide \cite{VivadoUgHLS}.
+More detailed information can be found in the respective Xilinx user guide \cite{VivadoUgHLS}.
 
@@
 HLS supports special data types, which are not available in standard \cplusplus.
-The focus of these data types is mainly to support an optimized translation into RTL, which requires the least amount of logic cells in the FPGA.\\
+The focus of these data types is mainly to support an optimized translation into RTL, to achieve the least amount of logic cells in the FPGA.\\
 
-Standard \cplusplus\ only supports variables in the bitwidth of byte-boundaries, so in multiples of 8 bit.
+Standard \cplusplus\ only supports variables in the bitwidth of byte-boundaries, so in multiples of 8~bit.
 However, this may be inefficient to fit into the available FPGA target hardware.
-One example therefor is the DSP48 macro that Xilinx FPGAs have built-in.
-They support the multiplication of only up to 18 bit.
-In an example application, the reachable number range of a variable might be limited to 17 bit for the multiplication.
-However, in order to represent that, using a 32 bit data type would be necessary in standard \cplusplus.
-The HLS compiler does not have this prior knowledge of the number range limit and would thus implement a 32 bit multiplication, which is a waste of resources in this case.
+One example therefor is the DSP48 macro cell that Xilinx FPGAs have built-in.
+They support the direct multiplication of two vectors, with a bitwidth of up to 18~bit each.
+In an example application, the achievable maximum value of a variable might be limited, and thus, able to fit into a range of 17~bit for the multiplication.
+However, in order to represent that, using a 32~bit data type would be necessary in standard \cplusplus.
+The HLS compiler does not have this prior knowledge of the number range limit and would thus implement a 32~bit multiplication, which is a waste of resources in this case.
 Because of this reason, a major feature of HLS is the possibility to define data types with arbitrary bitwidths.
-In that way, the above example can be optimized to use a data type that has a bitwidth of 17 bit, which is exactly the bitwidth that is required for the application.
+In that way, the above example can be optimized to use a data type that actually has a bitwidth of 17~bit, which is exactly the bitwidth that is required for the application.
 This results in a lower logic utilization, which in turn results in a higher maximum clock frequency that can be used.
-Consequently, more functionality an be implemented in the FPGA device.
+Consequently, more functionality can be implemented and packed into the FPGA device.
 The referenced user guide shows a design example where an area reduction of 75\% is achieved by just reducing the bitwidths.\\
@@
 These can be instantiated in arbitrary bitwidths, where integer and fractional bitwidths can be specified separately.
-The fixed point format is often described in the format \textit{integer.fractional}, like \textit{2.14}, which represents a 16 bit fixed point data type that has 2 integer bits and 14 fractional bits.\\
+The fixed point format is often described in the format \textit{integer.fractional}, like \textit{2.14}, which represents a 16~bit fixed point data type that has 2 integer bits and 14 fractional bits.\\
 
@@
 The HLS code can be instumented with optimization directives.
-This is used to provide the HLS compiler with additional information and to instruct it on how to implement certain parts of the code in RTL.\\
-
+This is used to provide additional information to the HLS compiler, to assist and instruct it on how to implement certain parts of the code in RTL.
 Directives can be applied to interfaces, functions, loops, arrays or code regions.
@@
 
-Since this topic is very comprehensive, it cannot be covered into more detail here.
+Since the topic of optimization is very comprehensive, it cannot be covered into more detail here.
 For further information, refer to the user guide \cite{VivadoUgHLS}.
@@
 Interfaces are an important asset to any IP design, since this is where they communicate with their surroundings.
-This may be another IP that processes data, but may also be a CPU that reads status information or configures the IPs' settings.\\
+This may be another IP that processes data, but may also be a CPU that reads status information or configures the IP's settings.\\
 
@@
 The block-level interface is a general interface to the IP, that can provide a handshaking functionality for when it is idle and ready to start a new operation, or when the operation is completed.
-The port-level interfaces on the other hand are specific to input and output ports.
+The port-level interfaces on the other hand are specific to input and output (IO) ports.
 They can also provide handshaking functionality, such as data-valid indicators, acknowledge signals, or implement a FIFO for the IOs.
 However, port-level interfaces can also be implemented as AXI interfaces, such as AXI4-Stream, AXI4-Lite Slave or an AXI4 Master.
-This is an especially useful function, since the implementation of the actual interface logic is abstracted from the user.
+This is an especially useful function, since the implementation of the actual interface logic is abstracted from the developer.
 Additionally, software drivers are automatically generated along with the respective interfaces, which provides a fast and simple integration into a larger system.\\
 
-The code example in Program~\ref{prog:axi4_lite} shows the specification of an AXI4-Lite memory-mapped slave interface.
+The code example in Program~\ref{prog:axi4_lite} shows the specification of an AXI4-Lite memory-mapped slave interface, by using pre-processor pragma definitions.
 The provided registers are separated into configuration registers and status registers.
-The configuration is readable and writeable, while the status registers are implemented as read-only, which requires less logic in the RTL implementation.\\
+The configuration is readable and writeable, while the status registers are implemented as read-only, which requires less logic in the RTL implementation.
+Within the code, the interface is used like regular \cplusplus\ variables.\\
 
-\begin{program}
+The second code example, Program~\ref{prog:axi4_stream}, demonstrates the equally simple definition of an AXI4-Stream interface.
+HLS stream interface variables can be accessed via read- and write-functions.
+One very important property of these functions is, that they are blocking calls.
+The HLS streams are implemented with an underlying FIFO.
+Thus, values can only be written into it when there is an empty spot, and can only be read if there is at least one value available.
+If the respective condition is not satisfied, the functions block the execution, until the FIFO state changes.
+During simulation in the testbench, this may result in an endlessly blocking behaviour.
+
+\begin{program}[H]
   \caption{Implementation of an AXI4-Lite memory-mapped slave with a separation of read-only status registers and read-write configuration registers.}
@@
 
-The second code example, Program~\ref{prog:axi4_stream}, demonstrates the equally simple definition of an AXI4-Stream interface.
-
 \begin{program}
-  \caption{Implementation of an AXI4-Stream interface.}
+  \caption{Implementation of an AXI4-Stream interface. The stream is accessed with read- and write-functions.}
   \label{prog:axi4_stream}
@@
 \end{program}
-
-HLS stream interface variables can be accessed via read- and write-functions.
-One very important property of these functions is, that they are blocking calls.
-The HLS streams are implemented with an underlying FIFO.
-Thus, values can only be written into it when there is an empty spot, and can only be read if there is a at least one value available.
-If the respective condition is not satisfied, the functions block the execution, until the FIFO state changes.
-During simulation in the testbench, this may result in an endlessly blocking behaviour.
 

### 2021-08-21 doc/thesis/chapters/implementation.tex
hls if and coding

@@
 
-The streams use the types \texttt{iq\_sample\_t} and \texttt{audio\_sample\_t}, which are structs of two 16-bit fixed-point variables and thus represent a width of 32 bit.
+The streams use the types \texttt{iq\_sample\_t} and \texttt{audio\_sample\_t}, which are structs of two 16-bit fixed point variables and thus represent a width of 32 bit.
 The configuration and status registers are also implemented as structs, that contain multiple variables to represent the separate registers.

### 2021-08-21 doc/thesis/chapters/dsp.tex
hls done

@@
 Simply looking at this equation, the instantaneous frequency varies in the range between $f_c \pm \Delta f$, which is sometimes also called the swing.
-Considering the relationship between frequency and angle, as described above, and after some simple substitutions which will not be described into detail here, the equation for a generic frequency modulated wave is the following
+Considering the relationship between frequency and angle, and after some simple substitutions which will not be described into detail here, the equation for a generic frequency modulated wave is the following
 \begin{equation}
@@
 
-The differentiated signal still consists of a FM part, as explained above, using Equation~\eqref{equ:fm_demod_discriminator}.
+The differentiated signal still consists of a FM part, as explained earlier, using Equation~\eqref{equ:fm_demod_discriminator}.
 In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.

### 2021-08-21 doc/thesis/chapters/hls.tex
hls done

@@
 
-The general significance of the simulation testbench for a HDL design is explained here, as well as the different levels of abstraction in which a HLS testbench can be executed.
+The general significance of the simulation testbench for an HDL design is explained here, as well as the different levels of abstraction in which an HLS testbench can be executed.
 These different levels of abstraction in the testbench execution are a major difference between a VHDL testbench and an HLS testbench.
@@
 
-Section~\ref{sec:hls:workflow} above already mentioned the topics of testbenches and simulation types.
-The following descriptions shall be seen as a comprehensive extension around these topics.
+Section~\ref{sec:hls:workflow} already mentioned the topics of testbenches and simulation types briefly, in the context of workflow.
+The following descriptions are a comprehensive extension around these topics.
 
@@
 
-Thus, in order to allow debugging and verification of HDL code, the technology of simulation was invented.
-Here, the HDL is also translated into a netlist that represents all the parallel processes just like in hardware, but the subsequent step of synthesis is not taken.
+Thus, in order to allow debugging and verification of HDL design code, the technology of simulation was invented.
+Here, the time-consuming step of synthesis is not taken.
+Instead, the HDL simulator software is able to simulate the code in a way, that can represent parallel processes, just like they exist in hardware.
 Therefore, simulation comes to a result much faster.
@@
 They would need to be compared in their binary bit-pattern format, at the correct point in time.
-Of course the HDL language provides support with different datatypes, that abstract the binary number representation and make the numbers human-readable.
-However, it remains an enormous effort to manually write testing code to verify a bus interface, for example.
-In order to simplify such complex operations, various testbench frameworks and libraries are available to developers.
+Of course the HDL language provides support with different data types, that abstract the binary number representation and make the numbers human-readable.
+However, it remains an enormous effort to manually write testing code to verify bus interfaces, for example, because they consist of a large number of signals.
+In order to simplify such complex verification operations, various testbench frameworks and libraries are available to developers.
 
@@
 The most important note here is, that the HLS code is \textit{not} transformed into RTL at this point.
-Instead, the HLS code is treated as a regular \cplusplus\ code, that uses some libraries, like the fixed-point library.
-This means, that the HLS IP cores' top-level function is used as a regular function call in the testbench.\\
+Instead, the HLS code is treated as a regular \cplusplus\ code, that uses some libraries, like the fixed point library.
+This means, that the HLS IP core's top-level function is used as a regular function call in the testbench.\\
 
@@
 \noindent
-This type of simulation takes a significantly longer time to execute, because timing information now actually needs to be considered for the RTL code.
+This type of simulation takes a significantly longer time to execute, because timing information now actually needs to be considered for the RTL code, which is done by using an HDL simulator.

### 2021-08-21 doc/thesis/chapters/implementation.tex
hls done

@@
 The Matlab model is therefore developed by keeping in mind this limitation.
-Matlab provides special datatypes for fixed point number representation.
+Matlab provides special data types for fixed point number representation.
 However, these were not used in this implementation.
-Instead, everything is calculated in floating point representation, using the standard \textit{double} datatype.
-Then, in certain positions in the DSP chain, the numbers are rounded to the resolution of a fixed point datatype.
+Instead, everything is calculated in floating point representation, using the standard \textit{double} data type.
+Then, in certain positions in the DSP chain, the numbers are rounded to the resolution of a fixed point data type.
 A fixed point format of \textit{2.14}, meaning 2 bit for integer- and 14 bit for fractional number representation is chosen.

### 2021-08-21 doc/thesis/chapters/system-architecture-and-concept.tex
sys concept begin

@@
 It specifically talks about the system level approach that is taken in the implementation of the project.
-The main concept, as well as the architecture are explained.
+The main concept, as well as the architecture is explained.
 
@@
 Major decisions can be made at this stage, before any implementation has begun.
-This can have a direct impact on the efficiency of the entire development, which concequently has an impact on the final quality of the product.
+This can have a direct impact on the efficiency of the entire development, which consequently has an impact on the final quality of the product.
 A thorough system level design can also prevent from issues that would otherwise arise while the development is already ongoing.
-As an example it could turn out that one of the functional parts is not able to fit into the integrated system, in the way it is implemented.
+As an example, it could turn out that one of the functional parts is not able to fit into the integrated system, in the way it is implemented.
 This leads to the necessity to re-engineer this entire functional block.
 As a result, it takes more effort to implement the functionality, which directly correlates with the development cost and the time-to-market, in order to deliver the final product.
-Summing up, a comprehensive system level architecture and concept can pave the way for a successful implementation of a project from start to finish.\\
+Summing up, a comprehensive system level architecture and -concept can pave the way for a successful implementation of a project from start to finish.\\
 
@@
 This includes the definition of its functional range.
-Once this is done, the target system or target hardware should be considered, because there are various ways to implement a certain functionality.
-All of these variants will eventually result in the same output, but will, however, heavily differ in their implementation effort, efficiency, or their applicability for the products' aim in general.
-Thus, things like the available hardware platform, implementation time, degree of optimization, output quality, and similar things are important factors at this stage.
+Once this is done, the target system or target hardware should be considered, because often there are various ways to implement a certain functionality.
+All of these ways will eventually result in the same output, but will, however, heavily differ in their implementation effort, efficiency, or their applicability for the product's aim in general.
+Thus, things like the available hardware platform, estimated implementation effort and time, degree of optimization, output quality, and similar things are important factors at this stage.
 Of course, there are many more factors to consider when a product is to be developed.
@@
 
-The aim of this thesis is develop a system architecture and concept, that allows the implementation of an FM radio receiver in multiple different ways, while providing an elegant way to compare the different solutions.
-The final goal thereby is always be the same, that is: Listen to the music of a radio station that is transmitted over FM broadcast radio.\\
+The aim of this thesis is develop a system architecture and -concept, that allows the implementation of an FM radio receiver in multiple different ways, while providing an elegant way to compare the different solutions.
+The ultimate goal thereby is always the same, that is: Listen to the music of a radio station that is transmitted over FM broadcast radio.\\
 
@@
 It shows the front end, which processes the received antenna signal, so that it can be used in the FM radio receiver block.
-The radio receiver itself is implemented in multiple different variants.
-The produced audio output is sent to a speaker, since the ultimate goal is to listen to a radio stations' audio signal.
+The FM radio receiver itself is implemented in multiple different variants.
+The produced audio output is sent to a speaker, since the ultimate goal is to listen to a radio station's audio signal.
 
-\includepicture [1.0] [0] [H] {High-level overview of the system.} {bd_system_overview_highlevel} {img/draw.io/bd_system_overview_highlevel}
+\includepicture [1.0] [0] [H] {High-level overview of the system, representing the signal flow from the antenna, via the front end to the FM receiver, which is implemented in multiple variants, to the speakers for audio output.} {bd_system_overview_highlevel} {img/draw.io/bd_system_overview_highlevel}
 
@@
 Chapter~\ref{cha:SignalProcessingTheory} describes this part into detail.
-
 As already mentioned above, multiple ways of technology and implementation can lead to a similar result.
 Thus, the next step is to define a set of possible ways which will actually be implemented.
-For this thesis, the decision was made to implement the FM radio receiver by using the following methods.
+For this thesis, the decision was made to implement the FM radio receiver by using the following methods.\\
 
@@
       This abstracts the inner workings of each functional block and thus provides an implementation approach on a very high level of abstraction.
-      GNU Radio supports interfaces to external RF hardware \cite{SoftwareGnuRadio}.
+      GNU Radio supports interfaces to external RF hardware \cite{SoftwareGnuRadio}.\\
   \item \textbf{Matlab}\\
       Matlab is a software platform that can work with matrices and compute numerical solutions.
-      Matlab also includes various toolboxes, that support developers with complex calculations and tasks on datasets.
+      Matlab also includes various toolboxes, that support developers in complex calculations and tasks on datasets.
       This includes a range of toolboxes for signal processing, such as filter designers or FFT functions, as well as convenient ways to visualize data \cite{SoftwareMatlab}.
@@
       This can be done using high-level synthesis.
-      This thesis focuses in \cplusplus and does not implement a SystemC variant, because Xilinx has deprecated SystemC support for synthesis in their latest HLS tools.
+      This thesis focuses on \cplusplus\ HLS and does not implement a SystemC variant, because Xilinx has deprecated SystemC support for synthesis in their latest HLS tools.
       This, and the details of High-level synthesis (HLS) are explained in Chapter~\ref{cha:HLS}.
@@
       Similar to the HLS approach, this is a variant to write a hardware design for an FPGA.
-      The language VHDL is used in this approach.
+      The language VHDL is used in this approach.\\
 \end{itemize}
@@
 In order to implement an FM radio receiver, all of the listed methods require deep knowledge about the neccessary details of signal processing.
-However, the implementation of the various different methods require more or less knowledge and effort on the exact underlying implementation of each signal processing part.
+However, the implementation of the various different methods require more or less knowledge and effort on the exact underlying implementation of the algorithm of each signal processing part.
 This depends on their respective level of abstraction.

### 2021-08-21 doc/thesis/chapters/system-architecture-and-concept.tex
chapter sys-arch-and-design completed

@@
 
-The overall system concept has to be developed in a way, to fit the respective projects' needs and requirements.
+The overall system concept has to be developed in a way, to fit the respective project's needs and requirements.
 There should be a clear structure, in order to help to create a working system successfully.
@@
 
-In the apparent projects' case, a single block of the system is implemented in multiple ways, which should, however, all compute an equal output.
+In the apparent project's case, a single block of the system is implemented in multiple ways, which all should, however, compute an equal output.
 Thus, it is of advantage to define the data source to be common for all of these variants, to create a unified situation at the input.
-Additionally, the data source may be chosen to be recorded, to have a known, reproducable set of input data.
+Additionally, the data source is chosen to be recorded, to have a known, reproducable set of input data.
 This significantly eases the verification of the functionality in the final step, since all variants are expected to create the same output in the optimum, fully functioning case.
@@
 An example of a system concept and its architecture is developed and shown in Figure~\ref{fig:bd_system_overview_detailed}.
-Here, a common data source is provided by the block called \textit{Front End}.
+This is also the architecture, which is actually used in the realization of this project.
+A common data source is provided by the block called \textit{Front End}.
 It records data from an antenna and saves it to a file, in a format that the consequent blocks require at their input.
@@
 Its output is used as the verification data later in the process.
-This is because Matlab was used to research decoding algorithms and to develop the DSP chain, which is used in the GNU Radio, HLS and VHDL implementations.
+This is because Matlab is used to research decoding algorithms and to develop the DSP chain, which is used in the GNU Radio, HLS and VHDL implementations.
 In addition to the verification data, the Matlab model also writes common parameters to a file.
@@
 This is especially important, since there is no possible way to run all these different variants in a single verification environment.
-Instead, each methodology writes an output file, which is a supported functionality that is available anywhere.
+Instead, each methodology takes an input file and writes an output file, which is a supported functionality that is available anywhere.
 
-\includepicture [1.0] [0] {Block diagram of the overall system concept.} {bd_system_overview_detailed} {img/draw.io/bd_system_overview_detailed}
+\includepicture [1.0] [0] {Block diagram of the overall system concept.
+A common input source is used for the multiple implementations.
+The generated output files, including the used parameters, are verified in a common entity.} {bd_system_overview_detailed} {img/draw.io/bd_system_overview_detailed}
 

### 2021-08-21 doc/thesis/chapters/implementation.tex
implementation begin

@@
 
-This chapter describes the most important details of each implementation variant.
+This chapter describes the most important details of each implementation variant, as they are developed in the FM receiver project in this thesis.
 The implementation variants in High-Level Synthesis and VHDL are highlighted especially, since they are the main focus of this thesis.
@@
 The theory behind the demodulation of an FM signal is already explained in Chapter~\ref{cha:SignalProcessingTheory}.
-The demodulation of an FM broadcast channel specifically, is described in Section~\ref{sec:demodulation-of-broadcast-fm}.
-In this section however, the DSP chain is described into more detail and with more information, as it is implemented in the various methods.\\
+The demodulation of an FM broadcast channel specifically, is described in Section~\ref{sec:demodulation-of-a-broadcast-fm-channel}.
+In this section, the implementation details of the DSP chain are described, as they are implemented in the various methods.\\
 
 \noindent
-Figure~\ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram.\\
+Figure~\ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram.
+The following list explains implementation details about the respective DSP blocks.\\
 
@@
 
-\noindent
-The following list explains implementation details about the respective DSP blocks.
 \begin{itemize}
@@
       Matlab is used to generate the data source file, which is used as an input to the system.
-      An audio file is read and modulated as an FM channel signal.
+      An audio file is read and modulated as an FM channel signal.\\
 
   \item \textbf{FM Demodulator}\\
-       TODO\\
-       % TODO
-      The FM demodulator is implemented as a Baseband Delay Modulator, as described in Section~\ref{sec:BasebandDelayDemodulator}.
+      The FM demodulator is implemented according to the explanations in Section~\ref{sec:algorithms-for-digital-fm-demodulation}, with a frequency discriminator and a subsequent envelope detector.\\
 
   \item \textbf{De-Emphasis Filter}\\
-      The above described Matlab source generation does not include an Emphasis filter.
-      Thus, a de-emphasis filter is not implemented as it is not required.
+      Pre-emphasis in the transmitter and its counterpart, the de-emphasis in the receiver, is a strategy to minimize the noise that is induced due to effects in the transmission, by attenuating and amplifying certain frequency parts of the signal.
+      Both processes can be implemented, by using filters in the transmitter, as well as the receiver, respectively.
+      The processes are implemented in the Matlab model, but are disabled in the generation of the input file that is used for the system.
+      Thus, a de-emphasis filter is not implemented in the hardware, as it is not required for the show-case of this thesis.\\
 
-  \item \textbf{Downsample}\\
+  \item \textbf{Downsampling}\\
       Downsampling is implemented in a simplified version, without an anti-aliasing filter.
@@
       The input is band-limited at 75~kHz, which is the maximum frequency in an FM channel, as explained in Section \ref{sec:FrequencySpectrumOfABroadcastFmChannel}.
-      Thus, the input has an oversampling rate of 12.8, while downsampling is only performed with a factor of 8.
+      Thus, the input has an oversampling rate of 12.8, while downsampling is only performed with a factor of 8.\\
 
@@
       The 15~kHz lowpass (LP) filter is instantiated twice, with the exact same coefficients.
+      The two bandpass (BP) filters have different coefficients.\\
 
@@
       FIR filters introduce a delay on the output signal, with respect to the input.
-      The group delay in number of samples can be computed with the following formula.
+      The group delay, in number of samples, depends on the filter order, which is directly related to the number of coefficients N, and can be computed with the following formula.
       \begin{equation}
-        group\ delay = \frac{filter\ order}{2} = \frac{(N\ coefficients - 1)}{2}
+        group\ delay = \frac{filter\ order}{2} = \frac{(N_{coefficients} - 1)}{2}
       \end{equation}
-      This is due to the property of the FIR filters' linear phase response \cite{GaziOrhan2018UDSP}.
+      This is due to the property of the FIR filter's linear phase response \cite{GaziOrhan2018UDSP}.
       Since phase is directly proportional to time, the term \textit{group delay} is often used.
       This delay needs to be accounted for, depending on where the signal is used subsequently.
+      Looking at the block diagram in Figure~\ref{fig:bd_dsp_detailed}, the group delay needs to be compensated at two places.
 
-      Looking at the block diagram, the group delay needs to be compensated at two places.
       \begin{itemize}
-        \item \textbf{Between the branches \textit{Recover LR diff} and \textit{Recover Mono}:}\\
-          The \textit{Mono} branch only contains a single LP filter, while the \textit{LR diff} branch contains two filters, namely a BP and an LP filter.
-          The latter branches' signal is therefore delayed, with respect to the \textit{Mono} branch.\\
+        \item \textbf{Between the branches \textit{Recover LR Diff} and \textit{Recover Mono}:}\\
+          The \textit{Mono} branch only contains a single LP filter, while the \textit{LR Diff} branch contains two filters, namely a BP and an LP filter.
+          The latter branch's signal is therefore delayed, with respect to the \textit{Mono} branch.\\
           The issue here is, that both signals are added and substracted at the end of the processing chain, to compute the left and right channel signals, respectively.
           Thus, they need to be aligned in time at that point.
-          Consequently, a delay block needs to be inserted in the \textit{Mono} branch, to compensate the group delay of the BP.
+          Consequently, a delay block needs to be inserted in the \textit{Mono} branch, to compensate the group delay of the BP in the \textit{LR Diff} branch.
 
-        \item \textbf{Between the branches \textit{Recover LR diff} and \textit{Recover Carriers}:}\\
+        \item \textbf{Between the branches \textit{Recover LR Diff} and \textit{Recover Carriers}:}\\
           Both branches only contain a single BP filter.
-          However, they have properties, like the transition bandwidth and the passband.
+          However, they have different properties, like the transition bandwidth and the passband.
           Thus, their filter order may be different as well, which leads to a non-matching group delay.
-          Since the two signals are multiplicated in the subsequent mixer, and an important factor is the phase-coherency of the recovered carrier (as explained in Section~\ref{sec:demodulation-of-broadcast-fm}), the group delay compensation becomes necessary here.\\
-
-          To overcome the issue at this point in the actual hardware implementation, the filters are manually adapted to have a matching order.
+          Since the two signals are multiplicated in the subsequent mixer, and an important factor is the phase-coherency of the recovered carrier, the group delay compensation becomes necessary here.\\
       \end{itemize}
 
+      To overcome the group delay issue at this point in the actual hardware implementation, the filters are manually adapted to have a matching order.\\
   \item \textbf{Carrier Recovery}\\
       The 38~kHz carrier can be recovered from the 19~kHz pilot tone.
-      It is important that the recovered carrier is phase-coherent with the received input, as described in Section~\ref{sec:demodulation-of-broadcast-fm}.
+      It is important that the recovered carrier is phase-coherent with the received input, as described in Section~\ref{subsec:demod_stereo_audio}.
       This can be achieved, by simply multiplying the pilot tone with itself, which is the way it is implemented here.\\
@@
 Matlab is used to create a model of the receiver, which is used as a reference for the other implementation variants.
-Here, DSP algorithms are tested and evaluated, to find a method that is suitable to implement in hardware, specifically in an FPGA.\\
+Here, the DSP algorithms are tested and evaluated, to find a method that is suitable to implement in hardware, specifically in an FPGA.\\
 
-The Matlab script consists of a transmitter, the receiver model, and additional code for analysis.
+The Matlab script consists of a transmitter model, the receiver model, and additional code for analysis.
 

### 2021-08-21 doc/thesis/chapters/system-architecture-and-concept.tex
implementation begin

@@
       Matlab also includes various toolboxes, that support developers in complex calculations and tasks on datasets.
-      This includes a range of toolboxes for signal processing, such as filter designers or FFT functions, as well as convenient ways to visualize data \cite{SoftwareMatlab}.
+      This includes a range of toolboxes for signal processing, such as filter designers or FFT functions, as well as convenient ways to visualize data \cite{SoftwareMatlab}.\\
   \item \textbf{FPGA: \cplusplus\ High-Level Synthesis}\\
@@
       This thesis focuses on \cplusplus\ HLS and does not implement a SystemC variant, because Xilinx has deprecated SystemC support for synthesis in their latest HLS tools.
-      This, and the details of High-level synthesis (HLS) are explained in Chapter~\ref{cha:HLS}.
+      This, and the details of High-level synthesis (HLS) are explained in Chapter~\ref{cha:HLS}.\\
 
@@
 
-\includepicture [1.0] [0] {Block diagram of the overall system concept.
+\includepicture [0.95] [0, trim=14 0 0 4,clip] {Block diagram of the overall system concept, with common input source, multiple receiver implementations, and a common verification entity.} {bd_system_overview_detailed} {img/draw.io/bd_system_overview_detailed}
+%Block diagram of the overall system concept.
 A common input source is used for the multiple implementations.
-The generated output files, including the used parameters, are verified in a common entity.} {bd_system_overview_detailed} {img/draw.io/bd_system_overview_detailed}
+The generated output files, including the used parameters, are verified in a common entity.
 

### 2021-08-21 doc/thesis/chapters/implementation.tex
matlab model

@@
 In the implemented version, an audio file is modulated into the FM channel.
-The audio file contains data that represents speech with the words \textit{Left channel... Right channel...}.
-In terms of audio channels, the words exactly match the respective channels' output -- the first part is modulated on the left audio channel only, while the second part only targets the right audio channel.
-This is particulary useful for verification, especially to verify the functionality of channel separation.
+The audio file contains data that represents speech with the words \textit{Left channel... Right channel...} of a male voice.
+In terms of audio channels, the words exactly match the respective channel's output -- the first part is modulated on the left audio channel only, while the second part only targets the right audio channel.
+This is particulary useful for verification, especially to verify the functionality of the stereo channel separation.
 From the experience during development, the channel separation is a good indicator of whether the system is functioning correctly.
 This left-right indication also simplifies the verification process, because a developer can recognize it with their visual and acoustic senses.
-It is possible to audibly verify it by listening to the output.
-It is also possible to just visually inspect the output data, where a clear separation is visible between the channels.\\
+It is possible to audibly verify it by listening to the output, and also to just visually inspect the output data, where a clear separation is visible between the channels.\\
 
@@
 
-The achieved sample rate of 960~kHz could actually be chosen much lower, since an FM channels' maximum frequency is 75~kHz (see Section~\ref{sec:FrequencySpectrumOfABroadcastFmChannel}), and according to the Nyquist theorem the sample rate thus only needs to have 150~kHz (see Section~\ref{sec:SamplingTheorem}).
-However, the higher rate was chosen, mainly because a higher number of signal samples beautify any plots that are created during analysis.
-Since a thorough optimization of the DSP chain is not the focus of this thesis, this was an acceptable trade-off.\\
+The achieved sample rate of 960~kHz could actually be chosen much lower, since an FM channel's maximum frequency is 75~kHz (see Section~\ref{sec:FrequencySpectrumOfABroadcastFmChannel}), and according to the Nyquist theorem the sample rate thus only needs to have 150~kHz (see Section~\ref{sec:SamplingTheorem}).
+However, the higher rate is chosen, mainly because a higher number of signal samples beautify any plots that are created during analysis.
+Since a thorough optimization of the DSP chain is not the focus of this thesis, this is an acceptable trade-off.\\
 
@@
 The LR Diff part is shifted to 38~kHz in the spectrum.
-This is done by modulating the signal with its sub-carrier of 38~kHz.
+This is done by modulating the signal with the 38~kHz sub-carrier.
 A subsequent BP filter serves to remove the artifacts of the modulation, such as the image replica.
-It band-limits the spectrum of the LR Diff part between 23 and 53~kHz.
+It band-limits the spectrum of the LR~Diff part between 23 and 53~kHz.
 The BP filter introduces a constant group delay of the signal.
 In order to maintain the alignment of all signals, this side-effect needs to be accounted for and be compensated in the other signal branches.
-Thus, the Mono signal, as well as the Pilot Tone need to be delayed by the amount of samples defined by the BP filters' group delay.
+Thus, the Mono signal, as well as the Pilot Tone need to be delayed by the amount of samples defined by the BP filter's group delay.
 The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone \cite{FmMultiplexingForStereo}.\\
@@
 One part in this block diagram is the so-called Hinz Triller, which was part of the Autofahrer-Rundfunk-Information service (ARI, German for Automotive-Driver's-Broadcasting-Information).
-It was used as an indicator for traffic announcements.
+It was historically used as an indicator for traffic announcements.
 It is now obsolete and is replaced by the more modern Radio Data System (RDS).
-However, the Hinz Triller is still transmitted in many radio stations and can often be heard before and after the traffic announcements \cite{HinzTriller}.
+However, the Hinz Triller is still transmitted in many radio stations and can often be heard before and after the traffic announcements, as a distinctive beep tone \cite{HinzTriller}.
 In the implemented transmitter, the Hinz Triller is implemented, but is disabled, since it only represents a temporarily transmitted signal.\\
@@
 This signal is then processed by a complex baseband mixer, which results in inphase-, and quadrature components in the baseband.
-An LP filter band-limits the signal and removes artifacts of the mixer.
+A lowpass filter band-limits the signal and removes artifacts of the mixer.
 As a final step, the sample rate is decreased again, back to the chosen sampling rate of 960~kHz.

### 2021-08-21 doc/thesis/chapters/implementation.tex
shift around for newline

@@
 Matlab is used to create a model of the receiver, which is used as a reference for the other implementation variants.
-Here, the DSP algorithms are tested and evaluated, to find a method that is suitable to implement in hardware, specifically in an FPGA.\\
-
+Here, the DSP algorithms are tested and evaluated, to find a method that is suitable to implement in hardware, specifically in an FPGA.
 The Matlab script consists of a transmitter model, the receiver model, and additional code for analysis.
@@
 This left-right indication also simplifies the verification process, because a developer can recognize it with their visual and acoustic senses.
-It is possible to audibly verify it by listening to the output, and also to just visually inspect the output data, where a clear separation is visible between the channels.\\
+It is possible to audibly verify it by listening to the output, and also to just visually inspect the output data, where a clear separation is visible between the channels, as presented in Figure~\ref{fig:matlab_analysis_time_domain}.\\
 
@@
 The audio file contains data in a sample rate of 48~kHz.
-Next, the sample rate is increased by an oversampling factor of 20, to achieve a sufficient sample rate for the subsequent DSP procedures.\\
-
+Next, the sample rate is increased by an oversampling factor of 20, to achieve a sufficient sample rate for the subsequent DSP procedures.
 The achieved sample rate of 960~kHz could actually be chosen much lower, since an FM channel's maximum frequency is 75~kHz (see Section~\ref{sec:FrequencySpectrumOfABroadcastFmChannel}), and according to the Nyquist theorem the sample rate thus only needs to have 150~kHz (see Section~\ref{sec:SamplingTheorem}).
@@
 Any kind of computation can be implemented within these logic cells.
-However, depending on its detailed implementation, the amount of required logic cells may differ by a large factor.
+However, depending on the detailed implementation of these computations, the amount of required logic cells may differ by a large factor.
 In order to overcome this issue, mathematical operations are often implemented in fixed point arithmetic, rather than floating point.

### 2021-08-21 doc/thesis/chapters/implementation.tex
more of impl section

@@
 The transmitter is implemented to provide a reproducable source of data, which can be used to feed the receiver.
-Another option would be to use external hardware as a software-defined radio and record some antenna data.
+Another option would be to use external hardware as a software-defined radio (SDR) and record some antenna data.
 However, the main advantage of a locally implemented transmitter over a recording is, that the raw data input is known.
@@
 The LR Diff part is shifted to 38~kHz in the spectrum.
-This is done by modulating the signal with the 38~kHz sub-carrier.
+This is done by modulating the signal with the 38~kHz subcarrier.
 A subsequent BP filter serves to remove the artifacts of the modulation, such as the image replica.
@@
 
-The receiver is implemented, following the block diagram in Figure~\ref{fig:bd_dsp_detailed}.
-Further assumptions and simplifications, such as the fixed point arithmetic are described into detail above.
+The Matlab receiver is implemented according to the block diagram in Figure~\ref{fig:bd_dsp_detailed}.
+Further assumptions and simplifications, such as the fixed point arithmetic and the filter design are already described into detail in the previous chapters.
 
@@
 
-The Matlab implementation also includes a comprehensive range of analysis, especially in terms of visual plots.
-These plots are useful during the development of new DSP algorithms, in order to verify the correct thereof.\\
+The Matlab implementation also includes a comprehensive range of analysis functionalities, especially in terms of visual plots.
+These plots are useful during the development of new DSP algorithms, in order to verify the correct functionality thereof.\\
 
@@
 The bottom diagrams then represent the result after the entire signal processing chain.
-It can be seen that the DSP chain is not able to demodulate the original signal without any error.
-However, the result shows a clear separation of the left and right channels, as resembles the original input very closely, which is good enough for this case study.\\
+It can be seen that the DSP chain is not able to demodulate the signal without any error, as the output signal does not exactly match the original input signal.
+However, the result shows a clear separation of the left and right channels, and resembles the original input very closely, which is good enough for this case study.\\
 
-\includepicture [1.0] [0] {Matlab analysis in the time domain.} {matlab_analysis_time_domain} {img/matlab/matlab_analysis_time_domain}
+\includepicture [1.0] [0] {Matlab analysis in the time domain.
+The stereo audio signal of the voice saying \textit{Left channel, Right channel} can be recognized in the signal, which allows an audible and visual verification of the receiver's functionality.} {matlab_analysis_time_domain} {img/matlab/matlab_analysis_time_domain}
+%TODO: latex font for this diagram!
 
@@
 The spectrum after the demodulation through a receiver is shown in the lower diagram.
-Spectral parts, such as the pilot tone at 19~kHz, or the LR Diff part around 38~kHz can be seen clearly.
+Spectral parts, such as the pilot tone at 19~kHz, or the LR Diff part around 38~kHz can be seen clearly, as presented in the theory in Figure~\ref{fig:fm_channel_baseband_freqs}.
 The demodulated spectrum shows an artifact at 57~kHz, which does not exist in the original transmitter spectrum.
@@
 
-\includepicture [1.0] [0] {Matlab analysis in the frequency domain.} {matlab_analysis_freq_domain} {img/matlab/matlab_analysis_freq_domain}
+\includepicture [1.0] [0] {Matlab analysis in the frequency domain.
+Spectral parts, like the mono audio, the pilot tone, or the LR difference signal around the 38~kHz subcarrier can clearly be seen.} {matlab_analysis_freq_domain} {img/matlab/matlab_analysis_freq_domain}
+%TODO: latex font for this diagram!
 
@@
 A transmitter, as well as a receiver are implemented using this software.
-Both can be deployed on actual hardware, so a continuous transmitter or receiver is built in a real environment.
+Both can be deployed on actual hardware, so a continuous, live transmitter or receiver is built in a real environment.
 
@@
 Multiple blocks are connected to a block diagram, that is very similar to the one in Figure~\ref{fig:bd_matlab_transmitter}.
-The USRP device is used to broadcast the RF signal over the air.\\
+An external device, the USRP B200mini from Ettus Research, is used to broadcast the RF signal over the air.\\
 
@@
 The frequency spectrum is strongly limited by the radio regulations and may be monitored by the authorities.
-Please make sure to adapt your setup to the respective local regulations, to avoid any legal issues.
+The transmitter setup needs to be adapted carefully to the respective local regulations, to avoid any legal issues.
 Some of the regulations are mentioned in Section~\ref{sec:FrequencyBand}, but may differ in various countries, however.\\
@@
 \includepicture [0.9] [0] [H] {Block diagram of the GNU Radio transmitter setup.} {bd_gnuradio_setup_transmitter} {img/draw.io/bd_gnuradio_setup_transmitter}
+%TODO: transmitted waves in reverse order (getting larger as they distance..)
 
@@
 
-The receiver is also implemented, by using standard DSP blocks that are available in GNU Radio.
+The receiver is also implemented by using standard DSP blocks that are available in GNU Radio.
 The implemented structure however does not represent a stereo receiver.
@@
 The antenna signal is FM-demodulated, decimated and then directly sent to the speakers as the output, without any further processing.
-This simple receiver was intended as a proof-of-concept and a first example of the cooperation of the external hardware and the GNU Radio software.
-Thus, the mono receiver was sufficient, as the focus of the GNU Radio application is put on the transmitter side.\\
+This simple receiver is intended as a proof-of-concept and a first example of the cooperation of the external hardware and the GNU Radio software.
+Therefor, the mono receiver is sufficient, as the focus of the GNU Radio application is put on the transmitter side.\\
 
-This receiver works with both devices, the USRP b200mini and the RTL-SDR.
-Figure~\ref{fig:bd_gnuradio_setup_receiver} shows a block diagram of the receiver application.
+This receiver is tested with two different external devices, the previously mentioned USRP B200mini and any RTL-SDR device.
+The devices are explained into more detail later in Section \ref{sec:deployment-on-hw:gnu-radio-devices}.
+A block diagram of the receiver application is shown in Figure~\ref{fig:bd_gnuradio_setup_receiver}.
 
+%TODO: b200mini --> B200mini (upper-case 'B')
 \includepicture [0.9] [0] [H] {Block diagram of the GNU Radio receiver setup.} {bd_gnuradio_setup_receiver} {img/draw.io/bd_gnuradio_setup_receiver}
@@
 The usage of a software like GNU Radio in combination with external hardware is an interesting tool for a use-case like this project.
-An FM receiver is to be implemented in multiple different versions.
-All of them need to be verified somehow -- this is where the SDR transmitter created with GNU Radio and external hardware opens up an entire new possibility from the viewpoint of system level design integration.\\
+An FM receiver is to be implemented in multiple different versions and all of them need to be verified somehow -- this is where the SDR transmitter, created with GNU Radio and external hardware, opens up an entire new possibility from the viewpoint of system level design integration.\\
 
-Previously it was explained that it is not possible to use recorded antenna data for verification.
-The reason was, that the underlying original data is not known and therefor a comparison of the original versus the received, decoded data is not feasible.
+Previously it was explained, that it is not possible to use recorded antenna data for verification.
+The reason was, that the underlying original data is not known and therefor a comparison of the original, versus the received, decoded data is not feasible.
 Consequently, the Matlab transmitter model was justified by the fact, that its originally encoded audio data is well-known.\\
 
-However, with the current approach of using a GNU Radio SDR transmitter, the encoded data \textit{is} well-known, because it is generated locally.
+However, with the current approach of using a GNU Radio SDR transmitter, the originally encoded data of the received signal \textit{is} well-known, because the transmitted signal is generated locally.
 Thus, the SDR can be used to transmit a verification signal.
-This signal could be recorded to file by another SDR, or sourced into the DUT (device under test) directly.
+This signal can be recorded to file by another SDR, or sourced into the DUT hardware directly.
 

### 2021-08-21 doc/thesis/chapters/implementation.tex
cocotb

@@
 
-Multiple different simulators are available on the market, such as Mentor Graphics ModelSim or QuestaSim, Synopsys VCS or the Cadence Simulator.
+Multiple different simulators are available on the market, such as Mentor Graphics \mbox{ModelSim} or QuestaSim, Synopsys VCS or the Cadence Simulator.
 The issue with most simulators is, that they are subject to a license fee, depending on the required set of features, or simulation speed.\\
@@
 An example therefor is GHDL, which is an analyzer, compiler and simulator for VHDL and is developed as an open-source project.
-GHDL translates the VHDL code to a binary file that contains machine code, which can be executed.
+GHDL translates the VHDL code, including the testbench code, to an executable binary file.
 This is supported for a all common operating systems such as Linux, Windows and macOS.
@@
 
-One of the main advantages of cocotb is, that test cases for the respective RTL code are written in the language Python.
+One of the main advantages of cocotb is, that the testbench for the respective RTL code is written in the language Python.
 It is therefore also called a co-simulation framework.
-Using Python provides a large factor of flexibility and possibilities, since it is a comprehensive language, considering all its extension libraries.
+Using Python provides a large factor of flexibility and possibilities, since it is a comprehensive language, considering all of its extension libraries.
 Additionally, Python may be more familiar to engineers in comparison to VHDL and thus, writing testbenches is made simpler with cocotb.\\
@@
 In the simulator, the device under test (DUT) is instantiated.
-The RTL simulation itself is managed by the simulators' scheduler.
+The RTL simulation itself is managed by the simulator's scheduler.
 

### 2021-08-21 doc/thesis/chapters/implementation.tex
impl vhdl done

@@
 
-The architecture for this testbench is created under the assumptions and according to the guidelines that are developed in Chapter~\ref{cha:SystemArchitectureAndConcept}.\\
+The architecture for this testbench is created according to the guidelines and assumptions that are developed in Chapter~\ref{cha:SystemArchitectureAndConcept}.\\
 
 Figure~\ref{fig:bd_testbench_architecture_vhdl} shows a block diagram of the implemented architecture.
-The split between the Python environment and the VHDL implementation is clearly visible but is represented from the developers perspective.
-The actual simulator interfaces, as described in Section~\ref{sub:VHDL:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.\\
+The split between the Python environment and the VHDL implementation is clearly visible and is represented from the developer's perspective.
+The actual simulator interfaces, as described in Section~\ref{sub:VHDL:FrameworkCocotb}, are hidden to maintain a better overview.\\
 
 \includepicture [1.0] [0]  {Block diagram of the implemented testbench architecture.} {bd_testbench_architecture_vhdl} {img/draw.io/bd_testbench_architecture_vhdl}
+%TODO: remove the yellow color
 
-The test cases are using the external verification data, that was previously generated by the Matlab model.
+The test cases are using the verification data, that was previously generated by the Matlab model.
 This data is supplied to the input of the DUT.
-The DUTs' input and output is implemented as an AXI Stream interface, which is an interface that features a valid-, as well as a ready-signal along with the data.
+The DUT's input and output is implemented as an AXI Stream interface, which is an interface that features a valid-, as well as a ready-signal along with the data.
 The cocotb framework supports both an AXI Stream Master, and an AXI Stream Slave class, to write and read the input and output, respectively.
@@
 The Python classes \textit{tb\_analyzer\_helper}, \textit{tb\_data\_handler} and \textit{fm\_receiver\_model} are implemented to create a better structure of the testbench and to provide code that is re-useable for the testbench of the HLS implementation.
-The \textit{tb\_analyzer\_helper} class provides functions to analyze the data, such as comparison against defined error thresholds or the creation of graphical plots with Pythons' matplotlib library.
+The \textit{tb\_analyzer\_helper} class provides functions to analyze the data, such as the comparison against defined error thresholds or the creation of graphical plots with Python's library matplotlib.
 The \textit{tb\_data\_handler} only works as a data storage class, that encapsulates the data that is captured from the DUT.
@@
 This could be implemented differently, since Python can be used to actually implement an FM receiver including the entire DSP functionality, just like in Matlab.
-Because the Matlab model already exists, this is not done here -- loading the Matlab data is sufficient as it is been seen as the single source of truth in this project.\\
+Because the Matlab model already exists, this is not done here -- loading the Matlab data is sufficient as it is seen as the single source of truth in this project.\\
 
-At the end of the test cases, the DUTs' output data is written to a file.
+At the end of the test cases, the DUT's output data is written to a file.
 This data is later used to directly compare the different implementation variants against each other.
@@
 Unit tests specifically target a single unit, while integration tests verify the combination of multiple units as a whole.
-Both of these stragegies are important.
+Both of these strategies are important.
 Given all unit tests being successful, it is ensured that all these single units are implemented correctly.
 However, their combination and interaction may still be subject to error -- this is where integration tests unfold their potential.
-Implementation errors in the glue logic between the units, as well as the interaction of the respective interfaces can be found in these integration tests.\\
+Implementation errors in the glue logic between the units, as well as the interaction of the respective interfaces can be revealed with these integration tests.\\
 
-The testbench architecture presented above represents an integration test, since the DUT is tested as an entire block.
+The testbench architecture presented in Figure~\ref{fig:bd_testbench_architecture_vhdl} represents an integration test, since the entire DUT is tested as a single block, with all the single units already being integrated.
 The advantage thereof is, that it tests the DUT in the same way as it will be integrated within the target hardware, the FPGA.
@@
 The source of error may be hidden anywhere within the nested and interconnected units.
-Thus, in addition to the DUTs' interfaces, further signals are logged from the inside of the DUT.
+Thus, in addition to the DUT's interface signals, further signals from the inside of the DUT are logged.
 These internal signals can be accessed through the Python testbench.\\
@@
 In this context, there are further keywords that need to be mentioned -- black box testing and white box testing.
-These are testing strategies, and describe how the DUT is treated.
-In the case of black box, the DUTs internals, such as implementation details, are not known.
+They are testing strategies, and describe how a DUT is treated in a test procedure.
+In the case of black box testing, the DUT's internals, such as implementation details, are not known.
 Thus, only the exposed interfaces can be tested.
-In the white box as a contrary, every detail of the internal structure is known.
-Thus, even single signals can be accessed, which is what is done in this Python testbench.
+In white box testing, in contrast, every detail of the internal structure is known.
+Thus, any single internal signal can be accessed, which is what is done in this Python testbench.
 The advantage of the white box is, that these internals can be tested.
 However, any time the implementation changes, even if it is just the naming of a signal, the testbench code needs to be adapted as well.
-In black box testing, the testbench code is independent of the detailed implementation of the DUT, as long as it maintains its required behaviour at the exposed interfaces.
-However, the disadvantage is that internal states cannot be observed.\\
+In black box testing, the testbench code is independent of the detailed implementation of the DUT, as long as it maintains its defined behaviour at the exposed interfaces.
+However, the disadvantage is that internal states can not be observed directly.\\
 
-The Python testbench that is implemented combines the advantages of both worlds and uses black box testing, as well as white box testing.
+The Python testbench that is implemented, combines the advantages of both worlds and uses black box testing, as well as white box testing.
 
@@
 The reason therefor is, that it is the most complex unit within the entire design and is used in multiple instances.
-All the other design units mostly consist of simple delay lines, additions or substractions and thus a special unit test is not implemented.
+All the other design units mostly consist of simple delay lines, additions or substractions and thus, special unit tests are not implemented there.
 
@@
 Ideally, this situation should be handled by the ready signal, since this is exactly what it is intended for.
-However, in order to simplify the implementation a strobe generator is used, which is a counter that generates a repetitive pulse with a defined period.
-This pulse, in combination with some additional logic, is used as the ready signal, to guarantee enough spacing between two consecutive input samples.
+However, in order to simplify the implementation, a strobe generator is used, which is a counter that generates a repetitive pulse with a defined period.
+This pulse, in combination with some additional logic, is used as the ready signal, to guarantee enough spacing between two consecutive input samples and therefore limit the input dataflow.
 
@@
 The VHDL code for this register interface is generated by a Python script which is named \textit{Register Engine}.
-The registers can be defined in a YAML-file.
+The registers can be defined in a YAML file.
 Additionally, documentation in the form of a Markdown language file is generated, and -- more importantly -- a C/\cplusplus\ code header, which defines all the register addresses and can be used in the application software.\\

### 2021-08-21 doc/thesis/chapters/implementation.tex
impl done

@@
 A testbench is equally important in the HLS implementation, as it is for the VHDL implementation.
-The HDL code for the IP is written in \cplusplus\ and thus the testbench is also written in this language.
-The execution of the testbench is done in the Xilinx HLS software.
+The HDL code for the IP is written in \cplusplus\ and thus, the testbench is also written in this language.
+The execution of the testbench is done in the Xilinx Vivado HLS software.
 A detailed explanation of how an HLS testbench works in general is given in Chapter~\ref{cha:HLS}.\\
@@
 The actual implementation to write the HLS signal data to a file, is implemented as a file-writer \cplusplus\ class within the HLS design code.
-These code parts are only evaluated during simulation, which is achieved by using pre-processor defines.
-The following code shows an example of this file-writer class.\\
+These code parts are only compiled for simulation, which is achieved by using pre-processor defines.
+The following code shows an example of the usage of this file-writer class.\\
 
@@
 
-The testbench is especially designed around the \textit{Design For Reuse} segment of the SIMILAR concept, which is introduced in Section~\ref{sec:SimilarProcess}.
+The testbench is especially designed around the \textit{Design For Reuse} segment of the \mbox{SIMILAR} concept, which is introduced in Section~\ref{sec:SimilarProcess}.
 This is the case because of two major parts.
@@
 
-\includepicture [0.8] [0] [H] {Overview of the testbench and the reuse of the application software.
+\includepicture [0.8] [0] [H] {Overview of the testbench and the reuse of the application software in testbench and hardware.
 The HLS IP is displayed with dotted lines, to point out that it is the same in testbench and hardware.} {testbench_hls_overview_reuse} {img/draw.io/testbench_hls_overview_reuse}
@@
 The streams use the types \texttt{iq\_sample\_t} and \texttt{audio\_sample\_t}, which are structs of two 16-bit fixed point variables and thus represent a width of 32 bit.
-The configuration and status registers are also implemented as structs, that contain multiple variables to represent the separate registers.
+The configuration- and status registers are also implemented as structs, that contain multiple variables to represent the separate registers.
 For the registers, a special detail is the distinction between the register types read-write and read-only.
@@
 The \texttt{bundle} keyword in the pragma-directives combine the two register groups into a single AXI4-Lite bus interface, instead of creating one bus each.
-This may be of advantage in the application software later, since it only needs to handle a single base address.
-The interface for the LEDs that can be connected is implemented as a data-only type, which does not have any protocol assigned.
+This is of advantage in the application software later, since it only needs to handle a single base address.
+The interface for the LEDs that can be connected, is implemented as a data-only type, which does not have any protocol assigned.
 It simply represents the data value that is written to the \cplusplus\ variable.
@@
 In the VHDL implementation, this is implemented with a counter that counts the number of incoming samples.
-If the counter reaches the number that matches the downsampling rate, is forwards the current sample.
+If the counter reaches the number that matches the downsampling rate, it forwards the current sample.
 Otherwise, the incoming samples are discarded.\\

### 2021-08-22 doc/thesis/chapters/results.tex
results begin

@@
 
-This chapter presents the results that are achieved in the development of the FM Radio Receiver project, with the chosen different methods.
-However, the main focus is to compare the results that are achieved with HLS and VHDL, since these are the major implementation variants of the project.
+This chapter presents the results that are achieved in the development of the FM radio receiver project, with the chosen implementation methods.
+The main focus is to compare the results that are achieved specifically with HLS and VHDL, since these are the major implementation variants of the project.
 
@@
 
-A main objective of this thesis is to implement an FM radio receiver in multiple different methods.
+The main objective of this thesis is to implement an FM radio receiver in multiple different methods, within a common system design.
 This includes GNU Radio, Matlab, VHDL and HLS, which all have their advantages and disadvantages and surface different challenges to the developer.
@@
 However, the level of detail that is required in the implementation, and the resulting effort and time it takes to implement the respective variant, differs by a large factor.
-This is mainly due to the different levels of abstraction, so that the low-level algorithms do not need to be known.\\
+This is mainly due to the different levels of abstraction, so that the low-level algorithms do not always need to be known into detail.\\
 
@@
 Generally, both tools -- HLS and VHDL -- provide the capabilities to implement any functionality in one or the other method.
+In other words, anything that is implemented in HLS, can also be implemented in VHDL directly.
 However, the implementation is done in a different level.
@@
 
-The implementation is split into two main parts, the communication interfaces and the DSP.
+The implementation of the FM receiver is split into two main parts, the communication interfaces and the DSP chain.
 The results are presented in the following sections.
-
-% Functionality same?
-% DSP yes, AXI no (FSM of axi-stream)
-% unit tests
 
@@
 The AXI4-Lite memory-mapped bus is implemented to have an exactly equal behaviour in both variants.
-There are read-only and read-writeable registers, which are all mapped onto a single base address.\\
+There are read-only and read-writeable registers, which are all mapped into a common register-map, that starts from a single base address, as they are accessible through this bus interface.\\
 
-The AXI stream interface however does not show the exact same behaviour.
+The AXI stream interface, however, does not show the exact same behaviour.
 Here, the HLS variant is using the AXI stream according to its protocol, while the VHDL variant is implemented differently.
@@
 The audio output of the HLS and VHDL variants is compared with the Matlab model, which serves as the reference.
-Additionally, the results of the testbench are compared with their respective result in the actual hardware.
-In summary, it can be stated that the DSP chain produces a very similar audio output in all the compared data sets.
+This includes the comparison with the testbench simulation, as well as the results from the execution on actual hardware.
+In summary it can be stated, that the DSP chain produces a very similar audio output in all the compared data sets.
 However, differences remain, which are presented in the following paragraphs and diagrams.
@@
 %%%%%
-\subsubsection{Comparing against the Matlab reference}
+\subsubsection{Comparison against the Matlab reference}
 
@@
 There seems to be an issue in the signal separation between the left and right audio channels.
-This is visible in the analysis of the audio signal, as shown is Figure~\ref{fig:audio_output_compare_tb_vs_hw}.
+This is visible in the analysis of the audio signal, as shown in Figure~\ref{fig:audio_output_compare_tb_vs_hw}.
 The signal strength of the left channel in the upper diagram is weaker than expected, while the right channel in the lower diagram is stronger.
@@
 The cause for this issue may be located in several components in the design.
-This includes the FM demodulator, the carrier recovery, i.e. the 38~kHz carrier, as well as a sample timing shift in the final summation and substraction to recover the left and right channels.
+This includes the carrier recovery, i.e. the 38~kHz subcarrier, as well as a sample time-shift in the final summation and substraction to recover the left and right channels.
 The FIR filter has a successfully passing unit test and is thus assumed to be correct.
@@
 Due to the time limitation in the elaboration of this thesis, this issue can not be traced down to the root cause and thus still persists in the current implementation.
-However, from the system-level point of view of this thesis, the IP works and can be integrated into the final system.
+However, from the system-level point of view of this thesis, the IP works and can be integrated into the final system successfully.
 
@@
 %%%%%
-\subsubsection{Comparing simulation- against hardware results}
+\subsubsection{Comparison of simulation- against hardware results}
 
-Here, the results of the simulation testbenches are compared against the values that are read from the FPGA directly.
+The results of the simulation testbenches are compared against the values that are read from the FPGA directly.
 Ideally, the values should be exactly the same.
-However, Figure~\ref{fig:audio_output_compare_tb_vs_hw} shows that this is not always the case.
-Again, the HLS implementation does have a matching result, whereas the VHDL variant has deviating values.
-The explanation here may be linked with above suspicions, but may also be linked to the initial reset values that exist in the hardware.
+However, this is not always the case, as Figure~\ref{fig:audio_output_compare_tb_vs_hw} demonstrates.
+The HLS implementation does have a matching result between simulation and hardware, whereas the VHDL variant has deviating values.
+The explanation here may be linked with the previously mentioned suspicions, but may also be linked to the initial reset values that exist in the hardware.
 In the HLS implementation, the reset logic of several registers is implemented automatically, whereas in VHDL these conditions need to be implemented manually.
-Therefore, the IPs internal state, i.e. the register values, may be different at the beginning, which leads to an error propagation throughout the design.
+Therefore, the IPs' internal state, i.e. the register values, may be different at the beginning, which leads to an error propagation throughout the design.
 The entire DSP chain is built like a pipeline and therefor it takes a number of samples to process, before the wrong, intermediate values are flushed out.
@@
 However, the end-to-end system design enables the developer to analyze an issue like this in a convenient way, in quick iterations.
-Any adaptions in the VHDL design can be simulated and integrated into the FPGA with automated scripts.
-Both results can then be analyzed and compared, which may lead to further adaptions.\\
+Any adaptions in the VHDL design can be simulated in the comprehensive testbench, and subsequently be integrated into the FPGA.
+All these processes are automated with scripts, so that, besides the actual work of interest in the HDL design, no manual actions are necessary.
+The new results can then be analyzed and compared again, which may lead to further adaptions and eventually resolve any existing issue.
 
@@
 It is a subjective metric though, since it is heavily influenced by the respective developer's experience in the specific application.
-However, both metrics are applied to the FM Radio Receiver project.\\
+However, both metrics are applied to the FM radio receiver project.\\
 
@@
 Therefore, the numbers that are presented here may vary widely by modifying any of these factors, i.e. using a different testbench framework for VHDL.
-Thus, a generalization in the art of \textit{HLS always requires less code than VHDL} can not be stated.
-However, the results in this section can give an idea and overview of this topic and provide a detailed comparison for this specific project.
+Thus, a generalization in a manner of \textit{HLS always requires less code than VHDL} can not be stated.
+However, the results in this section can give an idea and overview of this topic and provide a detailed and valid comparison for this specific project.
 
@@
 
-Several parts of the system design are taken into account to analyze the lines of code, to give an impression over the largest code parts in the project.
+Several parts of the system design are taken into account for the analysis of the lines of code, to give an impression over the largest code parts in the project.
 Figure~\ref{fig:lines_of_code_pie_charts} shows multiple pie charts, which represent the proportions of the lines of code in the respective system design parts.
-Table~\ref{tab:lines_of_code} displays the actual numbers of code lines, to quantify the size of the respective parts.\\
+Table~\ref{tab:lines_of_code} displays the actual numbers of code lines, to quantify the size of these respective parts.\\
 
@@
 Specifically, the code proportions between IP design and testbench are pointed out.
-The charts show that HLS requires about the same amount of code for IP design and testbench.
-VHDL requires the largest part of code in the IP design, while the testbench code amount is relatively small.
+The charts show, that HLS requires about the same amount of code for IP design and testbench.
+VHDL on the other hand, requires the largest part of code in the IP design, while the testbench code amount is relatively small.
 It is to be noted, that the common testbench consists of 4\%, which equals 630 lines of Python code.
@@
   \centering
-  \subfloat[\centering Overview of the proportions between all the implemented code parts of the entire project.]{{\includegraphics[trim=10 40 10 10,clip, width=0.65\textwidth]{img/matlab/lines_of_code_pie_chart_py_all} }}%
+  \subfloat[\centering Overview of the proportions between all the implemented code parts of the entire project.]{{\includegraphics[trim=10 40 10 10,clip, width=0.68\textwidth]{img/matlab/lines_of_code_pie_chart_py_all} }}%
   \\
   %\hspace{\fill}
-  \subfloat[\centering HLS in detail.]{{\includegraphics[trim=10 10 10 10,clip, width=0.49\textwidth]{img/matlab/lines_of_code_pie_chart_py_hls} }}%
+  \subfloat[\centering HLS in detail.]{{\includegraphics[trim=50 10 15 10,clip, width=0.49\textwidth]{img/matlab/lines_of_code_pie_chart_py_hls} }}%
   %\hspace{\fill}
   %\vspace{3cm}
-  \subfloat[\centering VHDL in detail.]{{\includegraphics[trim=10 10 10 10,clip, width=0.49\textwidth]{img/matlab/lines_of_code_pie_chart_py_vhdl} }}%
+  \subfloat[\centering VHDL in detail.]{{\includegraphics[trim=50 10 15 10,clip, width=0.49\textwidth]{img/matlab/lines_of_code_pie_chart_py_vhdl} }}%
 
-  \caption{Lines of code of the respective code parts in the entire project, presented in multiple pie charts.
-The percentages of code in IP design and testbench for the HLS and VHDL implementations are shown in detail in the lower diagrams.}%
+  \caption{Lines of code of the respective parts in the entire project, presented in multiple pie charts.
+The percentages of code in IP design and testbench for the HLS and VHDL implementations are presented in detail in the lower diagrams, where the legends show the respective implementation languages of the sub-parts in the inner circles.}%
   \label{fig:lines_of_code_pie_charts}%
@@
 The metric of implementation time -- the time it takes to implement a specific feature -- is a subjective measure, since it heavily depends on the respective developer's experience, as already noted above.
-In the present case, there is a strong background of knowledge and practical experience in VHDL design, while there is only very little of such in HLS design.
-However, this metric is used in this section to specifically compare HLS with VHDL in this project.
-Certain features are chosen for this comparison and their implementation times were logged during the actual implementation process.
-The implementation is always considered as the HDL description, including the corresponding testbench code, to verify the respective correct functionality.
-Table~\ref{tab:implementation_time_logging} shows the direct comparison between HLS and VHDL implementation times for these features.\\
+In the present case, there is a strong background of knowledge and practical experience in VHDL design, while there is only very little thereof in HLS design.
+Thus, the logged HLS times also include the time that is spent in research, which is required to implement a specific feature.
+However, this metric is used in this section to specifically compare HLS with VHDL for this project.
+Certain features are chosen for this comparison and their implementation times are logged during the actual implementation process.
+The implementation is always considered as the HDL description, including the corresponding testbench code, to verify the respective correct functionality.\\
+
+Table~\ref{tab:implementation_time_logging} shows the direct comparison between HLS and VHDL implementation times for these features.
+The following paragraphs present an insight into how these implementation times of the listed code parts arise.
 
@@
 \end{table}
-
-The following paragraphs present an insight into how the implementation times of the listed code parts arise.
 
@@
 With regards to latency, the two implementation variants appear to have a similar behaviour.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{GNU Radio}
+
+This section is specifically (TODO)

### 2021-08-22 doc/thesis/chapters/results.tex
res: interfaces and sim time

@@
 \subsection{Lines of Code}
+\label{sec:results:lines-of-code}
 
@@
 This is the main interface for the data input and -output of the DSP chain.
-In HLS it can be automatically inferred by adding pragma directives to the respective function parameter variables.
-Also the testbench code is implemented like a regular function call.
+In HLS, it can be automatically inferred by adding pragma directives to the respective function parameter variables, and the testbench code is implemented like a regular function call.
 Therefore, the actual interface logic is completely hidden from the developer.
@@
 In the time comparison, the larger amount of manual implementation in VHDL is clearly visible.
-It is to note that this comparison does not compare functionally equivalent implementations, because the actually implemented logic in VHDL does not use the ready-flag correctly, as explained in the previous chapters.
-According to this fact, a further increase in implementation hours should actually be considered for VHDL.
-However, also the HLS implementation has its particular characteristics that need to be understood for a fully functional implementation.
+It is to note, that this comparison does not compare functionally equivalent implementations though, because the actually implemented logic in VHDL does not use the ready-flag correctly, as explained in the previous chapters.
+According to this fact, an even further increase in implementation hours should actually be considered for VHDL.
+However, also the HLS implementation has its particular characteristics that need to be understood for a fully functional implementation, which requires time in research.
 Following the respective user guides and implementation examples provides a good starting point for that.
@@
 Similar to the previous AXI Stream interface, also this AXI4-Lite memory-mapped interface is automatically generated by HLS, while this has to be done manually in VHDL.
-To create this process more efficiently and extendable, a Python script is used, which can also automatically generate VHDL code for the registers' memory map.
+To create this process more efficiently and extendable, a Python script -- the so-called Register Engine -- is used, which can also automatically generate VHDL code for the registers' memory map.
 This turns out to be a multiplicator in productivity for the implementation of registers.
-The register generator script was previously used in projects and is therefor already well-known and tested before it is used in this project.
+The register generator script was already used in previous projects and is therefore well-known and tested, before it is used in this project.
 However, the VHDL code only represents one part of the interface -- the second part is the firmware driver that is used in the application software.
-Even though the register generator script also generates a \cplusplus\ header file that includes several address offset definitions of the registers, the actual driver functions need to be implemented manually.
-Several files need to be integrated into a pre-defined folder structure, in order to enable a correct integration of the IP in a Vivado block diagram.\\
+Even though the register generator script already generates a \cplusplus\ header file, which includes several address offset definitions of the registers, the actual driver functions need to be implemented manually.
+Furthermore, several files need to be integrated into a pre-defined folder structure that Vivado understands, in order to enable a correct integration of the IP in a Vivado block diagram.\\
 
@@
 The latter holds information about the build date and -time of the IPs, and is particularly time-consuming because of the implementation of the respective drivers to convert the date format in both variants.
-In contrast, the addition of the LED control and DSP mode registers is simple and fast, because of the auto-generated code in HLS and VHDL with usage of the register engine.
+In contrast, the addition of the LED control and DSP mode registers is simple and fast, because of the auto-generated code in HLS and VHDL, with usage of the register engine.
 The single bits in the LED control register directly represent the LEDs' values on the hardware.
-The DSP mode register can select the regular FM radio mode and a pass-through mode, which simply forwards the streaming data without any processing.
+The DSP mode register can select between the regular FM radio mode and a pass-through mode, which simply forwards the streaming data without any processing.\\
+
 Summing up, the addition of registers to the design can be time-consuming in both variants, depending on the required driver functionality.
-However, the general tendency is that VHDL requires more time, because of this driver issue.
+However, the general tendency is, that VHDL requires more time, because of this driver issue.
 
@@
 
-For the \cplusplus\ version, a code example for an FIR filter, which is provided by Xilinx HLS, is taken and modified for the present purpose.
+For the HLS version, a \cplusplus\ code example of an FIR filter, which is provided by Xilinx HLS, is taken and modified for the present purpose.
 The VHDL variant is taken from an older DSP project, which was previously developed during a university course.
-However, the integration and testbench code is developed for this project, specifically.
+However, the integration- and testbench code is developed for this project, specifically.
 Here, the HLS implementation took significantly longer, because of the limited knowledge regarding the usage of an AXI stream interface.
 It is the first time in the project, that this type of interface is applied in the IP design and used in the testbench.
-The VHDL implementation was straightforward in comparison.
+The VHDL implementation is straightforward, in comparison.
 The testbenches compare the FIR filters' output with the Matlab model's pilot tone output.
@@
 This is mainly due to the limited knowledge of the functionality of the AXI stream interface at this stage in the project's implementation, just as previously mentioned with the FIR filter.
-The initial version was implemented like the VHDL version -- a counter that only triggers the subsequent DSP blocks upon a certain number of samples was received.
+The initial version was implemented like the VHDL version -- a counter that only triggers the subsequent DSP blocks upon a certain number of samples is received.
 However, this led to an endlessly blocking AXI stream in the testbench execution.
@@
 Because the sample decimator did not write a sample every time, the subsequent read-function blocked indefinitely.
-A thorough investigation in several user guides, in combination with step-by-step debugging, the issue was found and led to the current, working implementation, which processes a number of samples, before it forwards one sample.
+By a thorough investigation in several user guides, in combination with step-by-step debugging, the issue was found and led to the current, working implementation, which processes a number of samples, before it forwards one sample.
 The process of finding this issue explains the largely varying implementation times, as listed in Table~\ref{tab:implementation_time_logging}.
@@
 
-The implementation of the receiver structure requires the combination of several previously implemented blocks, such as the FIR filter and any interfaces.
+The implementation of the receiver structure requires the combination of several previously implemented blocks, such as the FIR filter instances and any interfaces.
 This is done according to the Matlab model, to resemble its functionality.
 The VHDL version is significantly more time consuming in comparison to HLS.
-However, it is to note at this point that the VHDL version is implemented first, and thus some issues are already solved by the time of the HLS implementation.
+However, it is to note at this point, that the VHDL version is implemented first, and thus some issues are already solved by the time of the HLS implementation.
 Nevertheless, the VHDL implementation takes longer, as it requires more detailed care in the signal level.
 This specifically includes the clock-alignment of any signals that need to be used in a combination.
-For example, if two parallel FIR filters produce output signals that are to be added, they need to be aligned so that they are valid at the time of the summation.
+For example, if two parallel FIR filters, which have a different filter order and thus, a different group delay, produce output signals that are to be added, they need to be aligned, so that they are valid at the time of the summation.
 This often requires analysis and debugging in a signal wave form.
 In HLS this is not required, since the validity of a signal is determined internally by the HLS compiler, according to the usage of the respective data variables.
-Another factor that adds up to the implementation time is the number of lines of code, as presented in Table~\ref{tab:lines_of_code}.
+Another factor that adds up to the implementation time, is the number of lines of code, as presented in Section~\ref{sec:results:lines-of-code}.
 The VHDL IP design requires 3000 lines, while the HLS IP design only takes 920 for the same functionality.
@@
 
-In the final summation over several implementation times of all the listed features, the two variants HLS and VHDL only differ by 10 hours, which is around 15\%.
+In the final summation over several implementation times of all the listed features, the two variants HLS and VHDL only differ by 10 hours, which is about 15\%.
 However, as the previous paragraphs explain, prior knowledge regarding HLS was limited and had to be gathered during the ongoing implementation.
-Nevertheless, even consiering this fact, the HLS implementation is faster than the VHDL equivalent.
+Nevertheless, even considering this fact, the HLS implementation is faster than the VHDL equivalent.
 Thus, with an equal amount of prior knowledge and experience, it is assumed that an HLS implementation can lead to a successfully functioning result significantly faster.
@@
 The run-phase then actually applies this test data to the DUT and stores the respective output results.
-All time measurements are taken without the result analysis, since this is the exact same, shared code between all variants and thus takes the same amount of time everywhere.
+All time measurements are taken without the result analysis, since this is the exact same, shared code between all variants and thus, takes the same amount of time everywhere.
 Several measurements are done multiple times, to achieve an average result, that is independent of the current processor load of the host PC.\\
@@
 The diagram shows an immense difference of simulation speed between the variants.
-It is important to note that the y-axis has a logarithmic scale, as a method to enable the presentation of the values in multiple orders of magnitude.
+It is important to note, that the y-axis has a logarithmic scale, as a method to enable the presentation of the values in multiple orders of magnitude in a single diagram.
 VHDL only achieves around 20 samples per second.
-This is by far the slowest method, which is expected because of the way of how VHDL is simulated.
+This is by far the slowest method, which is expected, because of the way of how VHDL is simulated.
 This can not be explained in detail here, but one main feature of a VHDL simulator is its capability to simulate parallel processes, as they occur in actual parallel hardware.
 The VHDL simulator therefor needs to evaluate several concurrent processes and statements, whenever any of the interacting signals change their value, and this is time-consuming.
-In the current case, there is an additional interface between the VHDL simulator and the Python testbench framework, which likely additionally slows down the simulation.
-The HLS testbench in comparison achieves around 50000 samples per second, which is a speed-up factor of 2500.
-This is specifically talking about the HLS C/\cplusplus-simulation.
-The main reason for the speed-up therefor is, that in this type of simulation no timing behaviour is simulated -- any processes are treated as C/\cplusplus\ functions that runs like a regular executable on the host CPU.
-Also, the HLS compiler is using native \cplusplus\ data types to represent the signals, where possible, which results in the faster execution of the code.
+In the current case, there is also an additional interface between the VHDL simulator and the Python testbench framework, which likely slows down the simulation even further.
+The HLS testbench in comparison achieves around 50.000 samples per second, which is a speed-up factor of 2.500.
+This is specifically talking about the HLS C/\cplusplus\ simulation.
+The main reason for the speed-up is, that this type of simulation does not simulate any timing behaviour -- any processes are treated as C/\cplusplus\ functions, which run like a regular executable on the host CPU.
+Also, the HLS compiler is, where possible, using native \cplusplus\ data types to represent the signals, which results in the faster execution of the code.
 Finally, the simulation in Matlab achieves a further speed-up by a factor of 70.
-Matlab is specifically designed for applications which require matrix calculations and utilizes highly optimized algorithms in its underlying functions therefor.
-Thus, Matlab runs very efficiently and achieves the highest simulation speed of around 3 million samples per second.
+Matlab is specifically designed for applications, which require matrix calculations and utilizes highly optimized algorithms in its underlying functions, therefor.
+Consequently, Matlab runs very efficiently and achieves the highest simulation speed of around 3~million samples per second.
 

### 2021-08-22 doc/thesis/front/abstract.tex
abstract

@@
 
-This should be a 1-page (maximum) summary of your work in English.
+This thesis is elaborated as the final project in the master's degree program Embedded Systems Design, at the University of Applied Sciences, Campus Hagenberg.
+A comprehensive practical project is developed accompanying to the theoretical topics, which are elaborated in this thesis.\\
+
+The main topic of this thesis is the development of a system design for an embedded system, which combines all the knowledge of different disciplines and topics, that are accumulated over the course of the entire university program.
+A system-level design approach is taken to develop an FM radio receiver in multiple implementation variants, targeting a system-on-chip hardware, including an FPGA and a CPU.\\
+
+Specifically, high-level synthesis and manual VHDL implementation is used, to develop two versions of the FPGA design.
+This includes the digital signal processing chain, as well as multiple interfaces to communicate with the design.
+The two variants are compared, based on specific metrics, such as implementation time, lines of code, hardware utilization, and the general usability experience during the development process.
+At the end, a productive system is deployed on actual hardware, including a firmware application to provide a basic user interface.\\
+
+The entire development process is determined by the general point of view of system design.
+Therefore, a system model is developed in the first stage, by using Matlab.
+Furthermore, a prototype, which can be deployed on actual hardware, is established with GNU Radio.
+The entire system, with its two main implementation methods and the Matlab system model, is using a common verification environment, which guarantees a correctly functioning design.
+This verification is performed in software and hardware, which closes the loop of the system design approach.

### 2021-08-22 doc/thesis/front/abstract.tex
kurzfassung

@@
 The entire development process is determined by the general point of view of system design.
-Therefore, a system model is developed in the first stage, by using Matlab.
-Furthermore, a prototype, which can be deployed on actual hardware, is established with GNU Radio.
+Therefore, a system model is developed in the first stage, by using Matlab, which allows the elaboration of algorithms for digital signal processing and the creation of a system, which can be implemented in hardware.
+Furthermore, GNU Radio is used to establish a prototype, which can be deployed on actual hardware.
 The entire system, with its two main implementation methods and the Matlab system model, is using a common verification environment, which guarantees a correctly functioning design.
-This verification is performed in software and hardware, which closes the loop of the system design approach.
+This verification is performed in simulation and on the actual hardware, which closes the loop of the system design approach.

### 2021-08-22 doc/thesis/front/kurzfassung.tex
kurzfassung

@@
 \begin{german}
-An dieser Stelle steht eine Zusammenfassung der Arbeit, Umfang
-max.\ 1 Seite. 
-...
+
+Diese Masterarbeit wurde als Abschlussarbeit im Rahmen des Master Studiengangs Embedded Systems Design an der Fachhochschule Oberösterreich, Campus Hagenberg ausgearbeitet.
+Begleitend zu den theoretischen Themen dieser Arbeit wurde ein umfangreiches, praktisches Projekt entwickelt.\\
+
+Das Hauptthema dieser Arbeit ist es, ein System-Design für ein Embedded System zu entwickeln, welches alles Wissen über sämtliche Disziplinen und Themen umfasst, die im Rahmen des Studiums behandelt wurden.
+Mit einem Entwurfsansatz auf System-Ebene wird in mehreren Implementierungsvarianten ein FM Radio Empfänger entwickelt.
+Zielarchitekture ist ein System-on-Chip, das aus FPGA und CPU besteht.\\
+
+Im Speziellen werden zwei Varianten für die Implementierung eingesetzt -- High-Level Synthese und die manuelle Implementierung von VHDL code.
+Das betrifft hauptsächlich die digitale Signalverarbeitungskette, sowie mehrere Schnittstellen, anhand welcher der Hardware Entwurf mit seiner Umgebung kommunizieren kann.
+Die beiden Varianten werden anhand verschiedener Kennzahlen verglichen.
+Beispielsweise werden hierfür die Zeitdauer der Implementierung, die Anzahl der Programmzeilen, die Effizienz in der Nutzung der Hardware, sowie der generelle Eindruck bezüglich der Benutzbarkeit während der Entwicklung, herangezogen.
+Am Ende wird ein produktives System auf echter Hardware erstellt, inklusive einer Firmware Anwendung, um die Interaktion mit einem Benutzer zu ermöglichen.\\
+
+Der gesamte Entwickungsprozess wird immer in dem Gesichtspunkt des Systementwurfs betrachtet.
+Deshalb wird im ersten Schritt ein Systemmodell mit Matlab entwickelt.
+Dieses erlaubt es, Algorithmen für die digitale Signalverarbeitung, und generell ein System zu erarbeiten, welches später in Hardware implementiert werden kann.
+Weiters wird GNU Radio dazu verwendet, um einen Prototypen zu entwickeln, der bereits auf echter Hardware lauffähig ist.
+Das gesamte System, inklusive der beiden Implementierungsvarianten und dem Matlab Systemmodell, verwendet eine gemeinsame Strategie und Umgebung zur Verifikation der einwandfreien Funktion des Hardwareentwurfs.
+Diese Verifikation wird in Simulation, als auch direkt in der Hardware durchgeführt, wodurch sich der Kreis des Ansatzes als Systementwurf schließt.
+
 \end{german}

### 2021-08-22 doc/thesis/chapters/dsp.tex
implementation begin
move around newlines to avoid single lines

@@
 The spectrum consists of multiple parts.
-Therefore, it is also referred to as `multiplex signal' in the literature.
+Therefore, it is also referred to as \textit{multiplex signal} in the literature.
 Several of these parts are described in the following paragraphs, according to \cite{EbuFmRegulators},\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard} and \cite{ElectronicsNotesFmTutorial}.
@@
 It is used to generate a stereo audio signal, in combination with the mono signal.
-This process is explained into more detail in Chapter~\ref{sec:demodulation-of-broadcast-fm}.
+This process is explained into more detail in Chapter~\ref{sec:demodulation-of-a-broadcast-fm-channel}.
 
@@
 
-The method implemented in Figure~\ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as `Asynchronous Half-Wave Envelope Detector' \cite{DSPRelatedEnvelopeDetection}.
+The method implemented in Figure~\ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as \textit{Asynchronous Half-Wave Envelope Detector} \cite{DSPRelatedEnvelopeDetection}.
 It consists of a thresholding unit to remove negative values and a conventional lowpass filter.
@@
 Different methods may be implemented for the envelope detector.
-For example, an `Asynchronous Full-Wave Envelope Detector' can be used.
+For example, an \textit{Asynchronous Full-Wave Envelope Detector} can be used.
 Therefor, the previous method only needs to swap the thresholding unit with an unit that calculates the absolut value.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Demodulation of Broadcast FM}
-\label{sec:demodulation-of-broadcast-fm}
+\section{Demodulation of a Broadcast FM Channel}
+\label{sec:demodulation-of-a-broadcast-fm-channel}
 
@@
 \subsection{Stereo Audio}
+\label{subsec:demod_stereo_audio}
 

### 2021-08-22 doc/thesis/chapters/implementation.tex
move around newlines to avoid single lines

@@
       The processes are implemented in the Matlab model, but are disabled in the generation of the input file that is used for the system.
-      Thus, a de-emphasis filter is not implemented in the hardware, as it is not required for the show-case of this thesis.\\
+      Thus, a de-emphasis filter is not implemented in the hardware, as it is not required for the show-case of this thesis.
 
@@
           Thus, they need to be aligned in time at that point.
-          Consequently, a delay block needs to be inserted in the \textit{Mono} branch, to compensate the group delay of the BP in the \textit{LR Diff} branch.
+          Consequently, a delay block needs to be inserted in the \textit{Mono} branch, to compensate the group delay of the BP in the \textit{LR Diff} branch.\\
 
@@
 
-      To overcome the group delay issue at this point in the actual hardware implementation, the filters are manually adapted to have a matching order.\\
+      To overcome the group delay issue at this point in the actual hardware implementation, the filters are manually adapted to have a matching order.
+
   \item \textbf{Carrier Recovery}\\
@@
       It is important that the recovered carrier is phase-coherent with the received input, as described in Section~\ref{subsec:demod_stereo_audio}.
-      This can be achieved, by simply multiplying the pilot tone with itself, which is the way it is implemented here.\\
+      This can be achieved, by simply multiplying the pilot tone with itself, which is the way it is implemented here.
 \end{itemize}
@@
 Thus, the Mono signal, as well as the Pilot Tone need to be delayed by the amount of samples defined by the BP filter's group delay.
-The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone \cite{FmMultiplexingForStereo}.\\
+The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone \cite{FmMultiplexingForStereo}.
 
@@
 Summing up, the Matlab model is not implemented in fixed point arithmetic, but is implemented in a way, that is close enough to the hardware implementation.
-Therefore, it can serve as a reference for the hardware implementation.
+Therefore, it can serve as a reference for the hardware implementation.\\
 
@@
 The Matlab implementation also includes a comprehensive range of analysis functionalities, especially in terms of visual plots.
-These plots are useful during the development of new DSP algorithms, in order to verify the correct functionality thereof.\\
-
+These plots are useful during the development of new DSP algorithms, in order to verify the correct functionality thereof.
 An example of a plot that is created by the analysis is depicted in Figure~\ref{fig:matlab_analysis_time_domain}.
@@
 However, the result shows a clear separation of the left and right channels, and resembles the original input very closely, which is good enough for this case study.\\
-
-\includepicture [1.0] [0] {Matlab analysis in the time domain.
-The stereo audio signal of the voice saying \textit{Left channel, Right channel} can be recognized in the signal, which allows an audible and visual verification of the receiver's functionality.} {matlab_analysis_time_domain} {img/matlab/matlab_analysis_time_domain}
-%TODO: latex font for this diagram!
 
@@
 The artifact is very low in energy, because it is attenuated by a BP filter in the down-conversion process, and can thus be ignored for further processing, even though it can still be seen in the spectral analysis.
+
+\includepicture [1.0] [0] {Matlab analysis in the time domain.
+The stereo audio signal of the voice saying \textit{Left channel, Right channel} can be recognized in the signal, which allows an audible and visual verification of the receiver's functionality.} {matlab_analysis_time_domain} {img/matlab/matlab_analysis_time_domain}
+%TODO: latex font for this diagram!
 

### 2021-08-22 sim/matlab/fm_analysis.m
matlab latex font in diagram

@@ -122,6 +122,8 @@ end
 
 %% Plots
 
+set(0,'defaulttextinterpreter','latex');
+
 disp('-- Plots');
 
 fig_title = 'Time domain signal';
@@ -131,30 +133,30 @@ ax2='';
 if EnableSenderSourceCreateSim
     ax1 = subplot(6,1,1);
     plot(tn/fs, audioDataL, 'r', 'DisplayName', 'audioDataL');
-    grid on; legend();
+    grid on; legend('Interpreter','latex');
     ax2 = subplot(6,1,2);
     plot(tn/fs, audioDataR, 'g', 'DisplayName', 'audioDataR');
-    grid on; legend();
+    grid on; legend('Interpreter','latex');
 end
 ax3 = subplot(6,1,3);
 hold on;
 plot(tnAudio/fs_audio, rx_audio_lrdiff, 'b', 'DisplayName', 'rx\_audio\_lrdiff');
 plot(tnAudio/fs_audio, rx_audio_mono, 'r', 'DisplayName', 'rx\_audio\_mono');
 ylabel('amplitude');
-grid on; legend();
+grid on; legend('Interpreter','latex');
 ax4 = subplot(6,1,4);
 plot(tnAudio/fs_audio, rx_audio_mono, 'b', 'DisplayName', 'rx\_audio\_mono');
-grid on; legend();
+grid on; legend('Interpreter','latex');
 ax5 = subplot(6,1,5);
 plot(tnAudio/fs_audio, rx_audio_L, 'r', 'DisplayName', 'rx\_audio\_L');
 ymax = max(rx_audio_L);
 ylim([-ymax,ymax]);
-grid on; legend();
+grid on; legend('Interpreter','latex');
 ax6 = subplot(6,1,6);
 plot(tnAudio/fs_audio, rx_audio_R, 'g', 'DisplayName', 'rx\_audio\_R');
 ylim([-ymax,ymax]);
 xlabel('time [s]');
-grid on; legend();
+grid on; legend('Interpreter','latex');
 linkaxes([ax1,ax2,ax3,ax4,ax5,ax6],'x');
 title(ax1,fig_title);
 if EnableSavePlotsToPng
@@ -257,22 +259,22 @@ fig_tx_spec = figure('Name',fig_title);
 ax1 = subplot(2,1,1);
 title(fig_title);
 hold on; grid on;
-xline(19e3,'k--','19 kHz');
-xline(38e3,'k--','38 kHz');
[…]

### 2021-08-22 doc/thesis/back/supplementary-materials.tex
Thesis chapters text. (#30)

\chapter{Supplementary Materials}
\label{app:materials}

A list of supplementary data, that is included with this thesis.

\section{PDF Files}
\begin{FileList}{/}
\fitem{thesis.pdf} Master thesis
\end{FileList}

\section{FM Radio Receiver Project}
\begin{FileList}{/fm-radio}
\fitem{/doc} Documentation, Analysis tools and this thesis document.
\fitem{/hardware} Hardware development, HLS and VHDL.
\fitem{/sim} Simulation, including GNU Radio and Matlab.
\end{FileList}

\vspace{5 mm}
\noindent
The FM Radio Receiver project can also be found on GitHub:\\
\url{https://github.com/wurmmi/fm-radio}

### 2021-08-22 doc/thesis/chapters/conclusion-and-prospect.tex
Thesis chapters text. (#30)

@@
 
-This thesis covers a comprehensive variety of topics, that ranges from digital signal processing to system modeling and -design, IP design in different methods, verification, to hardware-software interaction in hardware.
-The implementation of the FM radio receiver in multiple different methods, namely GNU Radio, Matlab, VHDL and HLS so that they achieve the same functionality, is a challenge, especially with the latter methods.
+This thesis covers a comprehensive variety of topics, including digital signal processing, system modeling and system design, IP design in different methods, verification, and hardware-software interaction in an actual hardware environment.
+The implementation of the FM radio receiver in multiple different methods, namely GNU Radio, Matlab, VHDL and HLS, so that they all achieve the same functionality, is a challenge, especially with the latter methods.
 However, this challenge is solved within an end-to-end system design, that strongly supports the development process.\\
@@
 The usage of Matlab as a tool to research DSP algorithms and strategies to demodulate a signal, in order to ultimately find a system model that can be implemented in hardware, turns out to be a powerful asset.
-It allows the development in fast iterations, through short simulation times, to achieve a result quickly.
+Through short simulation times, it allows the development in fast iterations, which helps to achieve a result quickly.
 Based on this system model, the implementation of actual hardware IPs, with VHDL and HLS can be done more efficiently and in a direct way, because issues in the algorithm level are already fixed in the model at this point.
-In combination with the simulation testbenches that share major parts, such as the input source, as well as a the analysis tool, the IP development works smoothly and efficiently.
+In combination with the simulation testbenches, which share major parts, such as the input source, as well as the analysis tool, the IP development works smoothly and efficiently.
 The verification as a comparison against the Matlab system model gives instant feedback of whether the IP functions correctly.
-Even if the system model changes, the testbench automatically adapts, since the input source files are re-generated by the model and thus picked up at the next iteration of the testbench simulation, so the hardware engineer can also adapt the IP code.
-Once the IP is deployed on actual hardware, it is tested by the firmware, in order to check its correct functionality again, which closes the loop from the system model and the IP simulation to the hardware deployment.
+Even if the system model changes, the testbench automatically adapts, since the input source files are re-generated by the model.
+These files are picked up at the next execution of the testbench simulation.
+Consequently, the hardware engineer can also adapt the IP code accordingly.
+Once the IP is deployed on actual hardware, it is tested by the firmware application, in order to check its correct functionality again, which closes the loop from the system model and the IP simulation to the hardware deployment.
 This entire process proved itself to be a successful approach.\\
 
-Regarding the implementation of the VHDL and HLS IPs and the accompanying comparison, the result turns out to be unexpected.
-The initial expectation, that HLS is inefficient in terms of resource utilization, is contradicted.
-Also the amount of time spent in the implementation and the metric of lines of code create a positive picture of HLS.
+Regarding the implementation of the FM receiver in VHDL and HLS, and their subsequent comparison, the result turns out to be unexpected.
+The initial expectation, that HLS may be inefficient in terms of resource utilization, is contradicted.
+Also the relatively low amount of time that is required for the implementation, as well as the metric of lines of code, create a positive impression of HLS.
 Further, the automatic generation of several files that are required to integrate the IP into a system design, such as interface definitions or software drivers, support this attitude.
 VHDL may have its advantages in specific areas of application, but HLS is definitely a powerful tool that can achieve great results in less time.
-This is especially true, when it comes to the development of complex functionalities, such as image- or video processing, with the support of built-in libraries like OpenCV.\\
+This is especially true, when it comes to the development of complex functionalities, such as image- or video processing, with the support of built-in libraries like OpenCV.
+Furthermore, the project presents a show-case, how HLS enables the concurrent development of a firmware application, which can already interface with the IP and get tested, before any actual hardware is available.\\
 
 However, the often heard statement that HLS enables any software developer to develop hardware IPs, can not be supported.
-A developer that writes HLS code, still needs to have a strong hardware background, to be able to understand what is being generated in hardware.
+A developer that writes HLS code, still needs to have a strong hardware background, to be able to understand what is being generated from HLS code in hardware.
 It is important to know about the limitations of the target hardware, i.e. FPGA, in order to write an efficient design.
-Therefore, the connection between specific HLS \cplusplus\ code lines, as well as the general structure and algorithm, and the hardware target need to be understood.\\
+Therefore, the connection between specific HLS \cplusplus\ code lines, as well as the code's general structure and algorithm, and the hardware target need to be understood.\\
 
@@
 Looking at the HLS code, it could be evaluated how compiler-independent it is.
-The question could be, if it is possible to write code that can be compiled with multiple compilers for different target platforms and vendors, e.g. Xilinx and Intel FPGAs, at the same time.
-Furthermore, the various optimization strategies that Xilinx HLS provides could be evaluated on the existing HLS code.
-Also, currently manually implemented functions, such as the FIR filter, could be replaced with build-in library functions.
+The question could be, if it is possible to write code that can be compiled with multiple HLS compilers for different target platforms and vendors, e.g. Xilinx and Intel FPGAs, at the same time.
+Furthermore, the various optimization strategies that Xilinx HLS provides, could be evaluated on the existing HLS code.
+Also, the currently manually implemented functions, such as the FIR filter, could be replaced with built-in library functions.
 Finally, the application software could be upgraded from the current bare-metal, real-time OS, to an embedded Linux OS which runs on the ARM CPU.
@@
 The amount of ideas for potential extensions to the project shows how comprehensive the project is and how many topics it covers.
-Summing up, it can be said that the project revealed a range of interesting and partially unexpected insights into multiple topics, and that a lot of knowledge was gained during the elaboration of this thesis.
+Summing up, it can be said, that the project revealed a range of very interesting and partially unexpected insights into multiple topics, and that a lot of knowledge was gained during the elaboration of this thesis.
 

### 2021-08-22 doc/thesis/chapters/deployment-on-hardware.tex
more of impl section
Thesis chapters text. (#30)

@@
 
-This chapter describes the additional parts of implementation that are necessary in order to run the implemented FM radio on actual hardware.
+This chapter describes the additional parts of implementation that are necessary, in order to run the implemented FM radio receiver on actual hardware.
 
@@
 \section{GNU Radio}
+\label{sec:deployment-on-hw:gnu-radio-devices}
 
 The GNU Radio implementations, both receiver and transmitter, are used in combination with external hardware.
-The specific hardware that is used is briefly described here.\\
-
-In order to interface with the real world, two different hardware devices are used.
-Both devices are supported by GNU Radio and are briefly described here.
+Two different hardware devices are used to interface with the real world, which are both supported by GNU Radio directly.
+The specific devices are briefly described here.
 
 %%%%%%%%%%%%%%%%%%%%
-\subsection{Ettus Research USRP b200mini}
+\subsection{Ettus Research USRP B200mini}
 
-The USRP b200mini is a device that can be used to create a software-defined radio.
+The USRP B200mini is a device that can be used to create a software-defined radio.
 It resembles the size of a business card and is developed by Ettus Research.
@@
 
-The RTL-SDR is a cheap USB dongle, that was originally developed as a DVB-T TV tuner.
+The RTL-SDR is a low-cost USB dongle, that was originally developed as a DVB-T TV tuner.
 It is based on the RTL2832U chipset, which includes an 8-bit ADC and a digital signal processor.
@@
 The ZedBoard is a development board that is developed by Xilinx.
-It is specifically designed around the Xilinx Zynq-7000 SoC, which contains a dual-core ARM Cortex A9 processor and a programmable logic fabric, the FPGA.
-The board further offers a range of features from external DDR3 memory, over various connectors, such as RJ45 Ethernet, USB, 3.5 mm audio, HDMI, Pmod, to buttons, LEDs and a display.\\
+It is specifically designed around the Xilinx Zynq-7000 SoC, which contains a dual-core ARM Cortex~A9 processor and a programmable logic fabric, the FPGA.
+The board further offers a range of features like external DDR3 memory, various connectors, such as RJ45 ethernet, USB, 3.5~mm audio in- and output, HDMI, Pmod, as well as user interface hardware such as buttons, LEDs and a display.\\
 
-The most important features for this project are the audio line-out connector and the SD card port.
-Additionally, the LEDs are used to display various status information during runtime of the FM receiver.
+The most important features for this project are the 3.5~mm audio line-out connector and the SD card port.
+Additionally, the LEDs are used to display various status information during the runtime of the FM receiver.
 
@@
 
-The implemented version of the FM Radio Receiver on the ZedBoard is a working prototype at this point.
+The implemented version of the FM radio receiver on the ZedBoard is a working prototype at this point.
 The implementation does not support live data streaming.
 Instead, antenna data is recorded to a file with the RTL-SDR hardware, using a regular PC.
-The antenna data is recorded as IQ data in the baseband.
+The data is recorded as IQ data in the baseband.
 The file is then stored onto an SD card, that is plugged into the ZedBoard.
-The CPU on the SoC then reads it and replays the antenna data to the FPGA.
+The CPU on the SoC reads it and replays the antenna data to the FPGA.
 In the future, this design may be extended to use an embedded Linux operating system.
@@
 The block diagram in Figure~\ref{fig:bd_impl_vivado} shows the implemented architecture that is used in the SoC.
-The various design domains in and around the SoC are depicted in different visual representations.
+The various design domains in, and around the SoC, are depicted in different visual representations.
 The FPGA domain blocks are highlighted in gray, with the main IPs being surrounded by emphasized lines.
 The processor system domain is shown as white blocks.
-Blocks that are external to the SoC are drawn with dashed lines.
+Blocks, which are external to the SoC, are drawn with dashed lines.
 
@@
 The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input stream.
-One of these duplicates is sourced to an AXI-Stream-To-I2S converter, which generates the I2S signals for the external audio codec chip.
+One of these duplicates is sourced to an AXI-Stream-To-I2S Converter, which generates the I2S signals for the external audio codec chip.
 The second duplicate is connected to a FIFO, which simply stores the generated output data.
 Once the FIFO is almost full, it asserts an interrupt to the CPU, which triggers it to read out the entire FIFO buffer.
-The CPU then stores this data on the external SD card again.
-In that way, the loop of data is closed, which enables the verification of the respective IPs' correct functionality.
+The CPU then stores this data on the external SD card.
+In that way, the loop of data is closed, which enables the verification of the respective IP's correct functionality.
 Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.
@@
 
-The application software is running on a single core of the dual-core ARM Cortex A9 CPU and is built upon a real-time operating system, namely FreeRTOS.
+The application software is running on a single core of the dual-core ARM Cortex~A9 CPU and is built upon a real-time operating system, namely FreeRTOS.
 Xilinx provides a ported version of this RTOS with the installation of their software development kit, the Xilinx SDK.
@@
 
-The respective software drivers for several IPs that are used in the FPGA design are automatically imported into the SDK project.
-They are embedded in the so-called hardware definition file (HDF) that is generated by Vivado after the synthesis process, and provided together with the bitstream file.
-It is a file with the extension \textit{*.hdf} and can be opened like a regular archive file, with any ZIP archive viewer.\\
+The respective software drivers for several IPs, which are used in the FPGA design, are automatically imported into the SDK project.
+They are embedded in the so-called hardware definition file (HDF) that is generated by Vivado after the synthesis process, and is provided together with the bitstream file.
+It is a file with the extension \textit{.hdf} and can be opened like a regular archive file, with any archive viewer.\\
 
@@
 The main tasks of the application are to configure several IPs in the FPGA, so that they can fulfill their respective purpose in the system, and to read the recorded data from the SD card, which is then streamed to the FPGA.
-It also provides a basic user interface over the serial console, with which a user can interact with the FM Radio Receiver product.\\
-
-However, software design is not the main focus of this thesis, and thus no further details are discussed here.
+It also provides a basic user interface over the serial console, which a user can use to interact with the FM Radio Receiver product.
+However, software design is not the main focus of this thesis, and thus, no further details are discussed here.

### 2021-08-22 doc/thesis/chapters/hls.tex
Thesis chapters text. (#30)

@@
   In this first step, the above explained, regular C simulation is executed.
-  Hereby, the inputs that are sent to the DUT are stored as \textit{input vectors}.
+  Hereby, the inputs that are sent to the DUT are stored as \textit{input vectors}.\\
+
   \item \textbf{Perform RTL simulation}\\
@@
   However, various other simulators like ModelSim are supported.
-  The generated outputs of this step are stored as \textit{output vectors}.
+  The generated outputs of this step are stored as \textit{output vectors}.\\
+
   \item \textbf{Verify the results}\\
@@
 \noindent
-This type of simulation takes a significantly longer time to execute, because timing information now actually needs to be considered for the RTL code, which is done by using an HDL simulator.
+The execution of this type of simulation takes significantly longer, because timing information now actually needs to be considered for the RTL code, which is done by using an HDL simulator.

### 2021-08-22 doc/thesis/chapters/results.tex
Thesis chapters text. (#30)

@@
 
-The system design architecture that is chosen in this project covers several parts of the system.
+The system design architecture, that is chosen in this project, covers several parts of the system.
 This includes a reference model, the development of the IP itself, including the simulation testbench, as well as the final integration into the productive system.
-Because of this modular and structured approach, that is given by this system design, it can be applied to almost any DSP project.
+Because of this modular and structured approach, given by this system design, it can be applied to almost any DSP project.
 The main processing IP can simply be replaced with another implementation, while the environment does not need to be modified.
 The reason therefor is, that DSP projects almost always use streaming interfaces for their inputs and outputs.
-Also a configuration interface is commonly used, to adapt certain parameters of the IP during runtime.
+Also, a configuration interface is commonly used, to adapt certain parameters of the IP during runtime.
 
@@
 The aim is to create IP blocks, that can be integrated into a block design in the Xilinx Vivado software.
-The IP blocks should be able to be connected with their AXI interfaces, in order to successfully communicate and interact with the embedded hard-processor, the ARM CPU.
+These IP blocks should be able to be connected with their AXI interfaces, in order to successfully communicate and interact with the embedded hard-processor, the ARM CPU, as well as other IPs that exist in the design.
 In the case of HLS, several files that are required for this integration are generated automatically.
@@
 There is an assistant, the IP Packager, which helps the developer to create the IP.
-However, at this point the folder structure already needs to be in place, so that the assistant works correctly, which again requires manual action.
+However, at this point, the folder structure already needs to be in place, so that the assistant works correctly, which again requires manual action.
 Regarding the top-level interfaces, the IP Packager supports an auto-detection of the signals that belong to the respective interface types, as long as they follow the expected naming guidelines.
@@
 The reason therefor was, that the top-level function of the IP did not have the exact same naming as the Vivado HLS project.
-For issues like that, the Eclipse-based SDK does not give any error output or any kind of information.
+This is a non-documented requirement, and furthermore, for issues like that, the Eclipse-based SDK does not give any error output or any kind of meaningful information.
 This however is a general issue with the SDK, where error handling is poor and error log outputs are not existent, or at least not useful to the developer, so that the SDK sometimes crashes without an obvious reason.
@@
 
-However, both IPs could successfully be integrated into the Vivado block design, and the tests on the hardware show that they function correctly within the system.
+However, both IPs could successfully be integrated into the Vivado block design in the end, and the tests on the hardware show that they function correctly within the system.
 
@@
 
-The hardware utilization of the HLS and VHDL implementations are compared in detail in this section.
-The numbers are taken from the final utilization report, that is created by Vivado at the end of the bitstream creation, since this is assumed to be the most accurate report that shows the actual utilization.\\
+The hardware utilization of the HLS and VHDL implementation is compared in detail in this section.
+The numbers are taken from the final utilization report, that is created by Vivado at the end of the bitstream creation, i.e.
+post-synthesis, since this is assumed to be the most accurate report, which shows the actual utilization, as it is being deployed to hardware.
 
 Figure~\ref{fig:hardware_utilization} shows the utilization numbers of the different FPGA resources.
-It reveals that the manual VHDL implementation clearly requires more resources in the FPGA.
-The study of the detailed utilization report however does not enable a deeper direct comparison of the respective single blocks, such as VHDL FIR filter against HLS FIR filter.
-This is because the HLS compiler automatically applies optimizations, so that the direct connection between code and resource report is partially lost.
+It reveals, that the manual VHDL implementation clearly requires more resources in the FPGA.
+The study of the detailed utilization report, however, does not enable a deeper direct comparison of the respective single blocks, such as VHDL FIR filter versus HLS FIR filter.
+This is because the HLS compiler performs some kind of name mangling, which re-names functions, and automatically applies optimizations, so that the direct connection between code and resource report is partially lost.
 Examples for optimizations are inlining of functions, re-ordering of operations and time-sharing.
 The latter has an especially large impact on the resource utilization.
-Generally, in hardware design there is a trade-off between resource utilization and throughput, which leads to time-sharing.
+Generally in hardware design, there is always a trade-off between resource utilization and throughput, which ultimately leads to time-sharing.
 For example, an entirely parallel-duplicated pipeline can process the double amount of data, but also requires the double amount of resources.
 On the other hand, time-sharing a resource may take the double amount of time, thus only processes half of the data, but also only requires half of the resource utilization.
-In HLS, this trade-off can be influenced and determined by the developer.
-Depending on the requirements of the design, the optimization may either be set towards a lower resource utilization, or towards a high throughput.
+In HLS, this trade-off can be influenced and determined by the developer, by the usage of optimization directives.
+Depending on the requirements of the design, the optimization may either be set towards a low resource utilization, or towards a high throughput.
 In the current HLS implementation, no measures are taken towards any optimization and thus, the compiler is free to decide.
@@
 
-Nevertheless, the resource report allows the comparison of the configuration interface, which is the AXI4-Lite interface.
-In VHDL, it requires 44 LUTs and 54 FFs, while HLS consumes only 36 LUTs and 37 FFs.
-So it seems that the manual VHDL implementation requires more resources in all the comparable parts.
+\includepicture [0.75] [0, trim=10 10 10 10, clip] {A comparison of the resource utilization of the VHDL and HLS implementations.
+The VHDL variant requires significantly more resources.} {hardware_utilization} {img/matlab/hardware_utilization}
+
+Nevertheless, the resource report allows the direct comparison of one specific unit, the configuration interface, which is the AXI4-Lite interface.
+In VHDL, it requires 44~LUTs and 54~FFs, while HLS only consumes 36~LUTs and 37~FFs.
+It seems, that the manual VHDL implementation requires more resources in all the comparable parts.
 By further investigating the report, the VHDL version apparently utilizes a large amount of LUTs for the delays, which are used to clock-align the FIR output signals.
@@
 However, for this project, both implementation variants are compared without any optimizations from the developer's side.
-
-\includepicture [0.75] [0, trim=10 10 10 10, clip] {A comparison of the resource utilization of the VHDL and HLS implementations.
-The VHDL variant requires significantly more resources.} {hardware_utilization} {img/matlab/hardware_utilization}
 
@@
 The report shows an estimated resource utilization, as well as the latency as a number of clock cycles.
-The latency is given with 2712 clock cycles.
+The latency is given with 2712~clock~cycles.
 For the VHDL version, a similar report is not available.
 Instead, the testbench of the integration test is executed, which produces a signal wave form.
-The latency is measured from this wave form and results in a number of 2470 clock cycles.
+The latency is measured from this wave form and results in a number of 2470~clock~cycles.
 With regards to latency, the two implementation variants appear to have a similar behaviour.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{GNU Radio}
+%\subsection{GNU Radio}
 
-This section is specifically (TODO)
+%This section is specifically (TODO)
+% - show that the transmitter works (show photo of radio display in my car)
+% - show that the receiver works (screenshot of decoded RDS data, etc.)

### 2021-08-22 doc/thesis/front/kurzfassung.tex
Thesis chapters text. (#30)

@@
 
-Diese Masterarbeit wurde als Abschlussarbeit im Rahmen des Master Studiengangs Embedded Systems Design an der Fachhochschule Oberösterreich, Campus Hagenberg ausgearbeitet.
+Diese Masterarbeit wurde als Abschlussarbeit im Rahmen des Masterstudiengangs Embedded Systems Design an der Fachhochschule Oberösterreich, Campus Hagenberg ausgearbeitet.
 Begleitend zu den theoretischen Themen dieser Arbeit wurde ein umfangreiches, praktisches Projekt entwickelt.\\
 
-Das Hauptthema dieser Arbeit ist es, ein System-Design für ein Embedded System zu entwickeln, welches alles Wissen über sämtliche Disziplinen und Themen umfasst, die im Rahmen des Studiums behandelt wurden.
+Das Hauptthema dieser Arbeit ist es, ein System-Design für ein Embedded System zu entwickeln, welches sämtliches Wissen über alle Disziplinen und Themen umfasst, die im Rahmen des Studiums behandelt wurden.
 Mit einem Entwurfsansatz auf System-Ebene wird in mehreren Implementierungsvarianten ein FM Radio Empfänger entwickelt.
-Zielarchitekture ist ein System-on-Chip, das aus FPGA und CPU besteht.\\
+Zielarchitektur ist ein System-on-Chip, das aus FPGA und CPU besteht.\\
 
-Im Speziellen werden zwei Varianten für die Implementierung eingesetzt -- High-Level Synthese und die manuelle Implementierung von VHDL code.
-Das betrifft hauptsächlich die digitale Signalverarbeitungskette, sowie mehrere Schnittstellen, anhand welcher der Hardware Entwurf mit seiner Umgebung kommunizieren kann.
+Im Speziellen werden zwei Varianten für die Implementierung eingesetzt -- High-Level Synthese und die manuelle Implementierung von VHDL Code.
+Dies betrifft hauptsächlich die digitale Signalverarbeitungskette, sowie mehrere Schnittstellen, anhand welcher der Hardwareentwurf mit seiner Umgebung kommunizieren kann.
 Die beiden Varianten werden anhand verschiedener Kennzahlen verglichen.
@@
 Deshalb wird im ersten Schritt ein Systemmodell mit Matlab entwickelt.
-Dieses erlaubt es, Algorithmen für die digitale Signalverarbeitung, und generell ein System zu erarbeiten, welches später in Hardware implementiert werden kann.
+Dieses erlaubt es, Algorithmen für die digitale Signalverarbeitung, und generell ein Gesamtsystem, zu erarbeiten, welches später in Hardware implementiert werden kann.
 Weiters wird GNU Radio dazu verwendet, um einen Prototypen zu entwickeln, der bereits auf echter Hardware lauffähig ist.

### 2021-08-22 doc/analysis/hardware-utilization/hardware_utilization.py

@@ -1,0 +1,52 @@
+# -------------------------------------------------------------------------
+# File        : hardware_utilization.py
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Create a diagram to represent the hardware utilization.
+# -------------------------------------------------------------------------
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.sans-serif": ["Helvetica"],
+    "font.size": 16,
+})
+
+# Data
+labels = ['LUT', 'FF', 'LUTRAM', 'SRL', 'RAMB18', 'DSP48']
+values_vhdl = [1203, 1361, 176, 0, 0, 8]
+values_hls = [650, 494, 0, 56, 7, 6]
+
+barwidth = 0.4
+
+# the label locations
+x1 = np.arange(len(labels)) * 1.5
+x2 = [x + barwidth for x in x1]
+
+# Plot
+fig, ax = plt.subplots()
+fig.set_size_inches(8, 5)
+ax.spines['top'].set_visible(False)
+ax.spines['right'].set_visible(False)
+
+rects1 = ax.bar(x1, values_vhdl, barwidth, label='VHDL', color=plt.cm.Greens(0.6), edgecolor='black')
+rects2 = ax.bar(x2, values_hls, barwidth, label='HLS', color=plt.cm.Blues(0.6), edgecolor='black')
+
+# Add some text for labels, title and custom x-axis tick labels, etc.
+# ax.set_ylabel('Scores')
+ax.set_title('Hardware Utilization')
+plt.xticks([r + barwidth / 2 for r in x1], labels)
+ax.set_xticklabels(labels)
+ax.legend(loc='right')
+#ax.legend(loc=(1.0, 0.2))
+#ax.legend(loc='center left', bbox_to_anchor=(1.0, 0.5))
+
+ax.bar_label(rects1, padding=3)
+ax.bar_label(rects2, padding=3)
+
+# show and save
+plt.tight_layout()
+plt.savefig('../../thesis/img/matlab/hardware_utilization.pdf')
+plt.show()

### 2021-08-22 doc/analysis/implementation-time/impl_time_analysis.py

@@ -1,0 +1,60 @@
+# -------------------------------------------------------------------------
+# File        : impl_time_analysis.py
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Create a diagram to represent the implementation time.
+# -------------------------------------------------------------------------
+
+import matplotlib.pyplot as plt
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.sans-serif": ["Helvetica"],
+    "font.size": 16,
+})
+
+variant = ['VHDL', 'HLS', 'Matlab']
+samples_per_second = [18, 45000, 3100000]
+
+x_pos = [i for i, _ in enumerate(variant)]
+
+fig = plt.figure()
+fig.set_size_inches(6, 5)
+ax = fig.add_subplot(1, 1, 1)
+
+ax.set_yscale('log')
+
+ax.spines['top'].set_visible(False)
+ax.spines['right'].set_visible(False)
+
+plt.bar(x_pos, samples_per_second, width=0.5, edgecolor='black')
+plt.xlabel("Implementation Variant")
+#plt.ylabel("Samples per second")
+plt.title("Processed Samples Per Second")
+
+plt.xticks(x_pos, variant)
+
+plt.annotate('x2500',
+             xy=(0.5, 10000),
+             xycoords='data',
+             xytext=(-70, -40),
+             textcoords='offset points',
+             fontsize=16,
+             horizontalalignment='left',
+             arrowprops=dict(arrowstyle='->', connectionstyle='arc3,rad=-0.2')
+             )
+plt.annotate('x70',
+             xy=(1.6, 1000000),
+             xycoords='data',
+             xytext=(-70, -40),
+             textcoords='offset points',
+             fontsize=16,
+             horizontalalignment='left',
+             arrowprops=dict(arrowstyle='->', connectionstyle='arc3,rad=-0.2')
+             )
+
+
+# show and save
+plt.tight_layout()
+plt.savefig('../../thesis/img/matlab/impl_time_analysis.pdf')
+plt.show()

### 2021-08-22 doc/analysis/lines-of-code/count_lines_of_code.sh

@@ -1,0 +1,152 @@
+#!/usr/bin/env bash
+#-------------------------------------------------------------------------------
+# File        : count_lines_of_code.sh
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Count lines of code in this repository and produce some statistics.
+#-------------------------------------------------------------------------------
+
+set -euo pipefail
+
+SCRIPT_PATH=$(dirname $(readlink -f $0))
+REPO_ROOT=$SCRIPT_PATH/../../../
+
+cd $REPO_ROOT
+
+###
+# NOTE:
+#  Analysis is done with CLOC.
+#  For more information on how to use CLOC, please visit: https://github.com/AlDanial/cloc
+#
+#  Choose a cloc version with the CLOC_CMD.
+#    (1) need to `sudo apt install cloc` (version 1.82 on Ubuntu 20.04)
+#    (2) need to have Docker installed on your system - always uses latest available version
+###
+#CLOC_CMD="cloc"
+CLOC_CMD="docker run --rm -v $REPO_ROOT:/tmp/ aldanial/cloc:latest"
+
+# Clean previous output products
+SCRIPT_PATH_RELATIVE=doc/analysis/lines-of-code/output
+mkdir -p $SCRIPT_PATH_RELATIVE
+rm -f $SCRIPT_PATH_RELATIVE/*.txt
+
+#-------------------------------------------------------------------------------
+
+create_report () {
+    $SCRIPT_PATH/get_total_columns.sh $OPTIONS | tee $REPORT_FILE.txt
+    echo -e "\n\n--- Detailed output ---\n" | tee -a $REPORT_FILE.txt
+    $CLOC_CMD $OPTIONS | tee -a $REPORT_FILE.txt
+}
+
+#-------------------------------------------------------------------------------
+# Overall
+#-------------------------------------------------------------------------------
+
+REPORT_FILE=$SCRIPT_PATH_RELATIVE/overall
+OPTIONS="./                                      \
+        --by-file-by-lang                        \
+        --fullpath                               \
+        --not-match-d=(auto-arrange-figs|vivado/ip|fixed_point|gnuradio|bd) \
+        --not-match-f=(RBDSExample.m|fixed_pkg*|fixed_float*|fm_radio_axi|filter_.*_pkg|filter_.*.h|component.xml|sqlite_formatter|.project) \
+        --exclude-ext=mat,cls,sty                \
+        --ignored=/tmp/${REPORT_FILE}_ignored.txt"
+
+create_report;
+
[…]

### 2021-08-22 doc/analysis/lines-of-code/get_total_columns.sh

@@ -1,0 +1,36 @@
+#!/usr/bin/env bash
+#-------------------------------------------------------------------------------
+# File        : get_total_columns.sh
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Custom output formatting of CLOC output.
+#               See https://github.com/AlDanial/cloc/tree/v1.90#wrapping-cloc-in-other-scripts-
+#-------------------------------------------------------------------------------
+
+###
+# NOTE: need to `sudo apt install sqlite3` to use this script.
+###
+
+SCRIPT_PATH=$(dirname $(readlink -f $0))
+REPO_ROOT=$SCRIPT_PATH/../../../
+cd $REPO_ROOT
+
+if test $# -eq 0 ; then
+    echo "Usage: $0  [cloc arguments]"
+    echo "       Run cloc to count lines of code with an additional"
+    echo "       output column for total lines (code+comment+blank)."
+    exit
+fi
+DBFILE=`tempfile`
+cloc --sql 1 --sql-project x $@ | sqlite3 ${DBFILE}
+SQL="select Language, count(File)   as files                       ,
+                      sum(nBlank)   as blank                       ,
+                      sum(nComment) as comment                     ,
+                      sum(nCode)    as code                        ,
+                      sum(nComment)+sum(nCode) as TotalWithoutBlank     ,
+                      sum(nBlank)+sum(nComment)+sum(nCode) as Total
+         from t group by Language order by code desc;
+"
+
+cd $SCRIPT_PATH
+echo ${SQL} | sqlite3 -header ${DBFILE} | ./sqlite_formatter
+rm ${DBFILE}

### 2021-08-22 doc/analysis/lines-of-code/lines_of_code_diagram.m

@@ -1,0 +1,70 @@
+%-------------------------------------------------------------------------
+% File        : lines_of_code_diagram.m
+% Author      : Michael Wurm <wurm.michael95@gmail.com>
+% Description : Create a diagram to represent the lines of code.
+%-------------------------------------------------------------------------
+
+%% Prepare environment
+clear;
+close all;
+clc;
+
+%% Settings
+dir_output_doc = "../../thesis/img/matlab";
+
+
+%% Plot
+
+% details
+%values    = [2730, 2743, 208, 708, 107, 27, 430+489, 224+129, 182, 119, 85];
+%value_txt = {'Matlab','VHDL IP Design (VHDL)', 'VHDL IP Design (Python)', 'VHDL Tb (Py)','VHDL Tb (make)', 'VHDL Tb (shell)', 'HLS IP Design (C++)', 'HLS Tb (C++)','HLS Tb (Tcl)','HLS Tb (make)','HLS Tb (Py)'};
+
+% hls, tb/design
+%values    = [430+489, 224+129+182+119+85];
+%value_txt = {'HLS IP Design', 'HLS Testbench'};
+
+% hls, tb
+%values    = [224+129,182,119,85];
+%value_txt = {'cpp','Tcl','make','Python'};
+
+% vhdl, tb/design
+%values    = [2743+208, 708+107+27];
+%value_txt = {'VHDL IP Design', 'VHDL Testbench'};
+
+% vhdl, tb
+%values    = [708, 107, 27];
+%value_txt = {'Python','make','shell'};
+
+% vhdl, design
+%values    = [2743,208];
+%value_txt = {'vhdl','Python'};
+
+% all, tb/design
+%values    = [2730, 2743+208, 708+107+27, 430+489, 224+129+182+119+85];
+%value_txt = {'Matlab','VHDL IP Design', 'VHDL Testbench', 'HLS IP Design', 'HLS Testbench'};
+
+% all
+%values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85, 490, 2800, 460];
+%value_txt = {'Matlab','VHDL', 'HLS', 'Vivado Scripts', 'C++ Application', 'Common Tb'};
+
+values    = [2730, 2743+208+708+107+27, 430+489+224+129+182+119+85, 2800, 490, 460];
+value_txt = {'Matlab','VHDL', 'HLS', 'C++ Application', 'Vivado Scripts', 'Common Tb'};
+
[…]

### 2021-08-22 doc/analysis/lines-of-code/lines_of_code_diagram.py

@@ -1,0 +1,247 @@
+# -------------------------------------------------------------------------
+# File        : lines_of_code_diagram.py
+# Author      : Michael Wurm <wurm.michael95@gmail.com>
+# Description : Create a diagram to represent the lines of code.
+# -------------------------------------------------------------------------
+
+import matplotlib.pyplot as plt
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.sans-serif": ["Helvetica"],
+    "font.size": 16,
+    #    "legend.fontsize": 10,
+    #    "axes.labelsize": 10,
+})
+
+
+# ------------------------- ALL -----------------------------------------------
+# Create data
+group_names = [
+    '32\%',
+    '23\%',
+    '23\%',
+    '14\%',
+    '4\%',
+    '4\%',
+]
+group_size = [
+    2743 + 208 + 708 + 107 + 27,
+    2730,
+    2800,
+    430 + 489 + 224 + 129 + 182 + 119 + 85,
+    490,
+    460
+]
+group_names_legs = [
+    'VHDL',
+    'Matlab',
+    'C++ Application',
+    'HLS',
+    'Common Testbench',
+    'Vivado Scripts',
+]
+
+# Create colors
+color_blue = plt.cm.Blues
+color_red = plt.cm.Reds
+color_green = plt.cm.Greens
+color_purples = plt.cm.Purples
+color_oranges = plt.cm.Oranges
+color_greys = plt.cm.Greys
+
+# First Ring (outside)
+fig, ax = plt.subplots()
+fig.set_size_inches(6, 4)
+ax.axis('equal')
+mypie, _ = ax.pie(group_size, radius=1.45,
+                  labels=group_names,
+                  labeldistance=0.65,
+                  colors=[
+                      color_blue(0.6),
+                      color_green(0.6),
+                      color_red(0.6),
+                      color_purples(0.6),
+                      color_oranges(0.6),
+                      color_greys(0.4),
+                  ],
+                  textprops=dict(rotation_mode='anchor', va='center', ha='center', size=13),
+                  startangle=10,
+                  # frame=True,
+                  )
+plt.setp(mypie, width=1.45 - 0.45, edgecolor='white')
+# fig.subplots_adjust(0.15, 0.0, 0.6, 1.0)  # left, bottom, right, top
+
+plt.margins(0, 0)
+plt.legend()
[…]

### 2021-08-22 doc/lines-of-code-analysis/count_lines_of_code.sh

@@ -1,152 +1,0 @@
-#!/usr/bin/env bash
-#-------------------------------------------------------------------------------
-# File        : count_lines_of_code.sh
-# Author      : Michael Wurm <wurm.michael95@gmail.com>
-# Description : Count lines of code in this repository and produce some statistics.
-#-------------------------------------------------------------------------------
-
-set -euo pipefail
-
-SCRIPT_PATH=$(dirname $(readlink -f $0))
-REPO_ROOT=$SCRIPT_PATH/../../
-
-cd $REPO_ROOT
-
-###
-# NOTE:
-#  Analysis is done with CLOC.
-#  For more information on how to use CLOC, please visit: https://github.com/AlDanial/cloc
-#
-#  Choose a cloc version with the CLOC_CMD.
-#    (1) need to `sudo apt install cloc` (version 1.82 on Ubuntu 20.04)
-#    (2) need to have Docker installed on your system - always uses latest available version
-###
-#CLOC_CMD="cloc"
-CLOC_CMD="docker run --rm -v $REPO_ROOT:/tmp/ aldanial/cloc:latest"
-
-# Clean previous output products
-SCRIPT_PATH_RELATIVE=doc/lines-of-code-analysis/output
-mkdir -p $SCRIPT_PATH_RELATIVE
-rm -f $SCRIPT_PATH_RELATIVE/*.txt
-
-#-------------------------------------------------------------------------------
-
-create_report () {
-    $SCRIPT_PATH/get_total_columns.sh $OPTIONS | tee $REPORT_FILE.txt
-    echo -e "\n\n--- Detailed output ---\n" | tee -a $REPORT_FILE.txt
-    $CLOC_CMD $OPTIONS | tee -a $REPORT_FILE.txt
-}
-
-#-------------------------------------------------------------------------------
-# Overall
-#-------------------------------------------------------------------------------
-
-REPORT_FILE=$SCRIPT_PATH_RELATIVE/overall
-OPTIONS="./                                      \
-        --by-file-by-lang                        \
-        --fullpath                               \
-        --not-match-d=(auto-arrange-figs|vivado/ip|fixed_point|gnuradio|bd) \
-        --not-match-f=(RBDSExample.m|fixed_pkg*|fixed_float*|fm_radio_axi|filter_.*_pkg|filter_.*.h|component.xml|sqlite_formatter|.project) \
-        --exclude-ext=mat,cls,sty                \
-        --ignored=/tmp/${REPORT_FILE}_ignored.txt"
-
-create_report;
-
[…]

### 2021-08-22 doc/lines-of-code-analysis/get_total_columns.sh

@@ -1,36 +1,0 @@
-#!/usr/bin/env bash
-#-------------------------------------------------------------------------------
-# File        : get_total_columns.sh
-# Author      : Michael Wurm <wurm.michael95@gmail.com>
-# Description : Custom output formatting of CLOC output.
-#               See https://github.com/AlDanial/cloc/tree/v1.90#wrapping-cloc-in-other-scripts-
-#-------------------------------------------------------------------------------
-
-###
-# NOTE: need to `sudo apt install sqlite3` to use this script.
-###
-
-SCRIPT_PATH=$(dirname $(readlink -f $0))
-REPO_ROOT=$SCRIPT_PATH/../../
-cd $REPO_ROOT
-
-if test $# -eq 0 ; then
-    echo "Usage: $0  [cloc arguments]"
-    echo "       Run cloc to count lines of code with an additional"
-    echo "       output column for total lines (code+comment+blank)."
-    exit
-fi
-DBFILE=`tempfile`
-cloc --sql 1 --sql-project x $@ | sqlite3 ${DBFILE}
-SQL="select Language, count(File)   as files                       ,
-                      sum(nBlank)   as blank                       ,
-                      sum(nComment) as comment                     ,
-                      sum(nCode)    as code                        ,
-                      sum(nComment)+sum(nCode) as TotalWithoutBlank     ,
-                      sum(nBlank)+sum(nComment)+sum(nCode) as Total
-         from t group by Language order by code desc;
-"
-
-cd $SCRIPT_PATH
-echo ${SQL} | sqlite3 -header ${DBFILE} | ./sqlite_formatter
-rm ${DBFILE}

### 2021-08-22 hardware/common/analysis/analyze_ip_output_data.m

@@ -12,7 +12,11 @@ close all;
 clc;
 
 %% Settings
-EnableZoomToBegin = false;
+dir_output_doc = "../../../doc/thesis/img/matlab";
+
+EnableSaveToFile                = true;
+EnableZoomToVhdlTimeRange       = true;
+EnableZoomToComparisonTimeRange = false;
 
 %=========================================================================
 %% Read data
@@ -61,7 +65,9 @@ audioDataLeft_VHDL_FPGA  = circshift(audioDataLeft_VHDL_FPGA, 3);
 audioDataRight_VHDL_FPGA = circshift(audioDataRight_VHDL_FPGA,3);
 
 % Plot
-fig_title = 'Audio Output';
+set(0,'defaulttextinterpreter','latex')
+
+fig_title = 'Audio Output Comparison';
 fig_audio_time = figure('Name',fig_title);
 sgtitle(fig_title);
 
@@ -78,8 +84,10 @@ h4 = plot(audioDataLeft_HLS_SIM,  'r--','DisplayName', 'HLS SIM');
 xline(length(audioDataLeft_HLS_FPGA),'k--','FPGA');
 xline(length(audioDataLeft_HLS_SIM), 'k--','HLS SIM');
 xline(length(audioDataLeft_VHDL_SIM),'k--','VHDL SIM');
-grid on; legend([h0,h1,h2,h3,h4],'Location','east');
+grid on; legend([h0,h1,h2,h3,h4],'Location','southeast');
 ylim([ymin,ymax]);
+xlabel('time [sample]');
+ylabel('amplitude');
 
 ax2 = subplot(2,1,2); hold on;
 title('Right channel');
@@ -91,13 +99,32 @@ h4 = plot(audioDataRight_HLS_SIM,  'r--','DisplayName', 'HLS SIM');
 xline(length(audioDataRight_HLS_FPGA),'k--','FPGA');
 xline(length(audioDataRight_HLS_SIM), 'k--','HLS SIM');
 xline(length(audioDataRight_VHDL_SIM),'k--','VHDL SIM');
-grid on; legend([h0,h1,h2,h3,h4],'Location','east');
+grid on; legend([h0,h1,h2,h3,h4],'Location','southeast');
 ylim([ymin,ymax]);
 
 xlabel('time [sample]');
+ylabel('amplitude');
 linkaxes([ax1,ax2],'xy');
-if EnableZoomToBegin
+
+% Adapt figure size
[…]

### 2021-08-22 hardware/hls/src/channel_decoder.cpp

@@ -26,9 +26,9 @@ audio_sample_t channel_decoder(hls::stream<sample_t>& in_sample) {
   sample_t audio_lrdiff;
 
   /** NOTE:
-   *  This loop performs decimation by N.
-   *  --> Processing N samples. Only the last sample is passed on
-   *      to the next processing steps.
+   *  This loop performs decimation by OSR_AUDIO.
+   *  --> Processing OSR_AUDIO samples. Only the last respective
+   *      sample is passed on to the next processing step.
    */
   for (uint32_t i = 0; i < OSR_AUDIO; i++) {
     sample_t sample = in_sample.read();

### 2021-08-22 hardware/hls/src/fm_receiver.cpp

@@ -43,9 +43,14 @@ audio_sample_t fm_receiver(hls::stream<iq_sample_t>& iq_in) {
 #pragma HLS STREAM depth = OSR_AUDIO variable = fm_channel_data
 
   /** NOTE:
-   *  This loop performs decimation by N.
-   *  --> Processing N samples. Only the last sample is passed on
-   *      to the next processing steps.
+   *  This loop performs decimation by OSR_RX.
+   *  --> Processing OSR_RX samples, where only the last sample
+   *      is passed on to the next processing step.
+   *  This is done OSR_AUDIO times, in order to fill the
+   *  FIFO of fm_channel_data.
+   *
+   *  At the end of this nested loop, the fm_channel_data FIFO
+   *  contains OSR_AUDIO number of samples.
    */
   sample_t fm_demod;
   for (uint32_t i = 0; i < OSR_AUDIO; i++) {

### 2021-08-22 hardware/vhdl/ip/tb/integration_test/tests.py

@@ -143,12 +143,12 @@ async def axi_stream_dsp_test(dut):
     #  This speedup factor needs to be chosen carefully. The IP needs to complete the calculation for one sample,
     #  before the next sample can be taken from the input. Therefore, set the factor small, to ensure enough
     #  low-cycles, to allow the IP to complete one entire calculation.
-    output_speedup_factor = 50
+    output_speedup_factor = 120
     strobe_num_cycles_high = 1
     strobe_num_cycles_low = tb.CLOCK_FREQ_MHZ * 1e6 // fm_global.fs_audio_c // output_speedup_factor - strobe_num_cycles_high
     ratio = strobe_num_cycles_low // strobe_num_cycles_high
     tb.backpressure_i2s.start(bit_toggler(repeat(strobe_num_cycles_high), repeat(strobe_num_cycles_low)))
-    assert ratio >= 9, "output_speedup_factor is set too high! --> IP won't have enough time to calculate a sample, before the next one arrives"
+    assert ratio >= 9, f"output_speedup_factor is set too high ({ratio}<9)! --> IP won't have enough time to calculate a sample, before the next one arrives"
 
     # Fork the 'receiving parts'
     fm_demod_output_fork = cocotb.fork(tb.read_fm_demod_output())

### 2021-08-22 sim/matlab/fm_analysis.m

@@ -122,43 +122,49 @@ end
 
 %% Plots
 
+set(0,'defaulttextinterpreter','latex');
+
 disp('-- Plots');
 
 fig_title = 'Time domain signal';
 fig_audio_time = figure('Name',fig_title);
-title(fig_title);
 ax1='';
 ax2='';
 if EnableSenderSourceCreateSim
     ax1 = subplot(6,1,1);
     plot(tn/fs, audioDataL, 'r', 'DisplayName', 'audioDataL');
-    grid on; legend();
+    grid on; legend('Interpreter','latex');
     ax2 = subplot(6,1,2);
     plot(tn/fs, audioDataR, 'g', 'DisplayName', 'audioDataR');
-    grid on; legend();
+    grid on; legend('Interpreter','latex');
 end
 ax3 = subplot(6,1,3);
 hold on;
 plot(tnAudio/fs_audio, rx_audio_lrdiff, 'b', 'DisplayName', 'rx\_audio\_lrdiff');
 plot(tnAudio/fs_audio, rx_audio_mono, 'r', 'DisplayName', 'rx\_audio\_mono');
 ylabel('amplitude');
-grid on; legend();
+grid on; legend('Interpreter','latex');
 ax4 = subplot(6,1,4);
 plot(tnAudio/fs_audio, rx_audio_mono, 'b', 'DisplayName', 'rx\_audio\_mono');
-grid on; legend();
+grid on; legend('Interpreter','latex');
 ax5 = subplot(6,1,5);
 plot(tnAudio/fs_audio, rx_audio_L, 'r', 'DisplayName', 'rx\_audio\_L');
 ymax = max(rx_audio_L);
 ylim([-ymax,ymax]);
-grid on; legend();
+grid on; legend('Interpreter','latex');
 ax6 = subplot(6,1,6);
 plot(tnAudio/fs_audio, rx_audio_R, 'g', 'DisplayName', 'rx\_audio\_R');
 ylim([-ymax,ymax]);
 xlabel('time [s]');
-grid on; legend();
+grid on; legend('Interpreter','latex');
 linkaxes([ax1,ax2,ax3,ax4,ax5,ax6],'x');
+title(ax1,fig_title);
 if EnableSavePlotsToPng
     saveas(fig_audio_time, sprintf("%s%s",dir_output, "time_audio.png"));
+end
+if EnableSavePlotsToPDF
+    fig_audio_time.Position(3:4) = [900 600];
[…]

### 2021-08-22 sim/matlab/fm_transceiver.m

@@ -26,13 +26,15 @@ end
 %% Settings
 
 % Paths
-dir_filters = "./filters/stored/";
-dir_output  = "./matlab_output/";
+dir_filters    = "./filters/stored/";
+dir_output     = "./matlab_output/";
+dir_output_doc = "../../doc/thesis/img/matlab/";
 
 % Simulation options
 EnableWriteDataFiles = true;
 EnablePlots          = true;
 EnableSavePlotsToPng = true;
+EnableSavePlotsToPDF = false;
 
 EnableSenderSourceRecordedFile = false;
 EnableSenderSourceCreateSim    = true;
@@ -82,7 +84,12 @@ fm_sender();
 %% Receiver
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+tic;
+
 fm_receiver();
+
+elapsed_time = toc;
+fprintf('fm_receiver() took %f seconds\n',elapsed_time);
 
 % TODO: implement a version that is actually using 16 bit fixed_point like hardware
 % fm_receiver_fixed_point();

### 2021-09-30 doc/thesis/front/abstract.tex
Incorporate feedback (#32)

@@
 
-The main topic of this thesis is the development of a system design for an embedded system, which combines all the knowledge of different disciplines and topics, that are accumulated over the course of the entire university program.
+The objective of this thesis is the development of a system design for an embedded system.
 A system-level design approach is taken to develop an FM radio receiver in multiple implementation variants, targeting a system-on-chip hardware, including an FPGA and a CPU.\\
